### 2016-02-01 paper/intro.tex
Moved analysis to scripts. Added initial paper draft

\section{Executive Summary}

This document provides a summary of the work done in the first year of the HimL project in the
evaluation work package 5. 
The objectives of this work package are to develop human and automatic accuracy-based evaluation strategies 
for machine translation, and to carefully evaluate the quality and impact of the innovations we deliver to 
NHS24 and Cochrane.

The first section in this report describes the creation of the HimL test set from our use case partners.
We then describe experiments where we apply these test sets to the HimL  machine translation systems,
and we report the first results for our models on in-domain data.

The next section describes the evaluation plan which we developed in close collaboration with NHS24 and Cochrane.
Work package 5 is responsible for the thorough evaluation of translation systems. 
Although automatic metrics give indications of the quality of the translations, they
are not good at capturing accuracy or usefulness. Accuracy needs to be investigated using human evaluation which is 
very time consuming and expensive. The
evaluation plan allows us to map out the schedule of these evaluations 
and defines the  types of evaluation that both the academic
partners and the use case partners will most benefit from.

The last section of this report describes the development of our human semantic evaluation method. 
  This evaluation uses semantic trees on the source sentence and asks
 bilingual annotators to label 
%what 
which
parts of the tree have been correctly translated. 
 In this way we are able to quantify how much of the meaning of the source sentence has been
 retained 
%on the target. 
in the translation.
Having multiple annotators for two language pairs allowed us to
investigate inter-annotator agreement, and results show that this approach is reliable. 
This human evaluation comprises the first year's user acceptance testing.

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|l|l|p{6.5cm}|}
      \hline

\bf{Task} & \bf{Description} & \bf{Planned Schedule}  & \bf{Status } \\
\hline                               
     5.1   & Test corpora for the required language pairs & M1-M6  & Complete  \\ 
     5.2   & Human Semantic MT Evaluation Metric &  M1-M12 & Complete \\ 
     5.3   & Automatic Semantic MT Evaluation Metric & M6-M18  &  Initiated \\ 
     5.4   & User acceptance testing &  M9-M12 annually  &  Year 1 testing completed as part of Task 5.2 \\ 
     5.5   & Evaluation of the impact &  M6-M36 & Planning complete, started data collection  \\

      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Tasks in workpackage 5, their dates and their status as of 31/Jan/2016
}
\label{tab:tasksschedule}
\end{table}

 Table~\ref{tab:tasksschedule} summarises the status of the tasks in the evaluation workpackage. 
We can see that all tasks are currently on schedule and that puts us in a strong position
to succeed in delivering the work promised in year two and three of the HimL project.

### 2016-02-01 paper/main.tex
Moved analysis to scripts. Added initial paper draft

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{times}
\usepackage{latexsym}
\usepackage{graphicx}

\newcommand\BibTeX{B{\sc ib}\TeX}

\title{Human accuracy evaluation using UCCA}

\author{Author 1\\
	    XYZ Company\\
	    111 Anywhere Street\\
	    Mytown, NY 10000, USA\\
	    {\tt author1@xyz.org}
	  \And
	Author 2\\
  	ABC University\\
  	900 Main Street\\
  	Ourcity, PQ, Canada A1A 1T2\\
  {\tt author2@abc.ca}}

\date{}

\begin{document}

\maketitle

\begin{abstract}
 This document contains instructions for preparing ACL 2016 submissions and camera-ready
  manuscripts.  The document itself conforms
  to its own specifications, and is therefore an example of what
  your manuscript should look like.  Papers are required to conform to
  all the directions reported in this document.  
By using the provided \LaTeX\ and \BibTeX\ styles ({\small\tt acl2016.sty}, {\small\tt acl2016.bst}), the required formatting will be enabled by default.
\end{abstract}

\section{ Human Semantic Evaluation}
\label{sec:sem-eval:human}
We focus on producing high accuracy machine translation systems, but common 
automatic MT metrics are not able to directly capture accuracy. Even previously suggested methods
for using humans to evaluate accuracy are highly problematic. We aim to  develop a human evaluation method 
which is reliable and affordable and apply it to the MT prototypes. 
%This 
The
work described
in this section relates to 
%the 
task
\emph{T5.2: Human semantic evaluation}.
% defined in the description of action.

In November 2015, we ran an evaluation task with 6 bilingual annotators, 2 from NHS 24 and 4 from Cochrane. 
We asked them to annotate about 350 sentences translated with the HimL year one systems 
and they had a budget of  up  to 40 hours each to perform this task. 
In this section we motivate and describe the experiment that we ran and we provide an initial analysis of
the results. In  Year 2 and Year 3 we will refine this evaluation task and use it to track the progress of our HimL prototypes.

\subsection{Overview}

Semantic evaluation of machine translation has typically been done at the sentence level and
we propose an  approach which breaks down the evaluation into basic semantic units, making evaluation 
simpler and more consistent. 
Our  assumption is that the semantic structure on the source sentence should be 
retained in the translation, and if it is not, then some essential part of the meaning 
%will be 
is
lost. 
The semantic framework that we base our evaluations on is called 
Universal Conceptual Cognitive Annotation \shortcite{abend-rappoport:2013:ACL2013}.  
UCCA has been developed using linguistic theories about 
what types of components and structures are universal across many different languages.

Our goal is to quantify how much of the meaning of the source sentence is preserved through translation.
There have been many approaches to evaluating the quality of machine translation, but most of them
have asked the annotator to give a score for the entire sentence. There are of course many ways 
that a translation can be incorrect and asking an annotator to provide a global score for a sentence
is a cognitively difficult task even if e.g. limited to a relative comparison
with another candidate translation. How serious is an error? What is the impact of multiple errors on global meaning?
By using UCCA structure to break the evaluation into meaningful components, we provide 
a more consistent and reliable method of evaluating translation accuracy.

The annotation proceeds as follows. Firstly, the source sentences (English, in our case) are annotated with UCCA trees. This
annotation is normally performed by computational linguists, and requires some training in UCCA, but the annotation can be 
reused for different target languages and different MT systems. 
We then create translations of the source sentences with the
MT system, collecting the word alignments from 
%
the
source sentence to 
the
translation provided by the system. These word
alignments are used to project the UCCA annotation from the source sentence to the the translation output, and then bilingual
annotators go through each projected UCCA node, assessing how well it is translated.  
We can estimate the impact of individual errors given their location in the semantic structure 
and we can thus extract a score for the whole sentence. More details on the procedure are provided 
below.

\subsection{Semantic Annotation}

The source sentences in this annotation scheme have been annotated with  a semantic structure defined as
Universal Conceptual Cognitive Annotation (UCCA).  UCCA was developed in the Computational Linguistics Lab of the Computer Science Department of the Hebrew University by Omri Abend and Ari Rappoport.
UCCA views the text as a collection of scenes (or events)
and their inter-relations and participants.

\begin{figure}[ht]
    \includegraphics[width=1\textwidth]{ucca-tree}
    \caption{UCCA Tree with scenes}
    \label{ucca-tree}
\end{figure}

%As you can see 
As can be seen
in Figure~\ref{ucca-tree}, the UCCA annotation results in a tree structure where each leaf is linked to
a word in the sentence at the bottom. A scene must contain a process (P) or a state (S). It can also contain
participants (A) and it can be linked to other scenes by a linker (Linker). Participants, processes and states can be
further analysed into elaborators (E), centres (C) and relators (R). These labels are very 
%high level 
high-level 
and relate to
cognitive concepts which should remain stable across languages.

The fact that in UCCA
the labels are cognitive concepts and that they are linked directly to words
are both advantages 
when considering which semantic formalism is appropriate for machine translation evaluation.

One of such alternative formalisms is
Abstract Meaning Representation 
%%% Macro broken?
%(AMR; \inparcite{banarescu2013amr}). 
\cite{banarescu2013amr}. 
AMR is being actively developed with 
%the view to 
a view towards
using it as a way of generating 
translations but AMR graphs are not aligned to the words in the sentences. Having more abstract semantic structures 
makes the link between source words, target words, and structures more complex and potentially less useful. 
Furthermore, AMR  has been developed mainly with English in mind, and it remains
to be seen how universal AMR graphs are. See
\shortcite{amr:interlingua:lrec:2014} for first observations of divergences
between English 
%vs.
vs.\ 
Chinese and
Czech AMRs.

Another possible semantic framework for this kind of MT evaluation is Semantic
Role Labelling 
%(SRL; \inparcite{palmer2010semantic}).
\cite{palmer2010semantic}.
SRL has been used in a human translation metric called HMEANT~\cite{lo2011structured}. 
HMEANT uses semantic role labels
to measure how much of the “who, why, when,
where” has been preserved in translation. Annotators are instructed to identify verbs as
heads of semantic frames. Then they attach role
fillers to the heads and finally they align heads
and role fillers in the candidate translation with
those in a reference translation. Using SRL for evaluating SMT has a number of
disadvantages as explored by
\shortcite{birch-EtAl:2013:WMT} for German, \shortcite{bojar:wu:ssst:2012} for
Czech and by \shortcite{chuchunkov-tarelkin-galinskaya:2014:SSST-8} for Russian.
The most important drawbacks are as follows:
\begin{itemize}
\item SRL frames are based around a verb which is particularly problematic for
copular verbs and  when verbs are translated correctly as nouns or correctly
omitted (the verb ``to be'' in some Russian constructions).
\item SRL frames do not cover the entire source sentence and the semantic structure is
therefore not completely defined, importantly links between frames are not
considered and prepositional phrases which attach to nouns are not marked.
\item Even considering a limited set of eleven roles (agent, patient,
experiencer, locative etc.) is problematic because we cannot assume that these roles will 
 remain stable across different languages. 
 When looking at an automatically parsed  English-Chinese  parallel corpora,
 it was shown that 8.7\% of the arguments do not preserve their semantic roles~\cite{fung2006automatic}. 
\end{itemize}

UCCA provides universal semantic structures which 
%has 
have
a minimal set of labels. It provides a complete semantic tree which does not rely on syntactic heads and the semantic structure is grounded directly to the words in the sentence. 
 Even though the set of UCCA labels are fairly restricted,  nevertheless they allow us to determine the most important components of the graph (for example it defines centres and linkers which would be likely to carry more weight than elaborators), and we can use this to better calculate the score. 
We think that UCCA is the
most promising representation for evaluating translation.

\subsection{Copied from emails}

UCCA-Eval in contrast to HMEANT:

- needs *very* trained UCCA annotators
- we were not very quick at explaining even the Eval annotations, but I guess we could have done better
+ relies only on one structure
+ relies on automatic alignment (in the form of hints only)
  These two positive design decisions are in line with what Dekai observed for HMEANT (see the abstract of http://repository.ust.hk/ir/Record/1783.1-66107): IAA of SRL drops from 90 to 61 when the two aligned structures are from two different annotators.

+ allows for multi-word and even non-contiguous elements
+ allows to indicate the quality of the predicates ("Actions")
+/- expects to cover the whole hypothesis with annotations, whereas HMEANT 'gives up' when the hypothesis is not intelligible, giving a zero score to the bad frame/whole sentence. This can be a good decision for HMEANT, since our way may artificially boost the chances of understanding the sentence.

HMEAN disadvantages:
. Importantly: it disregards copula clauses, nominalizations (which are often translated to verbal clauses), inter-clause relations (discourse markers, complement clauses etc.). UCCA-Eval addresses these.

Actually, I think this is a +/- . UCCA-Eval only needs the semantic evaluation done once for a given test source, whereas for the HMEANT set-up you have to keep annotating MT output. That's a + for UCCA-Eval. However we need bilingual annotators, which should be a -. Of course we could align translation and reference using monolingual annotators, but then we have the problem of alignment.
In the *MEANT papers they mainly test by either
A) correlating the metric to ranking judgements; or
B) using automated versions of the metric for tuning MT systems

I think neither is particularly convincing, but I'm not sure at the moment what we replace them with. Dekai also made a claim about the *MEANT metrics being "interpretable" in the sense that they could guide you towards semantic errors that the MT system may be making, however I don't recall seeing this done in his papers.

By the claim that HMEANT explains things, Dekai probably means that the score is decomposable, but so is UCCA-Eval and even BLEU. And UCCA-Eval has one benefit over HMEANT: it points to words in the MT output that are aligned with something but do not convey the meaning well (and we know which part of the source meaning). Pure HMEANT might say only that some part of the MT output has no clear frames whatsover (and we do not know which part of the source/reference, unless we add the word alignments).

Differences of our approach:

The annotation experiment that we just conducted differs in two main ways from HMEANT:
1) The semantic framework is different - UCCA vs SRL - and SRL has problems which Omri lists below, some of which Dekai dismissed when they were brought up. Maybe these problems do not matter for MT evaluation or tuning, but I think this would have to be determined empirically.
2) The annotation process is different. We semantically annotate the source, then automatically project it to the translation, and ask bilingual annotators to assess whether this projection shows a good or bad translation. In HMEANT they semantically annotate both the reference and the translation, then align these manually, and assess to what extent they match.

Advatages of our approach:
Here is an additional argument to support the need for target-only intelligibility:
http://www.aclweb.org/anthology/W14-4005
see Table  3 and the text in Section 5.2.
The more frames can people annotate in the hypothesis, the better the hypothesis overall.

\subsection{MT Evaluation Overview}

The basic strategy for annotating the semantics of the machine translations is to step through the source sentence semantic 
components, looking at the translation via the word alignments, and marking on the source structure
%, 
which parts have 
been correctly translated.
There are two kinds of components: a word or basic semantic unit, and a structural component which contains one or more sub-components.

\subsubsection{Lexical nodes}
Lexical nodes are usually comprised of individual words. They are the leaf nodes on the tree, the smallest 
%meaning bearing 
meaning-bearing
units of the tree.
Leaf nodes can be labelled as green (correct), orange (partially correct) and red (incorrect). 
The traffic light system makes the marking of the
lexical units as simple as possible.

\begin{itemize}
\item Green: The meaning of the word or phrase has been largely captured.
\item Orange: the essential meaning has been captured, but some part of the translation is wrong. This could often be due to the translated word having the wrong tense, or the wrong morphology. 
\item Red: The essential meaning of the unit has not been captured.
\end{itemize}

\subsubsection{Structural nodes}

Structural nodes contain other nodes which could be either lexical or structural nodes. 
These nodes are also called parent nodes and they can be labelled as ``Adequate'' or ``Bad''. 
What we are trying to gauge for structural nodes
%, 
is if the children of these nodes
relate to each other in the same way in the source sentence and in the translation. 
There are many ways that the relationship between the children might go wrong in translation:
\begin{itemize}
\item One of the child nodes is missing from the aligned translation. 
\item An extra word or phrase has been inserted into the aligned  translation.
\item The components of the translation are ordered differently 
%to 
from
the source.
\end{itemize}

If any of these changes have occurred and this damages the meaning of the aligned translation, then we mark the 
structural node as ``Bad''. 
However if the arrangement of the child nodes is essentially correct, except that one or
more of the child nodes has themselves been translated wrongly (and so is marked
orange or red), then the structural node should be 
annotated as ``Acceptable''.

\begin{figure}[ht]
    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
	English paraphrase).}
    \label{ucca-tree-mteval}
\end{figure}

%TODO describe figure 2 here

There are cases where we cannot usefully compare individual words in the source to words in the target.
We often need to compare translation at the phrase level. If this is the case, then we can 
 treat the structural node as a lexical unit, and mark it with the traffic light labels. This 
means that none of the children of this node will be examined in order determine the semantic score
for this sentence. 
In Figure~\ref{ucca-tree-mteval}, we can see that the phrase ``became a passion'' has been labelled as a lexical
node and it has been evaluated as ``Orange'', or partially correct, as the translation of this phrase ``loved'' 
only partially capture the semantics of the source.

We are separating lexical and structural evaluation in order to simplify evaluation and to localise errors
to their point of origin. For this reason, if a word has been translated incorrectly,
the parent node should still be correct if the number and relationship between its children are the same
as in the source. 
In Figure~\ref{ucca-tree-mteval}, we can see that even though an important child of the first scene is 
incorrectly translated (``Golf''), and the children of are ordered differently, 
the scene itself is marked as ``Acceptable''. The translation of the scene retains the structural 
relationships between the children.
\\

\newpage

\subsection{Machine Translation Evaluation Tool}

In this section, we will describe the tool that the annotators used to perform the annotation.

\begin{figure}[ht]
    \includegraphics[width=1\textwidth]{interface-annotations}
    \caption{Machine translation evaluation tool}
    \label{mttool}
\end{figure}

In Figure~\ref{mttool} you can see the MT evaluation tool. The sentence at the top shows the complete MT system output. Underneath the MT output is the  source sentence.  Underneath the source sentence we see its expandable semantic tree structure with both lexical and structural nodes. Lexical nodes only have traffic light annotation, whereas a structural node would normally be labelled ``A'' or ``B'' but could also be labelled as a lexical node, in the case where there is no word to word correspondence for the translation of its children.

 The child components of a structural node are marked with different coloured rectangles. When navigating through the different source sentence nodes, one can see the relevant sections of the translation because the aligned words in the translation are highlighted in the complete sentence above. Aligned translations are also shown in black alongside the source node. If the node is aligned to a set of discontiguous words in the translation, then the unaligned words that appear in between the aligned words are shown in red. Even if these words are not directly aligned to the source node, they 
will likely change the meaning of the translation and must be considered when marking 
the translation as correct or not. The alignments are meant just as a guide. The annotator should  look at the complete translation when deciding on the evaluation of a node. The translation could have content added before or after the node which changes its
 structure or meaning. If an extra component is prepended to a structural node, for example, it should be marked as ``Bad''.

%\subsection{Example}
%
%\lexi{TODO write this up}
%
%NODE1. (make sure) (you’ve got a solid chair)
%MT: (stellen Sie sicher, dass) (Sie haben einen festen Stuhl)
%Correct would be: (stellen Sie sicher, dass) (Sie einen festen Stuhl haben)
% 
%NODE2. (you) (’ve) (got) (a solid chair)
%MT: (Sie) (haben) (einen festen Stuhl)
%Correct would be: (Sie) (einen festen Stuhl) (haben)
%
%Node 1 is the parent node of Node 2. The error in MT is that the verb in German "haben" appears in the wrong place, it should be at the end and this makes the translation difficult to understand. The place to mark this error is where haben is a child node, ie. the error occurs in Node 2. This is where the relationship or the order between the child nodes (Sie)(haben) and (einen festen Stuhl) is wrong. The error propagates up to Node 1, but at Node 1 we are evaluating the relationship between (stellen Sie sicher, dass) and (Sie haben einen festen Stuhl) and this is correct and should be marked A.

\newpage

\subsection{Results}

%new_mteval_ro1: Christian
%new_mteval_ro2: Rodica  NHS
%new_mteval_pl1: Malgorzata
%new_mteval_pl2: Iwona (starting from the 29th in the sequence) NHS
%new_mteval_cs1: Martina
%new_mteval_de1: Juliane (starting from the 28th in the sequence)

In this section we report the results of the evaluation. We first report some statistics which give us an overall idea
of how the annotators fared. We then delve into the question of how reliable the annotations are by examining
inter-annotator agreement and looking at confusion matrices. Finally we calculate UCCA semantic scores
for the different language pairs.

\subsubsection{Overall Statistics}

In order to explore the results of our annotations, we first show some basic statistics
about the task. In Table~\ref{tab:stat-all}, we can see the total number of sentences annotated
%
with
the number of nodes. We also show the percentages of the main types of nodes.
Looking at this table we see that the number of sentences varies from 324 to 351 sentences,
except for Ro1 who did considerably fewer, only 230. 
All annotators were shown the same set of sentences, but their results were only
collected if they pressed the submit button. So these missing sentences were either mistakenly
not submitted, or they did not get to the end of the task. 
Looking at the types of nodes annotated, these seem to be distributed quite uniformly.
About 35\% of the nodes were considered to be structural, and 60\% to be lexical.
Some nodes were not evaluated, and these are called ``Missing''. 
The node may represent an English
word which is not required in the target language (often an article or a
pronoun).
As per instructions, the annotators were expected to mark such nodes with green
but some may have left the node unannotated (in which case it would be perhaps
appropriate to consider the parent as a lexical node).
With hindsight, we should have introduced a separate annotation category for this, and
also enforced the lexical/structural distinction in the tool. ``Missing''
annotations are not very common, but
one annotator left 6.2\% of the nodes unlabelled which is a bit too high and in 
future experiments we will alter the tool to disallow missing nodes, unless their parent is a lexical node.

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|c|c|c|c|c|}
      \hline
 & \bf{No. Sentences} & \bf{No. Nodes} & \bf{\%Structural} & \bf{\%Lexical}  & \bf{\%Missing} \\
\hline                               
    De1  &  339 &  9253 & 36.2 & 62.5 & 1.3 \\ 
    Cs1  & 324 &  8794 & 33.4 & 63.3 & 3.3\\ 
    Ro1  & 230 &  6152 & 36.4 & 62.7 & 0.8\\ 
    Ro2  & 337 &  9228 & 35.6 & 61.4 & 2.9 \\ 
    Pl1  & 351 &  9557 & 34.0 & 59.8 & 6.2 \\ 
    Pl2  & 340 &  9303 & 31.0 & 64.6 & 4.4 \\ 
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Number of annotated sentences and nodes and the percentages of each of the main kind of nodes
}
\label{tab:stat-all}
\end{table}

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|c|c|c|}
      \hline
   & \bf{No. Nodes} & \bf{\%A} & \bf{\%B}   \\
\hline                               
    De1  &  3350 &  68.2 & 31.7  \\ 
    Cs1  & 2939 &  73.4 & 26.6 \\ 
    Ro1  & 2239 &  68.2 & 31.8  \\ 
    Ro2  & 3286 &  65.4 & 34.6 \\ 
    Pl1  & 3246 &  50.3 & 49.7 \\ 
    Pl2  & 2884 &  71.5 & 28.5 \\ 
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Number of structural  nodes and the percentages of each of the types of nodes
}
\label{tab:stat-struct}
\end{table}

In Table~\ref{tab:stat-struct}, we can see the number of structural nodes, and their
breakdown in terms of acceptable  ``A'' or bad ``B''. Most annotators seem to 
label about 70\% of the structural nodes as correct. The exception is Pl1 who
seems to be following a different annotation strategy. It is possible that Pl1 
did not understand the guidelines correctly. Further investigation with a Polish
speaking colleague will be necessary.

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|c|c|c|c|}
      \hline
   & \bf{No. Nodes} & \bf{\%G} & \bf{\%O}  & \bf{\%R} \\
\hline                               
    De1  & 5784  &  70.8 & 13.0 & 26.6  \\ 
    Cs1  & 5568 &  62.8 & 18.4 & 18.8 \\ 
    Ro1  & 3859 &  75.9 & 8.4 & 15.7  \\ 
    Ro2  & 5671 &  68.7 & 16.9 & 14.3 \\ 
    Pl1  & 5717 &  60.8 & 13.1 & 26.1 \\ 
    Pl2  & 6007 &  49.6  & 17.3 & 33.0 \\ 
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Number of lexical  nodes and the percentages of each of the types of nodes
}
\label{tab:stat-lex}
\end{table}

In Table~\ref{tab:stat-lex} we can see the number of lexical nodes, and their
breakdown in terms of correct  ``G'', partial ``O'' or incorrect ``R''. 
There is more variation here between the annotators than for the
structural nodes. The percentage of correct nodes fluctuates from 49.6\%
to 75.9\%. One expects that  languages with different amounts of 
morphology and agglutination would report different numbers here,
but even amongst annotators of the same languages the results vary.
For Polish the number of correct lexical 
%node 
nodes
varies between 49.6\%
and 60.8\%.

\subsubsection{IAA}

In order to determine how reliable our human evaluation approach is, we need to
examine the agreement that we find between annotators.
For two of the language pairs, Romanian and Polish, we had 2 annotators. 
 Inter-annotator agreement can be calculated 
using the Kappa score which is more reliable than agreement because it takes into
account the probability of agreement by chance.

%Lexi TODO put Kappas in a separate table

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|cc|c|}
      \hline
\bf{MT Eval Label} & \bf{A} & \bf{B} & \bf{All} \\
\hline                               
    A   & 1096 &  285 & 1381\\ 
   B   & 101 & 507  & 608\\ 
   \hline
   All &  1197 &  792 & 1989\\
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Confusion Matrix for the structural nodes for the Romanian annotators with a Kappa of 0.58
}
\label{tab:iaa-ab-ro}
\end{table}

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|ccc|c|}
      \hline
\bf{MT Eval Label} & \bf{R} & \bf{O} & \bf{G} & \bf{All} \\
\hline                               
   R   & 2274 &  361 & 126 & 2761 \\ 
   O   & 92 & 164  & 37 & 293 \\ 
   G   & 82 & 76  & 358 & 516 \\ 
   \hline
   All &  2448 &  601 & 521 & 3570\\
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Confusion Matrix for the lexical nodes for the Romanian annotators with a Kappa of 0.50
}
\label{tab:iaa-rgo-ro}
\end{table}

In Tables~\ref{tab:iaa-ab-ro} and \ref{tab:iaa-rgo-ro} we see the confusion matrix of
the two main types of evaluation labels for Romanian. We separate structural and lexical nodes, because
there is very little disagreement between annotators about which node is a lexical and which
node is structural, and so if we calculate Kappa over all the values, this might artifically 
inflate our IAA scores.
For both structural and lexical nodes, the Kappa values 
of 0.58 and 0.50 respectively 
can
be considered to display moderate agreement. 
If we calculate Kappa scores across all five categories, it rises to 
0.69 because of the strong agreement about what is a lexical and a structural 
node.

%\lexi{TODO look at confusion between structural and lexical}

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|cc|c|}
      \hline
\bf{MT Eval Label} & \bf{A} & \bf{B} & \bf{All} \\
\hline                               
    A   & 1208 &  192 & 1400 \\ 
   B   & 681 & 574  & 1255 \\ 
   \hline
   All &  1889 &  766 & 2655 \\
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Confusion Matrix for the structural nodes for the Polish annotators with a Kappa of 0.33
}
\label{tab:iaa-ab-pl}
\end{table}

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|ccc|c|}
      \hline
\bf{MT Eval Label} & \bf{R} & \bf{O} & \bf{G} & \bf{All} \\
\hline                               
   R   & 2430 &  444 & 427 & 3301 \\ 
   O   & 119 & 398  & 198 & 715 \\ 
   G   & 161 & 109  & 1110 & 1380 \\ 
   \hline
   All &  2710 &  951 & 1735 & 5396 \\
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Confusion Matrix for the lexical nodes for the Polish annotators with a Kappa of 0.58
}
\label{tab:iaa-rgo-pl}
\end{table}

In Tables~\ref{tab:iaa-ab-pl} and \ref{tab:iaa-rgo-pl} we see the confusion matrix of
the two main types of evaluation labels for Polish. Here we see similar Kappa results for
the lexical nodes (0.56) but the Kappa for structural nodes drops to 0.33. 
This low agreement could be the result of one annotator misinterpreting the guidelines
and warrants further analysis.
If we calculate Kappa scores across all five categories, it rises to 
0.58.

Looking beyond Kappa scores, we did some analysis of the cases where
one annotator marked a word as Red and another marked it Green. 
For Polish  the most common English words about which the annotators
disagreed were: ``to'', ``you'', ``blood'' (which was aligned to ``krwi''), ``in'' and ``for''. 
For Romanian, the words were ``you'', ``to'', ``the'', and ``your''. 
Apart from ``blood'', all these words are short function words which can 
be translated in a very different fashion in the target and this could contribute to
 confusion about
whether they are correct or not.

Looking at the words which were marked Red most often again these were largely function words.
This is partially due to the fact that they occur more frequently than other words. But
words such as ``bias'', ``falls'', ``fibrillation'' and ``review'' also occur frequently in this list and these
are words with domain specific translations which we are possibly not capturing with
the unadapted HimL models.

\subsubsection{UCCA Scores}

\paragraph{Node score} This simple score reflects the percentage of correct MT evaluation nodes. 
A and G are correct, O counts as 50\% correct and R and B count as incorrect. Nodes which are missing an
evaluation are ignored.

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|c|c|c|}
      \hline
\bf{Annotator ID}  & \bf{Structural } & \bf{Lexical } & \bf{Overall }\\
\hline                               
    De1  &  68.20 & 77.40 & 74.03 \\ 
    Cs1  & 73.35 & 72.00 & 72.46 \\ 
    Ro1  & 68.15 & 80.09 & 75.71 \\ 
    Ro2  & 65.42 & 77.26  & 72.92 \\ 
    Pl1  & 50.33 & 67.33 & 61.17 \\ 
    Pl2  & 71.53 &58.26  & 62.56 \\ 
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Node scores: percentage of correct nodes. 
}
\label{tab:node_scores}
\end{table}

In Table~\ref{tab:node_scores} we can see the simple node scores for the different annotators. 
Results show that most systems are evaluated as having about 70\% correct overall, with the
notable exception of both Polish annotators. They gave scores to the HimL test set which
were about 10\% lower. English$\to$Polish is considered to be a very challenging language pair
for machine translation and so this result is not surprising.  Another trend is that the annotators gave the lexical scores of at least 
10\% higher than the structural scores for three out of five annotators. The two remaining
annotators behaved quite differently. Cs1 gave similar scores for structural and lexical nodes, but
Pl2 gave much better scores for structural nodes (71.53 vs. 58.26). 
The differences in behaviour displayed by the two Polish annotators in this table and the
low kappa reported in Table~\ref{tab:iaa-ab-pl}
brings into question the reliability of the Polish annotations.

\begin{figure}[ht]
    \includegraphics[width=1.1\textwidth]{tree_convert.png}
    \caption{UCCA tree on the source labelled with the MT Eval labels and aligned to the target translation. The 
    ``Acceptable'' structural nodes are yellow, the ``Bad'' ones are blue.}
    \label{fig:ucca-tree-evaluated}
\end{figure}

%\lexi{TODO: Any chance of an example with bad structural nodes ? }
In order to visualise the UCCA evaluation results and see how the node scores are calculated, we provide an  example from the experiment. 
Looking at Figure~\ref{fig:ucca-tree-evaluated}, we can see the machine translation at the top. Below this are the word alignments to the English source. 
Each of the English words (except punctuation) are linked to UCCA lexical nodes which have been evaluated as either
correct (green), partially correct (orange) or incorrect (red). The letters inside the circles correspond to the UCCA node types
which will be important in future versions of the score. Structural nodes are either acceptable (yellow) or bad (blue). 
So for this sentence given that we have 11 green nodes, 2 orange nodes, 6 yellow nodes and in total there are 24 nodes:

\begin{displaymath}
node score = \frac{ 11  + 0.5 ( 2 ) + 6  } { 24  } * 100\\
                    = 75
\end{displaymath}

\subsection{Discussion}

In this evaluation we have proposed an evaluation which has a number of advantages:
\begin{itemize}
\item  It shows moderate IAA agreement which means that our metric is relatively reliable. It is not directly comparable to other metrics because HMEANT papers do not report IAA and they had to rely on agreement statistics which can be misleading. 
\item  We can reuse the UCCA structures on the source multiple times, and as this is the most time consuming part
of the evaluation, this makes best use of expensive expert annotators. 
\item Translation annotators do not have to try to create semantic structures on erroneous machine translation output, in fact
they do not need to create any semantic structures at all as this can be done previously by experts.
\item Errors are isolated at the point where they occur due to the separation of structure and lexical content and there
is little ambiguity about whether an error in a leaf needs to be counted multiple times as we go up the tree. 
\item UCCA provides a minimal yet complete semantic framework which is based on universal cognitive concepts. 
The structure does not rely on syntactic heads and the semantic graph is grounded directly to the words in the sentence. 
\end{itemize}

The evaluation also brought up a number of issues that will need to be resolved by either better  training or
by improving the tool. Many of these issues were reported by the annotators:
\begin{itemize}
\item Isolating errors to the point where they occur is difficult to do for humans as there is a tendency to
penalise the parent nodes for errors which occur in their children. 
\item Initially some annotators struggled to understand the distinction between lexical and structural nodes. 
\item They also had some difficulty evaluating the 
%sentence 
sentences
without their context. 
\item The annotators were unsure about how to judge incorrect morphology or tense. 
\item  The alignments were meant only as a guide and some words are unaligned or doubly aligned and this
caused some confusion. 
\item Extraneous additions on the target might not be penalised unless they fall inside the source
semantic structure.
\end{itemize}

Even though there have been a number of issues which have arisen, the UCCA evaluation has shown to be a reliable and
efficient
method for evaluating the accuracy of HimL translation systems. 
We will continue to analyse the results from this experiment and we will run an improved version of
 this evaluation in year two
of the project.

%Comments from Juliane (Cochrane DE)
%- There weren't many, but a few mistakes in the source, that made it difficult to annotate.
%- Sometimes it was difficult to judge without context, but not very often - at least if you're familiar with the content.  
%- There were cases where the way the sentences were segmented made it difficult to annoate them in a useful way.
%- Many wrong declinations and conjugations.
%- German genus requires pronouns or articles to be repeated in lists of nouns in order to be correct, e.g. your strength and balance = Ihre Stärke und Ihr Gleichgewicht
%- There were a few out of domain vocabulary instance, e.g. "balance" being translated in the context of banking rather than the body, but only in one or two sentences, not in all where it occured. Similarly, the word "trial" got translated as "process", "study", and in the context of a court trial. 
%- A number of Cochrane or medical research specific terms was not well translated.
%- Abbreviations never got translated correctly.
%- Participle constructions often mess up the translation completely.
%- Some verbs kept getting translated as nouns or the other way round, e.g. in review author = the author of a review, review was always interpreted as a verb; falls (noun) was mostly translated as verb. 
%- Overall: both Cochrane and NHS sentences were more often bad than understandable. Only very few were acceptable for publication on our websites in my view. I guess that was expected :)
%What I mentioned earlier:
%- The past and future tense were often ignored, which is usally not a problem for future, but for past it is
%- Fairly often verbs were not translated at all (e.g. "cause" was never translated).  
%- If the verb is present, it is often in the wrong order with the subject, as it doesn't necessarily follow the same order as in English.
%- You is sometimes translated formally and sometimes informally, even in the same sentence.
%- Need a counter to know how many sentences completed
%
%Comments from Rodica (NHS Ro):
%One of the most recurrent confusions I had during my work was in connection with the word order in the MT.
%EG.  English: This review provides limited evidence that interventions...
%         Romanian: Acest studiu intervenții care furnizează dovezi limitate
%                               (This review interventions provide limited evidence)
%My question would have been how to mark the wrong or random word order in the MT translation, when the individual words appeared, so they couldn't be marked as red. 
%In most of the examples this occurred in the Romanian translations  because the default word order in Romanian is  noun + descriptive adjective(s), rather than the other way round and so, in the MT the descriptive adjectives would be first and sometimes other words would be added before the noun, which would make the sentence incomprehensible in Romanian.
%I think I mentioned to you the other questions during the work. The one that I should have asked earlier  was how far up the structure I should  mark a missing child. I understood it only once I spoke to you, but that was quite late.
%
%
%Comments from Iwona (NHS Polish)
%- Need a counter to know how many sentences completed
%1. Phrases split in wrong places, e.g. ‘depending on’, ‘there is’
%2. Certain phrases should be treated as one unit, e.g. ‘make sure’, so a leaf node for 
%marking as green, orange or red.
%3. Different sentence structure in English and Polish, e.g.
%‘our appetites often decrease’
%Machine translation: ‘our appetites often a fall’
%Correct translation:  ‘Often decreases for us appetite’
%Please note in the example above the meaning can be worked out and for that 
%reason I marked the structural node as A. However, this is bad Polish. There were a 
%number of sentences like this one, the translation of some of them made more sense 
%than others, on the whole it was difficult to annotate them. 
%4. The following is a common occurrence:
%‘before starting a new exercise programme’
%‘before’ – ‘przed rozpoczęciem’ (should be ‘przed’)
%‘starting’ – ‘przed rozpoczęciem’ (should be ‘rozpoczęciem’)
%Both leaf nodes have the same translation, but the translation appears only once in 
%the structural node. I marked the parent node as A, and the two leaf nodes as red. 
%5. Double negation in Polish causes problems at the structural level.

\subsection{Automatic Semantic Evaluation}

The results that we have gathered in this evaluation will be used as gold data to evaluate the performance of a number of automatic
metrics, to determine which of the existing metrics are most suited to evaluating accuracy.  We have also  initiated
work on developing an automatic version of the UCCA human evaluation task. For this we need an UCCA semantic parser
and data to train it, as the UCCA treebank is relatively small. We are looking at extracting further training examples from
the Prague treebank t-layer. This will be described in Deliverable D5.3, due at the end of July 2016.

\section{Conclusion}

\section*{Acknowledgments}

Do not number the acknowledgment section.
This section should not be presented for the submission version.

\bibliography{mt}
\bibliographystyle{acl2016}

\end{document}

### 2016-02-01 paper/build.sh

@@ -1,0 +1,15 @@
+#!/bin/sh
+
+case $1 in
+clean)
+  rm -f *.log *.aux *.toc *.blg *.pdf *.out *.bbl
+  ;;
+*)
+  pdflatex main.tex
+  bibtex bu1
+  bibtex bu2
+  bibtex main
+  pdflatex main.tex
+  pdflatex main.tex
+  ;;
+esac

### 2016-02-01 paper/convert.sh

@@ -1,0 +1,1 @@
+convert -alpha off tree.png tree_convert.png

### 2016-02-02 scripts/deploy.sh
draw all trees, and show on website

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+set -o xtrace
+
+rsync -av --rsh=ssh --delete \
+          sentences.csv \
+          nodes.csv \
+          index.html \
+          show_errors.py \
+          show_trees.py \
+          show_tree.py \
+          show_iaa_rg.py \
+          iaa-report.txt \
+          trees \
+          thor:/disk4/html/himl/internal/ucca-eval

### 2016-02-02 scripts/get_trees.py

@@ -1,0 +1,51 @@
+#!/usr/bin/env python
+from __future__ import print_function
+
+
+#
+# Draws all the annotated UCCA trees
+#
+
+import argparse
+import logging
+import math
+import os
+import os.path
+import pandas
+import pydot
+import sys
+
+from tree import Plotter
+
+LOG = logging.getLogger(__name__)
+   
+def main():
+  logging.basicConfig(format='%(asctime)s %(levelname)s: %(name)s: %(message)s', datefmt='%Y-%m-%d %H:%M:%S', level=logging.INFO)
+  parser = argparse.ArgumentParser()
+  parser.add_argument("-n", "--nodes-file", default="nodes.csv")
+  parser.add_argument("-s", "--sentences-file", default="sentences.csv")
+  parser.add_argument("-o", "--output-dir", default="trees")
+  args = parser.parse_args()
+
+  nodes = pandas.read_csv(args.nodes_file, converters={'node_id': str, 'parent' : str})
+  sentences = pandas.read_csv(args.sentences_file)
+  plotter = Plotter(sentences, nodes)
+
+  if not os.path.exists(args.output_dir):
+    os.makedirs(args.output_dir)
+  LOG.info("Writing trees to directory: {}/".format(args.output_dir))
+
+  #NOTE: Iterating through rows like this is not really the pandas way
+  for i,row in enumerate(sentences.iterrows()):
+    sent_id,annot_id = row[1].sent_id,row[1].annot_id
+    LOG.debug("Plotting sentence id: {} annotator id: {}".format(sent_id,annot_id))
+    graph_file = "{}/tree_{}_{}.png".format(args.output_dir,sent_id,annot_id)
+    graph = plotter.plot(sent_id,annot_id)
+    graph.write_png(graph_file)
+    if i  and i % 50 == 0:
+      LOG.info("Processed {} trees".format(i))
+
+if __name__ == "__main__":
+  main()
+
+

### 2016-02-02 scripts/run.sh

@@ -4,14 +4,4 @@ set -o xtrace
 
 ./get_csv.py -i data/raw/*dump
 ./iaa.py | tee iaa-report.txt
-
-if [ "$1" == "deploy" ] ; then
-  rsync -av --rsh=ssh --delete \
-            sentences.csv \
-            nodes.csv \
-            index.html \
-            show_errors.py \
-            show_iaa_rg.py \
-            iaa-report.txt \
-            thor:/disk4/html/himl/internal/ucca-eval
-fi
+#./get_trees.py

### 2016-02-02 scripts/show_tree.py

@@ -1,0 +1,46 @@
+#!/usr/bin/env python
+from __future__ import print_function
+
+import cgi
+import cgitb
+import pandas
+
+cgitb.enable()
+
+
+def main():
+  print("Content-Type: text/html")
+  print()
+
+
+  args = cgi.FieldStorage()
+  if "sent_id" not in args  or "annot_id" not in args:
+    print("<h1>Error</h1>")
+    print("Missing sentence id or annotator id<br>")
+    print(args)
+    return
+
+  annot_id = args["annot_id"].value
+  sent_id = args["sent_id"].value
+
+  print("<head><meta charset=\"UTF-8\">")
+  print("<TITLE>Tree: {}, Annotator: {}</TITLE>".format(sent_id,annot_id))
+  print("</head>")
+  print("<BODY>")
+
+  #print("<h2>Data</h2>")
+  #print("<table border=1>")
+  #def printrow(k,v):
+  #  print("<tr><td>{}</td><td>{}</td></tr>".format(k,v))
+  #printrow("SentenceId", sentence_id)
+  #printrow("AnnotatorId", annot_id)
+  #print("</table>")
+
+  #print("<h2>Tree</h2>")
+  print('<img src="trees/tree_{}_{}.png" width="2000">'.format(sent_id,annot_id))
+
+  print("</BODY>")
+
+if __name__ == "__main__":
+  main()
+

### 2016-02-02 scripts/show_trees.py

@@ -1,0 +1,77 @@
+#!/usr/bin/env python
+from __future__ import print_function
+
+import cgi
+import cgitb
+import pandas
+
+cgitb.enable()
+
+
+LANGS = {
+  "cs" : "Czech",
+  "de" : "German", 
+  "pl" : "Polish",
+  "ro" : "Romanian"
+}
+def main():
+  print("Content-Type: text/html")
+  print()
+
+
+  args = cgi.FieldStorage()
+  if "lang" not in args or args["lang"].value not in LANGS:
+    print("<h1>Error</h1>")
+    print("Missing or unknown language id")
+    print(args)
+    return
+  lang = args["lang"].value
+
+  sentences = pandas.read_csv("sentences.csv")
+  sentences['score'] = (sentences['mteval_A'] + sentences['mteval_G'] + sentences['mteval_O']/2) / \
+                        (sentences['ucca_node_count'] - sentences['mteval_M'])
+  sort = "sent_id"
+  if "sort" in args:
+    sort = args["sort"].value
+
+  has_mult_annot = (lang == "ro" or lang == "pl")
+
+  print("<head><meta charset=\"UTF-8\">")
+  print("<TITLE>All Annotated Trees: {}</TITLE>".format(LANGS[lang]))
+  print("</head>")
+  print("<BODY>")
+
+  
+
+  print("<table border=1>")
+  print("<tr>")
+  headings = [("SentenceId", "sent_id")]
+  if has_mult_annot:
+    headings.append(("AnnotId", "annot_id"))
+  headings += [("tree",""), ("bleu", "bleu"), ("Acceptable", "mteval_A") , ("Bad", "mteval_B"), ("Green", "mteval_G") \
+    , ("Orange", "mteval_O"), ("Red", "mteval_R"), ("Missing", "mteval_M"), ("UCCA Score", "score")]
+  for h,sortorder in headings:
+    if sortorder:
+      print("<th><a href=\"show_trees.py?lang={}&sort={}\">{}</th>".format(lang,sortorder,h))
+    else:
+      print("<th>{}</th>".format(h))
+  print("</tr>")
+
+  for i,row in sentences[sentences.lang == lang].sort_values(by=sort, ascending=False).iterrows():
+    fields = [row["sent_id"]]
+    if has_mult_annot:
+      fields.append(row["annot_id"])
[…]

### 2016-02-02 scripts/tree.py

@@ -16,8 +16,11 @@ COLORS = {
   "O" : "orange", 
   "G" : "green", 
   "A" : "yellow", 
-  "B" : "blue"
+  "B" : "blue",
+  "M" : "white"
 }
+
+LOG = logging.getLogger(__name__)
 
 class Plotter:
   
@@ -54,7 +57,8 @@ class Plotter:
     src_keys = set()
     for index,node in nodes.iterrows():
       if isinstance(node.source,str):
-        textnodes.append((node.pos,node.source))
+        #textnodes.extend(zip(node.pos.split(), node.source.split()))
+        textnodes.append((node.pos.split()[0],node.source))
     text = pydot.Cluster("src", rank = 'same', rankdir='LR', color="white")
     text.add_node(pydot.Node(style="invis", name = "src_start"))
     for pos,source in sorted(textnodes, key=lambda x: int((x[0]))):
@@ -69,9 +73,9 @@ class Plotter:
     graph.add_edge(pydot.Edge("tgt_end", "src_end", style="invis"))
 
     # Alignment
-    print (sentence['source'].iloc[0])
-    print (sentence['target'].iloc[0])
-    print (sentence['align'].iloc[0])
+    LOG.debug(sentence['source'].iloc[0])
+    LOG.debug(sentence['target'].iloc[0])
+    LOG.debug(sentence['align'].iloc[0])
     for align in sentence['align'].iloc[0].split():
       src_pos,tgt_pos = align.split("-")
       src_key = "src_" + str(src_pos)
@@ -86,7 +90,7 @@ class Plotter:
       graph.add_node(ucca)
       if isinstance(node.source,str):
         # Add link to word
-        textnode = text.get_node("src_" + node.pos)
+        textnode = text.get_node("src_" + node.pos.split()[0])
         graph.add_edge(pydot.Edge(textnode[0],ucca))
       if node.parent != "0": 
         # Link to parent
@@ -98,15 +102,15 @@ class Plotter:
 def main():
   logging.basicConfig(format='%(asctime)s %(levelname)s: %(name)s: %(message)s', datefmt='%Y-%m-%d %H:%M:%S', level=logging.DEBUG)
   parser = argparse.ArgumentParser()
+  parser.add_argument("-a", "--annotator-id", default="pl1")
+  parser.add_argument("-s", "--sentence-id", type=int, default=169)
   args = parser.parse_args()
 
[…]

### 2016-02-02 scripts/show_trees.py
titles

@@ -35,14 +35,16 @@ def main():
     sort = args["sort"].value
 
   has_mult_annot = (lang == "ro" or lang == "pl")
+  title = "All Annotated Trees: {}".format(LANGS[lang])
 
   print("<head><meta charset=\"UTF-8\">")
-  print("<TITLE>All Annotated Trees: {}</TITLE>".format(LANGS[lang]))
+  print("<TITLE>{}</TITLE>".format(title))
   print("</head>")
   print("<BODY>")
 
   
 
+  print("<h1>{}</h1>".format(title))
   print("<table border=1>")
   print("<tr>")
   headings = [("SentenceId", "sent_id")]

### 2016-02-08 scripts/tree.py
straight alignment lines

@@ -33,7 +33,23 @@ class Plotter:
     #print((self.nodes.columns))
     nodes = self.nodes[(self.nodes.sent_id == sent_id) & (self.nodes.annot_id == annot_id)]
     sentence = self.sentences[(self.sentences.sent_id == sent_id) & (self.sentences.annot_id == annot_id)]
-    graph = pydot.Dot(graph_type='graph')
+    graph = pydot.Dot(graph_type='graph', splines=False, rankdir="TB")
+
+    # Source nodes
+    textnodes = [] # (pos,text)
+    src_keys = set()
+    for index,node in nodes.iterrows():
+      if isinstance(node.source,str):
+        textnodes.append((node.pos.split()[0],node.source))
+    text = pydot.Cluster("src", rank = 'same', rankdir='LR', color="white")
+    text.add_node(pydot.Node(style="invis", name = "src_start"))
+    for pos,source in sorted(textnodes, key=lambda x: int((x[0]))):
+      key = "src_" + pos
+      src_keys.add(key)
+      textnode = pydot.Node(name = key, label=source, shape="rectangle")
+      text.add_node(textnode)
+    text.add_node(pydot.Node(style="invis", name = "src_end"))
+    graph.add_subgraph(text)
 
     # Target nodes
     target = pydot.Cluster("tgt", rank = 'same', rankdir='LR', color="white")
@@ -47,27 +63,10 @@ class Plotter:
       node = pydot.Node(name = key, label=token, shape="rectangle")
       target.add_node(node)
       #if prev_node:
-      #  target.add_edge(pydot.Edge(prev_node,node))
+      #  target.add_edge(pydot.Edge(prev_node,node, style="invis"))
       prev_node = node
     target.add_node(pydot.Node(style="invis", name = "tgt_end"))
     graph.add_subgraph(target)
-
-    # Collect source text nodes
-    textnodes = [] # (pos,text)
-    src_keys = set()
-    for index,node in nodes.iterrows():
-      if isinstance(node.source,str):
-        #textnodes.extend(zip(node.pos.split(), node.source.split()))
-        textnodes.append((node.pos.split()[0],node.source))
-    text = pydot.Cluster("src", rank = 'same', rankdir='LR', color="white")
-    text.add_node(pydot.Node(style="invis", name = "src_start"))
-    for pos,source in sorted(textnodes, key=lambda x: int((x[0]))):
-      key = "src_" + pos
[…]

### 2016-02-10 scripts/tree.py
escaping of text in trees

@@ -27,6 +27,13 @@ class Plotter:
   def __init__(self, sentences, nodes):
     self.nodes = nodes
     self.sentences = sentences
+  
+  def make_label(self,label):
+    #escape quotes
+    label = label.replace("\"", "\\\"")
+    # enclose in quotes
+    label = "\"{}\"".format(label)
+    return label
 
   def plot(self, sent_id, annot_id):
     """Plot an UCCA tree"""
@@ -46,7 +53,7 @@ class Plotter:
     for pos,source in sorted(textnodes, key=lambda x: int((x[0]))):
       key = "src_" + pos
       src_keys.add(key)
-      textnode = pydot.Node(name = key, label=source, shape="rectangle")
+      textnode = pydot.Node(name = key, label=self.make_label(source), shape="rectangle")
       text.add_node(textnode)
     text.add_node(pydot.Node(style="invis", name = "src_end"))
     graph.add_subgraph(text)
@@ -59,8 +66,7 @@ class Plotter:
     prev_node = None
     for pos,token in enumerate(target_tokens):
       key = "tgt_" + str(len(target_tokens) - pos - 1)
-      if token == ",": token = "COMMA"
-      node = pydot.Node(name = key, label=token, shape="rectangle")
+      node = pydot.Node(name = key, label=self.make_label(token), shape="rectangle")
       target.add_node(node)
       #if prev_node:
       #  target.add_edge(pydot.Edge(prev_node,node, style="invis"))

### 2016-02-10 scripts/get_trees.py
doubly annotated trees

@@ -40,7 +40,7 @@ def main():
     sent_id,annot_id = row[1].sent_id,row[1].annot_id
     LOG.debug("Plotting sentence id: {} annotator id: {}".format(sent_id,annot_id))
     graph_file = "{}/tree_{}_{}.png".format(args.output_dir,sent_id,annot_id)
-    graph = plotter.plot(sent_id,annot_id)
+    graph = plotter.plot_tree(sent_id,annot_id)
     graph.write_png(graph_file)
     if i  and i % 50 == 0:
       LOG.info("Processed {} trees".format(i))

### 2016-02-10 scripts/tree.py

@@ -35,44 +35,99 @@ class Plotter:
     label = "\"{}\"".format(label)
     return label
 
-  def plot(self, sent_id, annot_id):
-    """Plot an UCCA tree"""
+  def plot_tree(self, sent_id, annot_id):
+    """Plot an UCCA tree, with source and target sentences"""
     #print((self.nodes.columns))
     nodes = self.nodes[(self.nodes.sent_id == sent_id) & (self.nodes.annot_id == annot_id)]
     sentence = self.sentences[(self.sentences.sent_id == sent_id) & (self.sentences.annot_id == annot_id)]
     graph = pydot.Dot(graph_type='graph', splines=False, rankdir="TB")
 
+    source,target = self.plot_source_target(graph,sentence,nodes)
+
+    # Add the ucca tree
+    for index, node in nodes.iterrows():
+      ucca = pydot.Node(name = node.node_id, label = node.ucca_label,
+         fillcolor = COLORS[node.mt_label], style="filled", shape="circle")
+      graph.add_node(ucca)
+      if isinstance(node.source,str):
+        # Add link to word
+        textnode = source.get_node("src_" + node.pos.split()[0])
+        graph.add_edge(pydot.Edge(textnode[0],ucca))
+      if node.parent != "0": 
+        # Link to parent
+        #print("Creating {} -> {}".format(node.ucca_label, node.parent))
+        graph.add_edge(pydot.Edge(ucca, node.parent))
+    return graph
+
+  def make_split_label(self,ucca_label, mt_label1, mt_label2):
+      return """<
+      <table cellpadding="0" cellborder="0" cellspacing="0" border="1">
+      <tr>
+      <td bgcolor=\"{1}\">{0}&nbsp;</td>
+      <td bgcolor=\"{2}\"><font color=\"{2}\">{0}&nbsp;</font></td>
+      </tr>
+      </table>
+      >""".format(ucca_label, COLORS[mt_label1], COLORS[mt_label2])
+
+
+  def plot_doubly_annotated_tree(self, sent_id, annot_id1, annot_id2):
+    """Plot a doubly annotated UCCA tree with its source and target sentences."""
+    nodes1 = self.nodes[(self.nodes.sent_id == sent_id) & (self.nodes.annot_id == annot_id1)]
+    sentence1 = self.sentences[(self.sentences.sent_id == sent_id) & (self.sentences.annot_id == annot_id1)]
+    graph = pydot.Dot(graph_type='graph', splines=False, rankdir="TB")
+
[…]

### 2016-02-12 scripts/deploy.sh
listing of double trees

@@ -8,7 +8,9 @@ rsync -av --rsh=ssh --delete \
           index.html \
           show_errors.py \
           show_trees.py \
+          show_double_trees.py \
           show_tree.py \
+          show_double_tree.py \
           show_iaa_rg.py \
           iaa-report.txt \
           trees \

### 2016-02-12 scripts/get_trees.py

@@ -36,6 +36,20 @@ def main():
   LOG.info("Writing trees to directory: {}/".format(args.output_dir))
 
   #NOTE: Iterating through rows like this is not really the pandas way
+
+  # All doubly annotated trees
+  merged = sentences.merge(sentences, on = ["sent_id", "lang"])
+  agree = merged[(merged["annot_id_x"] < merged["annot_id_y"])]
+  for i,row in enumerate(agree.iterrows()):
+    sent_id,annot_id1,annot_id2 = row[1].sent_id,row[1].annot_id_x,row[1].annot_id_y
+    LOG.debug("Plotting sentence id: {} annotators: {} and {}".format(sent_id,annot_id1,annot_id2))
+    graph_file = "{}/tree_{}_{}_{}.png".format(args.output_dir,sent_id,annot_id1,annot_id2)
+    graph = plotter.plot_doubly_annotated_tree(sent_id,annot_id1,annot_id2)
+    graph.write_png(graph_file)
+    if i  and i % 50 == 0:
+      LOG.info("Processed {} doubly annotated trees".format(i))
+
+  # All trees
   for i,row in enumerate(sentences.iterrows()):
     sent_id,annot_id = row[1].sent_id,row[1].annot_id
     LOG.debug("Plotting sentence id: {} annotator id: {}".format(sent_id,annot_id))
@@ -44,6 +58,8 @@ def main():
     graph.write_png(graph_file)
     if i  and i % 50 == 0:
       LOG.info("Processed {} trees".format(i))
+
+
 
 if __name__ == "__main__":
   main()

### 2016-02-12 scripts/show_double_tree.py

@@ -1,0 +1,47 @@
+#!/usr/bin/env python
+from __future__ import print_function
+
+import cgi
+import cgitb
+import pandas
+
+cgitb.enable()
+
+
+def main():
+  print("Content-Type: text/html")
+  print()
+
+
+  args = cgi.FieldStorage()
+  if "sent_id" not in args  or "annot_id1" not in args or "annot_id2" not in args:
+    print("<h1>Error</h1>")
+    print("Missing sentence id or annotator id<br>")
+    print(args)
+    return
+
+  annot_id1 = args["annot_id1"].value
+  annot_id2 = args["annot_id2"].value
+  sent_id = args["sent_id"].value
+
+  print("<head><meta charset=\"UTF-8\">")
+  print("<TITLE>Tree: {}</TITLE>".format(sent_id))
+  print("</head>")
+  print("<BODY>")
+
+  #print("<h2>Data</h2>")
+  #print("<table border=1>")
+  #def printrow(k,v):
+  #  print("<tr><td>{}</td><td>{}</td></tr>".format(k,v))
+  #printrow("SentenceId", sentence_id)
+  #printrow("AnnotatorId", annot_id)
+  #print("</table>")
+
+  #print("<h2>Tree</h2>")
+  print('<img src="trees/tree_{}_{}_{}.png" width="2000">'.format(sent_id,annot_id1, annot_id2))
+
+  print("</BODY>")
+
+if __name__ == "__main__":
+  main()
+

### 2016-02-12 scripts/show_double_trees.py

@@ -1,0 +1,103 @@
+#!/usr/bin/env python
+from __future__ import print_function
+
+import cgi
+import cgitb
+import pandas
+
+cgitb.enable()
+
+
+LANGS = {
+  "cs" : "Czech",
+  "de" : "German", 
+  "pl" : "Polish",
+  "ro" : "Romanian"
+}
+
+def display(lang,sort):
+  sentences = pandas.read_csv("sentences.csv")
+  nodes = pandas.read_csv("nodes.csv", converters={'node_id': str, 'parent' : str})
+  sentences = sentences[sentences.lang == lang]
+  nodes = nodes[nodes.lang == lang]
+
+  # Create table of matching nodes
+  nodes = nodes.merge(nodes, on = ["node_id", "sent_id"])
+  nodes = nodes[(nodes["annot_id_x"] < nodes["annot_id_y"])]
+  nodes = nodes[(nodes["mt_label_x"] != "M") & (nodes["mt_label_y"] != "M")]
+  nodes['match'] = nodes['mt_label_x'] == nodes['mt_label_y']
+
+  # One record per sentence
+  annot_id1,annot_id2 = sorted(sentences['annot_id'].unique())
+  sentences = sentences[sentences.annot_id == annot_id1] #one record per sentence
+  del sentences['mteval_A']
+  del sentences['mteval_B']
+
+  # Join match counts into sentence table
+  grouped = nodes.groupby(['sent_id'], as_index=True)
+  matches = pandas.DataFrame({ \
+    'nodes' : grouped['match'].size(),
+    'nodes_agree' : grouped['match'].sum(),
+    'agreement' : grouped['match'].sum() / grouped['match'].size()
+  })
+  sentences = sentences.merge(matches, left_on="sent_id", right_index=True)
+  sentences['nodes_agree'] = sentences['nodes_agree'].astype('int')
+
+  # Source length
+  sentences['src_length'] = sentences['source'].str.split().apply(len)
+  
+  # Columns should be sentence id, nodes, nodes agree, agreement , source length, bleu
+
+  title = "All Doubly Annotated Trees: {}".format(LANGS[lang])
+  print("<head><meta charset=\"UTF-8\">")
+  print("<TITLE>{}</TITLE>".format(title))
+  print("</head>")
+  print("<BODY>")
+  print("<h1>{}</h1>".format(title))
+  print("<table border=1>")
+  print("<tr>")
[…]

### 2016-02-12 scripts/show_double_tree.py
more info on sentences

@@ -24,8 +24,19 @@ def main():
   annot_id2 = args["annot_id2"].value
   sent_id = args["sent_id"].value
 
+  sentences = pandas.read_csv("sentences.csv")
+  sentence = sentences[(sentences.annot_id == annot_id1) & (sentences.sent_id == int(sent_id))]
+
   print("<head><meta charset=\"UTF-8\">")
   print("<TITLE>Tree: {}</TITLE>".format(sent_id))
+  print("""
+    <style>
+    th, td {
+          padding: 10px;
+    }
+    </style>
+  """)
+
   print("</head>")
   print("<BODY>")
 
@@ -39,6 +50,23 @@ def main():
 
   #print("<h2>Tree</h2>")
   print('<img src="trees/tree_{}_{}_{}.png" width="2000">'.format(sent_id,annot_id1, annot_id2))
+
+  print("<br>")
+  print("<table>")
+  print("<tr>")
+  print("<td>Source</td>")
+  print("<td>{}</td>".format(sentence.source.iloc[0]))
+  print("</tr>")
+  print("<tr>")
+  print("<td>Target</td>")
+  print("<td>{}</td>".format(sentence.target.iloc[0]))
+  print("</tr>")
+  print("<tr>")
+  print("<td>Alignment</td>")
+  print("<td>{}</td>".format(sentence['align'].iloc[0]))
+  print("</tr>")
+
+  print("</table>")
 
   print("</BODY>")
 

### 2016-02-12 scripts/show_tree.py

@@ -23,8 +23,18 @@ def main():
   annot_id = args["annot_id"].value
   sent_id = args["sent_id"].value
 
+  sentences = pandas.read_csv("sentences.csv")
+  sentence = sentences[(sentences.annot_id == annot_id) & (sentences.sent_id == int(sent_id))]
+
   print("<head><meta charset=\"UTF-8\">")
   print("<TITLE>Tree: {}, Annotator: {}</TITLE>".format(sent_id,annot_id))
+  print("""
+  <style>
+  th, td {
+        padding: 10px;
+  }
+  </style>
+  """)
   print("</head>")
   print("<BODY>")
 
@@ -39,6 +49,22 @@ def main():
   #print("<h2>Tree</h2>")
   print('<img src="trees/tree_{}_{}.png" width="2000">'.format(sent_id,annot_id))
 
+  print("<br>")
+  print("<table>")
+  print("<tr>")
+  print("<td>Source</td>")
+  print("<td>{}</td>".format(sentence.source.iloc[0]))
+  print("</tr>")
+  print("<tr>")
+  print("<td>Target</td>")
+  print("<td>{}</td>".format(sentence.target.iloc[0]))
+  print("</tr>")
+  print("<tr>")
+  print("<td>Alignment</td>")
+  print("<td>{}</td>".format(sentence['align'].iloc[0]))
+  print("</tr>")
+
+  print("</table>")
   print("</BODY>")
 
 if __name__ == "__main__":

### 2016-03-06 paper/main.tex
abstract

@@
 
+\newcommand{\com}[1]{}
+\newcommand{\secref}[1]{Section \ref{#1}}
+\newcommand{\figref}[1]{Figure~\ref{#1}}
+\newcommand{\tabref}[1]{Table~\ref{#1}}
+\newcommand{\oa}[1]{\footnote{\color{red} #1}}
+
 \newcommand\BibTeX{B{\sc ib}\TeX}
 
-\title{Human accuracy evaluation using UCCA}
+\title{HUME: Human UCCA-based Machine-Translation Evaluation}
 
@@
 \begin{abstract}
- This document contains instructions for preparing ACL 2016 submissions and camera-ready
-  manuscripts.  The document itself conforms
-  to its own specifications, and is therefore an example of what
-  your manuscript should look like.  Papers are required to conform to
-  all the directions reported in this document.  
-By using the provided \LaTeX\ and \BibTeX\ styles ({\small\tt acl2016.sty}, {\small\tt acl2016.bst}), the required formatting will be enabled by default.
+
+  Semantic evaluation of Machine Translation (MT) systems has become
+  a necessity, given recent interest in semantics-aware MT systems.
+  We present a novel human semantic evaluation measure for MT, Human
+  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
+  Making progress over previous work, the proposed measure takes into account
+  a wider range of semantic phenomena and does not rely on semantic annotation
+  of the MT output, whose interpretation is often unclear. 
+  We experiment with four language pairs, demonstrating the broad applicability of
+  the measure, as well as its reliability, reflected by the inter-annotator agreement
+  rates.
+
 \end{abstract}
 
-\section{ Human Semantic Evaluation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}\label{sec:intro}
+
+Human judgement is the central (and perhaps the ultimate) criterion for the quality of an MT system.
+Nevertheless, common measures for MT evaluation, such as adequacy and frequency judgements or the relative
+ranking of several possible translations, are problematic in two ways.
+First, as the quality of translation is determined by multiple factors, it is difficult
+to assign a single number so to reflect the quality of the entire sentence, even for a
+trained annotator. This also shows up with the diminishing IAA scores as a function of the
+sentence length \cite{}.
+Second, a single number is not instructive enough for localizing the translation errors, so to inform
+the improvement of the MT system. This problem is partly addressed by using scores that decompose over
+the sub-parts of the sentence, such as BLEU, thereby localizing the translation error.
+However, decomposition into n-grams are most suitable for evaluating shallow methods,
+such as phrase-based translation, 
+and are less instructive for more structurally-aware methods,
+which are at the forefront of MT research today.
+
+These difficulties have motivated the definition of a human measure that decomposes over the semantic
+structure of the sentence, rather than its n-grams. The most notable attempt of this brand is the HMEANT
+measure \cite{LoWu2010,LoWu2011a,LoWu2011b,LoWu2012}, which is based on the similarity
+between the translation output and a reference, defined in terms of the verbal argument structures of the
+two sentences. While appealing in addressing the two major shortcomings detailed above, HMEANT 
+is unable to capture semantic differences which do not show up in the verb-argument structure,
+including several very frequent phenomena, such as inter-clause linkage, nominalizations and copula clauses
+(see \secref{}). For instance, consider the sentence ``A coronary angioplasty may not be technically possible'',
+which was translated by one of the systems we experimented with as
+``Eine koronare Angioplastie kann nicht technisch m{\"o}glich'', i.e., the main verb ``sein'' (``be'') is omitted.
+While this may be interpreted is a minor error, HMEANT will assign it a very low score, as it failed to
+translate the main verb. 
+Another problem with HMEANT is that it requires semantically annotating MT output, which is often unintelligible,
+as well as involves the annotation and comparison to a reference translation: an impractical if not unattainable
+solution, given the huge number of reasonable translations a sentence may have \cite{dryer_marcu:12}.
+
+In this work we propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
+a human evaluation measure that decomposes over local semantic structures.
+The measure is based on a decomposition of the source sentence into semantic units according
+to the UCCA semantic annotation, and the estimation of their translation quality.
+UCCA covers a wide range of fundamental linguistic phenomena,
+including verbal, nominal and adjectival argument structures, inter-clause relations, coorindation
+and others. Importantly, HUME only requires an annotation of the source sentence, thus avoiding the 
+annotatation machine-generated text. By basing the measure on a direct comparison
+of the source and translation we avoid the difficulties engendered by using
+reference translations, but impose the additional requirement
+that the evaluator be proficient in both the source
+and target languages.
+Further discussion of the strengths and weaknesses of HUME relative to previous approaches
+can be found in \secref{}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Human Semantic Evaluation}
 \label{sec:sem-eval:human}
@@
 
-\begin{figure}[ht]
+\begin{figure*}[t]
     \includegraphics[width=1\textwidth]{ucca-tree}
@@
     \label{ucca-tree}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[ht]
+\begin{figure*}[t]
     \includegraphics[width=1\textwidth]{ucca-tree-mteval}
@@
     \label{ucca-tree-mteval}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[ht]
+\begin{figure*}[t]
     \includegraphics[width=1\textwidth]{interface-annotations}
@@
     \label{mttool}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}[ht]
+\begin{figure*}[t]
     \includegraphics[width=1.1\textwidth]{tree_convert.png}
@@
     \label{fig:ucca-tree-evaluated}
-\end{figure}
+\end{figure*}
 

### 2016-03-06 paper/main.tex
xxx

--- 2016-02-01 paper/main.tex
@@
 \usepackage{graphicx}
+\usepackage{url}
+\usepackage{subfigure}
+\usepackage{pgfplotstable}
+\usepackage{hyperref}
+\usepackage{color}
+\usepackage{lipsum,adjustbox}
+
+\newcommand{\com}[1]{}
+\newcommand{\secref}[1]{Section \ref{#1}}
+\newcommand{\figref}[1]{Figure~\ref{#1}}
+\newcommand{\tabref}[1]{Table~\ref{#1}}
+\newcommand{\oa}[1]{\footnote{\color{red} #1}}
 
@@
 
-\title{Human accuracy evaluation using UCCA}
+\title{HUME: Human UCCA-based Machine-Translation Evaluation}
 
@@
 \begin{abstract}
- This document contains instructions for preparing ACL 2016 submissions and camera-ready
-  manuscripts.  The document itself conforms
-  to its own specifications, and is therefore an example of what
-  your manuscript should look like.  Papers are required to conform to
-  all the directions reported in this document.  
-By using the provided \LaTeX\ and \BibTeX\ styles ({\small\tt acl2016.sty}, {\small\tt acl2016.bst}), the required formatting will be enabled by default.
+
+  Semantic evaluation of Machine Translation (MT) systems has become
+  a necessity, given recent interest in semantics-aware MT systems.
+  We present a novel human semantic evaluation measure for MT, Human
+  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
+  Making progress over previous work, the proposed measure takes into account
+  a wider range of semantic phenomena and does not rely on semantic annotation
+  of the MT output, whose interpretation is often unclear. 
+  We experiment with four language pairs, demonstrating the broad applicability of
+  the measure, as well as its reliability, reflected by the inter-annotator agreement
+  rates.
+
 \end{abstract}
 
-\section{ Human Semantic Evaluation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}\label{sec:intro}
+
+Human judgement is the central, and perhaps ultimate, criterion for estimating the
+quality of an MT system.
+Nevertheless, common measures for MT evaluation, such as adequacy and frequency judgements
+or the relative ranking of several possible translations, are problematic in two ways.
+First, as the quality of translation is determined by multiple factors, it is difficult
+to assign a single number that quantifies the quality of the entire sentence, as is reflected
+by the diminishing inter-annotator agreement rates of human ranking measures
+as a function of the sentence length \cite{Bojar:2011}.
+Second, a global quality score does not localize the errors in the translation output,
+and hence cannot inform measures to overcome these errors.
+This problem can be addressed by using scores that decompose over
+the sub-parts of the sentence. 
+However, common decompositions into n-grams or even individual words, as employed by the
+popular BLEU \cite{Papineni:2002} and METEOR measures \cite{Banerjee:2005},
+are most suitable for evaluating shallow MT approaches and are
+less instructive for more structurally-aware methods, which are at
+the forefront of MT research today.
+
+These difficulties have motivated the definition of a human measure that decomposes over
+the semantic structure of the sentence, rather than its n-grams. The most notable attempt of
+this brand is the HMEANT
+measure \cite{LoWu2010,lo2011structured,LoWu2011b,LoWu2012}, which quantifies the
+similarity the verbal argument structures of the translation output and reference.
+While appealing in addressing the two major shortcomings
+detailed above, HMEANT is unable to capture semantic differences which do not show up in
+the verb-argument structure, including several very frequent phenomena, such as
+inter-clause linkage, nominalizations and copula clauses (see \secref{sec:hmeant_comp}).
+Another shortcoming of MEANT lies with its reliance on semantically annotating MT output,
+which is often unintelligible. Finally, measures based on a comparison to a reference translation
+is necessarily incomplete, given the huge number of reasonable translations
+a sentence may have \cite{dryer_marcu:12}.
+
+In this work we propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
+a human evaluation measure that decomposes over the sentence's semantic units.
+The units are defined according to the 
+UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
+due to its cross-linguistic applicability, support for rapid annotation, and coverage
+of many fundamental semantic phenomena, such as verbal, nominal and adjectival
+argument structures and the inter-relations between them.
+In contrast to previous proposals of semantic MT evaluation,
+
+HUME operates by aggregating human assessments of the translation quality of the different
+semantic units in the source sentence.
+In thus only requires the semantic annotation of the source sentence, thus avoiding the 
+annotatation of machine-generated text, a difficult if not ill-defined task.
+By basing the measure on a direct comparison
+of the source and translation we avoid the difficulties engendered by using
+reference translations, albeit at the cost of imposing the additional requirement
+that the evaluator be proficient in both the source and target languages.
+See \secref{sec:} for a discussion of HUME's advantages realative to other measures.
+
+We conduct experiments with four language pairs, English-German, English-Polish,
+English-Romanian and English-Czech, and report efficiency (time of annotation) and
+inter-annotator agreement scores, as well as qualitative comparison of resulting
+HUME scores and BLEU, and crowd-sourced adequacy judgements. Results show that HUME obtains 
+agreements which are in par with standard sentence-level ranking measures, but unlike with them,
+the agreement does not diminish with the sentence length.
+
+\oa{summary of contribution}
+
+\oa{what is the purpose of the BLEU binning?}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Background}\label{sec:background}
+
+% MT evaluation
+\paragraph{MT Evaluation.}
+
+% semantic representation
+\paragraph{Semantic Representation.}
+Much interest has been devoted recently to broad-coverage semantic analysis of text.
+The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
+shares much of UCCA's motivation for defining a more complete semantic annotation.
+However, it is not optimal for our purposes as they do not ground the semantic symbols
+in the text, thus not straightwardly allowing a decomposition of the sentence into sub-units,
+and are more fine-grained and consequently more difficult to annotate.
+
+Several approaches have been proposed for grounded semantic representations,
+generally viewing semantic structures as bi-lexical dependencies. Examples include
+the Prague Treebank tectogrammatical layer \cite{hajic2012announcing}, as well as dependencies derived
+from Minimal Recursion Semantics representations \cite{oepen2006discriminant}.
+Despite their appeal as deep and intuitive means for semantic representations,
+these approaches also provide fine-grained distinctions and require linguistic expertise to annotate.
+%\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
+%but to our knowledge no results of this brand have been reported.
+
+UCCA (Universal Conceptual Cognitive Annotation) is a natural selection for purposes of MT evaluation
+for a number of reasons. 
+First, it diverges from the above approaches in representing
+semantic structure as a hierarchy of units, defined using coarse-grained and intuitive definitions,
+and consequently can be reliably annotated by non-experts, after as little as two hours
+of training \cite{marinotti2014}.
+Second, UCCA is a cross-linguistically applicable scheme, seeking to represent what is shared between languages,
+by building on an established framework for typological description called
+``Basic Linguistic Theory'' \cite[BLT]{Dixon:10a,Dixon:10b,Dixon:12},
+one of the most widely used frameworks for linguistic description.
+Indeed, the scheme's cross-linguistic applicability has been demonstrated through its
+application to annotation in English, French, German and Czech. 
+Third, the scheme has been shown to be semantically stable
+in translation: UCCA annotations of translated text usually contain the same set of relationships
+\cite{sulem2012}. This is an important feature for the semantic annotation, as it allows us to
+assume that the semantic units of the source indeed reflect a common layer of representation that
+should be preserved in translation, and so should be represented in the translation output too.
+The non-stability of verbal argument structures, which may be translated to a nominal or adjectival
+argument structure in another language, is one of the shortcomings of MEANT which HUME addresses.
+\oa{example?}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{The HUME Measure}\label{sec:hume}
+
+We do not distinguish fluency and adequacy judgements, as they are correlated anyway \cite{ccb}.
+
+\subsection{Comparison with HMEANT}\label{sec:hmeant_comp}
+
+For instance, consider the sentence ``A coronary angioplasty may not be technically possible'',
+which was translated by one of the systems we experimented with as
+``Eine koronare Angioplastie kann nicht technisch m{\"o}glich'', i.e., the main verb ``sein'' (``be'') is omitted.
+While this may be interpreted is a minor error, HMEANT will assign it a very low score, as it failed to
+translate the main verb.
+
+\oa{add statistics about frequency of nominalization and copula clauses and linkers annotation}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experiments}\label{sec:experiments}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conlusion}\label{sec:conclusion}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Human Semantic Evaluation}
+
+\oa{mention TER: using insertions, deletions, substitutions and shift of the entire sentence}
+
 \label{sec:sem-eval:human}
@@
 
-\bibliography{mt}
+\bibliography{main}
 \bibliographystyle{acl2016}

### 2016-03-07 paper/main.tex
zzz

@@
 argument structures and the inter-relations between them.
-In contrast to previous proposals of semantic MT evaluation,
 
@@
 
+While keeping a human in the loop is never a scalable solution, human evaluation is the
+fundamental measure of quality of MT output, and we therefore begin our inquiries here.
+
 \oa{summary of contribution}
-
-\oa{what is the purpose of the BLEU binning?}
 
@@
 
-% MT evaluation
-\paragraph{MT Evaluation.}
+\paragraph{Machine Translation Evaluation.}
+MT evaluation measures has been traditionally pursued through human-based
+evaluation that quantifies the adequacy of the translation in preserving the meaning
+of the source sentence, and its fluency as a sentence in the target language.
+Human evaluation is still accepted as the standard evaluation measure, and is
+estimated either in terms of the relative ranking of the outputs of multiple systems
+(as is standard in the WMT shared tasks \cite[inter alia]{wmt14}, or by assigning
+indiviaul adequacy/fluency scores for each sentence, recently streamlined by \newcite{graham2015}.
+The problem with these measures is that they are costly to conduct on large scale, as
+they require a human in the loop.
+
+While providing the cleanest signal, human-based evaluation are not ideal for cases where
+a wide variety of systems needs to be evaluated (e.g., with many different settings of hyper-parameters),
+which has motivated the construction of cheaper methods that correlate with human
+judgement. A common approach is to compare the MT output with reference translations,
+created or verified by an expert. Methods often rely on some sort of string matching
+methods, such as overlap of n-grams, as with the BLEU and METEOR measures, or edit
+distance. While these measures greatly supported MT research by allowing to cheply
+conduct many MT experiments, they suffer from two principal problems.
+First, they rely on comparison to reference translations, which reflect a small
+number out of the enormous number of ways to translate a sentence and are thus
+inherently incomplete. Second, they decompose over the words and n-grams of the sentence,
+and not over their lexical units, and are thus better suited for localizing errors
+when evaluating shallow MT methods, and are less telling for identifying the semantic
+nature of translation errors in the output.
+
+To address the second issue, more structure-aware methods were proposed, with the
+purpose of quantifying how different the reference and translation output are in terms
+of their {\it structure}. \newcite{LiuGildea2005} took a syntactic approach, checking
+syntactic similarity between the reference and translation\oa{verify!}. \newcite{Owczarzak2008}
+took a similar approach using lexical-functional grammar structures.
+
+Comparison of semantic structures with a reference translation
+was conducted by \newcite{gimenezMarquez2007,gimenezMarquez2008},
+who incorapoted them as features that were combined with lexical and morphosyntactic features.
+Perhaps the most notable attempt at a semantic MT evaluation measure is MEANT \cite{lo2011meant} and
+its human variant HMEANT, which compares the reference and translation in terms of their semantic
+role labeling structure. The measure shares much of HUME's motivation for providing a measure that
+decomposes over semantic units, rather than strings. Nevertheless, HMEANT suffers from several major
+shortcomings, in that it is not able to capture much frequent phenomena that is not expressed
+in the semantic structures, in addition to not addressing the problems incurred by using references.
+HUME addresses both of these shortcomings, see \secref{sec:hmeant_comp}.
 
@@
 \oa{add statistics about frequency of nominalization and copula clauses and linkers annotation}
+
+\oa{what is the purpose of the BLEU binning?}
 

### 2016-03-07 paper/main.tex
some iaa

@@
 \newcommand{\oa}[1]{\footnote{\color{red} #1}}
+\newcommand{\bh}[2]{\footnote{\color{blue} #1}}
 
@@
 \section{Experiments}\label{sec:experiments}
+\subsection{Data Sets and Translation Systems}
+\begin{itemize}
+  \item Briefly describe translation systems.
+  \item Source texts (no need to describe references)
+\end{itemize}
+
+\subsection{Annotation Process}
+\begin{itemize}
+  \item Numbers of annotators, IDs
+  \item Length of annotation, timings
+  \item Sentences annotated
+\end{itemize}
+
+\subsection{Inter-Annotator Agreement}
+\begin{itemize}
+  \item Multiple annotation stats
+  \item Kappas -- ROG, AB
+  \item Kappa versus length
+  \item Kappa versus bleu?
+\end{itemize}
+
+In order to assess the consistency of the annotation, we measure the Inter-Annotator
+Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
+For German, Polish and Romanian we used two annotators, so we report IAA on the overlap
+between these annotators. For Czech we used one main annotator, plus several other extra
+annotators to provide IAA.
+
+To calculate Kappa, we consider the annotation task as a classification task on 
+nodes. We only consider nodes where there are at least two annotations from different
+annotators and we treat nodes with more than two annotations as a set of paired
+annotations\bh{This may not be correct, so should/could we do Kappa across the 
+variable multiple annotations on Czech}. We consider Kappa separately on the lexical
+nodes (annotated as red, orange or green) and the structural nodes (annotated
+as acceptable or bad) since there should not be any confusion between these two types of
+nodes. The small number of nodes which one annotator had labelled as structural and 
+the other as lexical were ignored for the purposes of IAA, since such cases of annotator 
+error could be caught by improved tool support.  The Kappas are shown in \tabref{tab:iaa}
+\begin{table}[!ht]
+\begin{center}
+\begin{tabular}{l|cccc}
+ & cs & de & pl & ro \\
+\hline
+Sentences & & & & \\
+\hline
+Lexical nodes & & & & \\
+Kappa & & & & \\
+\hline
+Structural nodes & & & & \\
+Kappa & & & & \\
+\end{tabular}
+\caption{Inter-annotator agreement for the multiply-annotated portions of the data, as
+measured by Cohen's Kappa. }
+\label{tab:iaa}
+\end{center}
+\end{table}
+
+To assess whether, as we hoped, HUME would make it easier to reliably annotate long sentences,
+we binned the sentences according to length and measured Kappa on each of the bins.
 

### 2016-03-08 paper/main.tex
more

--- 2016-03-07 paper/main.tex
@@
 
-UCCA (Universal Conceptual Cognitive Annotation) is a natural selection for purposes of MT evaluation
+UCCA (Universal Conceptual Cognitive Annotation) is a natural pick for purposes of MT evaluation
 for a number of reasons. 
@@
 argument structure in another language, is one of the shortcomings of MEANT which HUME addresses.
-\oa{example?}
 
@@
 We do not distinguish fluency and adequacy judgements, as they are correlated anyway \cite{ccb}.
+
+\subsection{Annotation Guidelines}\label{sec:guidelines}
+
+\subsection{Composite Score}\label{sec:score}
+
+composite score (uniform, sub-sets, weighted rather not -- according to categories [or empirically according to agreement])
 
@@
 
-\oa{add statistics about frequency of nominalization and copula clauses and linkers annotation}
-
-\oa{what is the purpose of the BLEU binning?}
+We conducted an analysis of the English UCCA Wikipedia
+corpus\footnote{\url{http://www.cs.huji.ac.il/~oabend/ucca.html}}, comprising
+of 5324 sentences,
+in order to assess the frequency of some of the major phenomena HMEANT fails to address,
+Coppula cluases are treated in HMEANT simply as instances of the main verb ``be'', which
+generally does not convey the meaning of these clauses. They appear in 21.7\% of the sentences,
+according to a conservative estimates that only considers non-auxilliary instances of ``be''.
+Nominative argument structures, which are completely ignored under HMEANT are in fact highly
+pervasive, appearing in 48.7\% of the sentences. Linkers expressing an inter-relation between
+clauses (mainly discourse markers and coordinating conjunctions) appear in 56\% of the
+sentences\footnote{Argument structures and linkers are explicitly marked in UCCA. We identified
+  non-auxilliary instances of ``be'' and nouns using the NLTK standard tagger.}.
+We are not aware of any empirical evaluation that demonstrates that verb argument structure,
+taken alone, capture the crux of the sentence semantics meanings, and so the coverage of
+further phenomena by HUME will first allow to determine empirically
+(rather than stipulate), what semantic structures are important
+to preserve for MT translation, and which are less important
 
@@
 \section{Experiments}\label{sec:experiments}
+
+\subsection{Data}\label{sec:data}
+
+\subsection{Efficient of Annotation}\label{sec:efficiency}
+
+\subsection{Agreement across Languages }\label{sec:efficiency}
+
+\oa{what is the purpose of the BLEU binning? is it because low quality and high
+quality translations often diverge in their agreement scores?}
+
+\subsection{Discussion of Disagreements}\label{sec:disagreements}
+
+\subsection{Comparison with Crowdsourced Adequacy Judgements}\label{sec:adequacy}
 

### 2016-03-08 paper/main.tex

@@
 \section{Experiments}\label{sec:experiments}
+\subsection{Data Sets and Translation Systems}
+\begin{itemize}
+  \item Briefly describe translation systems.
+  \item Source texts (no need to describe references)
+\end{itemize}
+
+\subsection{Annotation Process}
+\begin{itemize}
+  \item Numbers of annotators, IDs
+  \item Length of annotation, timings
+  \item Sentences annotated
+\end{itemize}
 
@@
 
-\subsection{Efficient of Annotation}\label{sec:efficiency}
-
-\subsection{Agreement across Languages }\label{sec:efficiency}
+\subsection{Efficiency of Annotation}\label{sec:efficiency}
+
+\subsection{Inter-Annotator Agreement}
+\begin{itemize}
+  \item Multiple annotation stats
+  \item Kappas -- ROG, AB
+  \item Kappa versus length
+  \item Kappa versus bleu?
+\end{itemize}
 
@@
 quality translations often diverge in their agreement scores?}
+
+In order to assess the consistency of the annotation, we measure the Inter-Annotator
+Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
+For German, Polish and Romanian we used two annotators, so we report IAA on the overlap
+between these annotators. For Czech we used one main annotator, plus several other extra
+annotators to provide IAA.
+
+To calculate Kappa, we consider the annotation task as a classification task on 
+nodes. We only consider nodes where there are at least two annotations from different
+annotators and we treat nodes with more than two annotations as a set of paired
+annotations\bh{This may not be correct, so should/could we do Kappa across the 
+variable multiple annotations on Czech}. We consider Kappa separately on the lexical
+nodes (annotated as red, orange or green) and the structural nodes (annotated
+as acceptable or bad) since there should not be any confusion between these two types of
+nodes. The small number of nodes which one annotator had labelled as structural and 
+the other as lexical were ignored for the purposes of IAA, since such cases of annotator 
+error could be caught by improved tool support.  The Kappas are shown in \tabref{tab:iaa}
+\begin{table}[!ht]
+\begin{center}
+\begin{tabular}{l|cccc}
+ & cs & de & pl & ro \\
+\hline
+Sentences & & & & \\
+\hline
+Lexical nodes & & & & \\
+Kappa & & & & \\
+\hline
+Structural nodes & & & & \\
+Kappa & & & & \\
+\end{tabular}
+\caption{Inter-annotator agreement for the multiply-annotated portions of the data, as
+measured by Cohen's Kappa. }
+\label{tab:iaa}
+\end{center}
+\end{table}
+
+To assess whether, as we hoped, HUME would make it easier to reliably annotate long sentences,
+we binned the sentences according to length and measured Kappa on each of the bins.
 

### 2016-03-08 paper/main.tex
a bit about the system

@@
 
+For each of the four language pairs under consideration (i.e. English to 
+Czech, German, Polish and Romanian) we built phrase-based MT systems
+using Moses \cite{Koehn:2007}. We used large training sets extracted from OPUS \cite{tiedemann:2009}, and
+the data sets released for the WMT14 medical translation task \cite{bojar-EtAl:2014:W14-33}, to provide between
+45 and 85 million sentences of training data, depending on language pair.
+
+These translation systems were used to translate texts derived from both NHS 24\footnote{\url{http://www.nhs24.com/}} and 
+Cochrane\footnote{\url{http://www.cochrane.org/}} into the four languages. The first organisation is a public body
+providing healthcare and health-service related information in Scotland, whereas the second is an international NGO which 
+provides independent systematic reviews on health-related research. From NHS 24, we selected the segments from their
+``Health A-Z'' which is part of the NHS Inform website, and the Cochrane texts come from their plain language summaries
+and abstracts.\bh{How much more do we need? Perhaps in the intro we could justify our focus on public health text
+(accuracy!). The test corpora should have a public URL by the time we submit, although of course the source texts were 
+already public anyway, and Lexi did some selection of examples for use in this expt. -- do we have to explain this?}
+
 \subsection{Annotation Process}
@@
 \end{itemize}
+
+For the annotation process, we recruited seven annotators, two for each of the target languages apart from Czech. They were
+all native speakers of the target languages and fluent in English. For Czech, we also recruited a  group of additional annotators
+to annotate small numbers of sentences each for the purpose of calculating inter-annotator agreement. In \tabref{tab:annot}
+we show the total number of sentences and nodes annotated by each annotator, where we consider the additional Czech
+annotators as a single annotator.
 

### 2016-03-09 paper/main.tex
aaa

--- 2016-03-08 paper/main.tex
@@
 
-  Semantic evaluation of Machine Translation (MT) systems has become
-  a necessity, given recent interest in semantics-aware MT systems.
+  Semantic evaluation of Machine Translation (MT) systems, informative as for
+  semantic discrepencies between the translation and the source text,
+  has become a necessity, given recent interest in semantics-aware MT systems.
   We present a novel human semantic evaluation measure for MT, Human
@@
   the measure, as well as its reliability, reflected by the inter-annotator agreement
-  rates.
+  scores.
 
@@
 as a function of the sentence length \cite{Bojar:2011}.
-Second, a global quality score does not localize the errors in the translation output,
-and hence cannot inform measures to overcome these errors.
-This problem can be addressed by using scores that decompose over
-the sub-parts of the sentence. 
+Second, a sentence-level quality score does not indicate what parts of the sentence
+are erroneously translated in the output, and hence cannot inform developers in overcoming these errors.
+This problem can be addressed by using scores that decompose over the sub-parts of the sentence. 
 However, common decompositions into n-grams or even individual words, as employed by the
@@
 the semantic structure of the sentence, rather than its n-grams. The most notable attempt of
-this brand is the HMEANT
-measure \cite{LoWu2010,lo2011structured,LoWu2011b,LoWu2012}, which quantifies the
-similarity the verbal argument structures of the translation output and reference.
-While appealing in addressing the two major shortcomings
-detailed above, HMEANT is unable to capture semantic differences which do not show up in
+this type is the MEANT measure, and its human variant HMEANT
+\cite{lo2010evaluating,lo2011structured,Lo2011meant,lo2013meant}, which quantifies the
+similarity between the output translation and a reference translation in terms of the similarity
+between their verbal argument structures
+While the measure addresses the two major shortcomings
+detailed above, it is unable to capture semantic differences which do not show up in
 the verb-argument structure, including several very frequent phenomena, such as
 inter-clause linkage, nominalizations and copula clauses (see \secref{sec:hmeant_comp}).
-Another shortcoming of MEANT lies with its reliance on semantically annotating MT output,
+Another shortcoming of MEANT is its reliance on semantically annotating MT output,
 which is often unintelligible. Finally, measures based on a comparison to a reference translation
-is necessarily incomplete, given the huge number of reasonable translations
-a sentence may have \cite{dryer_marcu:12}.
+are necessarily incomplete, as they compare to a small number of references
+out of the the huge number of possible translations a sentence may have \cite{dryer_marcu:12}.
 
@@
 a human evaluation measure that decomposes over the sentence's semantic units.
-The units are defined according to the 
+Semantic units are defined according to the 
 UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
@@
 The semantic framework that we base our evaluations on is called 
-Universal Conceptual Cognitive Annotation \shortcite{abend-rappoport:2013:ACL2013}.  
+Universal Conceptual Cognitive Annotation \shortcite{abend2013universal}.  
 UCCA has been developed using linguistic theories about 

### 2016-03-09 paper/main.tex
bbb

@@
 are erroneously translated in the output, and hence cannot inform developers in overcoming these errors.
+
+These problems were partially addressed by measures that decompose over sub-parts of the sentence,
+most commonly over its words or n-grams, quantifying the overlap of the sub-parts with a reference
+translation. More recent work on MT evaluation, most notably the MEANT measure \cite{lo2011structured},
+proposed measures that decompose over semantically defined units,
+quantifying the similarity of the output and the reference in terms of their verb argument structures.
+While allowing a localization of the semantic errors in the translation,
+focusing on verb argument structures misses out on many very pervasive semantic phenomena,
+such as inter-clause linkage and nominal argument structures, and thus provides only a limited
+perspective on the semantic similarity between the output and reference.
+Moreover, assigning a semantic structure to MT output, which is often low in quality,
+is difficult and arguably ill-defined.
+
+In this work we propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
+a human evaluation measure that decomposes over the semantic units of the sentence.
+Semantic units are defined according to the 
+UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
+due to its cross-linguistic applicability, support for rapid annotation, and coverage
+of many fundamental semantic phenomena, such as verbal, nominal and adjectival
+argument structures and the inter-relations between them.
+
+HUME operates by aggregating human assessments of the translation quality of individual
+semantic units in the source sentence.
+Importantly, HUME only requires the semantic annotation of the source sentence, thus avoiding the 
+annotatation of machine-generated text, a difficult if not ill-defined task, and allowing the re-use
+of the source semantic annotation for measuring the quality of different translations of the same source sentence.
+By basing the measure on a direct comparison
+of the source and translation we avoid the difficulties engendered by using
+reference translations, albeit at the cost of imposing the additional requirement
+that the evaluator be proficient in both the source and target languages.
+See \secref{sec:hmeant_comp} for further discussion.
+
+We conduct experiments with four language pairs, English-German, English-Polish,
+English-Romanian and English-Czech, and report efficiency (time of annotation) and
+inter-annotator agreement scores, as well as qualitative comparison of the resulting
+HUME scores with crowd-sourced adequacy judgements. Results show that HUME obtains 
+agreements which are on par with those of standard sentence-level ranking measures,
+but unlike them, HUME's agreement does not diminish with the sentence length.
+
+While human evaluation is not a scalable solution,
+it provides the foundations and justification for defining a measure.
+The paper thus focuses on human evaluation, but concludes
+by outlining ongoing work to construct semi- and fully-automatic variants of HUME.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Background}\label{sec:background}
+
+While the measure addresses the two major shortcomings
+detailed above, it is unable to capture semantic differences which do not show up in
+the verb-argument structure, including several very frequent phenomena, 
+Another shortcoming of MEANT is its reliance on semantically annotating MT output,
+which is often unintelligible. Finally, measures based on a comparison to a
+
 This problem can be addressed by using scores that decompose over the sub-parts of the sentence. 
@@
 the verb-argument structure, including several very frequent phenomena, such as
-inter-clause linkage, nominalizations and copula clauses (see \secref{sec:hmeant_comp}).
+inter-clause linkage, nominalizations and copula clauses
+(see \secref{sec:hmeant_comp}).
 Another shortcoming of MEANT is its reliance on semantically annotating MT output,
-which is often unintelligible. Finally, measures based on a comparison to a reference translation
+which is often unintelligible. Finally, measures based on a comparison to a
+reference translation
 are necessarily incomplete, as they compare to a small number of references
-out of the the huge number of possible translations a sentence may have \cite{dryer_marcu:12}.
-
-In this work we propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
-a human evaluation measure that decomposes over the sentence's semantic units.
-Semantic units are defined according to the 
-UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
-due to its cross-linguistic applicability, support for rapid annotation, and coverage
-of many fundamental semantic phenomena, such as verbal, nominal and adjectival
-argument structures and the inter-relations between them.
-
-HUME operates by aggregating human assessments of the translation quality of the different
-semantic units in the source sentence.
-In thus only requires the semantic annotation of the source sentence, thus avoiding the 
-annotatation of machine-generated text, a difficult if not ill-defined task.
-By basing the measure on a direct comparison
-of the source and translation we avoid the difficulties engendered by using
-reference translations, albeit at the cost of imposing the additional requirement
-that the evaluator be proficient in both the source and target languages.
-See \secref{sec:} for a discussion of HUME's advantages realative to other measures.
-
-We conduct experiments with four language pairs, English-German, English-Polish,
-English-Romanian and English-Czech, and report efficiency (time of annotation) and
-inter-annotator agreement scores, as well as qualitative comparison of resulting
-HUME scores and BLEU, and crowd-sourced adequacy judgements. Results show that HUME obtains 
-agreements which are in par with standard sentence-level ranking measures, but unlike with them,
-the agreement does not diminish with the sentence length.
-
-While keeping a human in the loop is never a scalable solution, human evaluation is the
-fundamental measure of quality of MT output, and we therefore begin our inquiries here.
-
-\oa{summary of contribution}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Background}\label{sec:background}
+out of the the huge number of possible translations a sentence may have
+\cite{dreyer2012hyter}.
 

### 2016-03-09 scripts/iaa.py
add new de data

@@ -28,6 +28,7 @@ def print_overall_iaa(by_lang, args):
     cm = ConfusionMatrix(by_label['mt_label_x'], by_label['mt_label_y'], \
       true_name="annot_1", pred_name="annot_2")
     print(cm)
+    print(cm.stats())
     print("Kappa: %7.5f" % cm.stats()['overall']['Kappa'])
 
 def print_iaa_sentence_detail(agree, detail_file):

### 2016-03-09 scripts/reader.py

@@ -42,6 +42,9 @@ def get_sentences(filenames):
         if not line.startswith("==="):
           break # end of file
         filename = read_string(dfh)
+        # hack to remove end of file name in de2
+        if filename.endswith("_b4"):
+          filename = filename[:-3]
         lang = filename[-3:-1]
         annot_id = filename[-3:]
         sent_id = eval(dfh.readline())

### 2016-03-09 scripts/iaa.py
do my own kappa calculation

@@ -6,9 +6,29 @@ import logging
 import sys
 import pandas
 
+import numpy as np
 from  pandas_confusion import ConfusionMatrix
 
-LANGCODES = ("Romanian","ro"), ("Polish", "pl")
+LANGCODES = ("Romanian","ro"), ("Polish", "pl"), ("German", "de")
+
+def get_kappa(cm):
+  # The pandas_confusion kappa implementation crashes on some platforms,
+  # so I reimplement it here.
+  df = cm.to_dataframe()
+  N = df.sum().sum()
+  #print(N)
+  sum_diag = np.trace(df)
+  #print(sum_diag)
+  sum_row = df.sum(axis=0)
+  sum_col = df.sum(axis=1)
+  p_e = float(sum(sum_row*sum_col)) / (N*N)
+  #print(sum_row)
+  #print(sum_col)
+  #print (p_e)
+  p_o = float(sum_diag) / N
+  #print(p_o)
+  kappa = (p_o - p_e) / (1 - p_e)
+  return kappa
 
 def print_overall_stats(by_lang,args):
   print("Counts of doubly annotated nodes") 
@@ -28,8 +48,10 @@ def print_overall_iaa(by_lang, args):
     cm = ConfusionMatrix(by_label['mt_label_x'], by_label['mt_label_y'], \
       true_name="annot_1", pred_name="annot_2")
     print(cm)
-    print(cm.stats())
-    print("Kappa: %7.5f" % cm.stats()['overall']['Kappa'])
+    kappa = get_kappa(cm)
+    #print(cm.to_dataframe())
+    #print(cm.stats())
+    print("Kappa: %7.5f" % kappa)
 
 def print_iaa_sentence_detail(agree, detail_file):
 #TODO sentence id, matches per sentence, pc a/b, pc r/o/g. src, tgt, ucca stats

### 2016-03-09 scripts/iaa_binned.py

@@ -14,9 +14,9 @@ import sys
 import pandas
 
 from  pandas_confusion import ConfusionMatrix
+from iaa import get_kappa, LANGCODES
 
 LOG = logging.getLogger(__name__)
-LANGCODES = ("Romanian","ro"), ("Polish", "pl")
 
 def do_bins(agree, first_bin_start, bin_size, field_name, csv_file, graph_file, name):
   LOG.info("Calculating kappas, binning by " +name)
@@ -46,7 +46,7 @@ def do_bins(agree, first_bin_start, bin_size, field_name, csv_file, graph_file, 
             if len(match_labels) == 2:
               cm = ConfusionMatrix(selected['mt_label_x'], selected['mt_label_y'], \
                 true_name="annot_1", pred_name="annot_2")
-              kappa = cm.stats()['overall']['Kappa']
+              kappa = get_kappa(cm) #cm.stats()['overall']['Kappa']
               if kappa < 0:
                 print(cm)
             elif match_labels[0]:

### 2016-03-09 scripts/iaa_binned.py
add German

@@ -25,19 +25,23 @@ def do_bins(agree, first_bin_start, bin_size, field_name, csv_file, graph_file, 
     writer = csv.writer(ofh)
     writer.writerow(("lang","group","bin_start","bin_end","count","kappa"))
     for lang_name,lang in LANGCODES:
+      by_lang = agree[agree["lang"] == lang]
       LOG.info("Considering " + lang_name)
-      max_length = max(agree[agree['lang'] == lang][field_name])
+      max_length = max(by_lang[field_name])
       #min_length = min(agree[agree['lang'] == lang][field_name])
       kappas = { g: [] for g,_ in groups}
       bins = []
       bin_start = first_bin_start
       while bin_start < max_length:
         bin_end = bin_start + bin_size
+        if len(by_lang[by_lang[field_name] >= bin_start]) == 0: break
+        while len(by_lang[(by_lang[field_name] >= bin_start) & (by_lang[field_name] < bin_end)]) == 0:
+          bin_end += bin_size
         LOG.debug("Bin start: {}; Bin end: {}".format(bin_start,bin_end))
         for group_name, group in groups:
-          selected = agree[(agree["lang"] == lang) & \
-          (agree["mt_label_x"].isin(group)) & (agree["mt_label_y"].isin(group))  & \
-          (agree[field_name] >= bin_start) & (agree[field_name] < bin_end)]
+          selected = by_lang[ \
+          (by_lang["mt_label_x"].isin(group)) & (by_lang["mt_label_y"].isin(group))  & \
+          (by_lang[field_name] >= bin_start) & (by_lang[field_name] < bin_end)]
           LOG.debug("Selected {} nodes".format(len(selected)))
           if not len(selected):
             kappa = 0
@@ -57,7 +61,7 @@ def do_bins(agree, first_bin_start, bin_size, field_name, csv_file, graph_file, 
           writer.writerow((lang, group_name, str(bin_start), str(bin_end), len(selected), kappa))
           kappas[group_name].append(kappa)
         bins.append((bin_start,bin_end))
-        bin_start += bin_size
+        bin_start = bin_end
 
       # Plotting to file
       fig,ax = plt.subplots()

### 2016-03-09 scripts/iaa_binned.py
Only rog and ab agreement.

@@ -20,7 +20,8 @@ LOG = logging.getLogger(__name__)
 
 def do_bins(agree, first_bin_start, bin_size, field_name, csv_file, graph_file, name):
   LOG.info("Calculating kappas, binning by " +name)
-  groups = (("ab", ("A", "B")), ("rog", ("R", "O", "G")), ("all", ("A", "B","R", "O", "G")))
+  #groups = (("ab", ("A", "B")), ("rog", ("R", "O", "G")), ("all", ("A", "B","R", "O", "G")))
+  groups = (("struct", ("A", "B")), ("lex", ("R", "O", "G")))
   with open(csv_file, "w") as ofh:
     writer = csv.writer(ofh)
     writer.writerow(("lang","group","bin_start","bin_end","count","kappa"))
@@ -66,12 +67,13 @@ def do_bins(agree, first_bin_start, bin_size, field_name, csv_file, graph_file, 
       # Plotting to file
       fig,ax = plt.subplots()
       index = np.arange(len(bins))
-      cols = ('r', 'g', 'b')
+      cols = ('r', 'b', 'g')[:len(groups)]
       offset = 0.0
       rects = []
+      width = 1.0 / len(groups) - 0.03
       for (group,_),col in zip(groups,cols):
-        rects.append(ax.bar(index + offset, kappas[group], width=0.3, color = col))
-        offset += 0.3
+        rects.append(ax.bar(index + offset, kappas[group], width=width, color = col))
+        offset += width
       ax.set_yticks(np.arange(0,1.1,0.1))
       ax.set_xticks(index + 0.5)
       ax.set_xticklabels(["{}-{}".format(n,m) for n,m in bins])

### 2016-03-09 paper/main.tex
iaa figures
background

@@
 
-While the measure addresses the two major shortcomings
-detailed above, it is unable to capture semantic differences which do not show up in
-the verb-argument structure, including several very frequent phenomena, 
-Another shortcoming of MEANT is its reliance on semantically annotating MT output,
-which is often unintelligible. Finally, measures based on a comparison to a
-
-This problem can be addressed by using scores that decompose over the sub-parts of the sentence. 
-However, common decompositions into n-grams or even individual words, as employed by the
-popular BLEU \cite{Papineni:2002} and METEOR measures \cite{Banerjee:2005},
-are most suitable for evaluating shallow MT approaches and are
-less instructive for more structurally-aware methods, which are at
-the forefront of MT research today.
-
-These difficulties have motivated the definition of a human measure that decomposes over
-the semantic structure of the sentence, rather than its n-grams. The most notable attempt of
-this type is the MEANT measure, and its human variant HMEANT
-\cite{lo2010evaluating,lo2011structured,Lo2011meant,lo2013meant}, which quantifies the
-similarity between the output translation and a reference translation in terms of the similarity
-between their verbal argument structures
-While the measure addresses the two major shortcomings
-detailed above, it is unable to capture semantic differences which do not show up in
-the verb-argument structure, including several very frequent phenomena, such as
-inter-clause linkage, nominalizations and copula clauses
-(see \secref{sec:hmeant_comp}).
-Another shortcoming of MEANT is its reliance on semantically annotating MT output,
-which is often unintelligible. Finally, measures based on a comparison to a
-reference translation
-are necessarily incomplete, as they compare to a small number of references
-out of the the huge number of possible translations a sentence may have
-\cite{dreyer2012hyter}.
+%While the measure addresses the two major shortcomings 
+%detailed above, it is unable to capture semantic differences which do not show up in 
+%the verb-argument structure, including several very frequent phenomena, 
+%Another shortcoming of MEANT is its reliance on semantically annotating MT output,
+%which is often unintelligible. Finally, measures based on a comparison to a
+
+%These difficulties have motivated the definition of a human measure that decomposes over
+%the semantic structure of the sentence, rather than its n-grams. The most notable attempt of
+%this type is the MEANT measure, and its human variant HMEANT
+%\cite{lo2010evaluating,lo2011structured,Lo2011meant,lo2013meant}, which quantifies the
+%similarity between the output translation and a reference translation in terms of the similarity
+%between their verbal argument structures
+%While the measure addresses the two major shortcomings
+%detailed above, it is unable to capture semantic differences which do not show up in
+%the verb-argument structure, including several very frequent phenomena, such as
+%inter-clause linkage, nominalizations and copula clauses
+%(see \secref{sec:hmeant_comp}).
+%Another shortcoming of MEANT is its reliance on semantically annotating MT output,
+%which is often unintelligible. Finally, measures based on a comparison to a
+%reference translation
+%are necessarily incomplete, as they compare to a small number of references
+%out of the the huge number of possible translations a sentence may have
+%\cite{dreyer2012hyter}.
 
@@
 evaluation that quantifies the adequacy of the translation in preserving the meaning 
-of the source sentence, and its fluency as a sentence in the target language.
+of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
 Human evaluation is still accepted as the standard evaluation measure, and is
 estimated either in terms of the relative ranking of the outputs of multiple systems
-(as is standard in the WMT shared tasks \cite[inter alia]{wmt14}, or by assigning
-indiviaul adequacy/fluency scores for each sentence, recently streamlined by \newcite{graham2015}.
-The problem with these measures is that they are costly to conduct on large scale, as
-they require a human in the loop.
-
-While providing the cleanest signal, human-based evaluation are not ideal for cases where
-a wide variety of systems needs to be evaluated (e.g., with many different settings of hyper-parameters),
+as is standard in the WMT shared tasks, e.g., \cite{bojar2015findings}, or by assigning
+indiviaul adequacy/fluency scores for each sentence, a procedure recently improved
+by \newcite{graham2015improving}.
+
+While providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
+and is unfeasible for a wide variety of systems or hyper-parameter settings are to be evaluated.
 which has motivated the construction of cheaper methods that correlate with human
-judgement. A common approach is to compare the MT output with reference translations,
-created or verified by an expert. Methods often rely on some sort of string matching
-methods, such as overlap of n-grams, as with the BLEU and METEOR measures, or edit
-distance. While these measures greatly supported MT research by allowing to cheply
-conduct many MT experiments, they suffer from two principal problems.
-First, they rely on comparison to reference translations, which reflect a small
-number out of the enormous number of ways to translate a sentence and are thus
-inherently incomplete. Second, they decompose over the words and n-grams of the sentence,
-and not over their lexical units, and are thus better suited for localizing errors
+judgement.
+
+Scalability has been addressed by scores that decompose over the sub-parts of
+the translation, and quantify how many of these sub-parts appear in a reference translation,
+whose validity has been manually confirmed. The most commonly used measures decompose
+over n-grams or even individual words, with examples such as
+the BLEU \cite{Papineni:2002} and METEOR measures \cite{Banerjee:2005}.
+Other methods employ a string edit distance for determining the similarity between the
+reference and translation \cite{snover2006study}.
+While these measures stimulated much progress in MT research by allowing
+to evaluate experiments no a massive scale, they 
+decompose over the words and n-grams of the sentence,
+and are thus better suited for localizing errors
 when evaluating shallow MT methods, and are less telling for identifying the semantic
@@
 
-To address the second issue, more structure-aware methods were proposed, with the
+%These methods are most suitable for evaluating shallow MT approaches and are
+%less instructive as to what aspects of the reference's meaning were not preserved
+%in the translation output.
+%A common approach is to compare the MT output with reference translations,
+%created or verified by an expert. Methods often rely on some sort of string matching
+%methods, such as overlap of n-grams, as with the BLEU and METEOR measures, or edit
+%distance.
+%First, they rely on comparison to reference translations, which reflect a small
+%number out of the enormous number of ways to translate a sentence and are thus
+%inherently incomplete.
+
+To address this shortcoming, more structure-aware methods were proposed, with the 
 purpose of quantifying how different the reference and translation output are in terms 
-of their {\it structure}. \newcite{LiuGildea2005} took a syntactic approach, checking
-syntactic similarity between the reference and translation\oa{verify!}. \newcite{Owczarzak2008}
-took a similar approach using lexical-functional grammar structures.
-
-Comparison of semantic structures with a reference translation
-was conducted by \newcite{gimenezMarquez2007,gimenezMarquez2008},
-who incorapoted them as features that were combined with lexical and morphosyntactic features.
-Perhaps the most notable attempt at a semantic MT evaluation measure is MEANT \cite{lo2011meant} and
-its human variant HMEANT, which compares the reference and translation in terms of their semantic
-role labeling structure. The measure shares much of HUME's motivation for providing a measure that
-decomposes over semantic units, rather than strings. Nevertheless, HMEANT suffers from several major
-shortcomings, in that it is not able to capture much frequent phenomena that is not expressed
-in the semantic structures, in addition to not addressing the problems incurred by using references.
-HUME addresses both of these shortcomings, see \secref{sec:hmeant_comp}.
-
-% semantic representation
+of their {\it semantic structure}. \newcite{liu2005syntactic} took a syntactic approach, 
+quantifying the similarity between the syntactic structures of the reference and translation.
+\newcite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
+\newcite{gimenez2007linguistic} proposed an evaluation measure that incorapotes features
+that capture the overlap between the output and reference in terms of their
+semantic (argument structure phenomena), lexical and morphosyntactic features.
+
+Perhaps the most notable attempt at a semantic MT evaluation measure is MEANT and
+its human variant HMEANT, which compares the reference and translation in terms of the similarity
+of their semantic role labeling structure. HMEANT shares much of HUME's motivation for providing a
+measure that decomposes over semantic units, rather than strings, but focuses on verbal
+argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
+relations and copula clauses. HUME further diverges from both MEANT and the more shallow BLEU and
+METEOR measures, in not requiring a reference, as it compares the source directly with the output
+translation. We see this difference as crucial where semantic analysis is concerned, as it
+is often impossible to assign a semantic structure to a low quality MT output sentence.
+See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences between HUME
+and HMEANT.
+
 \paragraph{Semantic Representation.}
-Much interest has been devoted recently to broad-coverage semantic analysis of text.
-The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
-shares much of UCCA's motivation for defining a more complete semantic annotation.
-However, it is not optimal for our purposes as they do not ground the semantic symbols
-in the text, thus not straightwardly allowing a decomposition of the sentence into sub-units,
-and are more fine-grained and consequently more difficult to annotate.
-
-Several approaches have been proposed for grounded semantic representations,
-generally viewing semantic structures as bi-lexical dependencies. Examples include
-the Prague Treebank tectogrammatical layer \cite{hajic2012announcing}, as well as dependencies derived
-from Minimal Recursion Semantics representations \cite{oepen2006discriminant}.
-Despite their appeal as deep and intuitive means for semantic representations,
-these approaches also provide fine-grained distinctions and require linguistic expertise to annotate.
-%\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
-%but to our knowledge no results of this brand have been reported.
-
-UCCA (Universal Conceptual Cognitive Annotation) is a natural pick for purposes of MT evaluation
-for a number of reasons. 
-First, it diverges from the above approaches in representing
-semantic structure as a hierarchy of units, defined using coarse-grained and intuitive definitions,
+UCCA (Universal Conceptual Cognitive Annotation) is a cross-linguistically applicable, lightweight
+scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
+whose leaves correspond to the words of the text.
+The graph’s nodes, called {\sc units} , are either terminals or several elements jointly
+viewed as a single entity according to some semantic or cognitive consideration. Edges bear one
+or more categories, indicating the role of the sub-unit in the relation the unit represents. UCCA
+currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
+relations between them. The most basic notion is the {\sc scene}, which describes a movement, an
+action or a state which is persistent in time. Each Scene contains one main relation, as well
+as one or more participants. For example, the sentence ``After graduating, Tom moved to NYC''
+contains two Scenes, whose main relations are ``graduating'' and ``John''. The participant ``John''
+is a part of both Scenes, while ``NYC'' only of the latter. Further categories account for
+relations between Scenes and the internal structures of complex participants and relations.\oa{add figure}
+
+% UCCA Motivation
+UCCA is a natural candidate for defining a semantic
+MT evaluation measure for a number of reasons. 
+First, it defines a coarse-grained and intuitive set of distinctions definitions,
 and consequently can be reliably annotated by non-experts, after as little as two hours
@@
 Second, UCCA is a cross-linguistically applicable scheme, seeking to represent what is shared between languages,
-by building on an established framework for typological description called
-``Basic Linguistic Theory'' \cite[BLT]{Dixon:10a,Dixon:10b,Dixon:12},
-one of the most widely used frameworks for linguistic description.
-Indeed, the scheme's cross-linguistic applicability has been demonstrated through its
+by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
+\cite{Dixon:10a,Dixon:10b,Dixon:12}, one of the most widely used frameworks for linguistic description.
+Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
 application to annotation in English, French, German and Czech. 
@@
 in translation: UCCA annotations of translated text usually contain the same set of relationships
-\cite{sulem2012}. This is an important feature for the semantic annotation, as it allows us to
-assume that the semantic units of the source indeed reflect a common layer of representation that
+\cite{sulem2012}. This is an indication that the scheme reflect a common layer of representation that
 should be preserved in translation, and so should be represented in the translation output too.
-The non-stability of verbal argument structures, which may be translated to a nominal or adjectival
-argument structure in another language, is one of the shortcomings of MEANT which HUME addresses.
+We note that that verbal argument structures are less stable in this sense, as a verbal
+argument structure may be translated or paraphrased to a nominal or adjectival
+argument structure. For example, ``after he graduated, Tom moved to NYC'' conveys a similar meaning
+to ``after graduation, Tom moved to NYC''.
+
+While most broad-coverage semantic work has focused on shallow semantic structures,
+predominantly argument structures annotated with their semantic roles, much recent
+interest has been devoted to defining more complete semantic representation schemes.
+The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
+shares much of UCCA's motivation for defining a more complete semantic annotation.
+However, using AMR is not optimal for defining a decomposition of a sentence into semantic
+units as it does not ground its semantic symbols in the text,
+thus not straightwardly allowing a decomposition of the sentence into sub-units.
+Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.
+
+Other approaches have been proposed for semantic representations, which are grounded in the words
+and phrases of the text, and have generally represented semantic structures as bi-lexical dependencies.
+Examples include the Prague Treebank tectogrammatical layer \cite{hajic2012announcing}, as well
+as dependencies derived from Minimal Recursion Semantics representations \cite{oepen2006discriminant}.
+Nevertheless, despite their appeal as deep and intuitive means for semantic representations,
+these approaches require linguistic expertise for their annotatation, due to the intricate and
+fine-grained distinctions they provide, and are thus in our view less suitable and
+easily interpretable means for our purposes.
+%\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
+%but to our knowledge no results of this brand have been reported.
 

### 2016-03-09 paper/main.tex
some notes

--- 2016-03-09 paper/main.tex
@@
 \usepackage{acl2016}
+\usepackage{subfig}
 \usepackage{times}
@@
 \usepackage{url}
-\usepackage{subfigure}
+%\usepackage{subfigure}
 \usepackage{pgfplotstable}
@@
 \subsection{Annotation Guidelines}\label{sec:guidelines}
+In this section we summarise the procedure for annotating machine translation output.
 
@@
 \subsection{Annotation Process}
-\begin{itemize}
-  \item Numbers of annotators, IDs
-  \item Length of annotation, timings
-  \item Sentences annotated
-\end{itemize}
-
-For the annotation process, we recruited seven annotators, two for each of the target languages apart from Czech. They were
+
+For the annotation of the machine translation, we recruited seven annotators, two for each of the target languages apart from Czech. They were
 all native speakers of the target languages and fluent in English. For Czech, we also recruited a  group of additional annotators
@@
 annotators as a single annotator.
+
+\emph{Show numbers of annotations, and timings}
+
+\emph{Describe the UCCA annotation}
 
@@
 \hline
-Sentences & & & & \\
-\hline
-Lexical nodes & & & & \\
-Kappa & & & & \\
-\hline
-Structural nodes & & & & \\
-Kappa & & & & \\
+Sentences & & 102 & 334 & 217 \\
+\hline
+Lexical nodes & & 1724 & 5386 & 3570 \\
+Kappa & & 0.29 & 0.54 & 0.50 \\
+\hline
+Structural nodes & & 1040 & 2655 & 1989 \\
+Kappa & & 0.44 & 0.33 & 0.58 \\
 \end{tabular}
@@
 To assess whether, as we hoped, HUME would make it easier to reliably annotate long sentences,
-we binned the sentences according to length and measured Kappa on each of the bins.
+we binned the sentences according to length and measured Kappa on each of the bins. Each bin covers
+a length range of 10, and we show the variance of kappa with length in \figref{fig:iaalength}.\bh{Note that cs is missing}
+
+\def\iaafig #1{\includegraphics[width=7cm]{iaa_length_#1.png}}
+
+\begin{figure*}[ht!]
+\begin{tabular}{cc}
+
+\subfloat[English-Czech]{
+  \iaafig{de}
+}
+&
+\subfloat[English-German]{
+  \iaafig{de}
+
+}
+\\
+
+\subfloat[English-Polish]{
+  \iaafig{pl}
+
+}
+&
+\subfloat[English-Romanian]{
+  \iaafig{ro}
+
+}
+\end{tabular}
+\caption{Inter-annotator Agreement (Kappa) versus sentence length, separately for
+structural and lexical nodes. }
+\label{fig:iaallength}
+
+\end{figure*}
 

### 2016-03-09 paper/main.tex
rearrange annotation guideline section, remove results from deliverable

@@
 
+There are two manual steps in the annotation process. The first is to create UCCA annotations for the source sentence,
+following the UCCA guidelines. Since this is is fully explained elsewhere \cite{abend2013universal}, we will not go into detail
+here, except to say that in our case the source sentences were all in English, and  the annotation was performed by four
+computational linguists. After we have the UCCA annotation on 
+the source, we use the word-to-word alignments from the MT system in order to project this annotation to the MT output, and
+we used native speakers of the target language in order to annotate this output.
+
+The basic strategy for annotating the semantics of the MT output to step through the source sentence semantic 
+components (projected UCCA nodes), looking at the translation via the word alignments, and marking on the source structure
+which parts have 
+been correctly translated.
+There are two kinds of components: a word or basic semantic unit, and a structural component which contains one
+or more sub-components.
+
+\paragraph{Lexical nodes} are usually comprised of individual words. They are the leaf nodes on the tree, the smallest 
+meaning-bearing
+units of the tree.
+Leaf nodes can be labelled as green (correct), orange (partially correct) and red (incorrect). 
+The traffic light system makes the marking of the
+lexical units as simple as possible.
+
+\begin{description}
+\item[Green] The meaning of the word or phrase has been largely captured.
+\item[Orange] the essential meaning has been captured, but some part of the translation is wrong. This could often be due to the translated word having the wrong tense, or the wrong morphology. 
+\item[Red] The essential meaning of the unit has not been captured.
+\end{description}
+
+\paragraph{Structural nodes} contain other nodes which could be either lexical or structural nodes, 
+and they can be labelled as ``Adequate'' or ``Bad''. 
+What we are trying to gauge for structural nodes
+is if the children of these nodes
+relate to each other in the same way in the source sentence and in the translation. 
+There are many ways that the relationship between the children might go wrong in translation:
+\begin{itemize}
+\item One of the child nodes is missing from the aligned translation. 
+\item An extra word or phrase has been inserted into the aligned  translation.
+\item The components of the translation are ordered differently 
+%to 
+from
+the source.
+\end{itemize}
+
+If any of these changes have occurred and this damages the meaning of the aligned translation, then we mark the 
+structural node as ``Bad''. 
+However if the arrangement of the child nodes is essentially correct, except that one or
+more of the child nodes has themselves been translated wrongly (and so is marked
+orange or red), then the structural node should be 
+annotated as ``Acceptable''.
+
+\begin{figure*}[t]
+    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
+    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
+	English paraphrase).}
+    \label{ucca-tree-mteval}
+\end{figure*}
+
+%TODO describe figure 2 here
+
+There are cases where we cannot usefully compare individual words in the source to words in the target.
+We often need to compare translation at the phrase level. If this is the case, then we can 
+ treat the structural node as a lexical unit, and mark it with the traffic light labels. This 
+means that none of the children of this node will be examined in order determine the semantic score
+for this sentence. 
+In Figure~\ref{ucca-tree-mteval}, we can see that the phrase ``became a passion'' has been labelled as a lexical
+node and it has been evaluated as ``Orange'', or partially correct, as the translation of this phrase ``loved'' 
+only partially capture the semantics of the source.
+
+We are separating lexical and structural evaluation in order to simplify evaluation and to localise errors
+to their point of origin. For this reason, if a word has been translated incorrectly,
+the parent node should still be correct if the number and relationship between its children are the same
+as in the source. 
+In Figure~\ref{ucca-tree-mteval}, we can see that even though an important child of the first scene is 
+incorrectly translated (``Golf''), and the children of are ordered differently, 
+the scene itself is marked as ``Acceptable''. The translation of the scene retains the structural 
+relationships between the children.
+\\
+
+\begin{figure*}[t]
+    \includegraphics[width=1\textwidth]{interface-annotations}
+    \caption{Machine translation evaluation tool}
+    \label{mttool}
+\end{figure*}
+
+In Figure~\ref{mttool} you can see the MT evaluation tool.
+The sentence at the top shows the complete MT system output.
+Underneath the MT output is the  source sentence.
+Underneath the source sentence we see its expandable semantic tree structure with both lexical and structural nodes.
+Lexical nodes only have traffic light annotation, whereas a structural node would normally be labelled ``A'' or ``B'' but could also be labelled as a lexical node, in the case where there is no word to word correspondence for the translation of its children.
+
+ The child components of a structural node are marked with different coloured rectangles.
+When navigating through the different source sentence nodes, one can see the relevant sections of the translation because the aligned words in the translation are highlighted in the complete sentence above.
+Aligned translations are also shown in black alongside the source node.
+If the node is aligned to a set of discontiguous words in the translation, then the unaligned words that appear in between the aligned words are shown in red.
+Even if these words are not directly aligned to the source node, they
+will likely change the meaning of the translation and must be considered when marking 
+the translation as correct or not.
+The alignments are meant just as a guide.
+The annotator should  look at the complete translation when deciding on the evaluation of a node.
+The translation could have content added before or after the node which changes its
+ structure or meaning. If an extra component is prepended to a structural node, for example, it should be marked as ``Bad''.
+
 \subsection{Composite Score}\label{sec:score}
@@
 \subsection{Data Sets and Translation Systems}
-\begin{itemize}
-  \item Briefly describe translation systems.
-  \item Source texts (no need to describe references)
-\end{itemize}
 
@@
 
-\subsection{Annotation Process}
+\subsection{Annotation Statistics}
 
@@
 
-\subsection{MT Evaluation Overview}
-
-The basic strategy for annotating the semantics of the machine translations is to step through the source sentence semantic 
-components, looking at the translation via the word alignments, and marking on the source structure
-%, 
-which parts have 
-been correctly translated.
-There are two kinds of components: a word or basic semantic unit, and a structural component which contains one or more sub-components.
-
-\subsubsection{Lexical nodes}
-Lexical nodes are usually comprised of individual words. They are the leaf nodes on the tree, the smallest 
-%meaning bearing 
-meaning-bearing
-units of the tree.
-Leaf nodes can be labelled as green (correct), orange (partially correct) and red (incorrect). 
-The traffic light system makes the marking of the
-lexical units as simple as possible.
-
-\begin{itemize}
-\item Green: The meaning of the word or phrase has been largely captured.
-\item Orange: the essential meaning has been captured, but some part of the translation is wrong. This could often be due to the translated word having the wrong tense, or the wrong morphology. 
-\item Red: The essential meaning of the unit has not been captured.
-\end{itemize}
-
-\subsubsection{Structural nodes}
-
-Structural nodes contain other nodes which could be either lexical or structural nodes. 
-These nodes are also called parent nodes and they can be labelled as ``Adequate'' or ``Bad''. 
-What we are trying to gauge for structural nodes
-%, 
-is if the children of these nodes
-relate to each other in the same way in the source sentence and in the translation. 
-There are many ways that the relationship between the children might go wrong in translation:
-\begin{itemize}
-\item One of the child nodes is missing from the aligned translation. 
-\item An extra word or phrase has been inserted into the aligned  translation.
-\item The components of the translation are ordered differently 
-%to 
-from
-the source.
-\end{itemize}
-
-If any of these changes have occurred and this damages the meaning of the aligned translation, then we mark the 
-structural node as ``Bad''. 
-However if the arrangement of the child nodes is essentially correct, except that one or
-more of the child nodes has themselves been translated wrongly (and so is marked
-orange or red), then the structural node should be 
-annotated as ``Acceptable''.
-
-\begin{figure*}[t]
-    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
-    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
-	English paraphrase).}
-    \label{ucca-tree-mteval}
-\end{figure*}
-
-%TODO describe figure 2 here
-
-There are cases where we cannot usefully compare individual words in the source to words in the target.
-We often need to compare translation at the phrase level. If this is the case, then we can 
- treat the structural node as a lexical unit, and mark it with the traffic light labels. This 
-means that none of the children of this node will be examined in order determine the semantic score
-for this sentence. 
-In Figure~\ref{ucca-tree-mteval}, we can see that the phrase ``became a passion'' has been labelled as a lexical
-node and it has been evaluated as ``Orange'', or partially correct, as the translation of this phrase ``loved'' 
-only partially capture the semantics of the source.
-
-We are separating lexical and structural evaluation in order to simplify evaluation and to localise errors
-to their point of origin. For this reason, if a word has been translated incorrectly,
-the parent node should still be correct if the number and relationship between its children are the same
-as in the source. 
-In Figure~\ref{ucca-tree-mteval}, we can see that even though an important child of the first scene is 
-incorrectly translated (``Golf''), and the children of are ordered differently, 
-the scene itself is marked as ``Acceptable''. The translation of the scene retains the structural 
-relationships between the children.
-\\
-
-\newpage
-
-\subsection{Machine Translation Evaluation Tool}
-
-In this section, we will describe the tool that the annotators used to perform the annotation.
-
-\begin{figure*}[t]
-    \includegraphics[width=1\textwidth]{interface-annotations}
-    \caption{Machine translation evaluation tool}
-    \label{mttool}
-\end{figure*}
-
-In Figure~\ref{mttool} you can see the MT evaluation tool.
-The sentence at the top shows the complete MT system output.
-Underneath the MT output is the  source sentence.
-Underneath the source sentence we see its expandable semantic tree structure with both lexical and structural nodes.
-Lexical nodes only have traffic light annotation, whereas a structural node would normally be labelled ``A'' or ``B'' but could also be labelled as a lexical node, in the case where there is no word to word correspondence for the translation of its children.
-
- The child components of a structural node are marked with different coloured rectangles.
-When navigating through the different source sentence nodes, one can see the relevant sections of the translation because the aligned words in the translation are highlighted in the complete sentence above.
-Aligned translations are also shown in black alongside the source node.
-If the node is aligned to a set of discontiguous words in the translation, then the unaligned words that appear in between the aligned words are shown in red.
-Even if these words are not directly aligned to the source node, they
-will likely change the meaning of the translation and must be considered when marking 
-the translation as correct or not.
-The alignments are meant just as a guide.
-The annotator should  look at the complete translation when deciding on the evaluation of a node.
-The translation could have content added before or after the node which changes its
- structure or meaning. If an extra component is prepended to a structural node, for example, it should be marked as ``Bad''.
-
-%\subsection{Example}
-%
-%\lexi{TODO write this up}
-%
-%NODE1. (make sure) (you’ve got a solid chair)
-%MT: (stellen Sie sicher, dass) (Sie haben einen festen Stuhl)
-%Correct would be: (stellen Sie sicher, dass) (Sie einen festen Stuhl haben)
-% 
-%NODE2. (you) (’ve) (got) (a solid chair)
-%MT: (Sie) (haben) (einen festen Stuhl)
-%Correct would be: (Sie) (einen festen Stuhl) (haben)
-%
-%Node 1 is the parent node of Node 2.
-The error in MT is that the verb in German "haben" appears in the wrong place, it should be at the end and this makes the translation difficult to understand.
-The place to mark this error is where haben is a child node, ie.
-the error occurs in Node 2.
-This is where the relationship or the order between the child nodes (Sie)(haben) and (einen festen Stuhl) is wrong.
-The error propagates up to Node 1, but at Node 1 we are evaluating the relationship between (stellen Sie sicher, dass) and (Sie haben einen festen Stuhl) and this is correct and should be marked A.
-
-\newpage
-
-\subsection{Results}
-
-%new_mteval_ro1: Christian
-%new_mteval_ro2: Rodica  NHS
-%new_mteval_pl1: Malgorzata
-%new_mteval_pl2: Iwona (starting from the 29th in the sequence) NHS
-%new_mteval_cs1: Martina
-%new_mteval_de1: Juliane (starting from the 28th in the sequence)
-
-In this section we report the results of the evaluation. We first report some statistics which give us an overall idea
-of how the annotators fared. We then delve into the question of how reliable the annotations are by examining
-inter-annotator agreement and looking at confusion matrices. Finally we calculate UCCA semantic scores
-for the different language pairs.
-
-\subsubsection{Overall Statistics}
-
-In order to explore the results of our annotations, we first show some basic statistics
-about the task. In Table~\ref{tab:stat-all}, we can see the total number of sentences annotated
-%
-with
-the number of nodes. We also show the percentages of the main types of nodes.
-Looking at this table we see that the number of sentences varies from 324 to 351 sentences,
-except for Ro1 who did considerably fewer, only 230. 
-All annotators were shown the same set of sentences, but their results were only
-collected if they pressed the submit button. So these missing sentences were either mistakenly
-not submitted, or they did not get to the end of the task. 
-Looking at the types of nodes annotated, these seem to be distributed quite uniformly.
-About 35\% of the nodes were considered to be structural, and 60\% to be lexical.
-Some nodes were not evaluated, and these are called ``Missing''. 
-The node may represent an English
-word which is not required in the target language (often an article or a
-pronoun).
-As per instructions, the annotators were expected to mark such nodes with green
-but some may have left the node unannotated (in which case it would be perhaps
-appropriate to consider the parent as a lexical node).
-With hindsight, we should have introduced a separate annotation category for this, and
-also enforced the lexical/structural distinction in the tool. ``Missing''
-annotations are not very common, but
-one annotator left 6.2\% of the nodes unlabelled which is a bit too high and in 
-future experiments we will alter the tool to disallow missing nodes, unless their parent is a lexical node.
-
-\begin{table}[h!]
-\begin{center}
-      \begin{tabular}{|l|c|c|c|c|c|}
-      \hline
- & \bf{No. Sentences} & \bf{No. Nodes} & \bf{\%Structural} & \bf{\%Lexical}  & \bf{\%Missing} \\
-\hline                               
-    De1  &  339 &  9253 & 36.2 & 62.5 & 1.3 \\ 
-    Cs1  & 324 &  8794 & 33.4 & 63.3 & 3.3\\ 
-    Ro1  & 230 &  6152 & 36.4 & 62.7 & 0.8\\ 
-    Ro2  & 337 &  9228 & 35.6 & 61.4 & 2.9 \\ 
-    Pl1  & 351 &  9557 & 34.0 & 59.8 & 6.2 \\ 
-    Pl2  & 340 &  9303 & 31.0 & 64.6 & 4.4 \\ 
-      \hline
-    \end{tabular}
-\end{center}
-\normalsize
-\vspace*{-3ex}
-\caption{Number of annotated sentences and nodes and the percentages of each of the main kind of nodes
-}
-\label{tab:stat-all}
-\end{table}
-
-\begin{table}[h!]
-\begin{center}
-      \begin{tabular}{|l|c|c|c|}
-      \hline
-   & \bf{No. Nodes} & \bf{\%A} & \bf{\%B}   \\
-\hline                               
-    De1  &  3350 &  68.2 & 31.7  \\ 
-    Cs1  & 2939 &  73.4 & 26.6 \\ 
-    Ro1  & 2239 &  68.2 & 31.8  \\ 
-    Ro2  & 3286 &  65.4 & 34.6 \\ 
-    Pl1  & 3246 &  50.3 & 49.7 \\ 
-    Pl2  & 2884 &  71.5 & 28.5 \\ 
-      \hline
-    \end{tabular}
-\end{center}
-\normalsize
-\vspace*{-3ex}
-\caption{Number of structural  nodes and the percentages of each of the types of nodes
-}
-\label{tab:stat-struct}
-\end{table}
-
-In Table~\ref{tab:stat-struct}, we can see the number of structural nodes, and their
-breakdown in terms of acceptable  ``A'' or bad ``B''. Most annotators seem to 
-label about 70\% of the structural nodes as correct. The exception is Pl1 who
-seems to be following a different annotation strategy. It is possible that Pl1 
-did not understand the guidelines correctly. Further investigation with a Polish
-speaking colleague will be necessary.
-
-\begin{table}[h!]
-\begin{center}
-      \begin{tabular}{|l|c|c|c|c|}
-      \hline
-   & \bf{No. Nodes} & \bf{\%G} & \bf{\%O}  & \bf{\%R} \\
-\hline                               
-    De1  & 5784  &  70.8 & 13.0 & 26.6  \\ 
-    Cs1  & 5568 &  62.8 & 18.4 & 18.8 \\ 
-    Ro1  & 3859 &  75.9 & 8.4 & 15.7  \\ 
-    Ro2  & 5671 &  68.7 & 16.9 & 14.3 \\ 
-    Pl1  & 5717 &  60.8 & 13.1 & 26.1 \\ 
-    Pl2  & 6007 &  49.6  & 17.3 & 33.0 \\ 
-      \hline
-    \end{tabular}
-\end{center}
-\normalsize
-\vspace*{-3ex}
-\caption{Number of lexical  nodes and the percentages of each of the types of nodes
-}
-\label{tab:stat-lex}
-\end{table}
-
-In Table~\ref{tab:stat-lex} we can see the number of lexical nodes, and their
-breakdown in terms of correct  ``G'', partial ``O'' or incorrect ``R''. 
-There is more variation here between the annotators than for the
-structural nodes. The percentage of correct nodes fluctuates from 49.6\%
-to 75.9\%. One expects that  languages with different amounts of 
-morphology and agglutination would report different numbers here,
-but even amongst annotators of the same languages the results vary.
-For Polish the number of correct lexical 
-%node 
-nodes
-varies between 49.6\%
-and 60.8\%.
-
-\subsubsection{IAA}
-
-In order to determine how reliable our human evaluation approach is, we need to
-examine the agreement that we find between annotators.
-For two of the language pairs, Romanian and Polish, we had 2 annotators. 
- Inter-annotator agreement can be calculated 
-using the Kappa score which is more reliable than agreement because it takes into
-account the probability of agreement by chance.
-
-%Lexi TODO put Kappas in a separate table
-
-\begin{table}[h!]
-\begin{center}
-      \begin{tabular}{|l|cc|c|}
-      \hline
-\bf{MT Eval Label} & \bf{A} & \bf{B} & \bf{All} \\
-\hline                               
-    A   & 1096 &  285 & 1381\\ 
-   B   & 101 & 507  & 608\\ 
-   \hline
-   All &  1197 &  792 & 1989\\
-      \hline
-    \end{tabular}
-\end{center}
-\normalsize
-\vspace*{-3ex}
-\caption{Confusion Matrix for the structural nodes for the Romanian annotators with a Kappa of 0.58
-}
-\label{tab:iaa-ab-ro}
-\end{table}
-
-\begin{table}[h!]
-\begin{center}
-      \begin{tabular}{|l|ccc|c|}
-      \hline
-\bf{MT Eval Label} & \bf{R} & \bf{O} & \bf{G} & \bf{All} \\
-\hline                               
-   R   & 2274 &  361 & 126 & 2761 \\ 
-   O   & 92 & 164  & 37 & 293 \\ 
-   G   & 82 & 76  & 358 & 516 \\ 
-   \hline
-   All &  2448 &  601 & 521 & 3570\\
-      \hline
-    \end{tabular}
-\end{center}
-\normalsize
-\vspace*{-3ex}
-\caption{Confusion Matrix for the lexical nodes for the Romanian annotators with a Kappa of 0.50
-}
-\label{tab:iaa-rgo-ro}
-\end{table}
-
-In Tables~\ref{tab:iaa-ab-ro} and \ref{tab:iaa-rgo-ro} we see the confusion matrix of
-the two main types of evaluation labels for Romanian. We separate structural and lexical nodes, because
-there is very little disagreement between annotators about which node is a lexical and which
-node is structural, and so if we calculate Kappa over all the values, this might artifically 
-inflate our IAA scores.
-For both structural and lexical nodes, the Kappa values 
-of 0.58 and 0.50 respectively 
-can
-be considered to display moderate agreement. 
-If we calculate Kappa scores across all five categories, it rises to 
-0.69 because of the strong agreement about what is a lexical and a structural 
-node.
-
-%\lexi{TODO look at confusion between structural and lexical}
-
-\begin{table}[h!]
-\begin{center}
-      \begin{tabular}{|l|cc|c|}
-      \hline
-\bf{MT Eval Label} & \bf{A} & \bf{B} & \bf{All} \\
-\hline                               
-    A   & 1208 &  192 & 1400 \\ 
-   B   & 681 & 574  & 1255 \\ 
-   \hline
-   All &  1889 &  766 & 2655 \\
-      \hline
-    \end{tabular}
-\end{center}
-\normalsize
-\vspace*{-3ex}
-\caption{Confusion Matrix for the structural nodes for the Polish annotators with a Kappa of 0.33
-}
-\label{tab:iaa-ab-pl}
-\end{table}
-
-\begin{table}[h!]
-\begin{center}
-      \begin{tabular}{|l|ccc|c|}
-      \hline
-\bf{MT Eval Label} & \bf{R} & \bf{O} & \bf{G} & \bf{All} \\
-\hline                               
-   R   & 2430 &  444 & 427 & 3301 \\ 
-   O   & 119 & 398  & 198 & 715 \\ 
-   G   & 161 & 109  & 1110 & 1380 \\ 
-   \hline
-   All &  2710 &  951 & 1735 & 5396 \\
-      \hline
-    \end{tabular}
-\end{center}
-\normalsize
-\vspace*{-3ex}
-\caption{Confusion Matrix for the lexical nodes for the Polish annotators with a Kappa of 0.58
-}
-\label{tab:iaa-rgo-pl}
-\end{table}
-
-In Tables~\ref{tab:iaa-ab-pl} and \ref{tab:iaa-rgo-pl} we see the confusion matrix of
-the two main types of evaluation labels for Polish. Here we see similar Kappa results for
-the lexical nodes (0.56) but the Kappa for structural nodes drops to 0.33. 
-This low agreement could be the result of one annotator misinterpreting the guidelines
-and warrants further analysis.
-If we calculate Kappa scores across all five categories, it rises to 
-0.58.
-
-Looking beyond Kappa scores, we did some analysis of the cases where
-one annotator marked a word as Red and another marked it Green. 
-For Polish  the most common English words about which the annotators
-disagreed were: ``to'', ``you'', ``blood'' (which was aligned to ``krwi''), ``in'' and ``for''. 
-For Romanian, the words were ``you'', ``to'', ``the'', and ``your''. 
-Apart from ``blood'', all these words are short function words which can 
-be translated in a very different fashion in the target and this could contribute to
- confusion about
-whether they are correct or not.
-
-Looking at the words which were marked Red most often again these were largely function words.
-This is partially due to the fact that they occur more frequently than other words. But
-words such as ``bias'', ``falls'', ``fibrillation'' and ``review'' also occur frequently in this list and these
-are words with domain specific translations which we are possibly not capturing with
-the unadapted HimL models.
-
-\subsubsection{UCCA Scores}
-
-\paragraph{Node score} This simple score reflects the percentage of correct MT evaluation nodes. 
-A and G are correct, O counts as 50\% correct and R and B count as incorrect. Nodes which are missing an
-evaluation are ignored.
-
-\begin{table}[h!]
-\begin{center}
-      \begin{tabular}{|l|c|c|c|}
-      \hline
-\bf{Annotator ID}  & \bf{Structural } & \bf{Lexical } & \bf{Overall }\\
-\hline                               
-    De1  &  68.20 & 77.40 & 74.03 \\ 
-    Cs1  & 73.35 & 72.00 & 72.46 \\ 
-    Ro1  & 68.15 & 80.09 & 75.71 \\ 
-    Ro2  & 65.42 & 77.26  & 72.92 \\ 
-    Pl1  & 50.33 & 67.33 & 61.17 \\ 
-    Pl2  & 71.53 &58.26  & 62.56 \\ 
-      \hline
-    \end{tabular}
-\end{center}
-\normalsize
-\vspace*{-3ex}
-\caption{Node scores: percentage of correct nodes. 
-}
-\label{tab:node_scores}
-\end{table}
-
-In Table~\ref{tab:node_scores} we can see the simple node scores for the different annotators. 
-Results show that most systems are evaluated as having about 70\% correct overall, with the
-notable exception of both Polish annotators. They gave scores to the HimL test set which
-were about 10\% lower. English$\to$Polish is considered to be a very challenging language pair
-for machine translation and so this result is not surprising.  Another trend is that the annotators gave the lexical scores of at least 
-10\% higher than the structural scores for three out of five annotators. The two remaining
-annotators behaved quite differently. Cs1 gave similar scores for structural and lexical nodes, but
-Pl2 gave much better scores for structural nodes (71.53 vs. 58.26). 
-The differences in behaviour displayed by the two Polish annotators in this table and the
-low kappa reported in Table~\ref{tab:iaa-ab-pl}
-brings into question the reliability of the Polish annotations.
-
-\begin{figure*}[t]
-    \includegraphics[width=1.1\textwidth]{tree_convert.png}
-    \caption{UCCA tree on the source labelled with the MT Eval labels and aligned to the target translation. The 
-    ``Acceptable'' structural nodes are yellow, the ``Bad'' ones are blue.}
-    \label{fig:ucca-tree-evaluated}
-\end{figure*}
-
-%\lexi{TODO: Any chance of an example with bad structural nodes ? }
-In order to visualise the UCCA evaluation results and see how the node scores are calculated, we provide an  example from the experiment. 
-Looking at Figure~\ref{fig:ucca-tree-evaluated}, we can see the machine translation at the top. Below this are the word alignments to the English source. 
-Each of the English words (except punctuation) are linked to UCCA lexical nodes which have been evaluated as either
-correct (green), partially correct (orange) or incorrect (red). The letters inside the circles correspond to the UCCA node types
-which will be important in future versions of the score. Structural nodes are either acceptable (yellow) or bad (blue). 
-So for this sentence given that we have 11 green nodes, 2 orange nodes, 6 yellow nodes and in total there are 24 nodes:
-
-\begin{displaymath}
-node score = \frac{ 11  + 0.5 ( 2 ) + 6  } { 24  } * 100\\
-                    = 75
-\end{displaymath}
-
-\subsection{Discussion}
-
-In this evaluation we have proposed an evaluation which has a number of advantages:
-\begin{itemize}
-\item  It shows moderate IAA agreement which means that our metric is relatively reliable.
-It is not directly comparable to other metrics because HMEANT papers do not report IAA and they had to rely on agreement statistics which can be misleading.
-\item  We can reuse the UCCA structures on the source multiple times, and as this is the most time consuming part
-of the evaluation, this makes best use of expensive expert annotators. 
-\item Translation annotators do not have to try to create semantic structures on erroneous machine translation output, in fact
-they do not need to create any semantic structures at all as this can be done previously by experts.
-\item Errors are isolated at the point where they occur due to the separation of structure and lexical content and there
-is little ambiguity about whether an error in a leaf needs to be counted multiple times as we go up the tree. 
-\item UCCA provides a minimal yet complete semantic framework which is based on universal cognitive concepts. 
-The structure does not rely on syntactic heads and the semantic graph is grounded directly to the words in the sentence. 
-\end{itemize}
-
-The evaluation also brought up a number of issues that will need to be resolved by either better  training or
-by improving the tool. Many of these issues were reported by the annotators:
-\begin{itemize}
-\item Isolating errors to the point where they occur is difficult to do for humans as there is a tendency to
-penalise the parent nodes for errors which occur in their children. 
-\item Initially some annotators struggled to understand the distinction between lexical and structural nodes. 
-\item They also had some difficulty evaluating the 
-%sentence 
-sentences
-without their context. 
-\item The annotators were unsure about how to judge incorrect morphology or tense. 
-\item  The alignments were meant only as a guide and some words are unaligned or doubly aligned and this
-caused some confusion. 
-\item Extraneous additions on the target might not be penalised unless they fall inside the source
-semantic structure.
-\end{itemize}
-
-Even though there have been a number of issues which have arisen, the UCCA evaluation has shown to be a reliable and
-efficient
-method for evaluating the accuracy of HimL translation systems. 
-We will continue to analyse the results from this experiment and we will run an improved version of
- this evaluation in year two
-of the project.
-
-%Comments from Juliane (Cochrane DE)
-%- There weren't many, but a few mistakes in the source, that made it difficult to annotate.
-%- Sometimes it was difficult to judge without context, but not very often - at least if you're familiar with the content.  
-%- There were cases where the way the sentences were segmented made it difficult to annoate them in a useful way.
-%- Many wrong declinations and conjugations.
-%- German genus requires pronouns or articles to be repeated in lists of nouns in order to be correct, e.g. your strength and balance = Ihre Stärke und Ihr Gleichgewicht
-%- There were a few out of domain vocabulary instance, e.g.
-"balance" being translated in the context of banking rather than the body, but only in one or two sentences, not in all where it occured.
-Similarly, the word "trial" got translated as "process", "study", and in the context of a court trial.
-%- A number of Cochrane or medical research specific terms was not well translated.
-%- Abbreviations never got translated correctly.
-%- Participle constructions often mess up the translation completely.
-%- Some verbs kept getting translated as nouns or the other way round, e.g.
-in review author = the author of a review, review was always interpreted as a verb; falls (noun) was mostly translated as verb.
-%- Overall: both Cochrane and NHS sentences were more often bad than understandable. Only very few were acceptable for publication on our websites in my view. I guess that was expected :)
-%What I mentioned earlier:
-%- The past and future tense were often ignored, which is usally not a problem for future, but for past it is
-%- Fairly often verbs were not translated at all (e.g. "cause" was never translated).  
-%- If the verb is present, it is often in the wrong order with the subject, as it doesn't necessarily follow the same order as in English.
-%- You is sometimes translated formally and sometimes informally, even in the same sentence.
-%- Need a counter to know how many sentences completed
-%
-%Comments from Rodica (NHS Ro):
-%One of the most recurrent confusions I had during my work was in connection with the word order in the MT.
-%EG.  English: This review provides limited evidence that interventions...
-%         Romanian: Acest studiu intervenții care furnizează dovezi limitate
-%                               (This review interventions provide limited evidence)
-%My question would have been how to mark the wrong or random word order in the MT translation, when the individual words appeared, so they couldn't be marked as red. 
-%In most of the examples this occurred in the Romanian translations  because the default word order in Romanian is  noun + descriptive adjective(s), rather than the other way round and so, in the MT the descriptive adjectives would be first and sometimes other words would be added before the noun, which would make the sentence incomprehensible in Romanian.
-%I think I mentioned to you the other questions during the work.
-The one that I should have asked earlier  was how far up the structure I should  mark a missing child.
-I understood it only once I spoke to you, but that was quite late.
-%
-%
-%Comments from Iwona (NHS Polish)
-%- Need a counter to know how many sentences completed
-%1. Phrases split in wrong places, e.g. ‘depending on’, ‘there is’
-%2. Certain phrases should be treated as one unit, e.g. ‘make sure’, so a leaf node for 
-%marking as green, orange or red.
-%3. Different sentence structure in English and Polish, e.g.
-%‘our appetites often decrease’
-%Machine translation: ‘our appetites often a fall’
-%Correct translation:  ‘Often decreases for us appetite’
-%Please note in the example above the meaning can be worked out and for that 
-%reason I marked the structural node as A. However, this is bad Polish. There were a 
-%number of sentences like this one, the translation of some of them made more sense 
-%than others, on the whole it was difficult to annotate them. 
-%4. The following is a common occurrence:
-%‘before starting a new exercise programme’
-%‘before’ – ‘przed rozpoczęciem’ (should be ‘przed’)
-%‘starting’ – ‘przed rozpoczęciem’ (should be ‘rozpoczęciem’)
-%Both leaf nodes have the same translation, but the translation appears only once in 
-%the structural node. I marked the parent node as A, and the two leaf nodes as red. 
-%5. Double negation in Polish causes problems at the structural level.
-
-\subsection{Automatic Semantic Evaluation}
-
-The results that we have gathered in this evaluation will be used as gold data to evaluate the performance of a number of automatic
-metrics, to determine which of the existing metrics are most suited to evaluating accuracy.  We have also  initiated
-work on developing an automatic version of the UCCA human evaluation task. For this we need an UCCA semantic parser
-and data to train it, as the UCCA treebank is relatively small. We are looking at extracting further training examples from
-the Prague treebank t-layer. This will be described in Deliverable D5.3, due at the end of July 2016.
-
 \section{Conclusion}

### 2016-03-09 scripts/iaa_binned.py
show p_o and p_e

@@ -51,7 +51,7 @@ def do_bins(agree, first_bin_start, bin_size, field_name, csv_file, graph_file, 
             if len(match_labels) == 2:
               cm = ConfusionMatrix(selected['mt_label_x'], selected['mt_label_y'], \
                 true_name="annot_1", pred_name="annot_2")
-              kappa = get_kappa(cm) #cm.stats()['overall']['Kappa']
+              kappa,_,_ = get_kappa(cm) #cm.stats()['overall']['Kappa']
               if kappa < 0:
                 print(cm)
             elif match_labels[0]:

### 2016-03-09 paper/main.tex

@@
 \usepackage{acl2016}
+\usepackage{paralist}
 \usepackage{subfig}
@@
 the source, we use the word-to-word alignments from the MT system in order to project this annotation to the MT output, and
-we used native speakers of the target language in order to annotate this output.
+we used native speakers of the target language in order to annotate this output. Since these annotators have to
+judge the quality of the translation, they also needed to be fluent in English.\bh{Somewhere we could explain that we
+could do a similar annotation with monolinguals, although it entails aligning reference and MT output}
 
-The basic strategy for annotating the semantics of the MT output to step through the source sentence semantic 
+The basic strategy for annotating HUME on the  MT output to step through the source sentence semantic 
 components (projected UCCA nodes), looking at the translation via the word alignments, and marking on the source structure
@@
 been correctly translated.
-There are two kinds of components: a word or basic semantic unit, and a structural component which contains one
+There are two kinds of components (nodes): a word or basic semantic unit, and a structural component which contains one
 or more sub-components.
 
-\paragraph{Lexical nodes} are usually comprised of individual words. They are the leaf nodes on the tree, the smallest 
+\paragraph{Lexical nodes} are usually comprised of individual words. They are the leaf nodes of the tree, the smallest 
 meaning-bearing
-units of the tree.
+units.
 Leaf nodes can be labelled as green (correct), orange (partially correct) and red (incorrect). 
@@
 
-\begin{description}
+\begin{compactdesc}
 \item[Green] The meaning of the word or phrase has been largely captured.
-\item[Orange] the essential meaning has been captured, but some part of the translation is wrong. This could often be due to the translated word having the wrong tense, or the wrong morphology. 
+\item[Orange] the essential meaning has been captured, but some part of the translation is wrong. 
+For example, this could often be due to the translated word having the wrong tense, or the wrong morphology. 
 \item[Red] The essential meaning of the unit has not been captured.
-\end{description}
+\end{compactdesc}
 
@@
 There are many ways that the relationship between the children might go wrong in translation:
-\begin{itemize}
+\begin{compactitem}
 \item One of the child nodes is missing from the aligned translation. 
@@
 the source.
-\end{itemize}
+\end{compactitem}
 
@@
 \begin{figure*}[t]
+    \begin{center}
     \includegraphics[width=1\textwidth]{interface-annotations}
-    \caption{Machine translation evaluation tool}
+    \caption{The HUME annotation tool, based on the UCCA tool}
     \label{mttool}
+    \end{center}
 \end{figure*}
 
-In Figure~\ref{mttool} you can see the MT evaluation tool.
+In Figure~\ref{mttool} you we show the MT evaluation tool.
 The sentence at the top shows the complete MT system output.
@@
 \section{Experiments}\label{sec:experiments}
+
+In order to validate the HUME metric, we ran an annotation experiment with one source language (English), and four 
+target languages (Czech, German, Polish and Romanian), using text from the public health domain. We are 
+particularly interested in this domain since our ultimate goal is to provide high-quality translations of health
+information for speakers of languages other than English, and in this domain accurate transfer of meaning is
+paramount.
+
+Our annotation experiment sought to address the following questions:
+\begin{itemize}
+\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences.
+\item Can HUME capture translation issues that other annotation schemes (such as adequacy judgements and HMEANT
+cannot)?
+\item What can HUME annotation tell us about the strengths and weaknesses of MT?
+\end{itemize}
+
+\bh{Maybe we should remove the third one, we
+probably don't address it in this paper}
+
 \subsection{Data Sets and Translation Systems}
 
-For each of the four language pairs under consideration (i.e. English to 
-Czech, German, Polish and Romanian) we built phrase-based MT systems
+For each of the four language pairs under consideration  we built phrase-based MT systems
 using Moses \cite{Koehn:2007}. We used large training sets extracted from OPUS \cite{tiedemann:2009}, and

### 2016-03-09 paper/main.tex
fff

--- 2016-03-09 paper/main.tex
@@
 
-  Semantic evaluation of Machine Translation (MT) systems, informative as for
-  semantic discrepencies between the translation and the source text,
-  has become a necessity, given recent interest in semantics-aware MT systems.
+  Recent interest in semantics-aware MT systems underscores te importance of
+  semantic evaluation for Machine Translation (MT) systems. 
+  A semantic measure, sensitive to possible semantic discrepencies
+  between the translation output and the source, is likely support the construction
+  and tuning of structure-aware MT, much in the same way that string-based measures,
+  such as BLEU, support advances with shallower MT systems.  
   We present a novel human semantic evaluation measure for MT, Human
@@
   a wider range of semantic phenomena and does not rely on semantic annotation
-  of the MT output, whose interpretation is often unclear. 
+  of the MT output itself, whose interpretation is often unclear.
   We experiment with four language pairs, demonstrating the broad applicability of
-  the measure, as well as its reliability, reflected by the inter-annotator agreement
+  the measure, as well as its reliability, reflected in the inter-annotator agreement
   scores.
@@
 
-Human judgement is the central, and perhaps ultimate, criterion for estimating the
-quality of an MT system.
+Human judgement is the central, and perhaps most fundamental, criterion for
+estimating the quality of an MT system.
 Nevertheless, common measures for MT evaluation, such as adequacy and frequency judgements
@@
 First, as the quality of translation is determined by multiple factors, it is difficult
-to assign a single number that quantifies the quality of the entire sentence, as is reflected
-by the diminishing inter-annotator agreement rates of human ranking measures
+to assign a single number that quantifies the quality of the whole entire sentence. This
+is indeed reflected in the diminishing inter-annotator agreement rates of human ranking measures
 as a function of the sentence length \cite{Bojar:2011}.
 Second, a sentence-level quality score does not indicate what parts of the sentence
-are erroneously translated in the output, and hence cannot inform developers in overcoming these errors.
+are erroneously translated, and hence cannot inform developers in overcoming these errors.
 
@@
 most commonly over its words or n-grams, quantifying the overlap of the sub-parts with a reference
-translation. More recent work on MT evaluation, most notably the MEANT measure \cite{lo2011structured},
-proposed measures that decompose over semantically defined units,
-quantifying the similarity of the output and the reference in terms of their verb argument structures.
-While allowing a localization of the semantic errors in the translation,
-focusing on verb argument structures misses out on many very pervasive semantic phenomena,
-such as inter-clause linkage and nominal argument structures, and thus provides only a limited
-perspective on the semantic similarity between the output and reference.
-Moreover, assigning a semantic structure to MT output, which is often low in quality,
+translation. More recent work on MT evaluation proposed measures
+that decompose over semantically defined units,
+quantifying the similarity of the output and the reference in terms of
+their verb argument structure (the most notable of these measures is
+MEANT \cite{lo2011structured}).
+While allowing to localize semantic errors in the translation,
+these measures' focus on verb argument structures misses out on many pervasive semantic phenomena,
+such as inter-clause linkage and nominal argument structures, and thus only provides
+a limited perspective on the semantic similarity between the output and reference.
+Moreover, assigning semantic structure to MT output, especially when low in quality,
 is difficult and arguably ill-defined.
@@
 semantic units in the source sentence.
-Importantly, HUME only requires the semantic annotation of the source sentence, thus avoiding the 
-annotatation of machine-generated text, a difficult if not ill-defined task, and allowing the re-use
-of the source semantic annotation for measuring the quality of different translations of the same source sentence.
+Importantly, HUME only requires semantically annotating the source sentence,
+thus avoiding the annotatation of machine-generated text 
+and allowing re-use of the source semantic annotation for measuring the quality
+of different translations of the same source sentence.
 By basing the measure on a direct comparison
@@
 but unlike them, HUME's agreement does not diminish with the sentence length.
-
-While human evaluation is not a scalable solution,
-it provides the foundations and justification for defining a measure.
-The paper thus focuses on human evaluation, but concludes
-by outlining ongoing work to construct semi- and fully-automatic variants of HUME.
+The paper concludes by outlining ongoing work to construct semi- and fully-automatic variants of HUME.
 
@@
 \paragraph{Machine Translation Evaluation.}
-MT evaluation measures has been traditionally pursued through human-based
-evaluation that quantifies the adequacy of the translation in preserving the meaning 
-of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
-Human evaluation is still accepted as the standard evaluation measure, and is
+%MT evaluation has been traditionally pursued through human-based
+%evaluation that quantifies the adequacy of the translation in preserving the meaning 
+%of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
+Human evaluation is the standard measure, and is
 estimated either in terms of the relative ranking of the outputs of multiple systems
-as is standard in the WMT shared tasks, e.g., \cite{bojar2015findings}, or by assigning
-indiviaul adequacy/fluency scores for each sentence, a procedure recently improved
+e.g., in the WMT shared tasks \cite{bojar2015findings}, or by assigning
+indiviaul adequacy/fluency scores to each sentence, a procedure recently improved
 by \newcite{graham2015improving}.
-
-While providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
+However, while providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
 and is unfeasible for a wide variety of systems or hyper-parameter settings are to be evaluated.
@@
 
-Scalability has been addressed by scores that decompose over the sub-parts of
+Scalability is addressed by scores that decompose over the sub-parts of
 the translation, and quantify how many of these sub-parts appear in a reference translation,
 whose validity has been manually confirmed. The most commonly used measures decompose
-over n-grams or even individual words, with examples such as
-the BLEU \cite{Papineni:2002} and METEOR measures \cite{Banerjee:2005}.
+over n-grams or even individual words, e.g., 
+the BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR measures \cite{Banerjee:2005}.
 Other methods employ a string edit distance for determining the similarity between the
@@
 and are thus better suited for localizing errors
-when evaluating shallow MT methods, and are less telling for identifying the semantic
+in the output of shallow MT methods, and are less telling for identifying the semantic
 nature of translation errors in the output.
@@
 
-To address this shortcoming, more structure-aware methods were proposed, with the 
-purpose of quantifying how different the reference and translation output are in terms 
+To address this shortcoming, more more recent work quantifies
+the similarity of the reference and translation output in terms
 of their {\it semantic structure}. \newcite{liu2005syntactic} took a syntactic approach, 
@@
 
-Perhaps the most notable attempt at a semantic MT evaluation measure is MEANT and
-its human variant HMEANT, which compares the reference and translation in terms of the similarity
-of their semantic role labeling structure. HMEANT shares much of HUME's motivation for providing a
+Perhaps the most notable attempt at semantic MT evaluation is MEANT and
+its human variant HMEANT, which quantifies the similarity between
+the reference and translation in terms of the overlap in
+their verbal argument structures structures, and associated semantic roles.
+HMEANT shares much of HUME's motivation for providing a
 measure that decomposes over semantic units, rather than strings, but focuses on verbal
@@
 relations and copula clauses. HUME further diverges from both MEANT and the more shallow BLEU and
-METEOR measures, in not requiring a reference, as it compares the source directly with the output
+METEOR measures, in not requiring a reference. Instead,
+it compares the source directly with the output
 translation. We see this difference as crucial where semantic analysis is concerned, as it
@@
 
+\begin{figure}
+
+\begin{tikzpicture}[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
+      inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
+    \node (ROOT) [fill=black, circle] {}
+    child {node {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
+    child {node (H1) [fill=black,circle] {}
+      child {node {graduation} edge from parent node[right] {$P$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    }
+    child {node (H2) [fill=black,circle] {}
+      child {node {} edge from parent[draw=none]}
+      child {node {} edge from parent[draw=none]}
+      child {node (Tom) {Tom} edge from parent node[right] {$A$}}
+      child {node {moved} edge from parent node[right] {$P$}}
+      child {node [fill=black,circle] {}
+        child {node {to} edge from parent node[right] {$R$}}
+        child {node {NYC} edge from parent node[right] {$C$}}
+        edge from parent node[right] {$A$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    };
+    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+  \end{tikzpicture}
+\caption{\label{fig:ucca_example}
+  Sample UCCA annotation.}
+\end{figure}
+
 \paragraph{Semantic Representation.}
@@
 whose leaves correspond to the words of the text.
-The graph’s nodes, called {\sc units} , are either terminals or several elements jointly
-viewed as a single entity according to some semantic or cognitive consideration. Edges bear one
-or more categories, indicating the role of the sub-unit in the relation the unit represents. UCCA
+The graph’s nodes, called {\sc units}, are either terminals or several elements jointly
+viewed as a single entity according to some semantic or cognitive consideration. Edges bear
+a category, indicating the role of the sub-unit in the relation the unit represents. UCCA
 currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
@@
 is a part of both Scenes, while ``NYC'' only of the latter. Further categories account for
-relations between Scenes and the internal structures of complex participants and relations.\oa{add figure}
+relations between Scenes and the internal structures of complex participants and relations.
+See \figref{fig:ucca_example}.
 
@@
 Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
-application to annotation in English, French, German and Czech. 
+application to annotation in English, French, German and Czech\oa{Did we do experiments on Czech?}. 
 Third, the scheme has been shown to be semantically stable
 in translation: UCCA annotations of translated text usually contain the same set of relationships
-\cite{sulem2012}. This is an indication that the scheme reflect a common layer of representation that
-should be preserved in translation, and so should be represented in the translation output too.
+\cite{sulem2015conceptual}. This is an indication that the scheme reflects
+a common layer of representation that in a correct translation
+is mostly shared between the translation and the source. 
 We note that that verbal argument structures are less stable in this sense, as a verbal
 argument structure may be translated or paraphrased to a nominal or adjectival
-argument structure. For example, ``after he graduated, Tom moved to NYC'' conveys a similar meaning
-to ``after graduation, Tom moved to NYC''.
-
-While most broad-coverage semantic work has focused on shallow semantic structures,
+argument structure, and are thus less motivated as a basis for a semantic measure.
+For example, ``after he graduated, Tom moved to NYC''
+conveys a similar meaning to ``after graduation, Tom moved to NYC''.
+
+While most broad-coverage semantic work focused on shallow semantic structures,
 predominantly argument structures annotated with their semantic roles, much recent
@@
 
-Other approaches have been proposed for semantic representations, which are grounded in the words
+Other approaches have been proposed for semantic representation, which are grounded in the words
 and phrases of the text, and have generally represented semantic structures as bi-lexical dependencies.
@@
 as dependencies derived from Minimal Recursion Semantics representations \cite{oepen2006discriminant}.
-Nevertheless, despite their appeal as deep and intuitive means for semantic representations,
+Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
 these approaches require linguistic expertise for their annotatation, due to the intricate and
-fine-grained distinctions they provide, and are thus in our view less suitable and
-easily interpretable means for our purposes.
+fine-grained distinctions they provide, and are thus in our view less suitable for basing
+an interpretable MT evaluation on. 
 %\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
@@
 \subsection{Annotation Guidelines}\label{sec:guidelines}
-In this section we summarise the procedure for annotating machine translation output.
 
@@
 \subsection{Annotation Process}
-
-For the annotation of the machine translation, we recruited seven annotators, two for each of the target languages apart from Czech. They were
+\begin{itemize}
+  \item Numbers of annotators, IDs
+  \item Length of annotation, timings
+  \item Sentences annotated
+\end{itemize}
+
+For the annotation process, we recruited seven annotators, two for each of the target languages apart from Czech. They were
 all native speakers of the target languages and fluent in English. For Czech, we also recruited a  group of additional annotators
@@
 annotators as a single annotator.
-
-\emph{Show numbers of annotations, and timings}
-
-\emph{Describe the UCCA annotation}
 

### 2016-03-09 paper/main.tex

--- 2016-03-09 paper/main.tex
@@
 Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
-application to annotation in English, French, German and Czech. 
+application to annotation in English, French, German and Czech\oa{Did we do experiments on Czech?}. 
 Third, the scheme has been shown to be semantically stable
 in translation: UCCA annotations of translated text usually contain the same set of relationships
-\cite{sulem2012}. This is an indication that the scheme reflect a common layer of representation that
-should be preserved in translation, and so should be represented in the translation output too.
+\cite{sulem2015conceptual}. This is an indication that the scheme reflects
+a common layer of representation that in a correct translation
+is mostly shared between the translation and the source. 
 We note that that verbal argument structures are less stable in this sense, as a verbal
 argument structure may be translated or paraphrased to a nominal or adjectival
-argument structure. For example, ``after he graduated, Tom moved to NYC'' conveys a similar meaning
-to ``after graduation, Tom moved to NYC''.
+argument structure, and are thus less motivated as a basis for a semantic measure.
+For example, ``after he graduated, Tom moved to NYC''
+conveys a similar meaning to ``after graduation, Tom moved to NYC''.
 
-While most broad-coverage semantic work has focused on shallow semantic structures,
+While most broad-coverage semantic work focused on shallow semantic structures,
 predominantly argument structures annotated with their semantic roles, much recent

### 2016-03-09 paper/results.tex

\documentclass[11pt]{article}
\usepackage{graphicx}
%
% This is from the HimL deliverable
%

\begin{document}

\subsection{Results}

%new_mteval_ro1: Christian
%new_mteval_ro2: Rodica  NHS
%new_mteval_pl1: Malgorzata
%new_mteval_pl2: Iwona (starting from the 29th in the sequence) NHS
%new_mteval_cs1: Martina
%new_mteval_de1: Juliane (starting from the 28th in the sequence)

In this section we report the results of the evaluation. We first report some statistics which give us an overall idea
of how the annotators fared. We then delve into the question of how reliable the annotations are by examining
inter-annotator agreement and looking at confusion matrices. Finally we calculate UCCA semantic scores
for the different language pairs.

\subsubsection{Overall Statistics}

In order to explore the results of our annotations, we first show some basic statistics
about the task. In Table~\ref{tab:stat-all}, we can see the total number of sentences annotated
%
with
the number of nodes. We also show the percentages of the main types of nodes.
Looking at this table we see that the number of sentences varies from 324 to 351 sentences,
except for Ro1 who did considerably fewer, only 230. 
All annotators were shown the same set of sentences, but their results were only
collected if they pressed the submit button. So these missing sentences were either mistakenly
not submitted, or they did not get to the end of the task. 
Looking at the types of nodes annotated, these seem to be distributed quite uniformly.
About 35\% of the nodes were considered to be structural, and 60\% to be lexical.
Some nodes were not evaluated, and these are called ``Missing''. 
The node may represent an English
word which is not required in the target language (often an article or a
pronoun).
As per instructions, the annotators were expected to mark such nodes with green
but some may have left the node unannotated (in which case it would be perhaps
appropriate to consider the parent as a lexical node).
With hindsight, we should have introduced a separate annotation category for this, and
also enforced the lexical/structural distinction in the tool. ``Missing''
annotations are not very common, but
one annotator left 6.2\% of the nodes unlabelled which is a bit too high and in 
future experiments we will alter the tool to disallow missing nodes, unless their parent is a lexical node.

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|c|c|c|c|c|}
      \hline
 & \bf{No. Sentences} & \bf{No. Nodes} & \bf{\%Structural} & \bf{\%Lexical}  & \bf{\%Missing} \\
\hline                               
    De1  &  339 &  9253 & 36.2 & 62.5 & 1.3 \\ 
    Cs1  & 324 &  8794 & 33.4 & 63.3 & 3.3\\ 
    Ro1  & 230 &  6152 & 36.4 & 62.7 & 0.8\\ 
    Ro2  & 337 &  9228 & 35.6 & 61.4 & 2.9 \\ 
    Pl1  & 351 &  9557 & 34.0 & 59.8 & 6.2 \\ 
    Pl2  & 340 &  9303 & 31.0 & 64.6 & 4.4 \\ 
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Number of annotated sentences and nodes and the percentages of each of the main kind of nodes
}
\label{tab:stat-all}
\end{table}

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|c|c|c|}
      \hline
   & \bf{No. Nodes} & \bf{\%A} & \bf{\%B}   \\
\hline                               
    De1  &  3350 &  68.2 & 31.7  \\ 
    Cs1  & 2939 &  73.4 & 26.6 \\ 
    Ro1  & 2239 &  68.2 & 31.8  \\ 
    Ro2  & 3286 &  65.4 & 34.6 \\ 
    Pl1  & 3246 &  50.3 & 49.7 \\ 
    Pl2  & 2884 &  71.5 & 28.5 \\ 
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Number of structural  nodes and the percentages of each of the types of nodes
}
\label{tab:stat-struct}
\end{table}

In Table~\ref{tab:stat-struct}, we can see the number of structural nodes, and their
breakdown in terms of acceptable  ``A'' or bad ``B''. Most annotators seem to 
label about 70\% of the structural nodes as correct. The exception is Pl1 who
seems to be following a different annotation strategy. It is possible that Pl1 
did not understand the guidelines correctly. Further investigation with a Polish
speaking colleague will be necessary.

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|c|c|c|c|}
      \hline
   & \bf{No. Nodes} & \bf{\%G} & \bf{\%O}  & \bf{\%R} \\
\hline                               
    De1  & 5784  &  70.8 & 13.0 & 26.6  \\ 
    Cs1  & 5568 &  62.8 & 18.4 & 18.8 \\ 
    Ro1  & 3859 &  75.9 & 8.4 & 15.7  \\ 
    Ro2  & 5671 &  68.7 & 16.9 & 14.3 \\ 
    Pl1  & 5717 &  60.8 & 13.1 & 26.1 \\ 
    Pl2  & 6007 &  49.6  & 17.3 & 33.0 \\ 
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Number of lexical  nodes and the percentages of each of the types of nodes
}
\label{tab:stat-lex}
\end{table}

In Table~\ref{tab:stat-lex} we can see the number of lexical nodes, and their
breakdown in terms of correct  ``G'', partial ``O'' or incorrect ``R''. 
There is more variation here between the annotators than for the
structural nodes. The percentage of correct nodes fluctuates from 49.6\%
to 75.9\%. One expects that  languages with different amounts of 
morphology and agglutination would report different numbers here,
but even amongst annotators of the same languages the results vary.
For Polish the number of correct lexical 
%node 
nodes
varies between 49.6\%
and 60.8\%.

\subsubsection{IAA}

In order to determine how reliable our human evaluation approach is, we need to
examine the agreement that we find between annotators.
For two of the language pairs, Romanian and Polish, we had 2 annotators. 
 Inter-annotator agreement can be calculated 
using the Kappa score which is more reliable than agreement because it takes into
account the probability of agreement by chance.

%Lexi TODO put Kappas in a separate table

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|cc|c|}
      \hline
\bf{MT Eval Label} & \bf{A} & \bf{B} & \bf{All} \\
\hline                               
    A   & 1096 &  285 & 1381\\ 
   B   & 101 & 507  & 608\\ 
   \hline
   All &  1197 &  792 & 1989\\
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Confusion Matrix for the structural nodes for the Romanian annotators with a Kappa of 0.58
}
\label{tab:iaa-ab-ro}
\end{table}

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|ccc|c|}
      \hline
\bf{MT Eval Label} & \bf{R} & \bf{O} & \bf{G} & \bf{All} \\
\hline                               
   R   & 2274 &  361 & 126 & 2761 \\ 
   O   & 92 & 164  & 37 & 293 \\ 
   G   & 82 & 76  & 358 & 516 \\ 
   \hline
   All &  2448 &  601 & 521 & 3570\\
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Confusion Matrix for the lexical nodes for the Romanian annotators with a Kappa of 0.50
}
\label{tab:iaa-rgo-ro}
\end{table}

In Tables~\ref{tab:iaa-ab-ro} and \ref{tab:iaa-rgo-ro} we see the confusion matrix of
the two main types of evaluation labels for Romanian. We separate structural and lexical nodes, because
there is very little disagreement between annotators about which node is a lexical and which
node is structural, and so if we calculate Kappa over all the values, this might artifically 
inflate our IAA scores.
For both structural and lexical nodes, the Kappa values 
of 0.58 and 0.50 respectively 
can
be considered to display moderate agreement. 
If we calculate Kappa scores across all five categories, it rises to 
0.69 because of the strong agreement about what is a lexical and a structural 
node.

%\lexi{TODO look at confusion between structural and lexical}

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|cc|c|}
      \hline
\bf{MT Eval Label} & \bf{A} & \bf{B} & \bf{All} \\
\hline                               
    A   & 1208 &  192 & 1400 \\ 
   B   & 681 & 574  & 1255 \\ 
   \hline
   All &  1889 &  766 & 2655 \\
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Confusion Matrix for the structural nodes for the Polish annotators with a Kappa of 0.33
}
\label{tab:iaa-ab-pl}
\end{table}

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|ccc|c|}
      \hline
\bf{MT Eval Label} & \bf{R} & \bf{O} & \bf{G} & \bf{All} \\
\hline                               
   R   & 2430 &  444 & 427 & 3301 \\ 
   O   & 119 & 398  & 198 & 715 \\ 
   G   & 161 & 109  & 1110 & 1380 \\ 
   \hline
   All &  2710 &  951 & 1735 & 5396 \\
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Confusion Matrix for the lexical nodes for the Polish annotators with a Kappa of 0.58
}
\label{tab:iaa-rgo-pl}
\end{table}

In Tables~\ref{tab:iaa-ab-pl} and \ref{tab:iaa-rgo-pl} we see the confusion matrix of
the two main types of evaluation labels for Polish. Here we see similar Kappa results for
the lexical nodes (0.56) but the Kappa for structural nodes drops to 0.33. 
This low agreement could be the result of one annotator misinterpreting the guidelines
and warrants further analysis.
If we calculate Kappa scores across all five categories, it rises to 
0.58.

Looking beyond Kappa scores, we did some analysis of the cases where
one annotator marked a word as Red and another marked it Green. 
For Polish  the most common English words about which the annotators
disagreed were: ``to'', ``you'', ``blood'' (which was aligned to ``krwi''), ``in'' and ``for''. 
For Romanian, the words were ``you'', ``to'', ``the'', and ``your''. 
Apart from ``blood'', all these words are short function words which can 
be translated in a very different fashion in the target and this could contribute to
 confusion about
whether they are correct or not.

Looking at the words which were marked Red most often again these were largely function words.
This is partially due to the fact that they occur more frequently than other words. But
words such as ``bias'', ``falls'', ``fibrillation'' and ``review'' also occur frequently in this list and these
are words with domain specific translations which we are possibly not capturing with
the unadapted HimL models.

\subsubsection{UCCA Scores}

\paragraph{Node score} This simple score reflects the percentage of correct MT evaluation nodes. 
A and G are correct, O counts as 50\% correct and R and B count as incorrect. Nodes which are missing an
evaluation are ignored.

\begin{table}[h!]
\begin{center}
      \begin{tabular}{|l|c|c|c|}
      \hline
\bf{Annotator ID}  & \bf{Structural } & \bf{Lexical } & \bf{Overall }\\
\hline                               
    De1  &  68.20 & 77.40 & 74.03 \\ 
    Cs1  & 73.35 & 72.00 & 72.46 \\ 
    Ro1  & 68.15 & 80.09 & 75.71 \\ 
    Ro2  & 65.42 & 77.26  & 72.92 \\ 
    Pl1  & 50.33 & 67.33 & 61.17 \\ 
    Pl2  & 71.53 &58.26  & 62.56 \\ 
      \hline
    \end{tabular}
\end{center}
\normalsize
\vspace*{-3ex}
\caption{Node scores: percentage of correct nodes. 
}
\label{tab:node_scores}
\end{table}

In Table~\ref{tab:node_scores} we can see the simple node scores for the different annotators. 
Results show that most systems are evaluated as having about 70\% correct overall, with the
notable exception of both Polish annotators. They gave scores to the HimL test set which
were about 10\% lower. English$\to$Polish is considered to be a very challenging language pair
for machine translation and so this result is not surprising.  Another trend is that the annotators gave the lexical scores of at least 
10\% higher than the structural scores for three out of five annotators. The two remaining
annotators behaved quite differently. Cs1 gave similar scores for structural and lexical nodes, but
Pl2 gave much better scores for structural nodes (71.53 vs. 58.26). 
The differences in behaviour displayed by the two Polish annotators in this table and the
low kappa reported in Table~\ref{tab:iaa-ab-pl}
brings into question the reliability of the Polish annotations.

\begin{figure*}[t]
    \includegraphics[width=1.1\textwidth]{tree_convert.png}
    \caption{UCCA tree on the source labelled with the MT Eval labels and aligned to the target translation. The 
    ``Acceptable'' structural nodes are yellow, the ``Bad'' ones are blue.}
    \label{fig:ucca-tree-evaluated}
\end{figure*}

%\lexi{TODO: Any chance of an example with bad structural nodes ? }
In order to visualise the UCCA evaluation results and see how the node scores are calculated, we provide an  example from the experiment. 
Looking at Figure~\ref{fig:ucca-tree-evaluated}, we can see the machine translation at the top. Below this are the word alignments to the English source. 
Each of the English words (except punctuation) are linked to UCCA lexical nodes which have been evaluated as either
correct (green), partially correct (orange) or incorrect (red). The letters inside the circles correspond to the UCCA node types
which will be important in future versions of the score. Structural nodes are either acceptable (yellow) or bad (blue). 
So for this sentence given that we have 11 green nodes, 2 orange nodes, 6 yellow nodes and in total there are 24 nodes:

\begin{displaymath}
node score = \frac{ 11  + 0.5 ( 2 ) + 6  } { 24  } * 100\\
                    = 75
\end{displaymath}

\subsection{Discussion}

In this evaluation we have proposed an evaluation which has a number of advantages:
\begin{itemize}
\item  It shows moderate IAA agreement which means that our metric is relatively reliable. It is not directly comparable to other metrics because HMEANT papers do not report IAA and they had to rely on agreement statistics which can be misleading. 
\item  We can reuse the UCCA structures on the source multiple times, and as this is the most time consuming part
of the evaluation, this makes best use of expensive expert annotators. 
\item Translation annotators do not have to try to create semantic structures on erroneous machine translation output, in fact
they do not need to create any semantic structures at all as this can be done previously by experts.
\item Errors are isolated at the point where they occur due to the separation of structure and lexical content and there
is little ambiguity about whether an error in a leaf needs to be counted multiple times as we go up the tree. 
\item UCCA provides a minimal yet complete semantic framework which is based on universal cognitive concepts. 
The structure does not rely on syntactic heads and the semantic graph is grounded directly to the words in the sentence. 
\end{itemize}

The evaluation also brought up a number of issues that will need to be resolved by either better  training or
by improving the tool. Many of these issues were reported by the annotators:
\begin{itemize}
\item Isolating errors to the point where they occur is difficult to do for humans as there is a tendency to
penalise the parent nodes for errors which occur in their children. 
\item Initially some annotators struggled to understand the distinction between lexical and structural nodes. 
\item They also had some difficulty evaluating the 
%sentence 
sentences
without their context. 
\item The annotators were unsure about how to judge incorrect morphology or tense. 
\item  The alignments were meant only as a guide and some words are unaligned or doubly aligned and this
caused some confusion. 
\item Extraneous additions on the target might not be penalised unless they fall inside the source
semantic structure.
\end{itemize}

Even though there have been a number of issues which have arisen, the UCCA evaluation has shown to be a reliable and
efficient
method for evaluating the accuracy of HimL translation systems. 
We will continue to analyse the results from this experiment and we will run an improved version of
 this evaluation in year two
of the project.

%Comments from Juliane (Cochrane DE)
%- There weren't many, but a few mistakes in the source, that made it difficult to annotate.
%- Sometimes it was difficult to judge without context, but not very often - at least if you're familiar with the content.  
%- There were cases where the way the sentences were segmented made it difficult to annoate them in a useful way.
%- Many wrong declinations and conjugations.
%- German genus requires pronouns or articles to be repeated in lists of nouns in order to be correct, e.g. your strength and balance = Ihre Stärke und Ihr Gleichgewicht
%- There were a few out of domain vocabulary instance, e.g. "balance" being translated in the context of banking rather than the body, but only in one or two sentences, not in all where it occured. Similarly, the word "trial" got translated as "process", "study", and in the context of a court trial. 
%- A number of Cochrane or medical research specific terms was not well translated.
%- Abbreviations never got translated correctly.
%- Participle constructions often mess up the translation completely.
%- Some verbs kept getting translated as nouns or the other way round, e.g. in review author = the author of a review, review was always interpreted as a verb; falls (noun) was mostly translated as verb. 
%- Overall: both Cochrane and NHS sentences were more often bad than understandable. Only very few were acceptable for publication on our websites in my view. I guess that was expected :)
%What I mentioned earlier:
%- The past and future tense were often ignored, which is usally not a problem for future, but for past it is
%- Fairly often verbs were not translated at all (e.g. "cause" was never translated).  
%- If the verb is present, it is often in the wrong order with the subject, as it doesn't necessarily follow the same order as in English.
%- You is sometimes translated formally and sometimes informally, even in the same sentence.
%- Need a counter to know how many sentences completed
%
%Comments from Rodica (NHS Ro):
%One of the most recurrent confusions I had during my work was in connection with the word order in the MT.
%EG.  English: This review provides limited evidence that interventions...
%         Romanian: Acest studiu intervenții care furnizează dovezi limitate
%                               (This review interventions provide limited evidence)
%My question would have been how to mark the wrong or random word order in the MT translation, when the individual words appeared, so they couldn't be marked as red. 
%In most of the examples this occurred in the Romanian translations  because the default word order in Romanian is  noun + descriptive adjective(s), rather than the other way round and so, in the MT the descriptive adjectives would be first and sometimes other words would be added before the noun, which would make the sentence incomprehensible in Romanian.
%I think I mentioned to you the other questions during the work. The one that I should have asked earlier  was how far up the structure I should  mark a missing child. I understood it only once I spoke to you, but that was quite late.
%
%
%Comments from Iwona (NHS Polish)
%- Need a counter to know how many sentences completed
%1. Phrases split in wrong places, e.g. ‘depending on’, ‘there is’
%2. Certain phrases should be treated as one unit, e.g. ‘make sure’, so a leaf node for 
%marking as green, orange or red.
%3. Different sentence structure in English and Polish, e.g.
%‘our appetites often decrease’
%Machine translation: ‘our appetites often a fall’
%Correct translation:  ‘Often decreases for us appetite’
%Please note in the example above the meaning can be worked out and for that 
%reason I marked the structural node as A. However, this is bad Polish. There were a 
%number of sentences like this one, the translation of some of them made more sense 
%than others, on the whole it was difficult to annotate them. 
%4. The following is a common occurrence:
%‘before starting a new exercise programme’
%‘before’ – ‘przed rozpoczęciem’ (should be ‘przed’)
%‘starting’ – ‘przed rozpoczęciem’ (should be ‘rozpoczęciem’)
%Both leaf nodes have the same translation, but the translation appears only once in 
%the structural node. I marked the parent node as A, and the two leaf nodes as red. 
%5. Double negation in Polish causes problems at the structural level.

\subsection{Automatic Semantic Evaluation}

The results that we have gathered in this evaluation will be used as gold data to evaluate the performance of a number of automatic
metrics, to determine which of the existing metrics are most suited to evaluating accuracy.  We have also  initiated
work on developing an automatic version of the UCCA human evaluation task. For this we need an UCCA semantic parser
and data to train it, as the UCCA treebank is relatively small. We are looking at extracting further training examples from
the Prague treebank t-layer. This will be described in Deliverable D5.3, due at the end of July 2016.

\end{document}

### 2016-03-10 paper/main.tex
annotation stats

@@
 
-For the annotation of the machine translation, we recruited seven annotators, two for each of the target languages apart from Czech. They were
+For the annotation of MT output, we recruited seven annotators, two for each of the target languages apart from Czech. They were
 all native speakers of the target languages and fluent in English. For Czech, we also recruited a  group of additional annotators
@@
 we show the total number of sentences and nodes annotated by each annotator, where we consider the additional Czech
-annotators as a single annotator.
+annotators as a single annotator. Note that not all nodes in all sentences were annotated. Sometimes this was because the
+annotator did not finish a node, or accidentally missed a node, but in many cases it was also because it was an implicit 
+UCCA node, which was not shown to the annotators since there is no corresponding source segment.
+\begin{table}
+\begin{center}
+{\small
+\begin{tabular}{ll|cccc}
+& & cs & de & pl & ro \\
+\hline
+Sentences &  Annot. 1 & 324   & 339  & 351  & 230  \\
+ & Annot. 2 & -- & 104  & 340  & 337 \\
+\hline
+Nodes & Annot. 1 & 8794  & 9253 & 9557  & 6152 \\
+ &Annot. 2 & -- & 2906  & 9303  & 9228  \\
+\end{tabular}
+}
+\caption{Total numbers of sentences and nodes annotated on the MT output.}
+\label{tab:annot}
+\end{center}
+\end{table}
 
-\emph{Show numbers of annotations, and timings}
+\emph{Timing of annotations}
 
@@
 
-\subsection{Data}\label{sec:data}
+%\subsection{Data}\label{sec:data}
 
-\subsection{Efficiency of Annotation}\label{sec:efficiency}
+%\subsection{Efficiency of Annotation}\label{sec:efficiency}
 
 \subsection{Inter-Annotator Agreement}
+\label{sec:iaa}
 \begin{itemize}
@@
 \hline
-Sentences & & 102 & 334 & 217 \\
+Sentences & -- & 102 & 334 & 217 \\
 \hline
-Lexical nodes & & 1724 & 5386 & 3570 \\
-Kappa & & 0.29 & 0.54 & 0.50 \\
+Lexical nodes & -- & 1724 & 5386 & 3570 \\
+Kappa & -- & 0.29 & 0.54 & 0.50 \\
 \hline
 Structural nodes & & 1040 & 2655 & 1989 \\
-Kappa & & 0.44 & 0.33 & 0.58 \\
+Kappa & -- & 0.44 & 0.33 & 0.58 \\
 \end{tabular}

### 2016-03-14 scripts/extract_sentences_for_mturk.py
extract script for data needed by the Yvette Graham MTurk exp

@@ -1,0 +1,54 @@
+#!/usr/bin/env python
+from __future__ import print_function, division
+
+import sys
+import pandas
+from  pandas_confusion import ConfusionMatrix
+
+
+prefix="himl2015"
+
+def extract(alldata):
+  print ("Stats \n\n")
+  for lang in ('ro','de','cs','pl'):
+    data = alldata.loc[alldata["lang"] == lang]
+    ids = alldata['sent_id'].unique()
+    print ("Lang: " , lang)
+    lang_pair = "en-" + lang
+
+    ftrans = open(prefix + "." + lang_pair + ".ref." + lang, 'w')
+    fref = open(prefix + "." + lang_pair + ".trans." + lang, 'w')
+    fid = open(prefix + "." + lang_pair + ".uccaids" , 'w')
+    fsource = open(prefix + "." + lang_pair + ".en", 'w')
+
+    for id in sorted(ids):
+
+      datasent = data.loc[data["sent_id"] == id]
+      found = 0
+      for row_index, row in datasent.iterrows():
+        found = 1
+        fid.write('%s\n' % (row.sent_id))
+        fsource.write('%s\n' % (row.source))
+        fref.write('%s\n' % (row.reference))
+        ftrans.write('%s\n' % (row.target))
+        break
+      
+      if found == 0:
+        print ("missing row" + str(id)) 
+
+
+    ftrans.close()
+    fsource.close()
+    fref.close()
+    fid.close()
+
+
+def main():
+  alldata = pandas.read_csv("sentences.csv", converters={'node_id': str, 'parent' : str})
+  extract(alldata)
+
+
+
+if __name__ == "__main__":
+  main()
+

### 2016-03-14 paper/main.tex

@@
 
-  Recent interest in semantics-aware MT systems underscores te importance of
+  Recent interest in semantics-aware MT systems underscores the importance of
   semantic evaluation for Machine Translation (MT) systems. 
@@
   and tuning of structure-aware MT, much in the same way that string-based measures,
-  such as BLEU, support advances with shallower MT systems.  
+  such as BLEU, enabled progress in shallower MT.
   We present a novel human semantic evaluation measure for MT, Human
@@
   of the MT output itself, whose interpretation is often unclear.
-  We experiment with four language pairs, demonstrating the broad applicability of
-  the measure, as well as its reliability, reflected in the inter-annotator agreement
-  scores.
+  We experiment with four language pairs, demonstrating HUME's broad applicability,
+  as well as its reliability, reflected in its inter-annotator agreement rates.
 
@@
 estimating the quality of an MT system.
-Nevertheless, common measures for MT evaluation, such as adequacy and frequency judgements
+Nevertheless, common measures for MT evaluation, such as adequacy and fluency judgements
 or the relative ranking of several possible translations, are problematic in two ways.
 First, as the quality of translation is determined by multiple factors, it is difficult
-to assign a single number that quantifies the quality of the whole entire sentence. This
+to assign a single number that quantifies the quality of the entire sentence. This
 is indeed reflected in the diminishing inter-annotator agreement rates of human ranking measures
@@
 
-These problems were partially addressed by measures that decompose over sub-parts of the sentence,
-most commonly over its words or n-grams, quantifying the overlap of the sub-parts with a reference
-translation. More recent work on MT evaluation proposed measures
+These problems were partially addressed by measures that decompose over sub-parts of the evaluated
+translation sentence (henceforth, {\it translation}),
+most commonly over its words or n-grams,
+quantifying the overlap of its sub-parts relative a reference translation.
+More recent work on MT evaluation proposed measures
 that decompose over semantically defined units,
@@
 
-In this work we propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
+We propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
 a human evaluation measure that decomposes over the semantic units of the sentence.
@@
 that the evaluator be proficient in both the source and target languages.
-See \secref{sec:hmeant_comp} for further discussion.
+See \secref{sec:hmeant_comp}.
 
-We conduct experiments with four language pairs, English-German, English-Polish,
+We conduct experiments with four language pairs: English-German, English-Polish,
 English-Romanian and English-Czech, and report efficiency (time of annotation) and
@@
 HUME scores with crowd-sourced adequacy judgements. Results show that HUME obtains 
-agreements which are on par with those of standard sentence-level ranking measures,
-but unlike them, HUME's agreement does not diminish with the sentence length.
-The paper concludes by outlining ongoing work to construct semi- and fully-automatic variants of HUME.
+agreement scores which are on par with those of standard sentence-level ranking measures,
+but unlike them, HUME's scores do not decrease with logner sentences.
+The paper concludes by outlining ongoing work to construct semi- and
+fully-automatic variants of HUME.
 
@@
 %of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
-Human evaluation is the standard measure, and is
-estimated either in terms of the relative ranking of the outputs of multiple systems
+Human evaluation is generally determined
+either according to the relative ranking of the outputs of multiple systems
 e.g., in the WMT shared tasks \cite{bojar2015findings}, or by assigning
-indiviaul adequacy/fluency scores to each sentence, a procedure recently improved
+indiviaul adequacy/fluency scores to each translation, a procedure recently improved
 by \newcite{graham2015improving}.
 However, while providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
-and is unfeasible for a wide variety of systems or hyper-parameter settings are to be evaluated.
-which has motivated the construction of cheaper methods that correlate with human
-judgement.
+and is unfeasible where a wide variety of systems or parameter settings are to be evaluated.
+% has motivated the construction of cheaper methods that correlate with human
+%judgement.
 
-Scalability is addressed by scores that decompose over the sub-parts of
+Scalability is addressed by employing automatic and semi-automatic approximations of human
+judgements. Most commonly, such scores decompose over the sub-parts of
 the translation, and quantify how many of these sub-parts appear in a reference translation,
-whose validity has been manually confirmed. The most commonly used measures decompose
-over n-grams or even individual words, e.g., 
+whose validity has been manually confirmed. This decomposition further allows system developers
+to localize the errors within the translation.
+The most commonly used measures decompose over n-grams or even individual words, e.g., 
 the BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR measures \cite{Banerjee:2005}.
-Other methods employ a string edit distance for determining the similarity between the
-reference and translation \cite{snover2006study}.
+Another common approach is to determine the similarity between the reference and translation
+by applying string edit distance \cite{snover2006study}.
 While these measures stimulated much progress in MT research by allowing
-to evaluate experiments no a massive scale, they 
-decompose over the words and n-grams of the sentence,
-and are thus better suited for localizing errors
-in the output of shallow MT methods, and are less telling for identifying the semantic
-nature of translation errors in the output.
+the evaluation of massive-scale experiments,
+their decomposition over words and n-grams is better suited for localizing errors
+in the output of shallow MT methods, and is less instructive in identifying the semantic
+nature of the errors.
 
@@
 
-To address this shortcoming, more more recent work quantifies
-the similarity of the reference and translation output in terms
+In order to address this shortcoming, more recent work quantified
+the similarity of the reference and translation in terms
 of their {\it semantic structure}. \newcite{liu2005syntactic} took a syntactic approach, 
@@
 \newcite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
-\newcite{gimenez2007linguistic} proposed an evaluation measure that incorapotes features
-that capture the overlap between the output and reference in terms of their
-semantic (argument structure phenomena), lexical and morphosyntactic features.
+\newcite{gimenez2007linguistic} proposed to aggregate multiple types of information,
+capturing the overlap between the output and reference in terms of their
+semantic (predicate-argument structures), lexical and morphosyntactic features.
 
@@
 their verbal argument structures structures, and associated semantic roles.
-HMEANT shares much of HUME's motivation for providing a
-measure that decomposes over semantic units, rather than strings, but focuses on verbal
-argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
-relations and copula clauses. HUME further diverges from both MEANT and the more shallow BLEU and
-METEOR measures, in not requiring a reference. Instead,
-it compares the source directly with the output
-translation. We see this difference as crucial where semantic analysis is concerned, as it
-is often impossible to assign a semantic structure to a low quality MT output sentence.
-See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences between HUME
-and HMEANT.
+We discuss the differences between HMEANT and HUME in \secref{sec:hmeant_comp}.
 
@@
 MT evaluation measure for a number of reasons. 
-First, it defines a coarse-grained and intuitive set of distinctions definitions,
+First, it defines a coarse-grained and intuitive set of categories,
 and consequently can be reliably annotated by non-experts, after as little as two hours
@@
 argument structure may be translated or paraphrased to a nominal or adjectival
-argument structure, and are thus less motivated as a basis for a semantic measure.
+argument structure, and are thus less suitable as a basis for a semantic measure.
 For example, ``after he graduated, Tom moved to NYC''
@@
 thus not straightwardly allowing a decomposition of the sentence into sub-units.
-Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.
+Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.\oa{should we
+  add that it is less cross-linguistically stable? maybe that's an overkill..}
 
@@
 
-We do not distinguish fluency and adequacy judgements, as they are correlated anyway \cite{ccb}.
+%We do not distinguish fluency and adequacy judgements, as they are correlated anyway \cite{ccb}.
 
@@
 There are two manual steps in the annotation process. The first is to create UCCA annotations for the source sentence,
-following the UCCA guidelines. Since this is is fully explained elsewhere \cite{abend2013universal}, we will not go into detail
-here, except to say that in our case the source sentences were all in English, and  the annotation was performed by four
-computational linguists. After we have the UCCA annotation on 
-the source, we use the word-to-word alignments from the MT system in order to project this annotation to the MT output, and
-we used native speakers of the target language in order to annotate this output. Since these annotators have to
+following the
+UCCA guidelines\footnote{All UCCA-related resources can be found here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}. Source sentences are all in English, and the annotation was performed by four
+computational linguists\oa{were there no other annotators from Prague?}. After we have the UCCA annotation on 
+the source, we use the word-to-word alignments from the MT system in order to project this annotation to the MT output,
+and use native speakers of the target language in order to annotate this output. Since these annotators have to
 judge the quality of the translation, they also needed to be fluent in English.\bh{Somewhere we could explain that we
@@
 The basic strategy for annotating HUME on the  MT output to step through the source sentence semantic 
-components (projected UCCA nodes), looking at the translation via the word alignments, and marking on the source structure
-which parts have 
-been correctly translated.
-There are two kinds of components (nodes): a word or basic semantic unit, and a structural component which contains one
-or more sub-components.
+components (projected UCCA nodes), looking at the translation via the word alignments, and marking
+on the source structure which parts have been correctly translated.
+There are two kinds of components (nodes): a word or basic semantic unit, and a structural component
+which contains one or more sub-components.
 
-\paragraph{Lexical nodes} are usually comprised of individual words. They are the leaf nodes of the tree, the smallest 
-meaning-bearing
-units.
+\paragraph{Lexical nodes} are usually comprised of individual words. They are the leaf nodes of the tree,
+the smallest meaning-bearing units.
 Leaf nodes can be labelled as green (correct), orange (partially correct) and red (incorrect). 
@@
 
+HMEANT shares much of HUME's motivation for providing a
+measure that decomposes over semantic units, rather than strings, but focuses on verbal
+argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
+relations and copula clauses.
+HUME further diverges from both MEANT and the shallower BLEU and
+METEOR measures, in not requiring a reference.
+Instead, it compares the source directly with the output
+translation. We see this difference as crucial where semantic analysis is concerned, as it
+is often impossible to assign a semantic structure to a low quality MT output sentence.
+See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences between HUME
+and HMEANT.
+
 For instance, consider the sentence ``A coronary angioplasty may not be technically possible'',
@@
 We conducted an analysis of the English UCCA Wikipedia
-corpus\footnote{\url{http://www.cs.huji.ac.il/~oabend/ucca.html}}, comprising
+corpus, comprising
 of 5324 sentences,

### 2016-03-14 paper/main.tex
some comments and corrections on 1-3

@@
 \newcommand{\tabref}[1]{Table~\ref{#1}}
-\newcommand{\oa}[1]{\footnote{\color{red} #1}}
-\newcommand{\bh}[2]{\footnote{\color{blue} #1}}
+\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
+\newcommand{\bh}[2]{\footnote{\color{blue}BH: #1}}
 
@@
   between the translation output and the source, is likely support the construction
-  and tuning of structure-aware MT, much in the same way that string-based measures,
+  and tuning of structure-aware\bh{What is ``structure-aware MT''} MT, much in the same way that string-based measures,
   such as BLEU, enabled progress in shallower MT.
@@
 their verb argument structure (the most notable of these measures is
-MEANT \cite{lo2011structured}).
+MEANT \cite{lo2011structured})\bh{Should we just refer to HMEANT, as that is the comparable measure from the MEANT family?}.
 While allowing to localize semantic errors in the translation,
@@
 agreement scores which are on par with those of standard sentence-level ranking measures,
-but unlike them, HUME's scores do not decrease with logner sentences.
+but unlike them, HUME's scores do not decrease with longer sentences.\bh{We may have to focus
+more on qualitative aspects, if we just try to sell the measure on IAA, then the figures
+are not stunning.}
 The paper concludes by outlining ongoing work to construct semi- and
@@
 e.g., in the WMT shared tasks \cite{bojar2015findings}, or by assigning
-indiviaul adequacy/fluency scores to each translation, a procedure recently improved
+individual adequacy/fluency scores to each translation, a procedure recently improved
 by \newcite{graham2015improving}.
@@
 as one or more participants. For example, the sentence ``After graduating, Tom moved to NYC''
-contains two Scenes, whose main relations are ``graduating'' and ``John''. The participant ``John''
+contains two Scenes, whose main relations are ``graduating'' and ``John''\bh{Don't you mean ``graduating'' and
+``moved''? And where did ``John'' come from?}. The participant ``John''
 is a part of both Scenes, while ``NYC'' only of the latter. Further categories account for
@@
 the source, we use the word-to-word alignments from the MT system in order to project this annotation to the MT output,
-and use native speakers of the target language in order to annotate this output. Since these annotators have to
+and use native speakers of the target language in order to annotate this output.\bh{Is it obvious how this was done? What about $n$-$m$ alignments?} Since these annotators have to
 judge the quality of the translation, they also needed to be fluent in English.\bh{Somewhere we could explain that we
@@
 components (projected UCCA nodes), looking at the translation via the word alignments, and marking
-on the source structure which parts have been correctly translated.
-There are two kinds of components (nodes): a word or basic semantic unit, and a structural component
+on the source structure which parts have been correctly translated. A typical display from the HUME annotation tool is
+shown in \figref{mttool}, where we see the translation mapped onto UCCA nodes and corresponding source words.
+
+There are two kinds of nodes: a word or basic semantic unit, and a structural component
 which contains one or more sub-components.

### 2016-03-14 paper/main.tex
starting to go over sec 3

--- 2016-03-14 paper/main.tex
@@
 
+\oa{explain how semantic units represent relations}
+
 While most broad-coverage semantic work focused on shallow semantic structures,
@@
 
-\subsection{Annotation Guidelines}\label{sec:guidelines}
-In this section we summarise the procedure for annotating machine translation output.
+\subsection{Annotation Procedure}\label{sec:guidelines}
 
-There are two manual steps in the annotation process. The first is to create UCCA annotations for the source sentence,
-following the
-UCCA guidelines\footnote{All UCCA-related resources can be found here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}. Source sentences are all in English, and the annotation was performed by four
-computational linguists\oa{were there no other annotators from Prague?}. After we have the UCCA annotation on 
-the source, we use the word-to-word alignments from the MT system in order to project this annotation to the MT output,
-and use native speakers of the target language in order to annotate this output. Since these annotators have to
-judge the quality of the translation, they also needed to be fluent in English.\bh{Somewhere we could explain that we
-could do a similar annotation with monolinguals, although it entails aligning reference and MT output}
+This section summarises the procedure for annotation involved in computing HUME. We mark
+the source sentence as $s$ and the translation sentence as $t$. 
+The procedure involves two manual steps: (1) UCCA-annotating $s$, 
+(2) human judgements for each semantic unit of $s$,
+defined by its UCCA annotation, of whether its meaning has been preserved in $t$.
+We note that (1) is performed once for every source sentence, irrespective
+of the number of its translations we wish to evaluate.
 
-The basic strategy for annotating HUME on the  MT output to step through the source sentence semantic 
-components (projected UCCA nodes), looking at the translation via the word alignments, and marking
+\paragraph{UCCA Annotation.}
+We begin by creating UCCA annotations for the source sentence, following the
+UCCA guidelines\footnote{All UCCA-related resources can be found
+  here: http://www.cs.huji.ac.il/~oabend/ucca.html}.
+Let $L$ be the set of possible UCCA labels.
+Given a sentence $s = (w_1, \ldots,w_n)$, the UCCA annotation is a DAG $G=(V,E,\ell)$,
+where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
+For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
+The semantic units for a sentence $s$, induced by the annotation $G$
+is the set of yields of nodes in $V$. We denote it with $Units(s)$ for brevity.
+We note that the translation need not be annotated.
+
+\paragraph{Translation Evaluation.}
+The basic strategy for computing HUME is to traverse
+the semantic units of the source sentence, and to mark the extent to which
+they have been correctly translated. As semantic units correspond to the
+semantic relations expressed in the text and the inter-relations between them,
+HUME effectively aggregates the judgements of the users into a composite score.
+
+Human annotation of the semantic units requires first deciding whether
+a unit is an {\it atomic units}, i.e., that it has no meaning-bearing
+sub-units, or a {\it structural unit}. In most cases, atomic units would
+correspond to individual words, but they may also correspond to opaque
+multi-word expressions, i.e., expressions that do not have meaningful sub-units.
+When a multi-word unit is labeled as atomic, all its sub-units are discarded
+from the evaluation.
+
+Atomic units can be labelled as ``Green'' (correct), ``Orange'' (partially correct)
+and ``Red'' (incorrect). 
+A Green label means that the meaning of the word or phrase has been largely preserved.
+An Orange label means that the essential meaning of the unit has been preserved,
+but some part of the translation is wrong.
+This could often be due to the translated word having the wrong tense,
+or the wrong morphology, in a way that impacts little on the understandability of the sentence.
+A Red label means that the essential meaning of the unit has not been captured.
+
+Structural units have sub-units (children in the UCCA graph), which may either be
+atomic or structural, and represent the relation between the sub-units.
+For instance, a Scene unit 
+represents the relation between the main relation, participants and secondary
+relations in it. Structural units may be labeld as ``Adequate'' or ``Bad'', meaning
+that the relation between the sub-units went wrong.
+
+There are many ways that the relationship between the children might go
+wrong in translation:
+
+\begin{compactitem}
+\item The components of the translation are ordered differently from the source,
+  such that the relation between them changes. For instance, where
+  the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
+  the semantic unit corresponding to the Scene will be marked as Bad, while
+  the inidividual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
+\item A superfluous word or phrase has been inserted into
+  the translation. For instance, if ``dog bites man'' is translated into ``dog visciously bites man''.
+\item One of the sub-units is missing from the translation. For instance, if ``dog bites man''
+  is translated into ``dog bites''.
+\end{compactitem}
+
+We note that the defined labels are more similar in spirit to adequacy,
+rather than fluency judgements. Specifically, annotators are instructed to
+label a unit as Adequate if its translation is understandable and preserves
+the meaning of the source unit, even if its fluency is impaired.
+
+\subsection{Composite Score}\label{sec:score}
+
+composite score (uniform, sub-sets, weighted rather not -- according to categories [or empirically according to agreement])
+
+\subsection{Annotation Interface.}
+
+As translations may be long, we provide assistance to the annotator.
+Where the annotator is asked to judge the translation quality of a unit $u$,
+the part of by highlighting the part of the translation which is likely to be
+most relevant for the semantic unit in question.
+This is done by highlighting the words in the translation which are aligned
+
+using automatically derived word-to-word  in order to project
+this annotation to the MT output.
+
+looking at the translation via the word alignments, and marking
 on the source structure which parts have been correctly translated.
@@
 
-\paragraph{Lexical nodes} are usually comprised of individual words. They are the leaf nodes of the tree,
-the smallest meaning-bearing units.
-Leaf nodes can be labelled as green (correct), orange (partially correct) and red (incorrect). 
-The traffic light system makes the marking of the
-lexical units as simple as possible.
+In order 
+, and use native speakers of the target
+language to annotate this output.
 
-\begin{compactdesc}
-\item[Green] The meaning of the word or phrase has been largely captured.
-\item[Orange] the essential meaning has been captured, but some part of the translation is wrong. 
-For example, this could often be due to the translated word having the wrong tense, or the wrong morphology. 
-\item[Red] The essential meaning of the unit has not been captured.
-\end{compactdesc}
-
-\paragraph{Structural nodes} contain other nodes which could be either lexical or structural nodes, 
-and they can be labelled as ``Adequate'' or ``Bad''. 
-What we are trying to gauge for structural nodes
-is if the children of these nodes
-relate to each other in the same way in the source sentence and in the translation. 
-There are many ways that the relationship between the children might go wrong in translation:
-\begin{compactitem}
-\item One of the child nodes is missing from the aligned translation. 
-\item An extra word or phrase has been inserted into the aligned  translation.
-\item The components of the translation are ordered differently 
-%to 
-from
-the source.
-\end{compactitem}
+Since these annotators have to judge the quality of the translation,
+they also needed to be fluent in English.\bh{Somewhere we could explain that we
+could do a similar annotation with monolinguals, although it entails aligning reference and MT output}
 
@@
 
-\subsection{Composite Score}\label{sec:score}
-
-composite score (uniform, sub-sets, weighted rather not -- according to categories [or empirically according to agreement])
-
-\subsection{Comparison with HMEANT}\label{sec:hmeant_comp}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Comparison with HMEANT}\label{sec:hmeant_comp}
 
@@
 \section{Experiments}\label{sec:experiments}
+
+\paragraph{Annotation.}
+Source sentences are all in English, and the UCCA annotation was performed by four
+computational linguists\oa{were there no other annotators from Prague?}.
 
@@
 + relies on automatic alignment (in the form of hints only)
-  These two positive design decisions are in line with what Dekai observed for HMEANT (see the abstract of http://repository.ust.hk/ir/Record/1783.1-66107): IAA of SRL drops from 90 to 61 when the two aligned structures are from two different annotators.
+These two positive design decisions are in line with what Dekai observed for HMEANT
+(see the abstract of http://repository.ust.hk/ir/Record/1783.1-66107): IAA of SRL drops from 90
+to 61 when the two aligned structures are from two different annotators.
 
@@
 + allows to indicate the quality of the predicates ("Actions")
-+/- expects to cover the whole hypothesis with annotations, whereas HMEANT 'gives up' when the hypothesis is not intelligible, giving a zero score to the bad frame/whole sentence.
-This can be a good decision for HMEANT, since our way may artificially boost the chances of understanding the sentence.
++/- expects to cover the whole hypothesis with annotations, whereas HMEANT 'gives up' when the
+hypothesis is not intelligible, giving a zero score to the bad frame/whole sentence.
+This can be a good decision for HMEANT, since our way may artificially boost the chances of
+understanding the sentence.
 

### 2016-03-14 paper/main.tex
sections 3.1 and 3.2

@@
   between the translation output and the source, is likely support the construction
-  and tuning of structure-aware\bh{What is ``structure-aware MT''} MT, much in the same way that string-based measures,
+  and tuning of semantics-based MT, much in the same way that string-based measures,
   such as BLEU, enabled progress in shallower MT.
@@
 their verb argument structure (the most notable of these measures is
-MEANT \cite{lo2011structured})\bh{Should we just refer to HMEANT, as that is the comparable measure from the MEANT family?}.
+HMEANT \cite{lo2011structured}).
 While allowing to localize semantic errors in the translation,
@@
 agreement scores which are on par with those of standard sentence-level ranking measures,
-but unlike them, HUME's scores do not decrease with longer sentences.\bh{We may have to focus
-more on qualitative aspects, if we just try to sell the measure on IAA, then the figures
-are not stunning.}
+but unlike them, HUME's scores do not decrease with longer sentences.\bh{We may
+  have to focus more on qualitative aspects, if we just try to sell
+  the measure on IAA, then the figures are not stunning.}
 The paper concludes by outlining ongoing work to construct semi- and
@@
       inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
-    \node (ROOT) [fill=black, circle] {}
-    child {node {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
-    child {node (H1) [fill=black,circle] {}
-      child {node {graduation} edge from parent node[right] {$P$}}
+  \node (ROOT) [label={\#7}] [fill=black, circle] {}
+  child {node [label={\#1}] {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
+  child {node (H1) [label={\#8}] [fill=black,circle] {}
+    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
       %child {node {to Paris} edge from parent node[right] {$A$}}
@@
     }
-    child {node (H2) [fill=black,circle] {}
+    child {node (H2) [label={\#9}] [fill=black,circle] {}
       child {node  {} edge from parent[draw=none]}
       child {node  {} edge from parent[draw=none]}
-      child {node (Tom) {Tom} edge from parent node[right] {$A$}}
-      child {node {moved} edge from parent node[right] {$P$}}
-      child {node [fill=black,circle] {}
-        child {node {to} edge from parent node[right] {$R$}}
-        child {node {NYC} edge from parent node[right] {$C$}}
+      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
+      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
+      child {node [label={\#10}] [fill=black,circle] {}
+        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
+        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
         edge from parent node[right] {$A$}}
@@
 \caption{\label{fig:ucca_example}
-  Sample UCCA annotation.}
+  Sample UCCA annotation. Edge labels mark UCCA categories. \#<index> marks the node
+indices.}
 \end{figure}
@@
 
-This section summarises the procedure for annotation involved in computing HUME. We mark
-the source sentence as $s$ and the translation sentence as $t$. 
+This section summarises the procedure for annotation involved in computing HUME.
+We mark the source sentence as $s$ and the translation sentence as $t$. 
 The procedure involves two manual steps: (1) UCCA-annotating $s$, 
@@
   the translation. For instance, if ``dog bites man'' is translated into ``dog visciously bites man''.
-\item One of the sub-units is missing from the translation. For instance, if ``dog bites man''
-  is translated into ``dog bites''.
+\item One of the sub-units is missing from the translation. For instance, if ``dog bites man'' is translated into ``dog bites''.
 \end{compactitem}
@@
 
+For example, consider the sentence ``After graduation, Tom moved
+to NYC'' \figref{fig:ucca_example}. The atomic units are the inidividual words,
+nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation
+is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
+  be in English for presentation purposes. In our experiments $t$ is in German,
+  Czech, Polish or Romanian.}.
+Then ``after'', ``graduation'', ``Tom'' and ``to'' are marked Green, ``moved'' is marked
+Orange (as it corresponds to ``will move'' in $t$), and ``NYC''
+is marked as Red, as it was mistranslated to ``Paris''. Node \#9 is marked Adequate,
+as the Scene of Tom moving is properly reflected in the translation. Node \#10, which
+expresses the relation between the graduation and moving Scenes, is marked Bad,
+as their relation in $t$ is reversed.
+
 \subsection{Composite Score}\label{sec:score}
 
-composite score (uniform, sub-sets, weighted rather not -- according to categories [or empirically according to agreement])
+We proceed to detailing how judgements on the semantic units
+of the source are aggregated into a composite score.
+While future extensive empirical evaluation with HUME, and using it to
+tune MT systems may suggest more complex measure,
+as a starting point we take a parsimonious approach
+and define aggregation of the judgements.
+
+Let $Correct(s,t)$ be the number of atomic units in $s$ marked as Green relative to $t$ plus
+the number of structural units in $s$ marked as Adequate relative to $t$.
+Let $Partial(s,t)$ be the number of atomic units in $s$ marked as Orange relative
+to $t$. Then HUME's composite score is:
+
+\vspace{-.5cm}
+\begin{equation}
+  \small
+  HUME(s,t) = \frac{Correct(s,t) + 0.5\cdot Partial(s,t)}{|Units(s)|}
+\end{equation}
+
+%composite score (uniform, sub-sets, weighted rather not -- according to categories [or empirically according to agreement])
 

### 2016-03-15 scripts/get_times.py
time calculation

@@ -1,0 +1,56 @@
+#!/usr/bin/env python
+from __future__ import print_function
+
+
+#
+# Collects time stamps and intervals
+#
+
+import argparse
+import logging
+import math
+import os
+import os.path
+import pandas
+import sys
+
+import matplotlib.pyplot as plt
+import numpy as np
+
+from scipy import stats
+
+LOG = logging.getLogger(__name__)
+   
+def main():
+  logging.basicConfig(format='%(asctime)s %(levelname)s: %(name)s: %(message)s', datefmt='%Y-%m-%d %H:%M:%S', level=logging.INFO)
+  parser = argparse.ArgumentParser()
+  parser.add_argument("-s", "--sentences-file", default="sentences.csv")
+  args = parser.parse_args()
+
+  sentences = pandas.read_csv(args.sentences_file, converters={'timestamp' : pandas.to_datetime})
+  sentences.sort_values(by = 'timestamp',inplace=True )
+
+  annots = ("de1","cs1", "ro1", "pl1",  "de2", "ro2", "pl2")
+  fig, axes = plt.subplots(nrows=len(annots), ncols=1, figsize=(12,12))
+  fig.tight_layout()
+  for fig_id, annot_id in enumerate(annots):
+    by_annot = sentences[sentences.annot_id == annot_id]
+    diffs = by_annot.timestamp - by_annot.timestamp.shift()
+    diffs = diffs.dt.seconds
+    diffs = diffs.dropna()
+    #print (diffs.head())
+    bins = diffs.groupby(pandas.cut(diffs,np.arange(0,3600,20))).size()
+    #print("mode", bins.idxmax(), bins.max())
+    diffs = diffs[diffs < 3600]
+    ax = plt.subplot(len(annots),1,fig_id)
+    ax.set_title("Timestamp delta histogram for {}".format(annot_id))
+    diffs.hist(ax = ax, bins=180)
+    median = np.median(diffs[diffs < 500])
+    plt.axvline(median, color='r', linestyle='dashed')
+    print("Annot: {} Median: {}".format(annot_id,median))
+  graph_file_name = "times.png"
+  plt.savefig(graph_file_name)
+
+if __name__ == "__main__":
+  main()
+

### 2016-03-15 paper/main.tex
new iaa, time

@@
 \subsection{Annotation Statistics}
+\label{sec:annot_stats}
 
-For the annotation of MT output, we recruited seven annotators, two for each of the target languages apart from Czech. They were
-all native speakers of the target languages and fluent in English. For Czech, we also recruited a  group of additional annotators
-to annotate small numbers of sentences each for the purpose of calculating inter-annotator agreement. In \tabref{tab:annot}
-we show the total number of sentences and nodes annotated by each annotator, where we consider the additional Czech
-annotators as a single annotator. Note that not all nodes in all sentences were annotated. Sometimes this was because the
+For the annotation of the MT output, we recruited two annotators for each of German, Romanian
+and Polish and one main annotator for Czech. In order to allow us to calculate IAA also for
+the Czech annotations we recruited several further annotators to work on a small number of 
+sentences each. We treat these further annotators as one annotator, giving us two annotators
+for each language pair.
+The annotators were all native  speakers of the target languages and fluent in English.
+
+In \tabref{tab:annot}
+we show the total number of sentences and nodes annotated by each annotator. Note that not all nodes in all sentences were annotated. Sometimes this was because the
 annotator did not finish a node, or accidentally missed a node, but in many cases it was also because it was an implicit 
@@
 Sentences &  Annot. 1 & 324   & 339  & 351  & 230  \\
- & Annot. 2 & -- & 104  & 340  & 337 \\
+ & Annot. 2 & 205 & 104  & 340  & 337 \\
 \hline
 Nodes & Annot. 1 & 8794  & 9253 & 9557  & 6152 \\
- &Annot. 2 & -- & 2906  & 9303  & 9228  \\
+ &Annot. 2 & 5553 & 2906  & 9303  & 9228  \\
 \end{tabular}
@@
 
-\emph{Timing of annotations}
+The annotators worked remotely, at their own pace, so we did not collect precise timing
+information. Nevertheless we are able to estimate the annotation time using the timestamps
+provided by the annotation tool, which are recorded whenever an annotated sentence is
+submitted. Annotators are not able to re-open a sentence once submitted.
+
+To estimate the annotation time, we calculate the time difference between successive 
+sentences, and discard outlying times since we assume there was a break taken. From inspection
+of histograms of annotation times, we set the upper threshold at 500 seconds. We
+show median annotation times in Table~\ref{tab:annot_times}, indicating that the annotation
+of a sentence takes around 2--4 minutes, with some variation between annotators.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{l|cccc}
+& cs & de & pl & ro \\
+\hline
+Annot. 1 & 255 & 140  & 138 & 96 \\
+Annot. 2 & -- & 162 & 229 & 207 \\
+\end{tabular}
+\caption{Median annotation times per sentence, in seconds.
+We do not give times for the second Czech 
+``annotator'', since this was a collection of annotators, working in parallel.}
+\label{tab:annot_times}
+\end{center}
+\end{table}
 
@@
 \label{sec:iaa}
-\begin{itemize}
-  \item Multiple annotation stats
-  \item Kappas -- ROG, AB
-  \item Kappa versus length
-  \item Kappa versus bleu?
-\end{itemize}
 
@@
 Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
-For German, Polish and Romanian we used two annotators, so we report IAA on the overlap
-between these annotators. For Czech we used one main annotator, plus several other extra
-annotators to provide IAA.
+For German, Polish and Romanian we used two annotators, with a significant overlap
+between the sentences they annotated. For Czech we had one main annotator, plus
+several second annotators, so we treat all these second annotators as a single
+annotator.
 
 To calculate Kappa, we consider the annotation task as a classification task on 
-nodes. We only consider nodes where there are at least two annotations from different
-annotators and we treat nodes with more than two annotations as a set of paired
-annotations\bh{This may not be correct, so should/could we do Kappa across the 
-variable multiple annotations on Czech}. We consider Kappa separately on the lexical
+nodes, and consider all nodes where there are two annotations from different
+annotators We report the overall Kappa, as well as separate Kappas on lexical
 nodes (annotated as red, orange or green) and the structural nodes (annotated
-as acceptable or bad) since there should not be any confusion between these two types of
-nodes. The small number of nodes which one annotator had labelled as structural and 
-the other as lexical were ignored for the purposes of IAA, since such cases of annotator 
-error could be caught by improved tool support.  The Kappas are shown in \tabref{tab:iaa}
+as acceptable or bad). In general there should not be confusion between the two
+types of nodes, and in future work the separation between the nodes should be enforced
+by the tool.  The Kappas are shown in \tabref{tab:iaa}
 \begin{table}[!ht]
@@
 \hline
-Sentences & -- & 102 & 334 & 217 \\
+Sentences & 181 & 102 & 334 & 217 \\
 \hline
-Lexical nodes & -- & 1724 & 5386 & 3570 \\
-Kappa & -- & 0.29 & 0.54 & 0.50 \\
+All nodes & 4686   & 2793   & 8384   & 5604  \\
+Kappa & 0.64   & 0.61   & 0.58   & 0.69  \\
 \hline
-Structural nodes & & 1040 & 2655 & 1989 \\
-Kappa & -- & 0.44 & 0.33 & 0.58 \\
+Lexical nodes & 2982 & 1724 & 5386 & 3570 \\
+Kappa & 0.54 & 0.29 & 0.54 & 0.50 \\
+\hline
+Structural nodes & 1602 & 1040 & 2655 & 1989 \\
+Kappa & 0.31 & 0.44 & 0.33 & 0.58 \\
 \end{tabular}
@@
 we binned the sentences according to length and measured Kappa on each of the bins. Each bin covers
-a length range of 10, and we show the variance of kappa with length in \figref{fig:iaalength}.\bh{Note that cs is missing}
+a length range of 10, and we show the variance of kappa with length in \figref{fig:iaalength}.
 
@@
 \subfloat[English-Czech]{
-  \iaafig{de}
+  \iaafig{cs}
 }
@@
 structural and lexical nodes. }
-\label{fig:iaallength}
+\label{fig:iaalength}
 

### 2016-03-16 paper/main.tex
sss

--- 2016-03-14 paper/main.tex
@@
 
-%composite score (uniform, sub-sets, weighted rather not -- according to categories [or empirically according to agreement])
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Annotation Interface.}
 
-\subsection{Annotation Interface.}
+We turn to discuss the interface for collecting human judgements on the semantic units of the source,
+relating to the translation. UCCA annotations of the source are done offline, using the UCCA web
+application\footnote{See UCCA webpage.}.
+
+\figref{fig:interface} presents the user interface of HUME. The sentence at the top
+shows the complete MT system output. The source sentence appears at the top, and
+the source translation directly below it. Underneath the source sentence appears
+its list of semantic units. For each semantic unit, including the one for the
+entire source sentence, there is a textbox, displaying the yield of that unit.
+Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
+The user is asked to select a label for each unit,
+by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the left of the unit's box.
+
+The interface presents, for each unit, the segment of the translation which aligns to it.
+This allows the user, especially in long sentences, to focus her attention on the parts
+mostly likely to be relevant for her judgement. As the alignments are automatically derived,
+and therefore noisy, the annotator is instructed to treat the aligned text is a cue, but to ignore
+it and make a judgement according to the full translation sentence, in cases where the alignment
+is misleading.
+
+Concretely, let $s = (w_1,\ldots,w_n)$ be a source sentence, $t = (w'_1,\ldots,w'_n)$ a translation,
+and $A \subset 2^{s \times t}$ a many-to-many word alignment between them.
+If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
+$t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t$.
+As $a(u)$ may not be contiguous in the text, we mark all intervening units in red.
+
+The child components of a structural node are marked with different coloured rectangles.
+When navigating through the different
+source sentence nodes, one can see the relevant sections of the translation because the aligned words in the translation are
+highlighted in the complete sentence above. Aligned translations are also shown in black alongside the source node. If the node
+is aligned to a set of discontiguous words in the translation, then the unaligned words that appear in between the aligned words
+are shown in red. Even if these words are not directly aligned to the source node, they will likely change the meaning of the
+translation and must be considered when marking the translation as correct or not. The alignments are just meant as a guide. The
+annotator should look at the complete translation when deciding on the evaluation of a node. The translation could have content
+added before or after the node which changes its structure or meaning. If an extra component is prepended to a structural node,
+for example, it should be marked as “Bad”.
 

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

--- 2016-03-15 paper/main.tex
@@
 \newcommand{\com}[1]{}
-\newcommand{\secref}[1]{Section \ref{#1}}
+\newcommand{\secref}[1]{Section~\ref{#1}}
 \newcommand{\figref}[1]{Figure~\ref{#1}}
@@
 \newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
-\newcommand{\bh}[2]{\footnote{\color{blue}BH: #1}}
+\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
 
@@
   semantic evaluation for Machine Translation (MT) systems. 
-  A semantic measure, sensitive to possible semantic discrepencies
+  A semantic measure, sensitive to possible semantic discrepancies
   between the translation output and the source, is likely support the construction
@@
   We present a novel human semantic evaluation measure for MT, Human
-  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
+  UCCA-based MT Evaluation (HUME), building on the UCCA semantic representation scheme.
   Making progress over previous work, the proposed measure takes into account
@@
 
-We propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
+We propose the Human UCCA-based MT Evaluation (HUME) metric,
 a human evaluation measure that decomposes over the semantic units of the sentence.
@@
 inter-annotator agreement scores, as well as qualitative comparison of the resulting
-HUME scores with crowd-sourced adequacy judgements. Results show that HUME obtains 
+HUME scores with crowd-sourced adequacy judgements. The results show that HUME obtains 
 agreement scores which are on par with those of standard sentence-level ranking measures,
@@
 whose validity has been manually confirmed. This decomposition further allows system developers
-to localize the errors within the translation.
+to localize the errors or at least unconfirmed parts within the translation.
 The most commonly used measures decompose over n-grams or even individual words, e.g., 
@@
 UCCA guidelines\footnote{All UCCA-related resources can be found
-  here: http://www.cs.huji.ac.il/~oabend/ucca.html}.
+  here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}.
 Let $L$ be the set of possible UCCA labels.
@@
 
-We note that the defined labels are more similar in spirit to adequacy,
+XXX: Could someone review this for Ondrej:
+Note that for non-configuration languages, the relationship between the
+governing and the dependent units is often expressed using morphological
+properties of the dependents instead of their ordering (this holds especially for the verb and its
+modifiers, less so for components of noun phrases). Since errors in morphology
+should be expressed using the traffic lights on the dependent nodes, we would expect
+a higher proportion of ``Adequate'' structural units for Czech and Polish.
+
+We defined the labels to reflect more adequacy,
 rather than fluency judgements. Specifically, annotators are instructed to
@@
 For example, consider the sentence ``After graduation, Tom moved
-to NYC'' \figref{fig:ucca_example}. The atomic units are the inidividual words,
+to NYC'' \figref{fig:ucca_example}. The atomic units are the individual words,
 nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

@@
 \newcommand{\tabref}[1]{Table~\ref{#1}}
+\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
 \newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
@@
 
+\def\perscite#1{\newcite{#1}}
+\def\parcite#1{\cite{#1}}
+\def\inparcite#1{\newcite{#1}}
+
 \newcommand\BibTeX{B{\sc ib}\TeX}
 
-\title{HUME: Human UCCA-based Machine-Translation Evaluation}
+\title{HUME: Human UCCA-Based Evaluation of Machine Translation}
 
@@
 quantifying the similarity of the output and the reference in terms of
-their verb argument structure (the most notable of these measures is
-HMEANT \cite{lo2011structured}).
+their verb argument structure; the most notable of these measures is
+HMEANT \cite{lo2011structured}.
 While allowing to localize semantic errors in the translation,
@@
 Perhaps the most notable attempt at semantic MT evaluation is MEANT and
-its human variant HMEANT, which quantifies the similarity between
+its human variant HMEANT \parcite{lo2011structured}, which quantifies the similarity between
 the reference and translation in terms of the overlap in
-their verbal argument structures structures, and associated semantic roles.
+their verbal argument structures structures and associated semantic roles.
 We discuss the differences between HMEANT and HUME in \secref{sec:hmeant_comp}.
@@
 \caption{\label{fig:ucca_example}
-  Sample UCCA annotation. Edge labels mark UCCA categories. \#<index> marks the node
+  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
 indices.}
@@
 \paragraph{Semantic Representation.}
-UCCA (Universal Conceptual Cognitive Annotation) is a cross-linguistically applicable, lightweight
+UCCA (Universal Conceptual Cognitive Annotation, \inparcite{abend2013universal}) is a cross-linguistically applicable, lightweight
 scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
@@
 viewed as a single entity according to some semantic or cognitive consideration. Edges bear
-a category, indicating the role of the sub-unit in the relation the unit represents. UCCA
+a category, indicating the role of the sub-unit in the relation the unit
+represents. \XXX{Ondrej suggests to avoid the word 'relation' here. I usually use the
+term 'relation' for what is called 'role' here, so I got confused for a while.}
+UCCA
 currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
@@
 
-XXX: Could someone review this for Ondrej:
+\XXX{Could someone review this for Ondrej:}
 Note that for non-configuration languages, the relationship between the

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

@@
 \cite{Dixon:10a,Dixon:10b,Dixon:12}, one of the most widely used frameworks for linguistic description.
-Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
-application to annotation in English, French, German and Czech\oa{Did we do experiments on Czech?}. 
+%Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
+%application to annotation in English, French, German and Czech\oa{Did we do experiments on Czech?}. 
+The cross-linguistic applicability of UCCA has been tested in annotations of
+English, French, German and Czech.
 Third, the scheme has been shown to be semantically stable
@@
 is mostly shared between the translation and the source. 
-We note that that verbal argument structures are less stable in this sense, as a verbal
+We note that verbal argument structures are less stable in this sense, as a verbal
 argument structure may be translated or paraphrased to a nominal or adjectival

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

@@
 Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.\oa{should we
-  add that it is less cross-linguistically stable? maybe that's an overkill..}
+  add that it is less cross-linguistically stable? maybe that's an overkill..
+  OB: AMR has been applied to more languages, including Chinese, so I would not
+  dare saying that it is less stable.}
 
@@
 and phrases of the text, and have generally represented semantic structures as bi-lexical dependencies.
-Examples include the Prague Treebank tectogrammatical layer \cite{hajic2012announcing}, as well
-as dependencies derived from Minimal Recursion Semantics representations \cite{oepen2006discriminant}.
+Examples include the tectogrammatical layer \parcite{sgallhp:1986} used in the Prague dependency
+treebanks, e.g. the Czech-English one \parcite{hajic2012announcing}, as well
+as dependencies derived from Minimal Recursion Semantics representations \parcite{oepen2006discriminant}.
 Nevertheless, despite their appeal as deep and intuitive means for semantic representation,

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

@@
 Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
-these approaches require linguistic expertise for their annotatation, due to the intricate and
+these approaches require linguistic expertise for their annotation, due to the intricate and
 fine-grained distinctions they provide, and are thus in our view less suitable for basing
@@
 Human annotation of the semantic units requires first deciding whether
-a unit is an {\it atomic units}, i.e., that it has no meaning-bearing
+a unit is an {\it atomic unit}, i.e., that it has no meaning-bearing
 sub-units, or a {\it structural unit}. In most cases, atomic units would
@@
 multi-word expressions, i.e., expressions that do not have meaningful sub-units.
-When a multi-word unit is labeled as atomic, all its sub-units are discarded
+When a multi-word unit is labeled as atomic, all its sub-units' annotations are discarded
 from the evaluation.
@@
 represents the relation between the main relation, participants and secondary
-relations in it. Structural units may be labeld as ``Adequate'' or ``Bad'', meaning
+relations in it\XXX{again, could we avoid the word ``relation'' here? Would
+``event'' be an acceptable synonym?}. Structural units are labeled as ``Adequate'' or ``Bad'', meaning
 that the relation between the sub-units went wrong.
@@
   the semantic unit corresponding to the Scene will be marked as Bad, while
-  the inidividual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
+  the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
 \item A superfluous word or phrase has been inserted into
@@
 \XXX{Could someone review this for Ondrej:}
-Note that for non-configuration languages, the relationship between the
+Note that for non-configurational languages, the relationship between the
 governing and the dependent units is often expressed using morphological

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

@@
 should be expressed using the traffic lights on the dependent nodes, we would expect
-a higher proportion of ``Adequate'' structural units for Czech and Polish.
+a higher proportion of ``Adequate'' structural units for Czech and Polish
+compared to German and Romanian.
 
@@
 For example, consider the sentence ``After graduation, Tom moved
-to NYC'' \figref{fig:ucca_example}. The atomic units are the individual words,
+to NYC'' as the source $s$, see \figref{fig:ucca_example}. The atomic units are the individual words,
 nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation
+$t$
 is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
-  be in English for presentation purposes. In our experiments $t$ is in German,
+  be in English for presentation purposes. In our experiments, $t$ is in German,
   Czech, Polish or Romanian.}.

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

@@
 \usepackage{acl2016}
+\usepackage{amsmath}
 \usepackage{paralist}
@@
 \newcommand\BibTeX{B{\sc ib}\TeX}
+
+% better maths typography
+\def\setsize#1{\lvert #1 \rvert}
+\def\func#1{\text{\it #1}}  % much better kerning within function names
+\def\HUME{\func{HUME}}
+\def\Correct{\func{Correct}}
+\def\Partial{\func{Partial}}
+\def\Units{\func{Units}}
 
@@
 The semantic units for a sentence $s$, induced by the annotation $G$
-is the set of yields of nodes in $V$. We denote it with $Units(s)$ for brevity.
+is the set of yields of nodes in $V$. We denote it with $\Units(s)$ for brevity.
 We note that the translation need not be annotated.
@@
 of the source are aggregated into a composite score.
-While future extensive empirical evaluation with HUME, and using it to
-tune MT systems may suggest more complex measure,
+While future extensive empirical evaluation with HUME and its application to MT
+system tuning may suggest more complex definition,
 as a starting point we take a parsimonious approach
-and define aggregation of the judgements.
+and simply aggregate the judgements.
 
-Let $Correct(s,t)$ be the number of atomic units in $s$ marked as Green relative to $t$ plus
+Let $\Correct(s,t)$ be the number of atomic units in $s$ marked as Green relative to $t$ plus
 the number of structural units in $s$ marked as Adequate relative to $t$.
-Let $Partial(s,t)$ be the number of atomic units in $s$ marked as Orange relative
+Let $\Partial(s,t)$ be the number of atomic units in $s$ marked as Orange relative
 to $t$. Then HUME's composite score is:
@@
   \small
-  HUME(s,t) = \frac{Correct(s,t) + 0.5\cdot Partial(s,t)}{|Units(s)|}
+  \HUME(s,t) = \frac{\Correct(s,t) + 0.5\cdot \Partial(s,t)}{\setsize{\Units(s)}}
 \end{equation}

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

@@
 We note that (1) is performed once for every source sentence, irrespective
-of the number of its translations we wish to evaluate.
+of the number of its translations we wish to evaluate.\footnote{Instead of
+relying on the source,
+the evaluation could be based on the UCCA structure for the \emph{reference}
+translation. This would allow monolingual annotators to take part in the
+evaluation. On the downside, the translation would have to be aligned to the
+reference and more importantly, we would risk inaccurate judgements due to
+the naturally occurring differences between the source and its reference
+translations.}
 
@@
 Since these annotators have to judge the quality of the translation,
-they also needed to be fluent in English.\bh{Somewhere we could explain that we
-could do a similar annotation with monolinguals, although it entails aligning reference and MT output}
+they also needed to be fluent in English.
+% Ondrej did the following in a footnote in Section 3.1 Annotation Procedure
+%\bh{Somewhere we could explain that we
+%could do a similar annotation with monolinguals, although it entails aligning
+%reference and MT output}
 

### 2016-03-16 paper/main.tex
www

--- 2016-03-16 paper/main.tex
@@
 \newcommand{\com}[1]{}
-\newcommand{\secref}[1]{Section~\ref{#1}}
+\newcommand{\secref}[1]{Section \ref{#1}}
 \newcommand{\figref}[1]{Figure~\ref{#1}}
 \newcommand{\tabref}[1]{Table~\ref{#1}}
-\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
 \newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
-\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
-
-\def\perscite#1{\newcite{#1}}
-\def\parcite#1{\cite{#1}}
-\def\inparcite#1{\newcite{#1}}
+\newcommand{\bh}[2]{\footnote{\color{blue}BH: #1}}
 
@@
 
-\title{HUME: Human UCCA-Based Evaluation of Machine Translation}
+\title{HUME: Human UCCA-based Machine-Translation Evaluation}
 
@@
   semantic evaluation for Machine Translation (MT) systems. 
-  A semantic measure, sensitive to possible semantic discrepancies
+  A semantic measure, sensitive to possible semantic discrepencies
   between the translation output and the source, is likely support the construction
@@
   We present a novel human semantic evaluation measure for MT, Human
-  UCCA-based MT Evaluation (HUME), building on the UCCA semantic representation scheme.
+  UCCA-based MT evaluation (HUME), building on the UCCA semantic representation scheme.
   Making progress over previous work, the proposed measure takes into account
@@
 quantifying the similarity of the output and the reference in terms of
-their verb argument structure; the most notable of these measures is
-HMEANT \cite{lo2011structured}.
+their verb argument structure (the most notable of these measures is
+HMEANT \cite{lo2011structured}).
 While allowing to localize semantic errors in the translation,
@@
 
-We propose the Human UCCA-based MT Evaluation (HUME) metric,
+We propose the Human UCCA-based MT Evaluation ({\it HUME}) metric,
 a human evaluation measure that decomposes over the semantic units of the sentence.
@@
 inter-annotator agreement scores, as well as qualitative comparison of the resulting
-HUME scores with crowd-sourced adequacy judgements. The results show that HUME obtains 
+HUME scores with crowd-sourced adequacy judgements. Results show that HUME obtains 
 agreement scores which are on par with those of standard sentence-level ranking measures,
@@
 whose validity has been manually confirmed. This decomposition further allows system developers
-to localize the errors or at least unconfirmed parts within the translation.
+to localize the errors within the translation.
 The most commonly used measures decompose over n-grams or even individual words, e.g., 
@@
 Perhaps the most notable attempt at semantic MT evaluation is MEANT and
-its human variant HMEANT \parcite{lo2011structured}, which quantifies the similarity between
+its human variant HMEANT, which quantifies the similarity between
 the reference and translation in terms of the overlap in
-their verbal argument structures structures and associated semantic roles.
+their verbal argument structures structures, and associated semantic roles.
 We discuss the differences between HMEANT and HUME in \secref{sec:hmeant_comp}.
@@
 \caption{\label{fig:ucca_example}
-  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
+  Sample UCCA annotation. Edge labels mark UCCA categories. \#<index> marks the node
 indices.}
@@
 \paragraph{Semantic Representation.}
-UCCA (Universal Conceptual Cognitive Annotation, \inparcite{abend2013universal}) is a cross-linguistically applicable, lightweight
+UCCA (Universal Conceptual Cognitive Annotation) is a cross-linguistically applicable, lightweight
 scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
@@
 viewed as a single entity according to some semantic or cognitive consideration. Edges bear
-a category, indicating the role of the sub-unit in the relation the unit
-represents. \XXX{Ondrej suggests to avoid the word 'relation' here. I usually use the
-term 'relation' for what is called 'role' here, so I got confused for a while.}
-UCCA
+a category, indicating the role of the sub-unit in the relation the unit represents. UCCA
 currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
@@
 \cite{Dixon:10a,Dixon:10b,Dixon:12}, one of the most widely used frameworks for linguistic description.
-%Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
-%application to annotation in English, French, German and Czech\oa{Did we do experiments on Czech?}. 
-The cross-linguistic applicability of UCCA has been tested in annotations of
-English, French, German and Czech.
+Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
+application to annotation in English, French, German and Czech\oa{Did we do experiments on Czech?}. 
 Third, the scheme has been shown to be semantically stable
@@
 is mostly shared between the translation and the source. 
-We note that verbal argument structures are less stable in this sense, as a verbal
+We note that that verbal argument structures are less stable in this sense, as a verbal
 argument structure may be translated or paraphrased to a nominal or adjectival
@@
 Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.\oa{should we
-  add that it is less cross-linguistically stable? maybe that's an overkill..
-  OB: AMR has been applied to more languages, including Chinese, so I would not
-  dare saying that it is less stable.}
+  add that it is less cross-linguistically stable? maybe that's an overkill..}
 
@@
 and phrases of the text, and have generally represented semantic structures as bi-lexical dependencies.
-Examples include the tectogrammatical layer \parcite{sgallhp:1986} used in the Prague dependency
-treebanks, e.g. the Czech-English one \parcite{hajic2012announcing}, as well
-as dependencies derived from Minimal Recursion Semantics representations \parcite{oepen2006discriminant}.
+Examples include the Prague Treebank tectogrammatical layer \cite{hajic2012announcing}, as well
+as dependencies derived from Minimal Recursion Semantics representations \cite{oepen2006discriminant}.
 Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
-these approaches require linguistic expertise for their annotation, due to the intricate and
+these approaches require linguistic expertise for their annotatation, due to the intricate and
 fine-grained distinctions they provide, and are thus in our view less suitable for basing
@@
 
-This section summarises the procedure for annotation involved in computing HUME.
+This section summarises the manual annotation procedure performed in order
+to compute the HUME measure. 
 We mark the source sentence as $s$ and the translation sentence as $t$. 
 The procedure involves two manual steps: (1) UCCA-annotating $s$, 
-(2) human judgements for each semantic unit of $s$,
-defined by its UCCA annotation, of whether its meaning has been preserved in $t$.
-We note that (1) is performed once for every source sentence, irrespective
+(2) human judgements as to the translation quality of each semantic unit of $s$,
+defined by its UCCA annotation, relative to $t$.
+We note that UCCA annotation is performed once for every source sentence, irrespective
 of the number of its translations we wish to evaluate.
@@
 UCCA guidelines\footnote{All UCCA-related resources can be found
-  here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}.
+  here: http://www.cs.huji.ac.il/~oabend/ucca.html}.
 Let $L$ be the set of possible UCCA labels.
-Given a sentence $s = (w_1, \ldots,w_n)$, the UCCA annotation is a DAG $G=(V,E,\ell)$,
+Given a sentence $s = (w_1, \ldots,w_n)$, its UCCA annotation is a DAG $G=(V,E,\ell)$,
 where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
 For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
-The semantic units for a sentence $s$, induced by the annotation $G$
-is the set of yields of nodes in $V$. We denote it with $Units(s)$ for brevity.
-We note that the translation need not be annotated.
+The semantic units for a sentence $s$ induced by the annotation $G$,
+is the set of yields of nodes in $V$. %We denote it with $Units(s)$ for brevity.
+We note that the translation is not UCCA- annotated.
+
+\begin{figure*}[t]
+    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
+    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
+	English paraphrase).}
+    \label{ucca-tree-hume}
+\end{figure*}
 
@@
 Human annotation of the semantic units requires first deciding whether
-a unit is an {\it atomic unit}, i.e., that it has no meaning-bearing
-sub-units, or a {\it structural unit}. In most cases, atomic units would
-correspond to individual words, but they may also correspond to opaque
+a unit is {\it atomic}, i.e., that it has no meaning-bearing
+sub-units, or {\it structural}. In most cases, atomic units 
+correspond to individual words, but they may also correspond to unanalyzable
 multi-word expressions, i.e., expressions that do not have meaningful sub-units.
-When a multi-word unit is labeled as atomic, all its sub-units' annotations are discarded
+When a multi-word unit is labeled as atomic, all its sub-units are discarded
 from the evaluation.
@@
 represents the relation between the main relation, participants and secondary
-relations in it\XXX{again, could we avoid the word ``relation'' here? Would
-``event'' be an acceptable synonym?}. Structural units are labeled as ``Adequate'' or ``Bad'', meaning
+relations in it. Structural units may be labeld as ``Adequate'' or ``Bad'', meaning
 that the relation between the sub-units went wrong.
@@
   the semantic unit corresponding to the Scene will be marked as Bad, while
-  the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
+  the inidividual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
 \item A superfluous word or phrase has been inserted into
@@
 
-\XXX{Could someone review this for Ondrej:}
-Note that for non-configurational languages, the relationship between the
-governing and the dependent units is often expressed using morphological
-properties of the dependents instead of their ordering (this holds especially for the verb and its
-modifiers, less so for components of noun phrases). Since errors in morphology
-should be expressed using the traffic lights on the dependent nodes, we would expect
-a higher proportion of ``Adequate'' structural units for Czech and Polish
-compared to German and Romanian.
-
-We defined the labels to reflect more adequacy,
+We note that the defined labels are more similar in spirit to adequacy,
 rather than fluency judgements. Specifically, annotators are instructed to
@@
 For example, consider the sentence ``After graduation, Tom moved
-to NYC'' as the source $s$, see \figref{fig:ucca_example}. The atomic units are the individual words,
+to NYC'' \figref{fig:ucca_example}. The atomic units are the inidividual words,
 nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation
-$t$
 is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
-  be in English for presentation purposes. In our experiments, $t$ is in German,
+  be in English for presentation purposes. In our experiments $t$ is in German,
   Czech, Polish or Romanian.}.
@@
 as their relation in $t$ is reversed.
+
+Another example is given in \figref{fig:ucca-tree-hume}.
+%``Golf became an obsession for his oldest daughter, she took daily lessons and became very good'',
+The figure shows an annotation of a UCCA tree, and an example annotation for
+each of the units\oa{annotation for ``she'' is missing. and for ``very good'' is wrong}.
+We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
+in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
+individually.\oa{I'm not sure I agree with the UCCA annotation}
 
@@
 
-%composite score (uniform, sub-sets, weighted rather not -- according to categories [or empirically according to agreement])
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Annotation Interface}
 
-\subsection{Annotation Interface.}
+We turn to discuss the interface for collecting human judgements on the semantic units of the source,
+relating to the translation. UCCA annotations of the source are done offline, using the UCCA web
+application\footnote{See UCCA webpage.}.
 
-As translations may be long, we provide assistance to the annotator.
-Where the annotator is asked to judge the translation quality of a unit $u$,
-the part of by highlighting the part of the translation which is likely to be
-most relevant for the semantic unit in question.
-This is done by highlighting the words in the translation which are aligned
+\figref{fig:interface} presents the user interface of HUME. The sentence at the top
+shows the complete MT system output. The source sentence appears at the top, and
+the source translation directly below it. Underneath the source sentence appears
+its list of semantic units. For each semantic unit, including the one for the
+entire source sentence, there is a textbox, displaying the yield of that unit.
+Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
+The user is asked to select a label for each unit,
+by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the left of the unit's box.
 
-using automatically derived word-to-word  in order to project
-this annotation to the MT output.
+The interface presents, for each unit, the segment of the translation which aligns to it.
+This allows the user, especially in long sentences, to focus her attention on the parts
+mostly likely to be relevant for her judgement. As the alignments are automatically derived,
+and therefore noisy, the annotator is instructed to treat the aligned text is a cue, but to ignore
+it and make a judgement according to the full translation sentence, in cases where the alignment
+is misleading.
 
-looking at the translation via the word alignments, and marking
-on the source structure which parts have been correctly translated.
-There are two kinds of components (nodes): a word or basic semantic unit, and a structural component
-which contains one or more sub-components.
+Concretely, let $s = (w_1,\ldots,w_n)$ be a source sentence, $t = (w'_1,\ldots,w'_n)$ a translation,
+and $A \subset 2^{s \times t}$ a many-to-many word alignment between them.
+If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
+$t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
+In the event that $a(u)$ is discontiguous in the text, we mark words between the left
+and right boundaries of $a(u)$ which are not contained in it (intervening words)
+in a smaller red font. Intervening words are likely change the meaning of the translation
+of $u$, and thus should be paid attention to when considering whether the translation
+is correct or not.
 
-In order 
-, and use native speakers of the target
-language to annotate this output.
+For example, consider the sentence $s=$``The man watched the dog'' and the German
+translation ``t=Der Mann sieht den Hund an''. Then ``watched'' will be aligned to
+``sieht ... an'', and ``den Hund'' will be marked as intervening word. Another example
+can be seen in \figref{fig:interface}, where ``ongoing pregnancy'' is translated to
+``Schwangerschaft ... laufenden'' (lit. ``pregnancy ... ongoing'').
 
-Since these annotators have to judge the quality of the translation,
-they also needed to be fluent in English.\bh{Somewhere we could explain that we
-could do a similar annotation with monolinguals, although it entails aligning reference and MT output}
+\
 
-If any of these changes have occurred and this damages the meaning of the aligned translation, then we mark the 
-structural node as ``Bad''. 
-However if the arrangement of the child nodes is essentially correct, except that one or
-more of the child nodes has themselves been translated wrongly (and so is marked
-orange or red), then the structural node should be 
-annotated as ``Acceptable''.
-
-\begin{figure*}[t]
-    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
-    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
-	English paraphrase).}
-    \label{ucca-tree-mteval}
-\end{figure*}
-
-%TODO describe figure 2 here
-
-There are cases where we cannot usefully compare individual words in the source to words in the target.
-We often need to compare translation at the phrase level. If this is the case, then we can 
- treat the structural node as a lexical unit, and mark it with the traffic light labels. This 
-means that none of the children of this node will be examined in order determine the semantic score
-for this sentence. 
-In Figure~\ref{ucca-tree-mteval}, we can see that the phrase ``became a passion'' has been labelled as a lexical
-node and it has been evaluated as ``Orange'', or partially correct, as the translation of this phrase ``loved'' 
-only partially capture the semantics of the source.
-
-We are separating lexical and structural evaluation in order to simplify evaluation and to localise errors
-to their point of origin. For this reason, if a word has been translated incorrectly,
-the parent node should still be correct if the number and relationship between its children are the same
-as in the source. 
-In Figure~\ref{ucca-tree-mteval}, we can see that even though an important child of the first scene is 
-incorrectly translated (``Golf''), and the children of are ordered differently, 
-the scene itself is marked as ``Acceptable''. The translation of the scene retains the structural 
-relationships between the children.
-\\
+%There are cases where we cannot usefully compare individual words in the source to words in the target.
+%We often need to compare translation at the phrase level. If this is the case, then we can 
+% treat the structural node as a lexical unit, and mark it with the traffic light labels. This 
+%means that none of the children of this node will be examined in order determine the semantic score
+%for this sentence. 
+%In Figure~\ref{ucca-tree-mteval}, we can see that the phrase ``became a passion'' has been labelled as a lexical
+%node and it has been evaluated as ``Orange'', or partially correct, as the translation of this phrase ``loved'' 
+%only partially capture the semantics of the source.
 
@@
     \begin{center}
-    \includegraphics[width=1\textwidth]{interface-annotations}
-    \caption{The HUME annotation tool, based on the UCCA tool}
-    \label{mttool}
+    \includegraphics[width=1\textwidth]{hume_interface.png}
+    \caption{The HUME annotation tool, based on the UCCA tool. The faded orange box at the top
+      presents the translation, the large box below it presents the source. The tree of semantic units
+      for the source is given underneath, where the aligned text appears in italics, and intervening words
+      appear in red with a smaller font (e.g., in the first box).}
+    \label{fig:interface}
     \end{center}

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

--- 2016-03-16 paper/main.tex
@@
 
-\subsection{Annotation Interface.}
+\subsection{Annotation Interface}
 
@@
 HMEANT shares much of HUME's motivation for providing a
-measure that decomposes over semantic units, rather than strings, but focuses on verbal
+measure that decomposes over semantic units, rather than strings. On the other
+hand, HMEANT focuses on verbal
 argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
 relations and copula clauses.
-HUME further diverges from both MEANT and the shallower BLEU and
-METEOR measures, in not requiring a reference.
+HUME further diverges from both HMEANT (as well as the shallow automatic BLEU and
+METEOR measures) in not requiring a reference.
 Instead, it compares the source directly with the output
@@
 is often impossible to assign a semantic structure to a low quality MT output sentence.
-See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences between HUME
-and HMEANT.
+
+%See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences
+%between HUME
+%and HMEANT.
+% This is here!
 
@@
 While this may be interpreted is a minor error, HMEANT will assign it a very low score, as it failed to
-translate the main verb.
+translate the main verb. We understand that this is a design feature of HMEANT:
+if the translation per se cannot be annotated with verbal argument structure
+(for whatever reason, including incomprehensibility), it cannot convey the meaning.
+HMEANT thus makes a stricter check of whether the translation can be understood
+than we do.
 
+Several experiments with HMEANT applied to languages other than English
+\parcite{bojar:wu:ssst:2012,birch-EtAl:2013:WMT,chuchunkov-tarelkin-galinskaya:2014:SSST-8}
+indicated that HMEANT does not address some important phenomena.
 We conducted an analysis of the English UCCA Wikipedia
@@
 of 5324 sentences,
-in order to assess the frequency of some of the major phenomena HMEANT fails to address,
-Coppula cluases are treated in HMEANT simply as instances of the main verb ``be'', which
+in order to assess their frequency: \begin{inparadesc}
+\item[Copula clauses] are treated in HMEANT simply as instances of the main verb ``be'', which
 generally does not convey the meaning of these clauses. They appear in 21.7\% of the sentences,
 according to a conservative estimates that only considers non-auxilliary instances of ``be''.
-Nominative argument structures, which are completely ignored under HMEANT are in fact highly
-pervasive, appearing in 48.7\% of the sentences. Linkers expressing an inter-relation between
+\item[Nominative argument structures,] which are completely ignored under HMEANT are in fact highly
+pervasive, appearing in 48.7\% of the sentences.
+\item[Linkers] expressing an inter-relation between
 clauses (mainly discourse markers and coordinating conjunctions) appear in 56\% of the
@@
   non-auxilliary instances of ``be'' and nouns using the NLTK standard tagger.}.
+\end{inparadesc}
+
 We are not aware of any empirical evaluation that demonstrates that verb argument structure,
-taken alone, capture the crux of the sentence semantics meanings, and so the coverage of
-further phenomena by HUME will first allow to determine empirically
-(rather than stipulate), what semantic structures are important
-to preserve for MT translation, and which are less important
+taken alone, captures the crux of the sentence semantics meanings.
+The coverage of
+these and other phenomena by HUME will allow to determine empirically
+(rather than stipulate), which semantic structures are important
+to preserve for MT evaluation, and which are less important.
 

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

@@
 
-\paragraph{Annotation.}
-Source sentences are all in English, and the UCCA annotation was performed by four
-computational linguists\oa{were there no other annotators from Prague?}.
-
 In order to validate the HUME metric, we ran an annotation experiment with one source language (English), and four 
@@
 Our annotation experiment sought to address the following questions:
-\begin{itemize}
-\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences.
-\item Can HUME capture translation issues that other annotation schemes (such as adequacy judgements and HMEANT
-cannot)?
-\item What can HUME annotation tell us about the strengths and weaknesses of MT?
-\end{itemize}
+\begin{compactitem}
+\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences?
+\item Can HUME capture translation issues that other annotation schemes (such as
+adequacy judgements and HMEANT)
+cannot?
+%\item What can HUME annotation tell us about the strengths and weaknesses of MT?
+%\bh{Maybe we should remove the third one, we
+%probably don't address it in this paper}
+\end{compactitem}
 
-\bh{Maybe we should remove the third one, we
-probably don't address it in this paper}
+\section{UCCA Annotation}
+Source sentences are all in English, and the UCCA annotation was performed by four
+computational linguists\oa{were there no other annotators from Prague? Ondrej
+does not remember! Was it just Ondrej+Jana Sindlerova? Then it would be one
+computer scientist and one linguist for Czech.}.
 

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

@@
 
-\section{UCCA Annotation}
+\subsection{UCCA Annotation}
 Source sentences are all in English, and the UCCA annotation was performed by four

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

@@
 for each language pair.
-The annotators were all native  speakers of the target languages and fluent in English.
-
-In \tabref{tab:annot}
+The annotators were all native  speakers of the respective target languages and fluent in English.
+
+In \tabref{tab:annot},
 we show the total number of sentences and nodes annotated by each annotator. Note that not all nodes in all sentences were annotated. Sometimes this was because the
@@
 nodes, and consider all nodes where there are two annotations from different
-annotators We report the overall Kappa, as well as separate Kappas on lexical
+annotators \XXX{unfinished sentence?}
+We report the overall Kappa, as well as separate Kappas on lexical
 nodes (annotated as red, orange or green) and the structural nodes (annotated
@@
 a length range of 10, and we show the variance of kappa with length in \figref{fig:iaalength}.
+\XXX{We need to report the number of sentences in each bin, the last German bin
+seems to be an unstable result.}
 

### 2016-03-16 paper/main.tex

--- 2016-03-16 paper/main.tex
@@
 \usepackage{url}
-%\usepackage{subfigure}
 \usepackage{pgfplotstable}
@@
 viewed as a single entity according to some semantic or cognitive consideration. Edges bear
-a category, indicating the role of the sub-unit in the relation the unit
-represents. \XXX{Ondrej suggests to avoid the word 'relation' here. I usually use the
-term 'relation' for what is called 'role' here, so I got confused for a while.}
-UCCA
-currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
+a category, indicating the role of the sub-unit in the structure the unit
+represents. 
+UCCA currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
 relations between them. The most basic notion is the {\sc scene}, which describes a movement, an
@@
 
-%We do not distinguish fluency and adequacy judgements, as they are correlated anyway \cite{ccb}.
-
 \subsection{Annotation Procedure}\label{sec:guidelines}
 
-This section summarises the procedure for annotation involved in computing HUME.
+This section summarises the manual annotation procedure performed in order
+to compute the HUME measure. 
 We mark the source sentence as $s$ and the translation sentence as $t$. 
 The procedure involves two manual steps: (1) UCCA-annotating $s$, 
-(2) human judgements for each semantic unit of $s$,
-defined by its UCCA annotation, of whether its meaning has been preserved in $t$.
-We note that (1) is performed once for every source sentence, irrespective
-of the number of its translations we wish to evaluate.\footnote{Instead of
+(2) human judgements as to the translation quality of each semantic unit of $s$,
+defined by its UCCA annotation, relative to $t$.
+We note that UCCA annotation is performed once for every source sentence, irrespective
+of the number of its translations we wish to evaluate\footnote{Instead of
 relying on the source,
@@
 the naturally occurring differences between the source and its reference
-translations.}
+translations.}.
 
@@
 Let $L$ be the set of possible UCCA labels.
-Given a sentence $s = (w_1, \ldots,w_n)$, the UCCA annotation is a DAG $G=(V,E,\ell)$,
+Given a sentence $s = (w_1, \ldots,w_n)$, its UCCA annotation is a DAG $G=(V,E,\ell)$,
 where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
 For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
-The semantic units for a sentence $s$, induced by the annotation $G$
-is the set of yields of nodes in $V$. We denote it with $\Units(s)$ for brevity.
-We note that the translation need not be annotated.
+The semantic units for a sentence $s$ induced by the annotation $G$,
+is the set of yields of nodes in $V$. %We denote it with $Units(s)$ for brevity.
+We note that the translation is not UCCA-annotated.
+
+\begin{figure*}[t]
+    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
+    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
+	English paraphrase).}
+    \label{fig:ucca-tree-hume}
+\end{figure*}
 
@@
 Human annotation of the semantic units requires first deciding whether
-a unit is an {\it atomic unit}, i.e., that it has no meaning-bearing
-sub-units, or a {\it structural unit}. In most cases, atomic units would
-correspond to individual words, but they may also correspond to opaque
+a unit is {\it atomic}, i.e., that it has no meaning-bearing
+sub-units, or {\it structural}. In most cases, atomic units 
+correspond to individual words, but they may also correspond to unanalyzable
 multi-word expressions, i.e., expressions that do not have meaningful sub-units.
@@
 atomic or structural, and represent the relation between the sub-units.
-For instance, a Scene unit 
-represents the relation between the main relation, participants and secondary
-relations in it\XXX{again, could we avoid the word ``relation'' here? Would
-``event'' be an acceptable synonym?}. Structural units are labeled as ``Adequate'' or ``Bad'', meaning
+For instance, a Scene unit represents the relation between the
+Scene elements, such as the predicate (main relation), participants and adverbials.
+Structural units are labeled as ``Adequate'' or ``Bad'', meaning
 that the relation between the sub-units went wrong.
+We make a three-way distinction between translation qualities with structural
+nodes, as opposed to the two-way distinction with atomic units, as lexical units
+often can have slight errors, in which cases we do want to provide partial credit.
 
@@
 
-\XXX{Could someone review this for Ondrej:}
+\XXX{Could someone review this for Ondrej.}
 Note that for non-configurational languages, the relationship between the
@@
 modifiers, less so for components of noun phrases). Since errors in morphology
-should be expressed using the traffic lights on the dependent nodes, we would expect
+should be expressed on the atomic units, we would expect
 a higher proportion of ``Adequate'' structural units for Czech and Polish
-compared to German and Romanian.
+compared to German and Romanian\oa{is this because that error would already be marked
+  in the atomic units, and thus not in the structural ones?}.
 
-We defined the labels to reflect more adequacy,
+We defined the labels to reflect adequacy,
 rather than fluency judgements. Specifically, annotators are instructed to
@@
 as their relation in $t$ is reversed.
+
+Another example is given in \figref{fig:ucca-tree-hume}.
+%``Golf became an obsession for his oldest daughter, she took daily lessons and became very good'',
+The figure shows an annotation of a UCCA tree, and an example annotation for
+each of the units.
+We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
+in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
+individually.\oa{I'm not sure I agree with the UCCA annotation;
+also, translation annotation for ``she'' is missing. and for ``very good'' is wrong}
 
@@
 
-%composite score (uniform, sub-sets, weighted rather not -- according to categories [or empirically according to agreement])
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Annotation Interface}
 
-As translations may be long, we provide assistance to the annotator.
-Where the annotator is asked to judge the translation quality of a unit $u$,
-the part of by highlighting the part of the translation which is likely to be
-most relevant for the semantic unit in question.
-This is done by highlighting the words in the translation which are aligned
+We turn to discuss the interface for collecting human judgements on the semantic units of the source,
+relating to the translation. UCCA annotations of the source are done offline, using the UCCA web
+application\footnote{See UCCA webpage.}.
 
-using automatically derived word-to-word  in order to project
-this annotation to the MT output.
+\figref{fig:interface} presents the user interface of HUME. The sentence at the top
+shows the complete MT system output. The source sentence appears at the top, and
+the source translation directly below it. Underneath the source sentence appears
+its list of semantic units. For each semantic unit, including the one for the
+entire source sentence, there is a textbox, displaying the yield of that unit.
+Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
+The user is asked to select a label for each unit,
+by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the left of the unit's box.
 
-looking at the translation via the word alignments, and marking
-on the source structure which parts have been correctly translated.
-There are two kinds of components (nodes): a word or basic semantic unit, and a structural component
-which contains one or more sub-components.
+The interface presents, for each unit, the segment of the translation which aligns to it.
+This allows the user, especially in long sentences, to focus her attention on the parts
+mostly likely to be relevant for her judgement. As the alignments are automatically derived,
+and therefore noisy, the annotator is instructed to treat the aligned text is a cue, but to ignore
+it and make a judgement according to the full translation sentence, in cases where the alignment
+is misleading.
 
-In order 
-, and use native speakers of the target
-language to annotate this output.
+Concretely, let $s = (w_1,\ldots,w_n)$ be a source sentence, $t = (w'_1,\ldots,w'_n)$ a translation,
+and $A \subset 2^{s \times t}$ a many-to-many word alignment between them.
+If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
+$t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
+In the event that $a(u)$ is discontiguous in the text, we mark words between the left
+and right boundaries of $a(u)$ which are not contained in it (intervening words)
+in a smaller red font. Intervening words are likely change the meaning of the translation
+of $u$, and thus should be paid attention to when considering whether the translation
+is correct or not.
 
+For example, consider the sentence $s=$``The man watched the dog'' and the German
+translation ``t=Der Mann sieht den Hund an''. Then ``watched'' will be aligned to
+``sieht ... an'', and ``den Hund'' will be marked as intervening word. Another example
+can be seen in \figref{fig:interface}, where ``ongoing pregnancy'' is translated to
+``Schwangerschaft ... laufenden'' (lit. ``pregnancy ... ongoing'').
+
+\com{
 Since these annotators have to judge the quality of the translation,
 they also needed to be fluent in English.
+}
+
 % Ondrej did the following in a footnote in Section 3.1 Annotation Procedure
@@
 
-If any of these changes have occurred and this damages the meaning of the aligned translation, then we mark the 
-structural node as ``Bad''. 
-However if the arrangement of the child nodes is essentially correct, except that one or
-more of the child nodes has themselves been translated wrongly (and so is marked
-orange or red), then the structural node should be 
-annotated as ``Acceptable''.
-
-\begin{figure*}[t]
-    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
-    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
-	English paraphrase).}
-    \label{ucca-tree-mteval}
-\end{figure*}
-
-%TODO describe figure 2 here
-
-There are cases where we cannot usefully compare individual words in the source to words in the target.
-We often need to compare translation at the phrase level. If this is the case, then we can 
- treat the structural node as a lexical unit, and mark it with the traffic light labels. This 
-means that none of the children of this node will be examined in order determine the semantic score
-for this sentence. 
-In Figure~\ref{ucca-tree-mteval}, we can see that the phrase ``became a passion'' has been labelled as a lexical
-node and it has been evaluated as ``Orange'', or partially correct, as the translation of this phrase ``loved'' 
-only partially capture the semantics of the source.
-
-We are separating lexical and structural evaluation in order to simplify evaluation and to localise errors
-to their point of origin. For this reason, if a word has been translated incorrectly,
-the parent node should still be correct if the number and relationship between its children are the same
-as in the source. 
-In Figure~\ref{ucca-tree-mteval}, we can see that even though an important child of the first scene is 
-incorrectly translated (``Golf''), and the children of are ordered differently, 
-the scene itself is marked as ``Acceptable''. The translation of the scene retains the structural 
-relationships between the children.
-\\
+%There are cases where we cannot usefully compare individual words in the source to words in the target.
+%We often need to compare translation at the phrase level. If this is the case, then we can 
+% treat the structural node as a lexical unit, and mark it with the traffic light labels. This 
+%means that none of the children of this node will be examined in order determine the semantic score
+%for this sentence. 
+%In Figure~\ref{ucca-tree-mteval}, we can see that the phrase ``became a passion'' has been labelled as a lexical
+%node and it has been evaluated as ``Orange'', or partially correct, as the translation of this phrase ``loved'' 
+%only partially capture the semantics of the source.
 
@@
 
+\com{
 In Figure~\ref{mttool} you we show the MT evaluation tool.
@@
  structure or meaning. If an extra component is prepended to a structural node, for example, it should be marked as ``Bad''.
+}
 
@@
 HMEANT shares much of HUME's motivation for providing a
-measure that decomposes over semantic units, rather than strings. On the other
-hand, HMEANT focuses on verbal
+measure that decomposes over semantic units, rather than strings, but focuses on verbal
 argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
 relations and copula clauses.
-HUME further diverges from both HMEANT (as well as the shallow automatic BLEU and
-METEOR measures) in not requiring a reference.
+HUME further diverges from both MEANT and the shallower BLEU and
+METEOR measures, in not requiring a reference.
 Instead, it compares the source directly with the output
@@
 is often impossible to assign a semantic structure to a low quality MT output sentence.
-
-%See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences
-%between HUME
-%and HMEANT.
-% This is here!
+See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences between HUME
+and HMEANT.
 
@@
 While this may be interpreted is a minor error, HMEANT will assign it a very low score, as it failed to
-translate the main verb. We understand that this is a design feature of HMEANT:
-if the translation per se cannot be annotated with verbal argument structure
-(for whatever reason, including incomprehensibility), it cannot convey the meaning.
-HMEANT thus makes a stricter check of whether the translation can be understood
-than we do.
+translate the main verb.
 
-Several experiments with HMEANT applied to languages other than English
-\parcite{bojar:wu:ssst:2012,birch-EtAl:2013:WMT,chuchunkov-tarelkin-galinskaya:2014:SSST-8}
-indicated that HMEANT does not address some important phenomena.
 We conducted an analysis of the English UCCA Wikipedia
@@
 of 5324 sentences,
-in order to assess their frequency: \begin{inparadesc}
-\item[Copula clauses] are treated in HMEANT simply as instances of the main verb ``be'', which
+in order to assess the frequency of some of the major phenomena HMEANT fails to address,
+Coppula cluases are treated in HMEANT simply as instances of the main verb ``be'', which
 generally does not convey the meaning of these clauses. They appear in 21.7\% of the sentences,
 according to a conservative estimates that only considers non-auxilliary instances of ``be''.
-\item[Nominative argument structures,] which are completely ignored under HMEANT are in fact highly
-pervasive, appearing in 48.7\% of the sentences.
-\item[Linkers] expressing an inter-relation between
+Nominative argument structures, which are completely ignored under HMEANT are in fact highly
+pervasive, appearing in 48.7\% of the sentences. Linkers expressing an inter-relation between
 clauses (mainly discourse markers and coordinating conjunctions) appear in 56\% of the
@@
   non-auxilliary instances of ``be'' and nouns using the NLTK standard tagger.}.
-\end{inparadesc}
-
 We are not aware of any empirical evaluation that demonstrates that verb argument structure,
-taken alone, captures the crux of the sentence semantics meanings.
-The coverage of
-these and other phenomena by HUME will allow to determine empirically
-(rather than stipulate), which semantic structures are important
-to preserve for MT evaluation, and which are less important.
+taken alone, capture the crux of the sentence semantics meanings, and so the coverage of
+further phenomena by HUME will first allow to determine empirically
+(rather than stipulate), what semantic structures are important
+to preserve for MT translation, and which are less important
 
@@
 Our annotation experiment sought to address the following questions:
-\begin{compactitem}
-\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences?
-\item Can HUME capture translation issues that other annotation schemes (such as
-adequacy judgements and HMEANT)
-cannot?
-%\item What can HUME annotation tell us about the strengths and weaknesses of MT?
-%\bh{Maybe we should remove the third one, we
-%probably don't address it in this paper}
-\end{compactitem}
+\begin{itemize}
+\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences.
+\item Can HUME capture translation issues that other annotation schemes (such as adequacy judgements and HMEANT
+cannot)?
+\item What can HUME annotation tell us about the strengths and weaknesses of MT?
+\end{itemize}
 
-\section{UCCA Annotation}
-Source sentences are all in English, and the UCCA annotation was performed by four
-computational linguists\oa{were there no other annotators from Prague? Ondrej
-does not remember! Was it just Ondrej+Jana Sindlerova? Then it would be one
-computer scientist and one linguist for Czech.}.
+\bh{Maybe we should remove the third one, we
+probably don't address it in this paper}
 

### 2016-03-16 paper/main.tex

@@
 
-\com{
-Since these annotators have to judge the quality of the translation,
-they also needed to be fluent in English.
-}
-
 % Ondrej did the following in a footnote in Section 3.1 Annotation Procedure
@@
 While this may be interpreted is a minor error, HMEANT will assign it a very low score, as it failed to
-translate the main verb.
-
+translate the main verb. We understand that this is a design feature of HMEANT:
+if the translation per se cannot be annotated with verbal argument structure
+(for whatever reason, including incomprehensibility), it cannot convey the meaning.
+HMEANT thus makes a stricter check of whether the translation can be understood
+than we do.
+
+Several experiments with HMEANT applied to languages other than English
+\parcite{bojar:wu:ssst:2012,birch-EtAl:2013:WMT,chuchunkov-tarelkin-galinskaya:2014:SSST-8}
+indicated that HMEANT does not address some important phenomena.
 We conducted an analysis of the English UCCA Wikipedia
@@
 \section{Experiments}\label{sec:experiments}
-
-\paragraph{Annotation.}
-Source sentences are all in English, and the UCCA annotation was performed by four
-computational linguists\oa{were there no other annotators from Prague?}.
 
@@
 Our annotation experiment sought to address the following questions:
-\begin{itemize}
-\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences.
-\item Can HUME capture translation issues that other annotation schemes (such as adequacy judgements and HMEANT
-cannot)?
-\item What can HUME annotation tell us about the strengths and weaknesses of MT?
-\end{itemize}
-
-\bh{Maybe we should remove the third one, we
-probably don't address it in this paper}
+\begin{compactitem}
+\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences?
+\item Can HUME capture translation issues that other annotation schemes (such as
+adequacy judgements and HMEANT)
+cannot?
+%\item What can HUME annotation tell us about the strengths and weaknesses of MT?
+%\bh{Maybe we should remove the third one, we
+%probably don't address it in this paper}
+\end{compactitem}
+
+\subsection{UCCA Annotation}
+Source sentences are all in English, and the UCCA annotation was performed by four
+computational linguists\oa{were there no other annotators from Prague? Ondrej
+does not remember! Was it just Ondrej+Jana Sindlerova? Then it would be one
+computer scientist and one linguist for Czech.}.
 

### 2016-03-16 paper/main.tex
dd

@@
 \paragraph{Semantic Representation.}
-UCCA (Universal Conceptual Cognitive Annotation, \inparcite{abend2013universal}) is a cross-linguistically applicable, lightweight
+UCCA (Universal Conceptual Cognitive Annotation, \inparcite{abend2013universal}) is a
+cross-linguistically applicable, lightweight
 scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
 whose leaves correspond to the words of the text.
-The graph’s nodes, called {\sc units}, are either terminals or several elements jointly
+The graph's nodes, called {\sc units}, are either terminals or several elements jointly
 viewed as a single entity according to some semantic or cognitive consideration. Edges bear
@@
 represents.
-UCCA currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
-relations between them. The most basic notion is the {\sc scene}, which describes a movement, an
+
+UCCA's current inventory of distinctions focuses on annotating argument structures
+(adjectival, nominal, verbal and others) and relations between them.
+The most basic notion is the {\sc scene}, which describes a movement, an
 action or a state which is persistent in time. Each Scene contains one main relation, as well
-as one or more participants. For example, the sentence ``After graduating, Tom moved to NYC''
-contains two Scenes, whose main relations are ``graduating'' and ``John''\bh{Don't you mean ``graduating'' and
-``moved''? And where did ``John'' come from?}. The participant ``John''
-is a part of both Scenes, while ``NYC'' only of the latter. Further categories account for
+as one or more participants. For example, the sentence ``After graduation, Tom moved to NYC''
+contains two Scenes, whose main relations are ``graduation'' and ``Tom''.
+The participant ``John'' is a part of both Scenes, while ``NYC'' only of the
+latter. Further categories account for
 relations between Scenes and the internal structures of complex participants and relations.
@@
 MT evaluation measure for a number of reasons. 
-First, it defines a coarse-grained and intuitive set of categories,
-and consequently can be reliably annotated by non-experts, after as little as two hours
+First, it defines an intuitive set of categories
+that can be reliably annotated by non-experts, after as little as two hours
 of training \cite{marinotti2014}.
-Second, UCCA is a cross-linguistically applicable scheme, seeking to represent what is shared between languages,
+Second, UCCA is a cross-linguistically applicable scheme, seeking to
+represent what is shared between languages
 by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
 \cite{Dixon:10a,Dixon:10b,Dixon:12}, one of the most widely used frameworks for linguistic description.
-%Indeed, UCCA's cross-linguistic applicability has been demonstrated through its
-%application to annotation in English, French, German and Czech\oa{Did we do experiments on Czech?}. 
 The cross-linguistic applicability of UCCA has been tested in annotations of
@@
 is mostly shared between the translation and the source.
+
 We note that verbal argument structures are less stable in this sense, as a verbal
@@
 
-\oa{explain how semantic units represent relations}
-
-While most broad-coverage semantic work focused on shallow semantic structures,
-predominantly argument structures annotated with their semantic roles, much recent
-interest has been devoted to defining more complete semantic representation schemes.
+% While most broad-coverage semantic work focused on shallow semantic structures,
+% predominantly argument structures annotated with their semantic roles, much recent
+% interest has been devoted to defining more complete semantic representation schemes.
+% \oa{should we
+% add that it is less cross-linguistically stable? maybe that's an overkill..
+%  OB: AMR has been applied to more languages, including Chinese, so I would not
+%  dare saying that it is less stable.}
+
 The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
@@
 thus not straightwardly allowing a decomposition of the sentence into sub-units.
-Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.\oa{should we
-  add that it is less cross-linguistically stable? maybe that's an overkill..
-  OB: AMR has been applied to more languages, including Chinese, so I would not
-  dare saying that it is less stable.}
+Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.
 
@@
 an interpretable MT evaluation on.
+
 %\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
@@
   the translation. For instance, if ``dog bites man'' is translated into ``dog visciously bites man''.
-\item One of the sub-units is missing from the translation. For instance, if ``dog bites man'' is translated into ``dog bites''.
+\item One of the sub-units is missing from the translation.
+  For instance, if ``dog bites man'' is translated into ``dog bites''.
 \end{compactitem}

### 2016-03-16 paper/main.tex
ondrej's comments and fixes

--- 2016-03-16 paper/main.tex
@@
 \usepackage{url}
+%\usepackage{subfigure}
 \usepackage{pgfplotstable}
@@
 viewed as a single entity according to some semantic or cognitive consideration. Edges bear
-a category, indicating the role of the sub-unit in the structure the unit
-represents. 
-UCCA currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
+a category, indicating the role of the sub-unit in the relation the unit
+represents. \XXX{Ondrej suggests to avoid the word 'relation' here. I usually use the
+term 'relation' for what is called 'role' here, so I got confused for a while.}
+UCCA
+currently focuses on annotating argument structures (adjectival, nominal, verbal and others) and
 relations between them. The most basic notion is the {\sc scene}, which describes a movement, an
@@
 
+%We do not distinguish fluency and adequacy judgements, as they are correlated anyway \cite{ccb}.
+
 \subsection{Annotation Procedure}\label{sec:guidelines}
 
-This section summarises the manual annotation procedure performed in order
-to compute the HUME measure. 
+This section summarises the procedure for annotation involved in computing HUME.
 We mark the source sentence as $s$ and the translation sentence as $t$. 
 The procedure involves two manual steps: (1) UCCA-annotating $s$, 
-(2) human judgements as to the translation quality of each semantic unit of $s$,
-defined by its UCCA annotation, relative to $t$.
-We note that UCCA annotation is performed once for every source sentence, irrespective
-of the number of its translations we wish to evaluate\footnote{Instead of
+(2) human judgements for each semantic unit of $s$,
+defined by its UCCA annotation, of whether its meaning has been preserved in $t$.
+We note that (1) is performed once for every source sentence, irrespective
+of the number of its translations we wish to evaluate.\footnote{Instead of
 relying on the source,
@@
 the naturally occurring differences between the source and its reference
-translations.}.
+translations.}
 
@@
 Let $L$ be the set of possible UCCA labels.
-Given a sentence $s = (w_1, \ldots,w_n)$, its UCCA annotation is a DAG $G=(V,E,\ell)$,
+Given a sentence $s = (w_1, \ldots,w_n)$, the UCCA annotation is a DAG $G=(V,E,\ell)$,
 where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
 For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
-The semantic units for a sentence $s$ induced by the annotation $G$,
-is the set of yields of nodes in $V$. %We denote it with $Units(s)$ for brevity.
-We note that the translation is not UCCA-annotated.
-
-\begin{figure*}[t]
-    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
-    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
-	English paraphrase).}
-    \label{fig:ucca-tree-hume}
-\end{figure*}
+The semantic units for a sentence $s$, induced by the annotation $G$
+is the set of yields of nodes in $V$. We denote it with $\Units(s)$ for brevity.
+We note that the translation need not be annotated.
 
@@
 Human annotation of the semantic units requires first deciding whether
-a unit is {\it atomic}, i.e., that it has no meaning-bearing
-sub-units, or {\it structural}. In most cases, atomic units 
-correspond to individual words, but they may also correspond to unanalyzable
+a unit is an {\it atomic unit}, i.e., that it has no meaning-bearing
+sub-units, or a {\it structural unit}. In most cases, atomic units would
+correspond to individual words, but they may also correspond to opaque
 multi-word expressions, i.e., expressions that do not have meaningful sub-units.
@@
 atomic or structural, and represent the relation between the sub-units.
-For instance, a Scene unit represents the relation between the
-Scene elements, such as the predicate (main relation), participants and adverbials.
-Structural units are labeled as ``Adequate'' or ``Bad'', meaning
+For instance, a Scene unit 
+represents the relation between the main relation, participants and secondary
+relations in it\XXX{again, could we avoid the word ``relation'' here? Would
+``event'' be an acceptable synonym?}. Structural units are labeled as ``Adequate'' or ``Bad'', meaning
 that the relation between the sub-units went wrong.
-We make a three-way distinction between translation qualities with structural
-nodes, as opposed to the two-way distinction with atomic units, as lexical units
-often can have slight errors, in which cases we do want to provide partial credit.
 
@@
 
-\XXX{Could someone review this for Ondrej.}
+\XXX{Could someone review this for Ondrej:}
 Note that for non-configurational languages, the relationship between the
@@
 modifiers, less so for components of noun phrases). Since errors in morphology
-should be expressed on the atomic units, we would expect
+should be expressed using the traffic lights on the dependent nodes, we would expect
 a higher proportion of ``Adequate'' structural units for Czech and Polish
-compared to German and Romanian\oa{is this because that error would already be marked
-  in the atomic units, and thus not in the structural ones?}.
-
-We defined the labels to reflect adequacy,
+compared to German and Romanian.
+
+We defined the labels to reflect more adequacy,
 rather than fluency judgements. Specifically, annotators are instructed to
@@
 
-Another example is given in \figref{fig:ucca-tree-hume}.
-%``Golf became an obsession for his oldest daughter, she took daily lessons and became very good'',
-The figure shows an annotation of a UCCA tree, and an example annotation for
-each of the units.
-We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
-in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
-individually.\oa{I'm not sure I agree with the UCCA annotation;
-also, translation annotation for ``she'' is missing. and for ``very good'' is wrong}
-
 \subsection{Composite Score}\label{sec:score}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%composite score (uniform, sub-sets, weighted rather not -- according to categories [or empirically according to agreement])
+
 \subsection{Annotation Interface}
 
-We turn to discuss the interface for collecting human judgements on the semantic units of the source,
-relating to the translation. UCCA annotations of the source are done offline, using the UCCA web
-application\footnote{See UCCA webpage.}.
-
-\figref{fig:interface} presents the user interface of HUME. The sentence at the top
-shows the complete MT system output. The source sentence appears at the top, and
-the source translation directly below it. Underneath the source sentence appears
-its list of semantic units. For each semantic unit, including the one for the
-entire source sentence, there is a textbox, displaying the yield of that unit.
-Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
-The user is asked to select a label for each unit,
-by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the left of the unit's box.
-
-The interface presents, for each unit, the segment of the translation which aligns to it.
-This allows the user, especially in long sentences, to focus her attention on the parts
-mostly likely to be relevant for her judgement. As the alignments are automatically derived,
-and therefore noisy, the annotator is instructed to treat the aligned text is a cue, but to ignore
-it and make a judgement according to the full translation sentence, in cases where the alignment
-is misleading.
-
-Concretely, let $s = (w_1,\ldots,w_n)$ be a source sentence, $t = (w'_1,\ldots,w'_n)$ a translation,
-and $A \subset 2^{s \times t}$ a many-to-many word alignment between them.
-If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
-$t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
-In the event that $a(u)$ is discontiguous in the text, we mark words between the left
-and right boundaries of $a(u)$ which are not contained in it (intervening words)
-in a smaller red font. Intervening words are likely change the meaning of the translation
-of $u$, and thus should be paid attention to when considering whether the translation
-is correct or not.
-
-For example, consider the sentence $s=$``The man watched the dog'' and the German
-translation ``t=Der Mann sieht den Hund an''. Then ``watched'' will be aligned to
-``sieht ... an'', and ``den Hund'' will be marked as intervening word. Another example
-can be seen in \figref{fig:interface}, where ``ongoing pregnancy'' is translated to
-``Schwangerschaft ... laufenden'' (lit. ``pregnancy ... ongoing'').
-
+As translations may be long, we provide assistance to the annotator.
+Where the annotator is asked to judge the translation quality of a unit $u$,
+the part of by highlighting the part of the translation which is likely to be
+most relevant for the semantic unit in question.
+This is done by highlighting the words in the translation which are aligned
+
+using automatically derived word-to-word  in order to project
+this annotation to the MT output.
+
+looking at the translation via the word alignments, and marking
+on the source structure which parts have been correctly translated.
+There are two kinds of components (nodes): a word or basic semantic unit, and a structural component
+which contains one or more sub-components.
+
+In order 
+, and use native speakers of the target
+language to annotate this output.
+
+Since these annotators have to judge the quality of the translation,
+they also needed to be fluent in English.
 % Ondrej did the following in a footnote in Section 3.1 Annotation Procedure
@@
 
-%There are cases where we cannot usefully compare individual words in the source to words in the target.
-%We often need to compare translation at the phrase level. If this is the case, then we can 
-% treat the structural node as a lexical unit, and mark it with the traffic light labels. This 
-%means that none of the children of this node will be examined in order determine the semantic score
-%for this sentence. 
-%In Figure~\ref{ucca-tree-mteval}, we can see that the phrase ``became a passion'' has been labelled as a lexical
-%node and it has been evaluated as ``Orange'', or partially correct, as the translation of this phrase ``loved'' 
-%only partially capture the semantics of the source.
+If any of these changes have occurred and this damages the meaning of the aligned translation, then we mark the 
+structural node as ``Bad''. 
+However if the arrangement of the child nodes is essentially correct, except that one or
+more of the child nodes has themselves been translated wrongly (and so is marked
+orange or red), then the structural node should be 
+annotated as ``Acceptable''.
+
+\begin{figure*}[t]
+    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
+    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
+	English paraphrase).}
+    \label{ucca-tree-mteval}
+\end{figure*}
+
+%TODO describe figure 2 here
+
+There are cases where we cannot usefully compare individual words in the source to words in the target.
+We often need to compare translation at the phrase level. If this is the case, then we can 
+ treat the structural node as a lexical unit, and mark it with the traffic light labels. This 
+means that none of the children of this node will be examined in order determine the semantic score
+for this sentence. 
+In Figure~\ref{ucca-tree-mteval}, we can see that the phrase ``became a passion'' has been labelled as a lexical
+node and it has been evaluated as ``Orange'', or partially correct, as the translation of this phrase ``loved'' 
+only partially capture the semantics of the source.
+
+We are separating lexical and structural evaluation in order to simplify evaluation and to localise errors
+to their point of origin. For this reason, if a word has been translated incorrectly,
+the parent node should still be correct if the number and relationship between its children are the same
+as in the source. 
+In Figure~\ref{ucca-tree-mteval}, we can see that even though an important child of the first scene is 
+incorrectly translated (``Golf''), and the children of are ordered differently, 
+the scene itself is marked as ``Acceptable''. The translation of the scene retains the structural 
+relationships between the children.
+\\
 
@@
     \begin{center}
-    \includegraphics[width=1\textwidth]{hume_interface.png}
-    \caption{The HUME annotation tool, based on the UCCA tool. The faded orange box at the top
-      presents the translation, the large box below it presents the source. The tree of semantic units
-      for the source is given underneath, where the aligned text appears in italics, and intervening words
-      appear in red with a smaller font (e.g., in the first box).}
-    \label{fig:interface}
+    \includegraphics[width=1\textwidth]{interface-annotations}
+    \caption{The HUME annotation tool, based on the UCCA tool}
+    \label{mttool}
     \end{center}
@@
 
-\com{
 In Figure~\ref{mttool} you we show the MT evaluation tool.
@@
  structure or meaning. If an extra component is prepended to a structural node, for example, it should be marked as ``Bad''.
-}
 
@@
 \section{Comparison with HMEANT}\label{sec:hmeant_comp}
+
+\subsection{One vs. Two Structures}
+
+\subsection{Verbal Structures Only?}
+
+\subsection{Monolingual vs. Bilingual Evaluation}
++/- expects to cover the whole hypothesis with annotations, whereas HMEANT 'gives up' when the
+hypothesis is not intelligible, giving a zero score to the bad frame/whole sentence.
+This can be a good decision for HMEANT, since our way may artificially boost the chances of
+understanding the sentence.
+
+\subsection{Technical Differences}
+
++ allows for multi-word and even non-contiguous elements
+
++ allows to indicate the quality of the predicates ("Actions")
+
+\subsection{XXX to be moved}
 
@@
 to preserve for MT evaluation, and which are less important.
+
+\subsection{from e-mails}
+
+UCCA-Eval in contrast to HMEANT:
+
+- needs *very* trained UCCA annotators
+- we were not very quick at explaining even the Eval annotations, but I guess we could have done better
++ relies only on one structure
++ relies on automatic alignment (in the form of hints only)
+These two positive design decisions are in line with what Dekai observed for HMEANT
+(see the abstract of http://repository.ust.hk/ir/Record/1783.1-66107): IAA of SRL drops from 90
+to 61 when the two aligned structures are from two different annotators.
+
+Actually, I think this is a +/- .
+UCCA-Eval only needs the semantic evaluation done once for a given test source, whereas for the HMEANT set-up you have to keep annotating MT output.
+That's a + for UCCA-Eval.
+However we need bilingual annotators, which should be a -.
+Of course we could align translation and reference using monolingual annotators, but then we have the problem of alignment.
+In the *MEANT papers they mainly test by either
+A) correlating the metric to ranking judgements; or
+B) using automated versions of the metric for tuning MT systems
+
+I think neither is particularly convincing, but I'm not sure at the moment what we replace them with.
+Dekai also made a claim about the *MEANT metrics being "interpretable" in the sense that they could guide you towards semantic errors that the MT system may be making, however I don't recall seeing this done in his papers.
+
+By the claim that HMEANT explains things, Dekai probably means that the score is decomposable, but so is UCCA-Eval and even BLEU.
+And UCCA-Eval has one benefit over HMEANT: it points to words in the MT output that are aligned with something but do not convey the meaning well (and we know which part of the source meaning).
+Pure HMEANT might say only that some part of the MT output has no clear frames whatsover (and we do not know which part of the source/reference, unless we add the word alignments).
+
+Differences of our approach:
+
+The annotation experiment that we just conducted differs in two main ways from HMEANT:
+1) The semantic framework is different - UCCA vs SRL - and SRL has problems which Omri lists below, some of which Dekai dismissed when they were brought up.
+Maybe these problems do not matter for MT evaluation or tuning, but I think this would have to be determined empirically.
+2) The annotation process is different.
+We semantically annotate the source, then automatically project it to the translation, and ask bilingual annotators to assess whether this projection shows a good or bad translation.
+In HMEANT they semantically annotate both the reference and the translation, then align these manually, and assess to what extent they match.
+
+Advatages of our approach:
+Here is an additional argument to support the need for target-only intelligibility:
+http://www.aclweb.org/anthology/W14-4005
+see Table  3 and the text in Section 5.2.
+The more frames can people annotate in the hypothesis, the better the hypothesis overall.
 
@@
 
-\subsection{Copied from emails}
-
-UCCA-Eval in contrast to HMEANT:
-
-- needs *very* trained UCCA annotators
-- we were not very quick at explaining even the Eval annotations, but I guess we could have done better
-+ relies only on one structure
-+ relies on automatic alignment (in the form of hints only)
-These two positive design decisions are in line with what Dekai observed for HMEANT
-(see the abstract of http://repository.ust.hk/ir/Record/1783.1-66107): IAA of SRL drops from 90
-to 61 when the two aligned structures are from two different annotators.
-
-+ allows for multi-word and even non-contiguous elements
-+ allows to indicate the quality of the predicates ("Actions")
-+/- expects to cover the whole hypothesis with annotations, whereas HMEANT 'gives up' when the
-hypothesis is not intelligible, giving a zero score to the bad frame/whole sentence.
-This can be a good decision for HMEANT, since our way may artificially boost the chances of
-understanding the sentence.
-
-HMEAN disadvantages:
-.
-Importantly: it disregards copula clauses, nominalizations (which are often translated to verbal clauses), inter-clause relations (discourse markers, complement clauses etc.).
-UCCA-Eval addresses these.
-
-Actually, I think this is a +/- .
-UCCA-Eval only needs the semantic evaluation done once for a given test source, whereas for the HMEANT set-up you have to keep annotating MT output.
-That's a + for UCCA-Eval.
-However we need bilingual annotators, which should be a -.
-Of course we could align translation and reference using monolingual annotators, but then we have the problem of alignment.
-In the *MEANT papers they mainly test by either
-A) correlating the metric to ranking judgements; or
-B) using automated versions of the metric for tuning MT systems
-
-I think neither is particularly convincing, but I'm not sure at the moment what we replace them with.
-Dekai also made a claim about the *MEANT metrics being "interpretable" in the sense that they could guide you towards semantic errors that the MT system may be making, however I don't recall seeing this done in his papers.
-
-By the claim that HMEANT explains things, Dekai probably means that the score is decomposable, but so is UCCA-Eval and even BLEU.
-And UCCA-Eval has one benefit over HMEANT: it points to words in the MT output that are aligned with something but do not convey the meaning well (and we know which part of the source meaning).
-Pure HMEANT might say only that some part of the MT output has no clear frames whatsover (and we do not know which part of the source/reference, unless we add the word alignments).
-
-Differences of our approach:
-
-The annotation experiment that we just conducted differs in two main ways from HMEANT:
-1) The semantic framework is different - UCCA vs SRL - and SRL has problems which Omri lists below, some of which Dekai dismissed when they were brought up.
-Maybe these problems do not matter for MT evaluation or tuning, but I think this would have to be determined empirically.
-2) The annotation process is different.
-We semantically annotate the source, then automatically project it to the translation, and ask bilingual annotators to assess whether this projection shows a good or bad translation.
-In HMEANT they semantically annotate both the reference and the translation, then align these manually, and assess to what extent they match.
-
-Advatages of our approach:
-Here is an additional argument to support the need for target-only intelligibility:
-http://www.aclweb.org/anthology/W14-4005
-see Table  3 and the text in Section 5.2.
-The more frames can people annotate in the hypothesis, the better the hypothesis overall.
-
 \section{Conclusion}

### 2016-03-16 paper/main.tex

--- 2016-03-16 paper/main.tex
@@
 
+\subsection{One vs. Two Structures}
+
+\subsection{Verbal Structures Only?}
+
+\subsection{Monolingual vs. Bilingual Evaluation}
++/- expects to cover the whole hypothesis with annotations, whereas HMEANT 'gives up' when the
+hypothesis is not intelligible, giving a zero score to the bad frame/whole sentence.
+This can be a good decision for HMEANT, since our way may artificially boost the chances of
+understanding the sentence.
+
+\subsection{Technical Differences}
+
++ allows for multi-word and even non-contiguous elements
+
++ allows to indicate the quality of the predicates ("Actions")
+
+\subsection{XXX to be moved}
+
 HMEANT shares much of HUME's motivation for providing a

### 2016-03-16 paper/main.tex

@@
 Importantly, HUME only requires semantically annotating the source sentence,
-thus avoiding the annotatation of machine-generated text 
+thus avoiding the annotation of machine-generated text 
 and allowing re-use of the source semantic annotation for measuring the quality
@@
 
-\subsection{One vs. Two Structures}
-
-\subsection{Verbal Structures Only?}
-
-\subsection{Monolingual vs. Bilingual Evaluation}
-+/- expects to cover the whole hypothesis with annotations, whereas HMEANT 'gives up' when the
-hypothesis is not intelligible, giving a zero score to the bad frame/whole sentence.
-This can be a good decision for HMEANT, since our way may artificially boost the chances of
-understanding the sentence.
-
-\subsection{Technical Differences}
-
-+ allows for multi-word and even non-contiguous elements
-
-+ allows to indicate the quality of the predicates ("Actions")
-
-\subsection{XXX to be moved}
-
 HMEANT shares much of HUME's motivation for providing a
 measure that decomposes over semantic units, rather than strings. On the other
-hand, HMEANT focuses on verbal
+hand, there are important differences in design decisions, which we summarize in this
+section.
+
+First of all, the annotation process is different. We semantically annotate the
+source, then automatically project it to the translation, and ask bilingual
+annotators to assess whether this projection shows a good or bad translation. In
+HMEANT, they semantically annotate both the reference and the translation, then
+align these manually, and assess to what extent they match.
+
+\subsection{Verbal Structures Only?}
+
+HMEANT focuses on verbal
 argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
-relations and copula clauses.
-HUME further diverges from both HMEANT (as well as the shallow automatic BLEU and
-METEOR measures) in not requiring a reference.
-Instead, it compares the source directly with the output
-translation. We see this difference as crucial where semantic analysis is concerned, as it
-is often impossible to assign a semantic structure to a low quality MT output sentence.
-
-%See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences
-%between HUME
-%and HMEANT.
-% This is here!
-
-For instance, consider the sentence ``A coronary angioplasty may not be technically possible'',
-which was translated by one of the systems we experimented with as
-``Eine koronare Angioplastie kann nicht technisch m{\"o}glich'', i.e., the main verb ``sein'' (``be'') is omitted.
-While this may be interpreted is a minor error, HMEANT will assign it a very low score, as it failed to
-translate the main verb. We understand that this is a design feature of HMEANT:
-if the translation per se cannot be annotated with verbal argument structure
-(for whatever reason, including incomprehensibility), it cannot convey the meaning.
-HMEANT thus makes a stricter check of whether the translation can be understood
-than we do.
-
-Several experiments with HMEANT applied to languages other than English
-\parcite{bojar:wu:ssst:2012,birch-EtAl:2013:WMT,chuchunkov-tarelkin-galinskaya:2014:SSST-8}
-indicated that HMEANT does not address some important phenomena.
+relations and copula clauses. Some of the uncovered cases were already observed
+in previous applications of HMEANT to languages other than English (see \perscite{birch-EtAl:2013:WMT} for German, \perscite{bojar:wu:ssst:2012} for
+Czech and by \perscite{chuchunkov-tarelkin-galinskaya:2014:SSST-8} for Russian)
+but the problems are apparent in English as well.
+
 We conducted an analysis of the English UCCA Wikipedia
@@
 
-\subsection{from e-mails}
-
-UCCA-Eval in contrast to HMEANT:
-
-- needs *very* trained UCCA annotators
-- we were not very quick at explaining even the Eval annotations, but I guess we could have done better
-+ relies only on one structure
-+ relies on automatic alignment (in the form of hints only)
-These two positive design decisions are in line with what Dekai observed for HMEANT
-(see the abstract of http://repository.ust.hk/ir/Record/1783.1-66107): IAA of SRL drops from 90
-to 61 when the two aligned structures are from two different annotators.
-
-Actually, I think this is a +/- .
-UCCA-Eval only needs the semantic evaluation done once for a given test source, whereas for the HMEANT set-up you have to keep annotating MT output.
-That's a + for UCCA-Eval.
-However we need bilingual annotators, which should be a -.
-Of course we could align translation and reference using monolingual annotators, but then we have the problem of alignment.
+\subsection{Cost of Structural Annotation}
+
+HMEANT structural annotation (semantic role labelling, SRL), similar to e.g.
+PropBank
+\parcite{palmer2010semantic}, is designed to be very simple and easy to explain
+to lay annotators.
+HUME relies on UCCA, which is arguably more abstract, harder to
+explain, and thus more expensive to obtain.
+
+On the other hand, HUME 
+needs the semantic annotation
+done once for a given source sentence, whereas for the HMEANT set-up, every new
+MT output has to be structurally annotated.
+
+\subsection{One vs. Two Structures}
+
+HUME uses just one structural annotation (for the source) while HMEANT relies on
+two independently constructed structural annotations (for both the reference and
+the candidate translation).
+
+We see this difference as crucial where semantic analysis is concerned, as it
+is often impossible to assign a semantic structure to a low quality MT output sentence.
+
+For instance, consider the sentence ``A coronary angioplasty may not be technically possible'',
+which was translated by one of the systems we experimented with as
+``Eine koronare Angioplastie kann nicht technisch m{\"o}glich'', i.e., the main verb ``sein'' (``be'') is omitted.
+While this may be interpreted is a minor error, HMEANT will assign it a very low score, as it failed to
+translate the main verb.
+
+We understand that this is a design feature of HMEANT:
+if the translation per se cannot be annotated with verbal argument structure
+(for whatever reason, including incomprehensibility), it cannot convey the meaning.
+
+HMEANT thus makes a stricter check of whether the translation can be understood
+than we do. HUME may be artificially boosting the perceived understandability of
+the sentence by providing the source.
+
+\subsection{Importance of Alignment}
+
+In HMEANT, the alignment between the (source and candidate) structures is a key
+part of the manual annotation. If the alignment cannot be created, the candidate
+translation is heavily penalized in score.
+Previous experiments with HMEANT
+\parcite{bojar:wu:ssst:2012,chuchunkov-tarelkin-galinskaya:2014:SSST-8} have
+suggested that modifications to the two structures are desirable and appropriate
+at the alignment stage, to avoid unduly penalization due to spurious diverges in
+the structures.
+The same issue was observed by HMEANT authors themselves \parcite{XXXcite}
+\XXX{cite this: http://repository.ust.hk/ir/Record/1783.1-66107}: the
+inter-annotator agreement on SRL dropped from 90\% to 61\% 
+when the two aligned structures were from two different annotators.
+
+HUME uses only automatic (word-level) alignment, and it's only purpose is to
+make the annotation process easier. The alignment is indicative only and the
+user is expected to mentally correct the alignment as needed.
+
+\subsection{Monolingual vs. Bilingual Evaluation}
+HUME diverges from HMEANT (and also from the shallow automatic measures
+like BLEU or
+METEOR) in not requiring a reference.
+Instead, it compares the source directly with the output
+translation.
+
+This requires us to employ bilingual annotators, but has the benefit of avoiding
+the reference translation. It is known that the reference translations are not
+always literal enough and the MT output semantics can be composed differently
+than the semantics of the reference. While this is in principle not a problem
+for HUME (the whole sentence can be labelled as ``Green''), it prevents learning
+any fine-grained information about the MT output.
+
+Perhaps even more importantly,
+the reference translations were often not created by translating the source
+sentences. In some cases, the manual translation was reversed, or both the
+``source'' and ``reference'' are actually independent translations from a third
+language. The ``reference'' can thus easily contain more information than the
+``source''.
+
+\subsection{Technical Differences}
+
++ allows for multi-word and even non-contiguous elements
+
++ allows to indicate the quality of the predicates ("Actions")
+
+\subsection{XXX to be removed}
+
 In the *MEANT papers they mainly test by either
@@
 
-Differences of our approach:
-
-The annotation experiment that we just conducted differs in two main ways from HMEANT:
-1) The semantic framework is different - UCCA vs SRL - and SRL has problems which Omri lists below, some of which Dekai dismissed when they were brought up.
-Maybe these problems do not matter for MT evaluation or tuning, but I think this would have to be determined empirically.
-2) The annotation process is different.
-We semantically annotate the source, then automatically project it to the translation, and ask bilingual annotators to assess whether this projection shows a good or bad translation.
-In HMEANT they semantically annotate both the reference and the translation, then align these manually, and assess to what extent they match.
-
-Advatages of our approach:
-Here is an additional argument to support the need for target-only intelligibility:
-http://www.aclweb.org/anthology/W14-4005
-see Table  3 and the text in Section 5.2.
-The more frames can people annotate in the hypothesis, the better the hypothesis overall.
+%Here is an additional argument to support the need for target-only intelligibility:
+%http://www.aclweb.org/anthology/W14-4005
+%see Table  3 and the text in Section 5.2.
+%The more frames can people annotate in the hypothesis, the better the hypothesis overall.
 

### 2016-03-16 paper/main.tex
ddd

--- 2016-03-16 paper/main.tex
@@
 defined by its UCCA annotation, relative to $t$.
-We note that UCCA annotation is performed once for every source sentence, irrespective
-of the number of its translations we wish to evaluate\footnote{Instead of
-relying on the source,
+We note that UCCA annotation is performed once for every source sentence,
+irrespective of the number of its translations we wish
+to evaluate\footnote{Instead of relying on the source,
 the evaluation could be based on the UCCA structure for the \emph{reference}
@@
 For example, consider the sentence ``After graduation, Tom moved
-to NYC'' as the source $s$, see \figref{fig:ucca_example}. The atomic units are the individual words,
-nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation
-$t$
+to NYC'' as the source $s$, see \figref{fig:ucca_example}.
+The atomic units are the individual words,
+nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation $t$
 is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
@@
 
-We turn to discuss the interface for collecting human judgements on the semantic units of the source,
-relating to the translation. UCCA annotations of the source are done offline, using the UCCA web
-application\footnote{See UCCA webpage.}.
+We turn to discuss the interface for collecting human judgements on the semantic units
+of the source, relating to the translation. UCCA annotations of the source are done offline,
+using the UCCA web application\footnote{See UCCA webpage.}.
 
@@
 \section{Comparison with HMEANT}\label{sec:hmeant_comp}
-
-\subsection{One vs. Two Structures}
-
-\subsection{Verbal Structures Only?}
-
-\subsection{Monolingual vs. Bilingual Evaluation}
-+/- expects to cover the whole hypothesis with annotations, whereas HMEANT 'gives up' when the
-hypothesis is not intelligible, giving a zero score to the bad frame/whole sentence.
-This can be a good decision for HMEANT, since our way may artificially boost the chances of
-understanding the sentence.
-
-\subsection{Technical Differences}
-
-+ allows for multi-word and even non-contiguous elements
-
-+ allows to indicate the quality of the predicates ("Actions")
-
-\subsection{XXX to be moved}
 
@@
 to preserve for MT evaluation, and which are less important.
-
-\subsection{from e-mails}
-
-UCCA-Eval in contrast to HMEANT:
-
-- needs *very* trained UCCA annotators
-- we were not very quick at explaining even the Eval annotations, but I guess we could have done better
-+ relies only on one structure
-+ relies on automatic alignment (in the form of hints only)
-These two positive design decisions are in line with what Dekai observed for HMEANT
-(see the abstract of http://repository.ust.hk/ir/Record/1783.1-66107): IAA of SRL drops from 90
-to 61 when the two aligned structures are from two different annotators.
-
-Actually, I think this is a +/- .
-UCCA-Eval only needs the semantic evaluation done once for a given test source, whereas for the HMEANT set-up you have to keep annotating MT output.
-That's a + for UCCA-Eval.
-However we need bilingual annotators, which should be a -.
-Of course we could align translation and reference using monolingual annotators, but then we have the problem of alignment.
-In the *MEANT papers they mainly test by either
-A) correlating the metric to ranking judgements; or
-B) using automated versions of the metric for tuning MT systems
-
-I think neither is particularly convincing, but I'm not sure at the moment what we replace them with.
-Dekai also made a claim about the *MEANT metrics being "interpretable" in the sense that they could guide you towards semantic errors that the MT system may be making, however I don't recall seeing this done in his papers.
-
-By the claim that HMEANT explains things, Dekai probably means that the score is decomposable, but so is UCCA-Eval and even BLEU.
-And UCCA-Eval has one benefit over HMEANT: it points to words in the MT output that are aligned with something but do not convey the meaning well (and we know which part of the source meaning).
-Pure HMEANT might say only that some part of the MT output has no clear frames whatsover (and we do not know which part of the source/reference, unless we add the word alignments).
-
-Differences of our approach:
-
-The annotation experiment that we just conducted differs in two main ways from HMEANT:
-1) The semantic framework is different - UCCA vs SRL - and SRL has problems which Omri lists below, some of which Dekai dismissed when they were brought up.
-Maybe these problems do not matter for MT evaluation or tuning, but I think this would have to be determined empirically.
-2) The annotation process is different.
-We semantically annotate the source, then automatically project it to the translation, and ask bilingual annotators to assess whether this projection shows a good or bad translation.
-In HMEANT they semantically annotate both the reference and the translation, then align these manually, and assess to what extent they match.
-
-Advatages of our approach:
-Here is an additional argument to support the need for target-only intelligibility:
-http://www.aclweb.org/anthology/W14-4005
-see Table  3 and the text in Section 5.2.
-The more frames can people annotate in the hypothesis, the better the hypothesis overall.
 
@@
 
+\subsection{Copied from emails}
+
+UCCA-Eval in contrast to HMEANT:
+
+- needs *very* trained UCCA annotators
+- we were not very quick at explaining even the Eval annotations, but I guess we could have done better
++ relies only on one structure
++ relies on automatic alignment (in the form of hints only)
+These two positive design decisions are in line with what Dekai observed for HMEANT
+(see the abstract of http://repository.ust.hk/ir/Record/1783.1-66107): IAA of SRL drops from 90
+to 61 when the two aligned structures are from two different annotators.
+
++ allows for multi-word and even non-contiguous elements
++ allows to indicate the quality of the predicates ("Actions")
++/- expects to cover the whole hypothesis with annotations, whereas HMEANT 'gives up' when the
+hypothesis is not intelligible, giving a zero score to the bad frame/whole sentence.
+This can be a good decision for HMEANT, since our way may artificially boost the chances of
+understanding the sentence.
+
+HMEAN disadvantages:
+.
+Importantly: it disregards copula clauses, nominalizations (which are often translated to verbal clauses), inter-clause relations (discourse markers, complement clauses etc.).
+UCCA-Eval addresses these.
+
+Actually, I think this is a +/- .
+UCCA-Eval only needs the semantic evaluation done once for a given test source, whereas for the HMEANT set-up you have to keep annotating MT output.
+That's a + for UCCA-Eval.
+However we need bilingual annotators, which should be a -.
+Of course we could align translation and reference using monolingual annotators, but then we have the problem of alignment.
+In the *MEANT papers they mainly test by either
+A) correlating the metric to ranking judgements; or
+B) using automated versions of the metric for tuning MT systems
+
+I think neither is particularly convincing, but I'm not sure at the moment what we replace them with.
+Dekai also made a claim about the *MEANT metrics being "interpretable" in the sense that they could guide you towards semantic errors that the MT system may be making, however I don't recall seeing this done in his papers.
+
+By the claim that HMEANT explains things, Dekai probably means that the score is decomposable, but so is UCCA-Eval and even BLEU.
+And UCCA-Eval has one benefit over HMEANT: it points to words in the MT output that are aligned with something but do not convey the meaning well (and we know which part of the source meaning).
+Pure HMEANT might say only that some part of the MT output has no clear frames whatsover (and we do not know which part of the source/reference, unless we add the word alignments).
+
+Differences of our approach:
+
+The annotation experiment that we just conducted differs in two main ways from HMEANT:
+1) The semantic framework is different - UCCA vs SRL - and SRL has problems which Omri lists below, some of which Dekai dismissed when they were brought up.
+Maybe these problems do not matter for MT evaluation or tuning, but I think this would have to be determined empirically.
+2) The annotation process is different.
+We semantically annotate the source, then automatically project it to the translation, and ask bilingual annotators to assess whether this projection shows a good or bad translation.
+In HMEANT they semantically annotate both the reference and the translation, then align these manually, and assess to what extent they match.
+
+Advatages of our approach:
+Here is an additional argument to support the need for target-only intelligibility:
+http://www.aclweb.org/anthology/W14-4005
+see Table  3 and the text in Section 5.2.
+The more frames can people annotate in the hypothesis, the better the hypothesis overall.
+
 \section{Conclusion}

### 2016-03-16 paper/main.tex
ddd

@@
 \caption{\label{fig:ucca_example}
+  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
+indices.}
+\end{figure}
+
+\begin{figure}
+
+\begin{tikzpicture}[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
+      inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
+  \node (ROOT) [label={G}] [fill=green, circle] {}
+  child {node [label={R}] {Golf} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {became} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {a} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {passion} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {for} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {his} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {oldest} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {daughter} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {,} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {she} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {attended} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {daily} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {lesssons} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {and} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {greatly} edge from parent node[left]  {$L$}}
+  child {node [label={\#1}] {improved} edge from parent node[left]  {$L$}};
+
+  \com{
+  child {node (H1) [label={\#8}] [fill=black,circle] {}
+    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    }
+    child {node (H2) [label={\#9}] [fill=black,circle] {}
+      child {node  {} edge from parent[draw=none]}
+      child {node  {} edge from parent[draw=none]}
+      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
+      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
+      child {node [label={\#10}] [fill=black,circle] {}
+        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
+        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
+        edge from parent node[right] {$A$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    };
+    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+    }
+  \end{tikzpicture}
+\caption{\label{fig:ucca_example2}
   Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node

### 2016-03-16 paper/main.tex
ddd
figure

--- 2016-03-16 paper/main.tex
@@
 \end{figure}
+
+\begin{figure*}
+
+  \begin{tikzpicture}
+    [level 1/.style={sibling distance=7.5em},
+      level 2/.style={sibling distance=5em},
+      level 3/.style={sibling distance=5em},
+      level 4/.style={sibling distance=3em}]
+    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
+    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
+  \node (ROOT) [label={B}] [fill=blue, circle] {}
+  child {node (H3) [label={A}] [fill=yellow,circle] {}
+    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
+    child {node (H1) [label={O}] [fill=orange,circle] {}
+      child {node [label={}] {became} edge from parent node[left]  {}}
+      child {node [label={}] {a} edge from parent node[left]  {}}
+      child {node [label={}] {passion} edge from parent node[left]  {}}
+    }
+    child {node (H2) [label={A}] [fill=yellow,circle] {}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node [label={}] {for} edge from parent node[left]  {}}
+      child {node [label={}] {his} edge from parent node[left]  {}}
+      child {node [label={}] {oldest} edge from parent node[left]  {}}
+      child {node [label={}] {daughter} edge from parent node[left]  {}}
+    }
+  }
+  child {node {} edge from parent[draw=none] {}}
+  child {node (H4) [label={A}] [fill=yellow,circle] {}
+    child {node [label={}] {she} edge from parent node[left]  {}}
+      child {node [label={}] {attended} edge from parent node[left]  {}}
+      child {node [label={}] {daily} edge from parent node[left]  {}}
+      child {node [label={}] {lesssons} edge from parent node[left]  {}}
+  }
+  child {node [label={}] {and} edge from parent node[left]  {}}
+  child {node (H5) [label={A}] [fill=yellow,circle] {}  
+    child {node [label={}] {greatly} edge from parent node[left] {}}
+    child {node [label={}] {improved} edge from parent node[left] {}}
+  };
+
+  \com{
+  child {node (H1) [label={\#8}] [fill=black,circle] {}
+    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    }
+    child {node (H2) [label={\#9}] [fill=black,circle] {}
+      child {node  {} edge from parent[draw=none]}
+      child {node  {} edge from parent[draw=none]}
+      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
+      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
+      child {node [label={\#10}] [fill=black,circle] {}
+        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
+        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
+        edge from parent node[right] {$A$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    };
+    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+    }
+  \end{tikzpicture}
+\caption{\label{fig:ucca_example2}
+  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
+indices.}
+\end{figure*}
 

### 2016-03-16 paper/main.tex
figure

--- 2016-03-16 paper/main.tex
@@
 \end{figure}
+
+\begin{figure*}
+
+  \begin{tikzpicture}
+    [level 1/.style={sibling distance=7.5em},
+      level 2/.style={sibling distance=5em},
+      level 3/.style={sibling distance=5em},
+      level 4/.style={sibling distance=3em}]
+    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
+    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
+  \node (ROOT) [label={B}] [fill=blue, circle] {}
+  child {node (H3) [label={A}] [fill=yellow,circle] {}
+    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
+    child {node (H1) [label={[orange]:O}] [fill=orange,circle] {}
+      child {node [label={}] {became} edge from parent node[left]  {}}
+      child {node [label={}] {a} edge from parent node[left]  {}}
+      child {node [label={}] {passion} edge from parent node[left]  {}}
+    }
+    child {node (H2) [label={A}] [fill=yellow,circle] {}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
+      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
+      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
+      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
+    }
+  }
+  child {node {} edge from parent[draw=none] {}}
+  child {node (H4) [label={A}] [fill=yellow,circle] {}
+    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
+    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
+    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
+    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
+  }
+  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
+  child {node (H5) [label={A}] [fill=yellow,circle] {}  
+    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
+    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
+  };
+  \draw[dashed] (H5) -> (she) ;
+
+  \com{
+  child {node (H1) [label={\#8}] [fill=black,circle] {}
+    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    }
+    child {node (H2) [label={\#9}] [fill=black,circle] {}
+      child {node  {} edge from parent[draw=none]}
+      child {node  {} edge from parent[draw=none]}
+      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
+      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
+      child {node [label={\#10}] [fill=black,circle] {}
+        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
+        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
+        edge from parent node[right] {$A$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    };
+    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+    }
+  \end{tikzpicture}
+\caption{\label{fig:ucca_example2}
+  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
+indices.}
+\end{figure*}
 

### 2016-03-16 paper/main.tex

@@
 measure that decomposes over semantic units, rather than strings. On the other
-hand, HMEANT focuses on verbal
+hand, there are important differences in design decisions, which we summarize in this
+section.
+
+First of all, the annotation process is different. We semantically annotate the
+source, then automatically project it to the translation, and ask bilingual
+annotators to assess whether this projection shows a good or bad translation. In
+HMEANT, they semantically annotate both the reference and the translation, then
+align these manually, and assess to what extent they match.
+
+\subsection{Verbal Structures Only?}
+
+HMEANT focuses on verbal
 argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
-relations and copula clauses.
-HUME further diverges from both HMEANT (as well as the shallow automatic BLEU and
-METEOR measures) in not requiring a reference.
-Instead, it compares the source directly with the output
-translation. We see this difference as crucial where semantic analysis is concerned, as it
-is often impossible to assign a semantic structure to a low quality MT output sentence.
-
-%See \secref{sec:hmeant_comp} for a more elaborate discussion of the differences
-%between HUME
-%and HMEANT.
-% This is here!
-
-For instance, consider the sentence ``A coronary angioplasty may not be technically possible'',
-which was translated by one of the systems we experimented with as
-``Eine koronare Angioplastie kann nicht technisch m{\"o}glich'', i.e., the main verb ``sein'' (``be'') is omitted.
-While this may be interpreted is a minor error, HMEANT will assign it a very low score, as it failed to
-translate the main verb. We understand that this is a design feature of HMEANT:
-if the translation per se cannot be annotated with verbal argument structure
-(for whatever reason, including incomprehensibility), it cannot convey the meaning.
-HMEANT thus makes a stricter check of whether the translation can be understood
-than we do.
-
-Several experiments with HMEANT applied to languages other than English
-\parcite{bojar:wu:ssst:2012,birch-EtAl:2013:WMT,chuchunkov-tarelkin-galinskaya:2014:SSST-8}
-indicated that HMEANT does not address some important phenomena.
+relations and copula clauses. Some of the uncovered cases were already observed
+in previous applications of HMEANT to languages other than English (see \perscite{birch-EtAl:2013:WMT} for German, \perscite{bojar:wu:ssst:2012} for
+Czech and by \perscite{chuchunkov-tarelkin-galinskaya:2014:SSST-8} for Russian)
+but the problems are apparent in English as well.
+
 We conducted an analysis of the English UCCA Wikipedia
@@
 to preserve for MT evaluation, and which are less important.
+
+\subsection{Cost of Structural Annotation}
+
+HMEANT structural annotation (semantic role labelling, SRL), similar to e.g.
+PropBank
+\parcite{palmer2010semantic}, is designed to be very simple and easy to explain
+to lay annotators.
+HUME relies on UCCA, which is arguably more abstract, harder to
+explain, and thus more expensive to obtain.
+
+On the other hand, HUME 
+needs the semantic annotation
+done once for a given source sentence, whereas for the HMEANT set-up, every new
+MT output has to be structurally annotated.
+
+\subsection{One vs. Two Structures}
+
+HUME uses just one structural annotation (for the source) while HMEANT relies on
+two independently constructed structural annotations (for both the reference and
+the candidate translation).
+
+We see this difference as crucial where semantic analysis is concerned, as it
+is often impossible to assign a semantic structure to a low quality MT output sentence.
+
+For instance, consider the sentence ``A coronary angioplasty may not be technically possible'',
+which was translated by one of the systems we experimented with as
+``Eine koronare Angioplastie kann nicht technisch m{\"o}glich'', i.e., the main verb ``sein'' (``be'') is omitted.
+While this may be interpreted is a minor error, HMEANT will assign it a very low score, as it failed to
+translate the main verb.
+
+We understand that this is a design feature of HMEANT:
+if the translation per se cannot be annotated with verbal argument structure
+(for whatever reason, including incomprehensibility), it cannot convey the meaning.
+
+HMEANT thus makes a stricter check of whether the translation can be understood
+than we do. HUME may be artificially boosting the perceived understandability of
+the sentence by providing the source.
+
+\subsection{Importance of Alignment}
+
+In HMEANT, the alignment between the (source and candidate) structures is a key
+part of the manual annotation. If the alignment cannot be created, the candidate
+translation is heavily penalized in score.
+Previous experiments with HMEANT
+\parcite{bojar:wu:ssst:2012,chuchunkov-tarelkin-galinskaya:2014:SSST-8} have
+suggested that modifications to the two structures are desirable and appropriate
+at the alignment stage, to avoid unduly penalization due to spurious diverges in
+the structures.
+The same issue was observed by HMEANT authors themselves \parcite{XXXcite}
+\XXX{cite this: http://repository.ust.hk/ir/Record/1783.1-66107}: the
+inter-annotator agreement on SRL dropped from 90\% to 61\% 
+when the two aligned structures were from two different annotators.
+
+HUME uses only automatic (word-level) alignment, and it's only purpose is to
+make the annotation process easier. The alignment is indicative only and the
+user is expected to mentally correct the alignment as needed.
+
+\subsection{Monolingual vs. Bilingual Evaluation}
+HUME diverges from HMEANT (and also from the shallow automatic measures
+like BLEU or
+METEOR) in not requiring a reference.
+Instead, it compares the source directly with the output
+translation.
+
+This requires us to employ bilingual annotators, but has the benefit of avoiding
+the reference translation. It is known that the reference translations are not
+always literal enough and the MT output semantics can be composed differently
+than the semantics of the reference. While this is in principle not a problem
+for HUME (the whole sentence can be labelled as ``Green''), it prevents learning
+any fine-grained information about the MT output.
+
+Perhaps even more importantly,
+the reference translations were often not created by translating the source
+sentences. In some cases, the manual translation was reversed, or both the
+``source'' and ``reference'' are actually independent translations from a third
+language. The ``reference'' can thus easily contain more information than the
+``source''.
+
+\subsection{Technical Differences}
+
++ allows for multi-word and even non-contiguous elements
+
++ allows to indicate the quality of the predicates ("Actions")
+
+\subsection{XXX to be removed}
+
+In the *MEANT papers they mainly test by either
+A) correlating the metric to ranking judgements; or
+B) using automated versions of the metric for tuning MT systems
+
+I think neither is particularly convincing, but I'm not sure at the moment what we replace them with.
+Dekai also made a claim about the *MEANT metrics being "interpretable" in the sense that they could guide you towards semantic errors that the MT system may be making, however I don't recall seeing this done in his papers.
+
+By the claim that HMEANT explains things, Dekai probably means that the score is decomposable, but so is UCCA-Eval and even BLEU.
+And UCCA-Eval has one benefit over HMEANT: it points to words in the MT output that are aligned with something but do not convey the meaning well (and we know which part of the source meaning).
+Pure HMEANT might say only that some part of the MT output has no clear frames whatsover (and we do not know which part of the source/reference, unless we add the word alignments).
+
+%Here is an additional argument to support the need for target-only intelligibility:
+%http://www.aclweb.org/anthology/W14-4005
+%see Table  3 and the text in Section 5.2.
+%The more frames can people annotate in the hypothesis, the better the hypothesis overall.
 

### 2016-03-17 paper/main.tex
fixed diagram

@@
 \usepackage{pgfplotstable}
-\usepackage{hyperref}
+%\usepackage{hyperref}
 \usepackage{color}
@@
     inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
-  \node (ROOT) [label={\#7}] [fill=black, circle] {}
-  child {node [label={\#1}] {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
+  \node (ROOT) [fill=black, circle] {}
+  child {node {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
+  child {node (H1)  [fill=black,circle] {}
+    child {node {graduation} edge from parent node[right] {$P$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    }
+    child {node (H2)  [fill=black,circle] {}
+      child {node  {} edge from parent[draw=none]}
+      child {node  {} edge from parent[draw=none]}
+      child {node  (Tom) {Tom} edge from parent node[right] {$A$}}
+      child {node  {moved} edge from parent node[right] {$P$}}
+      child {node  [fill=black,circle] {}
+        child {node  {to} edge from parent node[right] {$R$}}
+        child {node  {NYC} edge from parent node[right] {$C$}}
+        edge from parent node[right] {$A$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    };
+    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+  \end{tikzpicture}
+\caption{\label{fig:ucca_example}
+  Sample UCCA annotation. Edge labels mark UCCA categories. }
+\end{figure}
+
+\begin{figure*}
+  \begin{adjustbox}{trim=2.3cm 0 0 0}
+  \begin{tikzpicture}
+    [level 1/.style={sibling distance=7.5em},
+      level 2/.style={sibling distance=5em},
+      level 3/.style={sibling distance=5em},
+      level 4/.style={sibling distance=3em}]
+    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
+    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
+  \node (ROOT) [label={B}] [fill=blue, circle] {}
+  child {node (H3) [label={A}] [fill=yellow,circle] {}
+    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
+    child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
+      child {node [label={}] {became} edge from parent node[left]  {}}
+      child {node [label={}] {a} edge from parent node[left]  {}}
+      child {node [label={}] {passion} edge from parent node[left]  {}}
+    }
+    child {node (H2) [label={A}] [fill=yellow,circle] {}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
+      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
+      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
+      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
+    }
+  }
+  child {node {} edge from parent[draw=none] {}}
+  child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
+    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
+    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
+    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
+    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
+  }
+  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
+  child {node (H5) [label={A}] [fill=yellow,circle] {}  
+    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
+    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
+  };
+  \draw[dashed] (H5) -> (she) ;
+
+  \com{
   child {node (H1) [label={\#8}] [fill=black,circle] {}
@@
     \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+    }
   \end{tikzpicture}
-\caption{\label{fig:ucca_example}
-  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
-indices.}
-\end{figure}
-
-\begin{figure*}
-
-  \begin{tikzpicture}
-    [level 1/.style={sibling distance=7.5em},
-      level 2/.style={sibling distance=5em},
-      level 3/.style={sibling distance=5em},
-      level 4/.style={sibling distance=3em}]
-    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
-    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
-  \node (ROOT) [label={B}] [fill=blue, circle] {}
-  child {node (H3) [label={A}] [fill=yellow,circle] {}
-    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
-    child {node (H1) [label={[orange]:O}] [fill=orange,circle] {}
-      child {node [label={}] {became} edge from parent node[left]  {}}
-      child {node [label={}] {a} edge from parent node[left]  {}}
-      child {node [label={}] {passion} edge from parent node[left]  {}}
-    }
-    child {node (H2) [label={A}] [fill=yellow,circle] {}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
-      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
-      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
-      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
-    }
-  }
-  child {node {} edge from parent[draw=none] {}}
-  child {node (H4) [label={A}] [fill=yellow,circle] {}
-    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
-  }
-  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
-  child {node (H5) [label={A}] [fill=yellow,circle] {}  
-    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
-    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
-  };
-  \draw[dashed] (H5) -> (she) ;
-
-  \com{
-  child {node (H1) [label={\#8}] [fill=black,circle] {}
-    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    }
-    child {node (H2) [label={\#9}] [fill=black,circle] {}
-      child {node  {} edge from parent[draw=none]}
-      child {node  {} edge from parent[draw=none]}
-      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
-      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
-      child {node [label={\#10}] [fill=black,circle] {}
-        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
-        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
-        edge from parent node[right] {$A$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    };
-    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
-    }
-  \end{tikzpicture}
+  \end{adjustbox}
 \caption{\label{fig:ucca_example2}
@@
 The procedure involves two manual steps: (1) UCCA-annotating $s$, 
-(2) human judgements as to the translation quality of each semantic unit of $s$,
-defined by its UCCA annotation, relative to $t$.
+(2) human judgements as to the translation quality of each semantic unit of $s$ relative to $t$,
+where units are defined according to the UCCA annotation.
 We note that UCCA annotation is performed once for every source sentence,
@@
 \begin{figure*}[t]
-    \includegraphics[width=1\textwidth]{ucca-tree-mteval}
+    \includegraphics[width=1\textwidth]{ucca-tree-mteval.jpg}
     \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
@@
 Human annotation of the semantic units requires first deciding whether
-a unit is {\it atomic}, i.e., that it has no meaning-bearing
-sub-units, or {\it structural}. In most cases, atomic units 
+a unit is {\it structural}, i.e., has meaning-bearing sub-units,
+or {\it atomic}. In most cases, atomic units
 correspond to individual words, but they may also correspond to unanalyzable
-multi-word expressions, i.e., expressions that do not have meaningful sub-units.
+multi-word expressions, i.e., expressions that do not have meaningful sub-units
+(e.g., idioms like ``takes two to tango'').
 When a multi-word unit is labeled as atomic, all its sub-units' annotations are discarded
@@
 Structural units have sub-units (children in the UCCA graph), which may either be
-atomic or structural, and represent the relation between the sub-units.
+atomic or structural. The labels assigned to such nodes reflect the extent to which
+the relation between the sub-units.
 For instance, a Scene unit represents the relation between the
@@
 
-For example, consider the sentence ``After graduation, Tom moved
-to NYC'' as the source $s$, see \figref{fig:ucca_example}.
-The atomic units are the individual words,
-nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation $t$
-is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
-  be in English for presentation purposes. In our experiments, $t$ is in German,
-  Czech, Polish or Romanian.}.
-Then ``after'', ``graduation'', ``Tom'' and ``to'' are marked Green, ``moved'' is marked
-Orange (as it corresponds to ``will move'' in $t$), and ``NYC''
-is marked as Red, as it was mistranslated to ``Paris''. Node \#9 is marked Adequate,
-as the Scene of Tom moving is properly reflected in the translation. Node \#10, which
-expresses the relation between the graduation and moving Scenes, is marked Bad,
-as their relation in $t$ is reversed.
+%For example, consider the sentence ``After graduation, Tom moved
+%to NYC'' as the source $s$, see \figref{fig:ucca_example}.
+%The atomic units are the individual words,
+%nodes \#1--6 and the structural nodes are \#7--10. Assume that the translation $t$
+%is ``After Tom will move to Paris, graduation''\footnote{We take $t$ to
+%  be in English for presentation purposes. In our experiments, $t$ is in German,
+%  Czech, Polish or Romanian.}.
+%Then ``after'', ``graduation'', ``Tom'' and ``to'' are marked Green, ``moved'' is marked
+%Orange (as it corresponds to ``will move'' in $t$), and ``NYC''
+%is marked as Red, as it was mistranslated to ``Paris''. Node \#9 is marked Adequate,
+%as the Scene of Tom moving is properly reflected in the translation. Node \#10, which
+%expresses the relation between the graduation and moving Scenes, is marked Bad,
+%as their relation in $t$ is reversed.
 
@@
 in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
-individually.\oa{I'm not sure I agree with the UCCA annotation;
-also, translation annotation for ``she'' is missing. and for ``very good'' is wrong}
+individually.
 
@@
 
-\figref{fig:interface} presents the user interface of HUME. The sentence at the top
-shows the complete MT system output. The source sentence appears at the top, and
-the source translation directly below it. Underneath the source sentence appears
+\figref{fig:interface} presents the user interface of HUME. 
+The translation appears at the top, and
+the source sentence directly below it. Underneath the source sentence appears
 its list of semantic units. For each semantic unit, including the one for the
@@
 
-The interface presents, for each unit, the segment of the translation which aligns to it.
+The interface presents, for each unit, the translation segment aligned with it.
 This allows the user, especially in long sentences, to focus her attention on the parts
-mostly likely to be relevant for her judgement. As the alignments are automatically derived,
+most likely to be relevant for her judgement. As the alignments are automatically derived,
 and therefore noisy, the annotator is instructed to treat the aligned text is a cue, but to ignore
@@
 $t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
-In the event that $a(u)$ is discontiguous in the text, we mark words between the left
+In the event that $a(u)$ is discontiguous in the text, words between the left
 and right boundaries of $a(u)$ which are not contained in it (intervening words)
-in a smaller red font. Intervening words are likely change the meaning of the translation
+are presented in a smaller red font.
+Intervening words are likely change the meaning of the translation
 of $u$, and thus should be paid attention to when considering whether the translation
@@
 \begin{figure*}[t]
-    \includegraphics[width=1\textwidth]{ucca-tree}
+    \includegraphics[width=1\textwidth]{ucca-tree.jpg}
     \caption{UCCA Tree with scenes}

### 2016-03-17 paper/main.tex
misc

@@
 to localize the errors or at least unconfirmed parts within the translation.
-The most commonly used measures decompose over n-grams or even individual words, e.g., 
+The most commonly used measures decompose over n-grams or individual words, e.g., 
 the BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR measures \cite{Banerjee:2005}.
@@
 the similarity of the reference and translation in terms
-of their {\it semantic structure}. \newcite{liu2005syntactic} took a syntactic approach, 
+of their structure. \newcite{liu2005syntactic} took a syntactic approach, 
 quantifying the similarity between the syntactic structures of the reference and translation.
@@
 \caption{\label{fig:ucca_example}
-  Sample UCCA annotation. Edge labels mark UCCA categories. }
+  Sample UCCA annotation of the sentence ``After graduation, Tom moved
+  to NYC''. Leaves correspond to words, whereas nodes correspond to units.
+  Edge labels mark UCCA categories, which are largely ignored in this work.}
 \end{figure}
-
-\begin{figure*}
-  \begin{adjustbox}{trim=2.3cm 0 0 0}
-  \begin{tikzpicture}
-    [level 1/.style={sibling distance=7.5em},
-      level 2/.style={sibling distance=5em},
-      level 3/.style={sibling distance=5em},
-      level 4/.style={sibling distance=3em}]
-    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
-    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
-  \node (ROOT) [label={B}] [fill=blue, circle] {}
-  child {node (H3) [label={A}] [fill=yellow,circle] {}
-    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
-    child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
-      child {node [label={}] {became} edge from parent node[left]  {}}
-      child {node [label={}] {a} edge from parent node[left]  {}}
-      child {node [label={}] {passion} edge from parent node[left]  {}}
-    }
-    child {node (H2) [label={A}] [fill=yellow,circle] {}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
-      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
-      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
-      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
-    }
-  }
-  child {node {} edge from parent[draw=none] {}}
-  child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
-    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
-  }
-  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
-  child {node (H5) [label={A}] [fill=yellow,circle] {}  
-    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
-    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
-  };
-  \draw[dashed] (H5) -> (she) ;
-
-  \com{
-  child {node (H1) [label={\#8}] [fill=black,circle] {}
-    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    }
-    child {node (H2) [label={\#9}] [fill=black,circle] {}
-      child {node  {} edge from parent[draw=none]}
-      child {node  {} edge from parent[draw=none]}
-      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
-      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
-      child {node [label={\#10}] [fill=black,circle] {}
-        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
-        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
-        edge from parent node[right] {$A$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    };
-    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
-    }
-  \end{tikzpicture}
-  \end{adjustbox}
-\caption{\label{fig:ucca_example2}
-  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
-indices.}
-\end{figure*}
 
@@
 
-UCCA's current inventory of distinctions focuses on annotating argument structures
+UCCA's current inventory of distinctions focuses on argument structures
 (adjectival, nominal, verbal and others) and relations between them.
-The most basic notion is the {\sc scene}, which describes a movement, an
+The most basic notion is the Scene, which describes a movement, an
 action or a state which is persistent in time. Each Scene contains one main relation, as well
@@
 The participant ``John'' is a part of both Scenes, while ``NYC'' only of the
-latter. Further categories account for
+latter (\figref{fig:ucca_example}). Further categories account for
 relations between Scenes and the internal structures of complex participants and relations.
-See \figref{fig:ucca_example}.
 
@@
 argument structure, and are thus less suitable as a basis for a semantic measure.
-For example, ``after he graduated, Tom moved to NYC''
-conveys a similar meaning to ``after graduation, Tom moved to NYC''.
+E.g., ``after he graduated, Tom moved to NYC'' contains two verbs,
+but is similar in meaning to the one-verbed
+``after graduation, Tom moved to NYC''.
 
@@
 Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.
-
-Other approaches have been proposed for semantic representation, which are grounded in the words
-and phrases of the text, and have generally represented semantic structures as bi-lexical dependencies.
+Other approaches pursued a grounded representation of semantics,
+and have generally represented semantic structures as bi-lexical dependencies.
 Examples include the tectogrammatical layer \parcite{sgallhp:1986} used in the Prague dependency
@@
 
-\begin{figure*}[t]
-    \includegraphics[width=1\textwidth]{ucca-tree-mteval.jpg}
-    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
-	English paraphrase).}
-    \label{fig:ucca-tree-hume}
+%\begin{figure*}[t]
+%    \includegraphics[width=1\textwidth]{ucca-tree-mteval.jpg}
+%    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
+%	English paraphrase).}
+%    \label{fig:ucca-tree-hume}
+%\end{figure*}
+
+\begin{figure*}
+  \begin{adjustbox}{trim=2.3cm 0 0 0}
+  \begin{tikzpicture}
+    [level 1/.style={sibling distance=7.5em},
+      level 2/.style={sibling distance=5em},
+      level 3/.style={sibling distance=5em},
+      level 4/.style={sibling distance=3em}]
+    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
+    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
+  \node (ROOT) [label={B}] [fill=blue, circle] {}
+  child {node (H3) [label={A}] [fill=yellow,circle] {}
+    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
+    child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
+      child {node [label={}] {became} edge from parent node[left]  {}}
+      child {node [label={}] {a} edge from parent node[left]  {}}
+      child {node [label={}] {passion} edge from parent node[left]  {}}
+    }
+    child {node (H2) [label={A}] [fill=yellow,circle] {}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
+      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
+      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
+      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
+    }
+  }
+  child {node {} edge from parent[draw=none] {}}
+  child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
+    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
+    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
+    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
+    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
+  }
+  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
+  child {node (H5) [label={A}] [fill=yellow,circle] {}  
+    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
+    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
+  };
+  \draw[dashed] (H5) -> (she) ;
+
+  \com{
+  child {node (H1) [label={\#8}] [fill=black,circle] {}
+    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    }
+    child {node (H2) [label={\#9}] [fill=black,circle] {}
+      child {node  {} edge from parent[draw=none]}
+      child {node  {} edge from parent[draw=none]}
+      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
+      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
+      child {node [label={\#10}] [fill=black,circle] {}
+        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
+        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
+        edge from parent node[right] {$A$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    };
+    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+    }
+  \end{tikzpicture}
+  \\
+  \end{adjustbox}
+  \adjustbox{width=\textwidth,center,margin=.5em}{\it His old daughter loved tennis, and took daily lessons however good he was}
+
+  \caption{\label{fig:ucca_example2}
+    Sample UCCA annotation of a source sentence, translation (below in italics), 
+    and translation judgements over the source units, marked with A,B,R,O,G labels. Yellow/blue marks
+    nodes annotated with A/B, respectively.
+    In our experiments, the translation is in German, Romaninan, Polish or Czech.
+    ``became a passion'' is marked as atomic, as it is a multi-word expression.
+    The dashed line indicates that ``she'' is also a child of the unit ``greatly improved''.
+    UCCA categories, and alignment between source and translation are omitted for brevity.}
+
 \end{figure*}

### 2016-03-17 paper/main.tex
misc
figure

@@
 For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
-The semantic units for a sentence $s$ induced by the annotation $G$,
-is the set of yields of nodes in $V$. %We denote it with $Units(s)$ for brevity.
+The semantic units for a sentence $s$ according to $G$
+is the set of yields of nodes in $V$. 
 We note that the translation is not UCCA-annotated.
@@
 \paragraph{Translation Evaluation.}
-The basic strategy for computing HUME is to traverse
-the semantic units of the source sentence, and to mark the extent to which
-they have been correctly translated. As semantic units correspond to the
-semantic relations expressed in the text and the inter-relations between them,
-HUME effectively aggregates the judgements of the users into a composite score.
+The basic strategy for computing HUME is to traverse the semantic units
+of the source sentence, which correspond to the different arguments and relations expressed
+in the text, and to mark the extent to which they have been correctly translated.
+HUME aggregates the judgements of the users into a composite score, 
+which reflects the overall extent to which the semantic content of $s$ is preserved in $t$.
 

### 2016-03-17 paper/main.tex
more about iaa

--- 2016-03-17 paper/main.tex
@@
 \newcommand\BibTeX{B{\sc ib}\TeX}
+
+\listfiles
 
@@
   \begin{adjustbox}{trim=2.3cm 0 0 0}
-  \begin{tikzpicture}
-    [level 1/.style={sibling distance=7.5em},
-      level 2/.style={sibling distance=5em},
-      level 3/.style={sibling distance=5em},
-      level 4/.style={sibling distance=3em}]
-    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
-    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
-  \node (ROOT) [label={B}] [fill=blue, circle] {}
-  child {node (H3) [label={A}] [fill=yellow,circle] {}
-    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
-    child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
-      child {node [label={}] {became} edge from parent node[left]  {}}
-      child {node [label={}] {a} edge from parent node[left]  {}}
-      child {node [label={}] {passion} edge from parent node[left]  {}}
-    }
-    child {node (H2) [label={A}] [fill=yellow,circle] {}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
-      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
-      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
-      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
-    }
-  }
-  child {node {} edge from parent[draw=none] {}}
-  child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
-    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
-  }
-  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
-  child {node (H5) [label={A}] [fill=yellow,circle] {}  
-    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
-    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
-  };
-  \draw[dashed] (H5) -> (she) ;
-
-  \com{
-  child {node (H1) [label={\#8}] [fill=black,circle] {}
-    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    }
-    child {node (H2) [label={\#9}] [fill=black,circle] {}
-      child {node  {} edge from parent[draw=none]}
-      child {node  {} edge from parent[draw=none]}
-      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
-      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
-      child {node [label={\#10}] [fill=black,circle] {}
-        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
-        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
-        edge from parent node[right] {$A$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    };
-    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
-    }
-  \end{tikzpicture}
+%BH: Not compiling for me, commented out
+%  \begin{tikzpicture}
+%    [level 1/.style={sibling distance=7.5em},
+%      level 2/.style={sibling distance=5em},
+%      level 3/.style={sibling distance=5em},
+%      level 4/.style={sibling distance=3em}]
+%    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
+%    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
+%  \node (ROOT) [label={B}] [fill=blue, circle] {}
+%  child {node (H3) [label={A}] [fill=yellow,circle] {}
+%    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
+%    child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
+%      child {node [label={}] {became} edge from parent node[left]  {}}
+%      child {node [label={}] {a} edge from parent node[left]  {}}
+%      child {node [label={}] {passion} edge from parent node[left]  {}}
+%    }
+%    child {node (H2) [label={A}] [fill=yellow,circle] {}
+%      child {node {} edge from parent[draw=none] {}}
+%      child {node {} edge from parent[draw=none] {}}
+%      child {node {} edge from parent[draw=none] {}}
+%      child {node {} edge from parent[draw=none] {}}
+%      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
+%      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
+%      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
+%      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
+%    }
+%  }
+%  child {node {} edge from parent[draw=none] {}}
+%  child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
+%    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
+%    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
+%    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
+%    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
+%  }
+%  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
+%  child {node (H5) [label={A}] [fill=yellow,circle] {}  
+%    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
+%    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
+%  };
+%  \draw[dashed] (H5) -> (she) ;
+%
+%  \com{
+%  child {node (H1) [label={\#8}] [fill=black,circle] {}
+%    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
+%      %child {node {to Paris} edge from parent node[right] {$A$}}
+%      edge from parent node[left] {$H$}
+%    }
+%    child {node (H2) [label={\#9}] [fill=black,circle] {}
+%      child {node  {} edge from parent[draw=none]}
+%      child {node  {} edge from parent[draw=none]}
+%      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
+%      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
+%      child {node [label={\#10}] [fill=black,circle] {}
+%        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
+%        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
+%        edge from parent node[right] {$A$}}
+%      %child {node {to Paris} edge from parent node[right] {$A$}}
+%      edge from parent node[left] {$H$}
+%    };
+%    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+%    }
+%  \end{tikzpicture}
   \end{adjustbox}
@@
 To calculate Kappa, we consider the annotation task as a classification task on 
-nodes, and consider all nodes where there are two annotations from different
-annotators \XXX{unfinished sentence?}
+nodes, and consider those nodes which have two annotations from different
+annotators.
 We report the overall Kappa, as well as separate Kappas on lexical
@@
 a length range of 10, and we show the variance of kappa with length in \figref{fig:iaalength}.
-\XXX{We need to report the number of sentences in each bin, the last German bin
-seems to be an unstable result.}
 
@@
 \caption{Inter-annotator Agreement (Kappa) versus sentence length, separately for
-structural and lexical nodes. }
+structural and lexical nodes. The numbers are the node counts in each bin. }
 \label{fig:iaalength}
-
 \end{figure*}
 
-\subsection{Discussion of Disagreements}\label{sec:disagreements}
-
-\subsection{Comparison with Crowdsourced Adequacy Judgements}\label{sec:adequacy}
+We see from \figref{fig:iaalength} that there is no reduction of IAA with sentence
+length, except perhaps in German. Also, from \tabref{tab:iaa} the overall IAA
+is quite similar (0.6-0.7) for all languages, showing good agreement, but there are
+differences observed when we break down by node type.  In particular, we see a contrast  between
+Czech and Polish (where the IAA is higher for lexical nodes than structural nodes) and German and Romanian
+(where the reverse  is true). We also observe low IAA (around 0.3) in some cases, for instance 
+in German lexical nodes, and for Polish and Czech structural nodes.
+
+\XXX{Discussion of the disagreements}
+
+%\subsection{Discussion of Disagreements}\label{sec:disagreements}
+
+%\subsection{Comparison with Crowdsourced Adequacy Judgements}\label{sec:adequacy}
 

### 2016-03-17 paper/main.tex

@@
 \caption{\label{fig:ucca_example}
-  Sample UCCA annotation. Edge labels mark UCCA categories. }
+  Sample UCCA annotation of the sentence ``After graduation, Tom moved
+  to NYC''. Leaves correspond to words, whereas nodes correspond to units.
+  Edge labels mark UCCA categories, which are largely ignored in this work.}
 \end{figure}
@@
 For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
-The semantic units for a sentence $s$ induced by the annotation $G$,
-is the set of yields of nodes in $V$. %We denote it with $Units(s)$ for brevity.
+The semantic units for a sentence $s$ according to $G$
+is the set of yields of nodes in $V$. 
 We note that the translation is not UCCA-annotated.
 
-\begin{figure*}[t]
-    \includegraphics[width=1\textwidth]{ucca-tree-mteval.jpg}
-    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
-	English paraphrase).}
-    \label{fig:ucca-tree-hume}
+%\begin{figure*}[t]
+%    \includegraphics[width=1\textwidth]{ucca-tree-mteval.jpg}
+%    \caption{UCCA Tree evaluated in comparison to an aligned ``translation'' (an
+%	English paraphrase).}
+%    \label{fig:ucca-tree-hume}
+%\end{figure*}
+
+\begin{figure*}
+  \begin{adjustbox}{trim=2.3cm 0 0 0}
+%  \begin{tikzpicture}
+%    [level 1/.style={sibling distance=7.5em},
+%      level 2/.style={sibling distance=5em},
+%      level 3/.style={sibling distance=5em},
+%      level 4/.style={sibling distance=3em}]
+%    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
+%    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
+%  \node (ROOT) [label={B}] [fill=blue, circle] {}
+%  child {node (H3) [label={A}] [fill=yellow,circle] {}
+%    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
+%    child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
+%      child {node [label={}] {became} edge from parent node[left]  {}}
+%      child {node [label={}] {a} edge from parent node[left]  {}}
+%      child {node [label={}] {passion} edge from parent node[left]  {}}
+%    }
+%    child {node (H2) [label={A}] [fill=yellow,circle] {}
+%      child {node {} edge from parent[draw=none] {}}
+%      child {node {} edge from parent[draw=none] {}}
+%      child {node {} edge from parent[draw=none] {}}
+%      child {node {} edge from parent[draw=none] {}}
+%      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
+%      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
+%      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
+%      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
+%    }
+%  }
+%  child {node {} edge from parent[draw=none] {}}
+%  child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
+%    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
+%    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
+%    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
+%    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
+%  }
+%  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
+%  child {node (H5) [label={A}] [fill=yellow,circle] {}  
+%    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
+%    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
+%  };
+%  \draw[dashed] (H5) -> (she) ;
+%
+%  \com{
+%  child {node (H1) [label={\#8}] [fill=black,circle] {}
+%    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
+%      %child {node {to Paris} edge from parent node[right] {$A$}}
+%      edge from parent node[left] {$H$}
+%    }
+%    child {node (H2) [label={\#9}] [fill=black,circle] {}
+%      child {node  {} edge from parent[draw=none]}
+%      child {node  {} edge from parent[draw=none]}
+%      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
+%      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
+%      child {node [label={\#10}] [fill=black,circle] {}
+%        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
+%        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
+%        edge from parent node[right] {$A$}}
+%      %child {node {to Paris} edge from parent node[right] {$A$}}
+%      edge from parent node[left] {$H$}
+%    };
+%    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+%    }
+%  \end{tikzpicture}
+  \\
+  \end{adjustbox}
+  \adjustbox{width=\textwidth,center,margin=.5em}{\it His old daughter loved tennis, and took daily lessons however good he was}
+
+  \caption{\label{fig:ucca_example2}
+    Sample UCCA annotation of a source sentence, translation (below in italics), 
+    and translation judgements over the source units, marked with A,B,R,O,G labels. Yellow/blue marks
+    nodes annotated with A/B, respectively.
+    In our experiments, the translation is in German, Romaninan, Polish or Czech.
+    ``became a passion'' is marked as atomic, as it is a multi-word expression.
+    The dashed line indicates that ``she'' is also a child of the unit ``greatly improved''.
+    UCCA categories, and alignment between source and translation are omitted for brevity.}
+
 \end{figure*}

### 2016-03-17 paper/main.tex
minor
www

--- 2016-03-17 paper/main.tex
@@
 \usepackage{lipsum,adjustbox}
+\listfiles
 
@@
 
-  \caption{\label{fig:ucca_example2}
+  \caption{\label{fig:hume_tree}
     Sample UCCA annotation of a source sentence, translation (below in italics), 
@@
 
-Structural units have sub-units (children in the UCCA graph), which may either be
-atomic or structural. The labels assigned to such nodes reflect the extent to which
-the relation between the sub-units.
+Structural units have sub-units (children in the UCCA graph), which may themselves
+either be atomic or structural. The labels assigned to such nodes reflect the extent to which
+the relation between the sub-units is preserved in $t$.
 For instance, a Scene unit represents the relation between the
@@
 that the relation between the sub-units went wrong.
-We make a three-way distinction between translation qualities with structural
-nodes, as opposed to the two-way distinction with atomic units, as lexical units
-often can have slight errors, in which cases we do want to provide partial credit.
+We use a three label system with atomic units, as
+opposed to a two label one in structural units,
+as lexical units often have slight errors, in which cases we do want to provide partial credit.
 
 There are many ways that the relationship between the children might go
-wrong in translation:
+wrong in translation
 
@@
 
-\XXX{Could someone review this for Ondrej.}
-Note that for non-configurational languages, the relationship between the
+We note that HUME labels reflect adequacy, rather than fluency judgements.
+Specifically, annotators are instructed to
+label a unit as Adequate if its translation is understandable and preserves
+the meaning of the source unit, even if its fluency is impaired.
+
+In non-configurational languages, the relationship between the
 governing and the dependent units is often expressed using morphological
-properties of the dependents instead of their ordering (this holds especially for the verb and its
-modifiers, less so for components of noun phrases). Since errors in morphology
+properties of the dependents instead of their ordering
+(this holds especially for the verb and its modifiers,
+less so for components of noun phrases). Since errors in morphology
 should be expressed on the atomic units, we would expect
 a higher proportion of ``Adequate'' structural units for Czech and Polish
-compared to German and Romanian\oa{is this because that error would already be marked
-  in the atomic units, and thus not in the structural ones?}.
-
-We defined the labels to reflect adequacy,
-rather than fluency judgements. Specifically, annotators are instructed to
-label a unit as Adequate if its translation is understandable and preserves
-the meaning of the source unit, even if its fluency is impaired.
+compared to German and Romanian\oa{but doesn't German also have a case system? Ondrej:
+  could you pinpoint what the difference exactly is?}.
+
+An example HUME annotation is given in \figref{fig:hume_tree},
+which presents an annotation of a UCCA tree, and a corresponding translation.
+We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
+in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
+individually.
 
@@
 %as their relation in $t$ is reversed.
-
-Another example is given in \figref{fig:ucca-tree-hume}.
-%``Golf became an obsession for his oldest daughter, she took daily lessons and became very good'',
-The figure shows an annotation of a UCCA tree, and an example annotation for
-each of the units.
-We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
-in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
-individually.
 

### 2016-03-17 paper/main.tex

--- 2016-03-17 paper/main.tex
@@
   \begin{adjustbox}{trim=2.3cm 0 0 0}
-%BH: Not compiling for me, commented out
+%%BH: Not compiling for me, commented out
 %  \begin{tikzpicture}

### 2016-03-17 paper/main.tex
uncomment pics

--- 2016-03-17 paper/main.tex
@@
 \usepackage{lipsum,adjustbox}
-\listfiles
 
@@
 
+\begin{figure*}
+  \begin{adjustbox}{trim=2.3cm 0 0 0}
+%BH: Not compiling for me, commented out
+  \begin{tikzpicture}
+    [level 1/.style={sibling distance=7.5em},
+      level 2/.style={sibling distance=5em},
+      level 3/.style={sibling distance=5em},
+      level 4/.style={sibling distance=3em}]
+    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
+    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
+  \node (ROOT) [label={B}] [fill=blue, circle] {}
+  child {node (H3) [label={A}] [fill=yellow,circle] {}
+    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
+    child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
+      child {node [label={}] {became} edge from parent node[left]  {}}
+      child {node [label={}] {a} edge from parent node[left]  {}}
+      child {node [label={}] {passion} edge from parent node[left]  {}}
+    }
+    child {node (H2) [label={A}] [fill=yellow,circle] {}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node {} edge from parent[draw=none] {}}
+      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
+      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
+      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
+      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
+    }
+  }
+  child {node {} edge from parent[draw=none] {}}
+  child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
+    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
+    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
+    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
+    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
+  }
+  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
+  child {node (H5) [label={A}] [fill=yellow,circle] {}  
+    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
+    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
+  };
+  \draw[dashed] (H5) -> (she) ;
+
+  \com{
+  child {node (H1) [label={\#8}] [fill=black,circle] {}
+    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    }
+    child {node (H2) [label={\#9}] [fill=black,circle] {}
+      child {node  {} edge from parent[draw=none]}
+      child {node  {} edge from parent[draw=none]}
+      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
+      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
+      child {node [label={\#10}] [fill=black,circle] {}
+        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
+        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
+        edge from parent node[right] {$A$}}
+      %child {node {to Paris} edge from parent node[right] {$A$}}
+      edge from parent node[left] {$H$}
+    };
+    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+    }
+  \end{tikzpicture}
+  \end{adjustbox}
+\caption{\label{fig:ucca_example2}
+  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
+indices.}
+\end{figure*}
+
 \paragraph{Semantic Representation.}
@@
 
-  \caption{\label{fig:hume_tree}
+  \caption{\label{fig:ucca_example2}
     Sample UCCA annotation of a source sentence, translation (below in italics), 
@@
 
-Structural units have sub-units (children in the UCCA graph), which may themselves
-either be atomic or structural. The labels assigned to such nodes reflect the extent to which
-the relation between the sub-units is preserved in $t$.
+Structural units have sub-units (children in the UCCA graph), which may either be
+atomic or structural. The labels assigned to such nodes reflect the extent to which
+the relation between the sub-units.
 For instance, a Scene unit represents the relation between the
@@
 that the relation between the sub-units went wrong.
-We use a three label system with atomic units, as
-opposed to a two label one in structural units,
-as lexical units often have slight errors, in which cases we do want to provide partial credit.
+We make a three-way distinction between translation qualities with structural
+nodes, as opposed to the two-way distinction with atomic units, as lexical units
+often can have slight errors, in which cases we do want to provide partial credit.
 
 There are many ways that the relationship between the children might go
-wrong in translation
+wrong in translation:
 
@@
 
-We note that HUME labels reflect adequacy, rather than fluency judgements.
-Specifically, annotators are instructed to
+\XXX{Could someone review this for Ondrej.}
+Note that for non-configurational languages, the relationship between the
+governing and the dependent units is often expressed using morphological
+properties of the dependents instead of their ordering (this holds especially for the verb and its
+modifiers, less so for components of noun phrases). Since errors in morphology
+should be expressed on the atomic units, we would expect
+a higher proportion of ``Adequate'' structural units for Czech and Polish
+compared to German and Romanian\oa{is this because that error would already be marked
+  in the atomic units, and thus not in the structural ones?}.
+
+We defined the labels to reflect adequacy,
+rather than fluency judgements. Specifically, annotators are instructed to
 label a unit as Adequate if its translation is understandable and preserves
 the meaning of the source unit, even if its fluency is impaired.
-
-In non-configurational languages, the relationship between the
-governing and the dependent units is often expressed using morphological
-properties of the dependents instead of their ordering
-(this holds especially for the verb and its modifiers,
-less so for components of noun phrases). Since errors in morphology
-should be expressed on the atomic units, we would expect
-a higher proportion of ``Adequate'' structural units for Czech and Polish
-compared to German and Romanian\oa{but doesn't German also have a case system? Ondrej:
-  could you pinpoint what the difference exactly is?}.
-
-An example HUME annotation is given in \figref{fig:hume_tree},
-which presents an annotation of a UCCA tree, and a corresponding translation.
-We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
-in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
-individually.
 
@@
 %as their relation in $t$ is reversed.
+
+Another example is given in \figref{fig:ucca-tree-hume}.
+%``Golf became an obsession for his oldest daughter, she took daily lessons and became very good'',
+The figure shows an annotation of a UCCA tree, and an example annotation for
+each of the units.
+We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
+in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
+individually.
 

### 2016-03-17 paper/main.tex
sss

--- 2016-03-17 paper/main.tex
@@
 system tuning may suggest more complex definition,
-as a starting point we take a parsimonious approach
-and simply aggregate the judgements.
+as a starting point we take a parsimonious approach,
+and compute an accuracy score.
 
@@
   \HUME(s,t) = \frac{\Correct(s,t) + 0.5\cdot \Partial(s,t)}{\setsize{\Units(s)}}
+  \nonumber
 \end{equation}
@@
 of the source, relating to the translation. UCCA annotations of the source are done offline,
-using the UCCA web application\footnote{See UCCA webpage.}.
+using the UCCA web application.
 
@@
 and therefore noisy, the annotator is instructed to treat the aligned text is a cue, but to ignore
-it and make a judgement according to the full translation sentence, in cases where the alignment
-is misleading.
+the alignment if it is misleading, and instead 
+make a judgement according to the full translation sentence.
 
@@
 $t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
-In the event that $a(u)$ is discontiguous in the text, words between the left
+Where $a(u)$ is discontiguous in the text, words between the left
 and right boundaries of $a(u)$ which are not contained in it (intervening words)
@@
 
-For example, consider the sentence $s=$``The man watched the dog'' and the German
-translation ``t=Der Mann sieht den Hund an''. Then ``watched'' will be aligned to
-``sieht ... an'', and ``den Hund'' will be marked as intervening word. Another example
-can be seen in \figref{fig:interface}, where ``ongoing pregnancy'' is translated to
+For example, consider the sentence $s=$``We're driving back on Friday'' and the German
+translation t=``Am Freitag fahren wir zur\"uck'' (lit. ``on Friday driving we back'').
+``driving back'' will be aligned to ``fahren ... zur\"uck'',
+and ``wir'' will be marked as intervening word. Another example
+appears in \figref{fig:interface}, where ``ongoing pregnancy'' is translated to
 ``Schwangerschaft ... laufenden'' (lit. ``pregnancy ... ongoing'').

### 2016-03-17 scripts/iaa.py
heatmaps

@@ -36,7 +36,7 @@ def print_overall_stats(by_lang,args):
   sentence_count = len(by_lang['sent_id'].value_counts())
   print("Sentence count: {}; Node count: {}".format(sentence_count, node_count))
 
-def print_overall_iaa(by_lang, args):
+def print_overall_iaa(by_lang, lang, args):
   groups = (("A", "B", "R", "O", "G"),)
   if args.separate_label_groups:
     groups = (("A", "B"), ("R", "O", "G"), ("R", "G"),("A", "B","R", "O", "G"))
@@ -55,6 +55,12 @@ def print_overall_iaa(by_lang, args):
     #print(cm.to_dataframe())
     #print(cm.stats())
     print("Kappa: %7.5f; P_o: %7.5f; P_e: %7.5f" % (kappa, p_o, p_e))
+    if args.create_heatmaps and group == ("A", "B","R", "O", "G"):
+      import matplotlib.pyplot as plt
+      import seaborn
+      heatmap = seaborn.heatmap(df)
+      plt.savefig("iaa_heatmap_{}.png".format(lang))
+      plt.clf()
 
 def print_iaa_sentence_detail(agree, detail_file):
 #TODO sentence id, matches per sentence, pc a/b, pc r/o/g. src, tgt, ucca stats
@@ -78,6 +84,7 @@ def main():
     help="CSV file containing the node data")
   parser.add_argument("--exclude-missing", default=True, action="store_true",
     help="Excluding nodes that either annotator has missed")
+  parser.add_argument("--create-heatmaps", default=False, action="store_true")
   parser.add_argument("--separate-label-groups", default=True, action="store_true",
     help="Treat A,B and R,G,O separately")
   parser.add_argument("--iaa-sentence-detail", help="Write detailed IAA for each sentence to the given file")
@@ -105,7 +112,7 @@ def main():
     by_lang = agree[agree['lang'] == code]
     print ("************{}*************".format(lang))
     print_overall_stats(by_lang,args)
-    print_overall_iaa(by_lang, args)
+    print_overall_iaa(by_lang, code, args)
     print ()
 
 

### 2016-03-17 paper/main.tex
this fig should have been removed
heatmaps
further on iaa

@@
 \end{figure}
-
-\begin{figure*}
-  \begin{adjustbox}{trim=2.3cm 0 0 0}
-%BH: Not compiling for me, commented out
-  \begin{tikzpicture}
-    [level 1/.style={sibling distance=7.5em},
-      level 2/.style={sibling distance=5em},
-      level 3/.style={sibling distance=5em},
-      level 4/.style={sibling distance=3em}]
-    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
-    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
-  \node (ROOT) [label={B}] [fill=blue, circle] {}
-  child {node (H3) [label={A}] [fill=yellow,circle] {}
-    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
-    child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
-      child {node [label={}] {became} edge from parent node[left]  {}}
-      child {node [label={}] {a} edge from parent node[left]  {}}
-      child {node [label={}] {passion} edge from parent node[left]  {}}
-    }
-    child {node (H2) [label={A}] [fill=yellow,circle] {}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
-      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
-      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
-      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
-    }
-  }
-  child {node {} edge from parent[draw=none] {}}
-  child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
-    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
-  }
-  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
-  child {node (H5) [label={A}] [fill=yellow,circle] {}  
-    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
-    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
-  };
-  \draw[dashed] (H5) -> (she) ;
-
-  \com{
-  child {node (H1) [label={\#8}] [fill=black,circle] {}
-    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    }
-    child {node (H2) [label={\#9}] [fill=black,circle] {}
-      child {node  {} edge from parent[draw=none]}
-      child {node  {} edge from parent[draw=none]}
-      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
-      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
-      child {node [label={\#10}] [fill=black,circle] {}
-        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
-        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
-        edge from parent node[right] {$A$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    };
-    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
-    }
-  \end{tikzpicture}
-  \end{adjustbox}
-\caption{\label{fig:ucca_example2}
-  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
-indices.}
-\end{figure*}
 
@@
 
+\def\iaafig #1{\includegraphics[width=7cm]{iaa_heatmap_#1.png}}
+
+\begin{figure*}[ht!]
+\begin{tabular}{cc}
+
+\subfloat[English-Czech]{
+  \iaafig{cs}
+}
+&
+\subfloat[English-German]{
+  \iaafig{de}
+
+}
+\\
+
+\subfloat[English-Polish]{
+  \iaafig{pl}
+
+}
+&
+\subfloat[English-Romanian]{
+  \iaafig{ro}
+
+}
+\end{tabular}
+\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
+\label{fig:heatmap}
+\end{figure*}
+
 We see from \figref{fig:iaalength} that there is no reduction of IAA with sentence
@@
 \XXX{Discussion of the disagreements}
+
+\begin{compactitem}
+  \item German has low overall proportion of orange annotation (1.2\% compared with 4.5-8.6)
+  \item German annotators give a lot of green (so skewed classes, high expected agreement), but 
+        differ more significantly on what should be green. 
+  \item For Polish, big difference in number of As (2000 vs 1500) given by each annotator
+  \item For Czech structural, both annotators give higher proportion of As, raising the expected agreement.
+\end{compactitem}
 

### 2016-03-17 scripts/iaa.py

@@ -22,6 +22,8 @@ def get_kappa(cm):
   sum_row = df.sum(axis=0)
   sum_col = df.sum(axis=1)
   p_e = float(sum(sum_row*sum_col)) / (N*N)
+  print("normalised sum_row x sum_col")
+  print(((sum_row*sum_col) / (N*N)))
   #print(sum_row)
   #print(sum_col)
   #print (p_e)

### 2016-03-17 paper/main.tex
dd

--- 2016-03-17 paper/main.tex
@@
 
-\begin{figure*}
-  \begin{adjustbox}{trim=2.3cm 0 0 0}
-%BH: Not compiling for me, commented out
-  \begin{tikzpicture}
-    [level 1/.style={sibling distance=7.5em},
-      level 2/.style={sibling distance=5em},
-      level 3/.style={sibling distance=5em},
-      level 4/.style={sibling distance=3em}]
-    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
-    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
-  \node (ROOT) [label={B}] [fill=blue, circle] {}
-  child {node (H3) [label={A}] [fill=yellow,circle] {}
-    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
-    child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
-      child {node [label={}] {became} edge from parent node[left]  {}}
-      child {node [label={}] {a} edge from parent node[left]  {}}
-      child {node [label={}] {passion} edge from parent node[left]  {}}
-    }
-    child {node (H2) [label={A}] [fill=yellow,circle] {}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
-      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
-      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
-      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
-    }
-  }
-  child {node {} edge from parent[draw=none] {}}
-  child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
-    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
-  }
-  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
-  child {node (H5) [label={A}] [fill=yellow,circle] {}  
-    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
-    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
-  };
-  \draw[dashed] (H5) -> (she) ;
-
-  \com{
-  child {node (H1) [label={\#8}] [fill=black,circle] {}
-    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    }
-    child {node (H2) [label={\#9}] [fill=black,circle] {}
-      child {node  {} edge from parent[draw=none]}
-      child {node  {} edge from parent[draw=none]}
-      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
-      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
-      child {node [label={\#10}] [fill=black,circle] {}
-        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
-        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
-        edge from parent node[right] {$A$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    };
-    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
-    }
-  \end{tikzpicture}
-  \end{adjustbox}
-\caption{\label{fig:ucca_example2}
-  Sample UCCA annotation. Edge labels mark UCCA categories. \#$<$index$>$ marks the node
-indices.}
-\end{figure*}
-
 \paragraph{Semantic Representation.}
@@
 
-HMEANT shares much of HUME's motivation for providing a
+HMEANT shares much of HUME's motivation in defining a
 measure that decomposes over semantic units, rather than strings. On the other
@@
 done once for a given source sentence, whereas for the HMEANT set-up, every new
-MT output has to be structurally annotated.
+MT output has to be structurally annotated.\oa{we can give some empirical data here.
+  In the WMT 13 experiment, Lexi and Barry report a little over 3 mins of annotation
+  per sentence + 3 translations. HUME annotation is about 2-4 minutes per sentence.
+  UCCA annotation by trained annotators is about 700 tokens/hour (we didn't
+  check it for individual sentences), but that's done only once so it should be amortized.}
 

### 2016-03-17 paper/main.tex

@@
 
+\begin{compactitem}
+  \item German has low overall proportion of orange annotation (1.2\% compared with 4.5-8.6)
+  \item German annotators give a lot of green (so skewed classes, high expected agreement), but 
+        differ more significantly on what should be green. 
+  \item For Polish, big difference in number of As (2000 vs 1500) given by each annotator
+  \item For Czech structural, both annotators give higher proportion of As, raising the expected agreement.
+\end{compactitem}
+
 %\subsection{Discussion of Disagreements}\label{sec:disagreements}

### 2016-03-17 paper/main.tex
sss

@@
 the candidate translation).
-
 We see this difference as crucial where semantic analysis is concerned, as it
@@
 
-For instance, consider the sentence ``A coronary angioplasty may not be technically possible'',
-which was translated by one of the systems we experimented with as
-``Eine koronare Angioplastie kann nicht technisch m{\"o}glich'', i.e., the main verb ``sein'' (``be'') is omitted.
-While this may be interpreted is a minor error, HMEANT will assign it a very low score, as it failed to
-translate the main verb.
-
-We understand that this is a design feature of HMEANT:
-if the translation per se cannot be annotated with verbal argument structure
-(for whatever reason, including incomprehensibility), it cannot convey the meaning.
-
-HMEANT thus makes a stricter check of whether the translation can be understood
+A design principle of HMEANT states that
+if the translation cannot be annotated with verbal argument structure
+(for whatever reason, including incomprehensibility), it cannot convey the meaning,
+while HUME does not share this theoretical commitment.
+For instance, consider the sentence ``a coronary angioplasty may not be technically possible'',
+and the translation ``eine koronare Angioplastie kann nicht technisch m{\"o}glich''
+(lit. ``a coronary angioplasty can not technically possible''),
+produced by an MT system in our experiments.
+The translation is correct, except that the main verb ``sein'' (``be'') is omitted.
+While this may be interpreted is a minor error, HMEANT will assign it a very low score,
+as it failed to translate the main verb.
+
+However, HMEANT's independent annotation of the source and translation makes a stricter check
+of whether the translation can be understood.
 than we do. HUME may be artificially boosting the perceived understandability of
-the sentence by providing the source.
+the sentence by allowing access to the source.
 

### 2016-03-17 paper/main.tex
sss

@@
 Let $\Partial(s,t)$ be the number of atomic units in $s$ marked as Orange relative
-to $t$. Then HUME's composite score is:
+to $t$. Let $\Units(s)$ be the total number of units marked with any of the five labels.
+Then HUME's composite score is:
 
@@
 at the alignment stage, to avoid unduly penalization due to spurious diverges in
-the structures.
+the structures.\oa{I didn't understand this}
 The same issue was observed by HMEANT authors themselves \parcite{XXXcite}
@@
 
-HUME uses only automatic (word-level) alignment, and it's only purpose is to
-make the annotation process easier. The alignment is indicative only and the
-user is expected to mentally correct the alignment as needed.
+HUME uses only automatic (word-level) alignment, which only serves as a cue for directing
+the attention of the annotators.
+The user is expected to mentally correct the alignment as needed.
 
@@
 HUME diverges from HMEANT (and also from the shallow automatic measures
-like BLEU or
-METEOR) in not requiring a reference.
-Instead, it compares the source directly with the output
-translation.
+like BLEU or METEOR) in not requiring a reference.
+Instead, it compares the source directly with the output translation.
 
@@
 always literal enough and the MT output semantics can be composed differently
-than the semantics of the reference. While this is in principle not a problem
-for HUME (the whole sentence can be labelled as ``Green''), it prevents learning
-any fine-grained information about the MT output.
-
-Perhaps even more importantly,
-the reference translations were often not created by translating the source
-sentences. In some cases, the manual translation was reversed, or both the
-``source'' and ``reference'' are actually independent translations from a third
-language. The ``reference'' can thus easily contain more information than the
-``source''.
+than the semantics of the reference. While HUME avoids this problem,
+imposing a structure only on the source does prevent any fine-grained semantic
+analysis of the translation.
+
+Perhaps more importantly, references are often not created by translating the source
+sentences. In some cases, the manual translation was reversed\oa{what does that mean?},
+and in others the ``source'' and ``reference'' are actually independent translations
+from a third language. ``References'' can thus easily contain information
+that does not appear in the ``source''.
 

### 2016-03-17 paper/main.tex
edits

@@
   \begin{adjustbox}{trim=2.3cm 0 0 0}
-  \begin{tikzpicture}
-    [level 1/.style={sibling distance=7.5em},
-      level 2/.style={sibling distance=5em},
-      level 3/.style={sibling distance=5em},
-      level 4/.style={sibling distance=3em}]
-    %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
-    %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
-  \node (ROOT) [label={B}] [fill=blue, circle] {}
-  child {node (H3) [label={A}] [fill=yellow,circle] {}
-    child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
-    child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
-      child {node [label={}] {became} edge from parent node[left]  {}}
-      child {node [label={}] {a} edge from parent node[left]  {}}
-      child {node [label={}] {passion} edge from parent node[left]  {}}
-    }
-    child {node (H2) [label={A}] [fill=yellow,circle] {}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node {} edge from parent[draw=none] {}}
-      child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
-      child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
-      child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
-      child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
-    }
-  }
-  child {node {} edge from parent[draw=none] {}}
-  child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
-    child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
-    child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
-  }
-  child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
-  child {node (H5) [label={A}] [fill=yellow,circle] {}  
-    child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
-    child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
-  };
-  \draw[dashed] (H5) -> (she) ;
-
-  \com{
-  child {node (H1) [label={\#8}] [fill=black,circle] {}
-    child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    }
-    child {node (H2) [label={\#9}] [fill=black,circle] {}
-      child {node  {} edge from parent[draw=none]}
-      child {node  {} edge from parent[draw=none]}
-      child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
-      child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
-      child {node [label={\#10}] [fill=black,circle] {}
-        child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
-        child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
-        edge from parent node[right] {$A$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    };
-    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
-    }
-  \end{tikzpicture}
+%   \begin{tikzpicture}
+%     [level 1/.style={sibling distance=7.5em},
+%       level 2/.style={sibling distance=5em},
+%       level 3/.style={sibling distance=5em},
+%       level 4/.style={sibling distance=3em}]
+%     %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
+%     %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
+%   \node (ROOT) [label={B}] [fill=blue, circle] {}
+%   child {node (H3) [label={A}] [fill=yellow,circle] {}
+%     child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
+%     child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
+%       child {node [label={}] {became} edge from parent node[left]  {}}
+%       child {node [label={}] {a} edge from parent node[left]  {}}
+%       child {node [label={}] {passion} edge from parent node[left]  {}}
+%     }
+%     child {node (H2) [label={A}] [fill=yellow,circle] {}
+%       child {node {} edge from parent[draw=none] {}}
+%       child {node {} edge from parent[draw=none] {}}
+%       child {node {} edge from parent[draw=none] {}}
+%       child {node {} edge from parent[draw=none] {}}
+%       child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
+%       child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
+%       child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
+%       child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
+%     }
+%   }
+%   child {node {} edge from parent[draw=none] {}}
+%   child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
+%     child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
+%     child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
+%     child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
+%     child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
+%   }
+%   child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
+%   child {node (H5) [label={A}] [fill=yellow,circle] {}  
+%     child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
+%     child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
+%   };
+%   \draw[dashed] (H5) -> (she) ;
+% 
+%   \com{
+%   child {node (H1) [label={\#8}] [fill=black,circle] {}
+%     child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
+%       %child {node {to Paris} edge from parent node[right] {$A$}}
+%       edge from parent node[left] {$H$}
+%     }
+%     child {node (H2) [label={\#9}] [fill=black,circle] {}
+%       child {node  {} edge from parent[draw=none]}
+%       child {node  {} edge from parent[draw=none]}
+%       child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
+%       child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
+%       child {node [label={\#10}] [fill=black,circle] {}
+%         child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
+%         child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
+%         edge from parent node[right] {$A$}}
+%       %child {node {to Paris} edge from parent node[right] {$A$}}
+%       edge from parent node[left] {$H$}
+%     };
+%     \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+%     }
+%   \end{tikzpicture}
   \\
@@
 The translation is correct, except that the main verb ``sein'' (``be'') is omitted.
-While this may be interpreted is a minor error, HMEANT will assign it a very low score,
+While this may be interpreted as a minor error, HMEANT will assign it a very low score,
 as it failed to translate the main verb.
@@
 However, HMEANT's independent annotation of the source and translation makes a stricter check
-of whether the translation can be understood.
+of whether the translation can be understood
 than we do. HUME may be artificially boosting the perceived understandability of

### 2016-03-17 paper/main.tex
iaa disagree

--- 2016-03-17 paper/main.tex
@@
 
-\XXX{Discussion of the disagreements}
-
-\begin{compactitem}
-  \item German has low overall proportion of orange annotation (1.2\% compared with 4.5-8.6)
-  \item German annotators give a lot of green (so skewed classes, high expected agreement), but 
-        differ more significantly on what should be green. 
-  \item For Polish, big difference in number of As (2000 vs 1500) given by each annotator
-  \item For Czech structural, both annotators give higher proportion of As, raising the expected agreement.
-\end{compactitem}
+Looking more closely at the areas of disagreement, we see that for the Polish structural nodes, the 
+proportion of As was quite different between the two annotators (53\% vs. 71\%), whereas for other
+languages the annotators agree in the proportions. We believe that this was because one of the Polish
+annotators did not fully understand the guidelines for structural nodes, and percolated
+errors up the tree, creating more Bs. For the German lexical nodes, and Czech structural nodes, where
+Kappa is also around 0.3, the proportion of such nodes being marked as ``correct'' is relatively 
+high, meaning that the class distribution is more skewed, so the expected agreement used in the
+Kappa calculation is high, lowering Kappa. Essentially, the annotation problem is harder when the classes
+are more skewed. Finally we note some evidence of domain-specific disagreements, for instance the German
+MT system normally translated ``review'' (as in ``systematic review'' -- a frequent term in the 
+Cochrane texts) as ``\"uberpr\"ufung'', which 
+one annotator marked correct, and the other (a Cochrane employee)  as incorrect.
 

### 2016-03-17 paper/main.tex
www

--- 2016-03-17 paper/main.tex
@@
 In order to validate the HUME metric, we ran an annotation experiment with one source language (English), and four 
-target languages (Czech, German, Polish and Romanian), using text from the public health domain. We are 
-particularly interested in this domain since our ultimate goal is to provide high-quality translations of health
-information for speakers of languages other than English, and in this domain accurate transfer of meaning is
-paramount.
-
-Our annotation experiment sought to address the following questions:
+target languages (Czech, German, Polish and Romanian), using text from the public health domain.
+The domain is particularly suitable for experimenting on MT evaluation, as being able to determine
+whether the translation accurately conveys the same information as the source is here paramount.
+
+%particularly interested in this domain since our ultimate goal is to provide high-quality translations of health
+%information for speakers of languages other than English, and in this domain accurate transfer of meaning is
+%paramount.
+
+The experiment seeks to address two questions:
+
 \begin{compactitem}
@@
 the data sets released for the WMT14 medical translation task \cite{bojar-EtAl:2014:W14-33}, to provide between
-45 and 85 million sentences of training data, depending on language pair.
+45 and 85 million sentences of training data, depending on language pair\oa{I think we should probably
+  say how much in each.}.
 
@@
 
-Looking more closely at the areas of disagreement, we see that for the Polish structural nodes, the 
-proportion of As was quite different between the two annotators (53\% vs. 71\%), whereas for other
-languages the annotators agree in the proportions. We believe that this was because one of the Polish
-annotators did not fully understand the guidelines for structural nodes, and percolated
-errors up the tree, creating more Bs. For the German lexical nodes, and Czech structural nodes, where
-Kappa is also around 0.3, the proportion of such nodes being marked as ``correct'' is relatively 
-high, meaning that the class distribution is more skewed, so the expected agreement used in the
-Kappa calculation is high, lowering Kappa. Essentially, the annotation problem is harder when the classes
-are more skewed. Finally we note some evidence of domain-specific disagreements, for instance the German
-MT system normally translated ``review'' (as in ``systematic review'' -- a frequent term in the 
-Cochrane texts) as ``\"uberpr\"ufung'', which 
-one annotator marked correct, and the other (a Cochrane employee)  as incorrect.
+\XXX{Discussion of the disagreements}
+
+\begin{compactitem}
+  \item German has low overall proportion of orange annotation (1.2\% compared with 4.5-8.6)
+  \item German annotators give a lot of green (so skewed classes, high expected agreement), but 
+        differ more significantly on what should be green. 
+  \item For Polish, big difference in number of As (2000 vs 1500) given by each annotator
+  \item For Czech structural, both annotators give higher proportion of As, raising the expected agreement.
+\end{compactitem}
 

### 2016-03-17 paper/main.tex
edits

@@
 
-\subsection{Verbal Structures Only?}
+\paragraph{Verbal Structures Only?}
 
@@
 
-\subsection{Cost of Structural Annotation}
+\paragraph{Cost of Structural Annotation}
 
@@
 
-\subsection{One vs. Two Structures}
+\paragraph{One vs. Two Structures}
 
@@
 
-\subsection{Importance of Alignment}
-
-In HMEANT, the alignment between the (source and candidate) structures is a key
+\paragraph{Importance of Alignment}
+
+In HMEANT, the alignment between the (reference and candidate) structures is a key
 part of the manual annotation. If the alignment cannot be created, the candidate
@@
 \parcite{bojar:wu:ssst:2012,chuchunkov-tarelkin-galinskaya:2014:SSST-8} have
-suggested that modifications to the two structures are desirable and appropriate
-at the alignment stage, to avoid unduly penalization due to spurious diverges in
-the structures.\oa{I didn't understand this}
-The same issue was observed by HMEANT authors themselves \parcite{XXXcite}
-\XXX{cite this: http://repository.ust.hk/ir/Record/1783.1-66107}: the 
+suggested that sometimes, the structures of the reference and candidate
+translations diverge without any good reason, for instance due to a different
+interpretation of a PP-attachment, the verb can have one additional modifier in
+one of the structures. It would be desirable to allow updates to the structural
+annotations also at the alignment stage
+to avoid unduly penalization due to such a spurious diverge.
+The same issue was observed by HMEANT authors themselves
+\parcite{lo:wu:reliability:2014}:
+the 
 inter-annotator agreement on SRL dropped from 90\% to 61\% 
@@
 
-\subsection{Monolingual vs. Bilingual Evaluation}
+\paragraph{Monolingual vs. Bilingual Evaluation}
 HUME diverges from HMEANT (and also from the shallow automatic measures
@@
 
-\subsection{Technical Differences}
+\paragraph{Technical Differences}
 
@@
 
-\subsection{XXX to be removed}
-
-In the *MEANT papers they mainly test by either
-A) correlating the metric to ranking judgements; or
-B) using automated versions of the metric for tuning MT systems
-
-I think neither is particularly convincing, but I'm not sure at the moment what we replace them with.
+% \subsection{ to be removed}
+% 
+% In the *MEANT papers they mainly test by either
+% A) correlating the metric to ranking judgements; or
+% B) using automated versions of the metric for tuning MT systems
+% 
+% I think neither is particularly convincing, but I'm not sure at the moment what we replace them with.
 Dekai also made a claim about the *MEANT metrics being "interpretable" in the sense that they could guide you towards semantic errors that the MT system may be making, however I don't recall seeing this done in his papers.
-
-By the claim that HMEANT explains things, Dekai probably means that the score is decomposable, but so is UCCA-Eval and even BLEU.
+% 
+% By the claim that HMEANT explains things, Dekai probably means that the score is decomposable, but so is UCCA-Eval and even BLEU.
 And UCCA-Eval has one benefit over HMEANT: it points to words in the MT output that are aligned with something but do not convey the meaning well (and we know which part of the source meaning).
 Pure HMEANT might say only that some part of the MT output has no clear frames whatsover (and we do not know which part of the source/reference, unless we add the word alignments).
+% 
+%
 
@@
 In order to validate the HUME metric, we ran an annotation experiment with one source language (English), and four 
-target languages (Czech, German, Polish and Romanian), using text from the public health domain.
-The domain is particularly suitable for experimenting on MT evaluation, as being able to determine
-whether the translation accurately conveys the same information as the source is here paramount.
-
-%particularly interested in this domain since our ultimate goal is to provide high-quality translations of health
-%information for speakers of languages other than English, and in this domain accurate transfer of meaning is
-%paramount.
-
-The experiment seeks to address two questions:
-
+target languages (Czech, German, Polish and Romanian), using text from the public health domain. We are 
+particularly interested in this domain since our ultimate goal is to provide high-quality translations of health
+information for speakers of languages other than English, and in this domain accurate transfer of meaning is
+paramount.
+
+Our annotation experiment sought to address the following questions:
 \begin{compactitem}
@@
 the data sets released for the WMT14 medical translation task \cite{bojar-EtAl:2014:W14-33}, to provide between
-45 and 85 million sentences of training data, depending on language pair\oa{I think we should probably
-  say how much in each.}.
+45 and 85 million sentences of training data, depending on language pair.
 

### 2016-03-17 paper/main.tex
ww

--- 2016-03-17 paper/main.tex
@@
 Underneath the source sentence we see its expandable semantic tree structure with both lexical and structural nodes.
-Lexical nodes only have traffic light annotation, whereas a structural node would normally be labelled ``A'' or ``B'' but could also be labelled as a lexical node, in the case where there is no word to word correspondence for the translation of its children.
+atomic nodes only have traffic light annotation, whereas a structural node would normally be labelled ``A'' or ``B'' but could also be labelled as a lexical node, in the case where there is no word to word correspondence for the translation of its children.
 
@@
 Source sentences are all in English, and the UCCA annotation was performed by four
-computational linguists\oa{were there no other annotators from Prague? Ondrej
-does not remember! Was it just Ondrej+Jana Sindlerova? Then it would be one
-computer scientist and one linguist for Czech.}.
+computational linguists and one linguist.
 
@@
 45 and 85 million sentences of training data, depending on language pair\oa{I think we should probably
-  say how much in each.}.
-
-These translation systems were used to translate texts derived from both NHS 24\footnote{\url{http://www.nhs24.com/}} and 
-Cochrane\footnote{\url{http://www.cochrane.org/}} into the four languages. The first organisation is a public body
-providing healthcare and health-service related information in Scotland, whereas the second is an international NGO which 
+  say how many in each.}.
+
+These translation systems were used to translate texts derived from both NHS 24\footnote{\url{http://www.nhs24.com/}} and Cochrane\footnote{\url{http://www.cochrane.org/}} into the four languages.
+The first organisation is a public body providing healthcare and health-service related information in Scotland, whereas the second is an international NGO which
 provides independent systematic reviews on health-related research. From NHS 24, we selected the segments from their
@@
 For German, Polish and Romanian we used two annotators, with a significant overlap
-between the sentences they annotated. For Czech we had one main annotator, plus
-several second annotators, so we treat all these second annotators as a single
-annotator.
-
-To calculate Kappa, we consider the annotation task as a classification task on 
-nodes, and consider those nodes which have two annotations from different
-annotators.
-We report the overall Kappa, as well as separate Kappas on lexical
-nodes (annotated as red, orange or green) and the structural nodes (annotated
-as acceptable or bad). In general there should not be confusion between the two
+between the sentences they annotated. For Czech we had one main annotator, who
+annotated the entire batch, as well multiple secondary annotators, where
+each annotated a sub-part of the batch. \oa{i editted it. is what i wrote correct?}
+
+In order to compute Kappa, we consider the annotation task as a classification task on 
+nodes, and compute agreement on nodes which have two annotations from different annotators.
+We report the overall Kappa, as well as separate Kappas on atomic
+nodes (annotated as Red, Orange or Green) and the structural nodes (annotated
+as Adequate or Bad). In general there should be little confusion between the two
 types of nodes, and in future work the separation between the nodes should be enforced
-by the tool.  The Kappas are shown in \tabref{tab:iaa}
+by the tool\oa{but there may be some disagreement in the margins on what nodes are atomic and
+  what are not}. The Kappas are shown in \tabref{tab:iaa}.
+
 \begin{table}[!ht]
@@
 \hline
-Lexical nodes & 2982 & 1724 & 5386 & 3570 \\
+Atomic nodes & 2982 & 1724 & 5386 & 3570 \\
 Kappa & 0.54 & 0.29 & 0.54 & 0.50 \\
@@
 \caption{Inter-annotator Agreement (Kappa) versus sentence length, separately for
-structural and lexical nodes. The numbers are the node counts in each bin. }
+structural and atomic nodes. The numbers are the node counts in each bin. }
 \label{fig:iaalength}
@@
 length, except perhaps in German. Also, from \tabref{tab:iaa} the overall IAA
-is quite similar (0.6-0.7) for all languages, showing good agreement, but there are
-differences observed when we break down by node type.  In particular, we see a contrast  between
-Czech and Polish (where the IAA is higher for lexical nodes than structural nodes) and German and Romanian
-(where the reverse  is true). We also observe low IAA (around 0.3) in some cases, for instance 
-in German lexical nodes, and for Polish and Czech structural nodes.
+is quite similar (0.6-0.7) for all languages, showing good agreement.
+However, there are differences observed when we break down by node type.
+In particular, we see a contrast  between
+Czech and Polish, where the IAA is higher for atomic than for structural nodes, and German and Romanian,
+where the reverse is true. We also observe low IAA (around 0.3) in the cases of
+German atomic nodes, and Polish and Czech structural nodes.
 

### 2016-03-17 paper/main.tex

--- 2016-03-17 paper/main.tex
@@
 the data sets released for the WMT14 medical translation task \cite{bojar-EtAl:2014:W14-33}, to provide between
-45 and 85 million sentences of training data, depending on language pair.
-
-These translation systems were used to translate texts derived from both NHS 24\footnote{\url{http://www.nhs24.com/}} and 
-Cochrane\footnote{\url{http://www.cochrane.org/}} into the four languages. The first organisation is a public body
-providing healthcare and health-service related information in Scotland, whereas the second is an international NGO which 
+45 and 85 million sentences of training data, depending on language pair\oa{I think we should probably
+  say how many in each.}.
+
+These translation systems were used to translate texts derived from both NHS 24\footnote{\url{http://www.nhs24.com/}} and Cochrane\footnote{\url{http://www.cochrane.org/}} into the four languages.
+The first organisation is a public body providing healthcare and health-service related information in Scotland, whereas the second is an international NGO which
 provides independent systematic reviews on health-related research. From NHS 24, we selected the segments from their

### 2016-03-17 paper/main.tex
some edits to intro

@@
 HMEANT \cite{lo2011structured}.
-While allowing to localize semantic errors in the translation,
-these measures' focus on verb argument structures misses out on many pervasive semantic phenomena,
-such as inter-clause linkage and nominal argument structures, and thus only provides
+These measures can localize semantic errors in the translation, but
+their focus on verb argument structures misses many pervasive semantic phenomena
+(such as inter-clause linkage and nominal argument structures) and thus they only provide
 a limited perspective on the semantic similarity between the output and reference.
@@
 English-Romanian and English-Czech, and report efficiency (time of annotation) and
-inter-annotator agreement scores, as well as qualitative comparison of the resulting
-HUME scores with crowd-sourced adequacy judgements. The results show that HUME obtains 
+inter-annotator agreement scores, as well showing how HUME can address a range of 
+semantic phenomena\bh{Do we?}. The results show that HUME obtains 
 agreement scores which are on par with those of standard sentence-level ranking measures,
-but unlike them, HUME's scores do not decrease with longer sentences.\bh{We may
-  have to focus more on qualitative aspects, if we just try to sell
-  the measure on IAA, then the figures are not stunning.}
+but unlike them, HUME's scores do not decrease with longer sentences.
 The paper concludes by outlining ongoing work to construct semi- and

### 2016-03-17 paper/main.tex
correction of examples

@@
 %of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
-Human evaluation is generally determined
-either according to the relative ranking of the outputs of multiple systems
+Human evaluation is generally done by ranking the outputs of multiple systems
 e.g., in the WMT shared tasks \cite{bojar2015findings}, or by assigning
@@
 However, while providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
-and is unfeasible where a wide variety of systems or parameter settings are to be evaluated.
+and is infeasible where a wide variety of systems or parameter settings are to be evaluated.
 % has motivated the construction of cheaper methods that correlate with human
@@
 judgements. Most commonly, such scores decompose over the sub-parts of
-the translation, and quantify how many of these sub-parts appear in a reference translation,
-whose validity has been manually confirmed. This decomposition further allows system developers
-to localize the errors or at least unconfirmed parts within the translation.
+the translation, and quantify how many of these sub-parts appear in a manually created reference translation.
+This decomposition further allows system developers
+to localize the errors, or at least unconfirmed parts within the translation.
 The most commonly used measures decompose over n-grams or individual words, e.g., 
-the BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR measures \cite{Banerjee:2005}.
+BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR  \cite{Banerjee:2005}.
 Another common approach is to determine the similarity between the reference and translation
@@
 as one or more participants. For example, the sentence ``After graduation, Tom moved to NYC''
-contains two Scenes, whose main relations are ``graduation'' and ``Tom''.
-The participant ``John'' is a part of both Scenes, while ``NYC'' only of the
+contains two Scenes, whose main relations are ``graduation'' and ``moved''.
+The participant ``Tom'' is a part of both Scenes, while ``NYC'' only of the
 latter (\figref{fig:ucca_example}). Further categories account for
@@
   \begin{adjustbox}{trim=2.3cm 0 0 0}
-%   \begin{tikzpicture}
-%     [level 1/.style={sibling distance=7.5em},
-%       level 2/.style={sibling distance=5em},
-%       level 3/.style={sibling distance=5em},
-%       level 4/.style={sibling distance=3em}]
-%     %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
-%     %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
-%   \node (ROOT) [label={B}] [fill=blue, circle] {}
-%   child {node (H3) [label={A}] [fill=yellow,circle] {}
-%     child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
-%     child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
-%       child {node [label={}] {became} edge from parent node[left]  {}}
-%       child {node [label={}] {a} edge from parent node[left]  {}}
-%       child {node [label={}] {passion} edge from parent node[left]  {}}
-%     }
-%     child {node (H2) [label={A}] [fill=yellow,circle] {}
-%       child {node {} edge from parent[draw=none] {}}
-%       child {node {} edge from parent[draw=none] {}}
-%       child {node {} edge from parent[draw=none] {}}
-%       child {node {} edge from parent[draw=none] {}}
-%       child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
-%       child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
-%       child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
-%       child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
-%     }
-%   }
-%   child {node {} edge from parent[draw=none] {}}
-%   child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
-%     child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
-%     child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
-%     child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
-%     child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
-%   }
-%   child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
-%   child {node (H5) [label={A}] [fill=yellow,circle] {}  
-%     child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
-%     child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
-%   };
-%   \draw[dashed] (H5) -> (she) ;
-% 
-%   \com{
-%   child {node (H1) [label={\#8}] [fill=black,circle] {}
-%     child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
-%       %child {node {to Paris} edge from parent node[right] {$A$}}
-%       edge from parent node[left] {$H$}
-%     }
-%     child {node (H2) [label={\#9}] [fill=black,circle] {}
-%       child {node  {} edge from parent[draw=none]}
-%       child {node  {} edge from parent[draw=none]}
-%       child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
-%       child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
-%       child {node [label={\#10}] [fill=black,circle] {}
-%         child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
-%         child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
-%         edge from parent node[right] {$A$}}
-%       %child {node {to Paris} edge from parent node[right] {$A$}}
-%       edge from parent node[left] {$H$}
-%     };
-%     \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
-%     }
-%   \end{tikzpicture}
+   \begin{tikzpicture}
+     [level 1/.style={sibling distance=7.5em},
+       level 2/.style={sibling distance=5em},
+       level 3/.style={sibling distance=5em},
+       level 4/.style={sibling distance=3em}]
+     %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
+     %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
+   \node (ROOT) [label={B}] [fill=blue, circle] {}
+   child {node (H3) [label={A}] [fill=yellow,circle] {}
+     child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
+     child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
+       child {node [label={}] {became} edge from parent node[left]  {}}
+       child {node [label={}] {a} edge from parent node[left]  {}}
+       child {node [label={}] {passion} edge from parent node[left]  {}}
+     }
+     child {node (H2) [label={A}] [fill=yellow,circle] {}
+       child {node {} edge from parent[draw=none] {}}
+       child {node {} edge from parent[draw=none] {}}
+       child {node {} edge from parent[draw=none] {}}
+       child {node {} edge from parent[draw=none] {}}
+       child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
+       child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
+       child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
+       child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
+     }
+   }
+   child {node {} edge from parent[draw=none] {}}
+   child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
+     child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
+     child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
+     child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
+     child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
+   }
+   child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
+   child {node (H5) [label={A}] [fill=yellow,circle] {}  
+     child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
+     child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
+   };
+   \draw[dashed] (H5) -> (she) ;
+
+   \com{
+   child {node (H1) [label={\#8}] [fill=black,circle] {}
+     child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
+       %child {node {to Paris} edge from parent node[right] {$A$}}
+       edge from parent node[left] {$H$}
+     }
+     child {node (H2) [label={\#9}] [fill=black,circle] {}
+       child {node  {} edge from parent[draw=none]}
+       child {node  {} edge from parent[draw=none]}
+       child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
+       child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
+       child {node [label={\#10}] [fill=black,circle] {}
+         child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
+         child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
+         edge from parent node[right] {$A$}}
+       %child {node {to Paris} edge from parent node[right] {$A$}}
+       edge from parent node[left] {$H$}
+     };
+     \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+     }
+   \end{tikzpicture}
   \\

### 2016-03-17 paper/main.tex
aaa

--- 2016-03-17 paper/main.tex
@@
 that the evaluator be proficient in both the source and target languages.
-See \secref{sec:hmeant_comp}.
+An elaborate comparision of HUME and HMEANT is given in \secref{sec:hmeant_comp}.
 
@@
 English-Romanian and English-Czech, and report efficiency (time of annotation) and
-inter-annotator agreement scores, as well as qualitative comparison of the resulting
-HUME scores with crowd-sourced adequacy judgements. The results show that HUME obtains 
-agreement scores which are on par with those of standard sentence-level ranking measures,
-but unlike them, HUME's scores do not decrease with longer sentences.\bh{We may
+inter-annotator agreement scores (\secref{sec:exps}). Efficiency results that HUME
+can be annotated in 2--4 minutes, once the UCCA annotation of the source is given,
+which is only about five times slower than MEANT, and hence still practical to
+employ on a moderate scale. HUME's agreement results are on par with those
+of standard sentence-level ranking measures, but unlike them,
+HUME's scores do not decrease with longer sentences.\bh{We may
   have to focus more on qualitative aspects, if we just try to sell
-  the measure on IAA, then the figures are not stunning.}
-The paper concludes by outlining ongoing work to construct semi- and
-fully-automatic variants of HUME.
+  the measure on IAA, then the figures are not stunning. OA: is that OK?}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conlusion}\label{sec:conclusion}
+\section{Conclusion}\label{sec:conclusion}
+
+The paper concludes by outlining ongoing work to construct semi- and
+fully-automatic variants of HUME.
 

### 2016-03-17 paper/main.tex

--- 2016-03-17 paper/main.tex
@@
 English-Romanian and English-Czech, and report efficiency (time of annotation) and
-inter-annotator agreement scores, as well showing how HUME can address a range of 
-semantic phenomena\bh{Do we?}. The results show that HUME obtains 
-agreement scores which are on par with those of standard sentence-level ranking measures,
-but unlike them, HUME's scores do not decrease with longer sentences.
-The paper concludes by outlining ongoing work to construct semi- and
-fully-automatic variants of HUME.
+inter-annotator agreement scores (\secref{sec:exps}). Efficiency results that HUME
+can be annotated in 2--4 minutes, once the UCCA annotation of the source is given,
+which is only about five times slower than MEANT, and hence still practical to
+employ on a moderate scale. HUME's agreement results are on par with those
+of standard sentence-level ranking measures, but unlike them,
+HUME's scores do not decrease with longer sentences.\bh{We may
+  have to focus more on qualitative aspects, if we just try to sell
+  the measure on IAA, then the figures are not stunning. OA: is that OK?}
 

### 2016-03-17 paper/main.tex
mi

@@
 English-Romanian and English-Czech, and report efficiency (time of annotation) and
-inter-annotator agreement scores (\secref{sec:exps}). Efficiency results that HUME
+inter-annotator agreement scores (\secref{sec:experiments}). Efficiency results that HUME
 can be annotated in 2--4 minutes, once the UCCA annotation of the source is given,

### 2016-03-17 paper/main.tex
background edits

@@
 units as it does not ground its semantic symbols in the text,
-thus not straightwardly allowing a decomposition of the sentence into sub-units.
-Moreover, AMR is more fine-grained than UCCA and consequently more difficult to annotate.
+thus does not provide clear decomposition of the sentence into sub-units.
+Also, AMR is more fine-grained than UCCA and consequently more difficult to annotate.
 Other approaches pursued a grounded representation of semantics,

### 2016-03-17 paper/main.tex
edits to annotation section

@@
 opposed to a two label one in structural units,
-as lexical units often have slight errors, in which cases we do want to provide partial credit.
+as atomic units often have slight errors, in which cases we do want to provide partial credit.
 
@@
 \item A superfluous word or phrase has been inserted into
-  the translation. For instance, if ``dog bites man'' is translated into ``dog visciously bites man''.
+  the translation. For instance, if ``dog bites man'' is translated into ``dog viciously bites man''.
 \item One of the sub-units is missing from the translation.
-  For instance, if ``dog bites man'' is translated into ``dog bites''.
+  For instance, if ``dog bites man'' is translated into ``dog bites''. \bh{Note that in this case we also 
+  get an error in the atomic unit, assuming that ``man'' is a semantic unit in the source, but for the other
+  two cases there is no atomic unit error.}
 \end{compactitem}
@@
 An example HUME annotation is given in \figref{fig:hume_tree},
-which presents an annotation of a UCCA tree, and a corresponding translation.
+which presents an annotation of an UCCA tree, and a corresponding translation.
 We see that ``became a passion'' is deemed atomic, as it is translated as a single unit

### 2016-03-17 paper/main.tex
www

@@
 
-In order to validate the HUME metric, we ran an annotation experiment with one source language (English), and four 
-target languages (Czech, German, Polish and Romanian), using text from the public health domain.
-The domain is particularly suitable for experimenting on MT evaluation, as being able to determine
-whether the translation accurately conveys the same information as the source is here paramount.
+In order to validate the HUME metric, we ran an annotation experiment with one source language (English),
+and four target languages (Czech, German, Polish and Romanian), using text from the public health domain.
+The domain is particularly suitable for experimenting on MT evaluation, as semantically
+accurate translation is paramount here, underscoring the importance of accurate
+semantic evaluation measures.
 
@@
 \emph{Describe the UCCA annotation}
-
-%\subsection{Data}\label{sec:data}
-
-%\subsection{Efficiency of Annotation}\label{sec:efficiency}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\com{
 \section{Human Semantic Evaluation}
@@
 most promising representation for evaluating translation. 
+}
 
@@
 
-\section*{Acknowledgments}
-
-Do not number the acknowledgment section.
-This section should not be presented for the submission version.
+%\section*{Acknowledgments}
+
+%Do not number the acknowledgment section.
+%This section should not be presented for the submission version.
 

### 2016-03-17 paper/main.tex
www
Added alternative images for UCCA and HUME annotations

@@
   to NYC''. Leaves correspond to words, whereas nodes correspond to units.
-  Edge labels mark UCCA categories, which are largely ignored in this work.}
+  Edge labels mark UCCA categories, which are described in the table.}
+%which are largely ignored in this work
+\end{figure}
+
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=0.45\textwidth]{ucca-tree-v2.png}
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+L & Linker &  A & Participant \\
+H & Scene & R & Relator \\
+P & Process & C & Centre \\
+\hline
+\end{tabular}
+    \end{center}
+\caption{\label{fig:ucca_example_v2}
+  Sample UCCA annotation where leaves correspond to words and nodes correspond to units.
+  Edge labels mark UCCA categories, which are described in the table.}
+%which are largely ignored in this work
 \end{figure}
@@
 
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=0.45\textwidth]{ucca-tree-mteval-v2.png}
+    \end{center}
+  \caption{\label{fig:hume_tree_v2}
+     HUME annotation of an UCCA tree with a word aligned example translation shown below. 
+Atomic units are labelled using traffic lights and structural nodes are marked A or B.}
+\end{figure}
+
+In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME annotation, with the translation shown in English for ease of comprehension.
+When evaluating ``to NYC'' the annotator looks at the translation and sees the word ``stateside''.
+This word captures the whole phrase and so we mark the structural node with an atomic label.
+Here we choose ``Orange'' but the annotator could equally have decided that this translation is wrong and marked it as ``Red''.
+The ability to mark structural nodes with atomic labels allows the annotator to account for translations which only correspond at the phrase level.
+Another feature highlighted in this example is that the root node is ``Adequate''.
+This might seem confusing, as the relationship between the scenes has changed in the translation.
+However, an important feature of HUME is that penalises errors where they occur.
+In HUME, the error is caused by a wrong translation of the linker ``After'' which translates as ``by'' and this is marked as ``Red''.
+The root node still has two scene linked by one linker and there relationship between child nodes is maintained and therefore it is marked ``Adequate''.
+In HMEANT it is not clear whether errors should be propegated up the SRL hierarchy.
+
 \begin{figure*}
@@
     \begin{center}
-    \includegraphics[width=1\textwidth]{hume_interface.png}
+    \includegraphics[width=1\textwidth]{hume_interface2.jpg}
     \caption{The HUME annotation tool, based on the UCCA tool. The faded orange box at the top

### 2016-03-18 paper/main.tex
edits

@@
 Perhaps more importantly, references are often not created by translating the source
-sentences. In some cases, the manual translation was reversed\oa{what does that mean?},
-and in others the ``source'' and ``reference'' are actually independent translations
+sentences. In some cases, the manual translation was actually carried out in the
+other direction, from ``reference'' to ``source''
+and in others the ``source'' and ``reference'' are independent translations
 from a third language. ``References'' can thus easily contain information
@@
 
-% \subsection{ to be removed}
+\oa{I would consider adding further examples from the Google Doc. For instnace ``your risk of falling'' -- ``Ihr Risiko'' or
+  ``Two {\bf review} authors assessed the studies independently and extracted outcome data'' -- ``Zwei Autoren bewertete die Studien unabh\"angig zu {\bf \"uberpr\"ufen}''}
+
+% \Subsectionngigngig zu  Zu { to be removed}
 % 

### 2016-03-18 paper/main.tex
edits

@@
 
-\paragraph{Technical Differences}
-
-+ allows for multi-word and even non-contiguous elements
-
-+ allows to indicate the quality of the predicates ("Actions")
+\paragraph{Annotation of Predicates}
+
+HUME allows to annotate multi-word and even
+non-contiguous elements as a single unit. HMEANT, on the other hand, always
+required exactly one word to be annotated as the predicate (``Action''), which
+is often linguistically inadequate.
+
+Another problem with HMEANT identified by \parcite{bojar:wu:ssst:2012} was the
+
+inability to indicate that the predicate is somewhat mistranslated (e.g.
+accidentally negated or in a slightly wrong tense). HMEANT
+annotation allows either to mark the source and reference frames as matching
+(which implies that the corresponding predicates are correct translations of each
+other), or as non-matching (which leads to a zero score for the whole frame) but
+nothing in between. In HUME, the verb is still present as a sub-unit of the
+frame and has the traffic-lights annotation.
 

### 2016-03-18 paper/main.tex
edits

@@
 
+It is also relatively common in translation that verbal constructions get
+nominalized or vice versa. An example from our dataset:
+  ``Two {\bf review} authors assessed the studies independently\dots{}'' --
+  ``Zwei Autoren bewertete die Studien unabh\"angig zu {\bf
+  \"uberpr\"ufen}\dots{}'' (lit. ``Two authors assessed the studies
+  independently to be reviewed\dots").\XXX{This is Omri's example, with
+  Ondrej's gloss for German. Ondrej is
+  not totally sure that the meaning is well preserved.}
+
 We are not aware of any empirical evaluation that demonstrates that verb argument structure,
@@
 Another problem with HMEANT identified by \parcite{bojar:wu:ssst:2012} was the
-
 inability to indicate that the predicate is somewhat mistranslated (e.g.
@@
 
-\oa{I would consider adding further examples from the Google Doc. For instnace ``your risk of falling'' -- ``Ihr Risiko'' or
-  ``Two {\bf review} authors assessed the studies independently and extracted outcome data'' -- ``Zwei Autoren bewertete die Studien unabh\"angig zu {\bf \"uberpr\"ufen}''}
+\XXX{Omri: I would consider adding further examples from the Google Doc. For instnace
+``your risk of falling'' -- ``Ihr Risiko''.
+Ondrej: I don't see any HMEANT problem here, the whole NP 'your risk of falling'
+is aligned with 'Ihr Risiko' and HMEANT allows to indicate that the match was
+partial, to that's probably what everyone would choose. I've put the other
+example higher, but please have a look, I'm not sure if we should keep it there
+at all.}
 

### 2016-03-18 paper/main.tex
edits

@@
 \def\inparcite#1{\newcite{#1}}
+
+\def\equo#1{``{#1}''}  % curly quotes
 
@@
 
+Here is \XXX{another} example with a missing linker: the source \equo{However,
+this review\dots} was translated as \equo{Diese
+Überprüfung\dots} (lit. \equo{This review}). The omission is not by HMEANT at
+all.
+
 We are not aware of any empirical evaluation that demonstrates that verb argument structure,
@@
 
-Another problem with HMEANT identified by \parcite{bojar:wu:ssst:2012} was the
+Another problem with HMEANT identified by \perscite{bojar:wu:ssst:2012} was the
 inability to indicate that the predicate is somewhat mistranslated (e.g.

### 2016-03-18 paper/main.tex
edits

@@
    \begin{tikzpicture}
-     [level 1/.style={sibling distance=7.5em},
-       level 2/.style={sibling distance=5em},
+     [ level distance=12mm,
+	   level 1/.style={sibling distance=7.5em,level distance=7mm},
+       level 2/.style={sibling distance=5em,level distance=12mm},
        level 3/.style={sibling distance=5em},
@@
   \end{adjustbox}
-  \adjustbox{width=\textwidth,center,margin=.5em}{\it His old daughter loved tennis, and took daily lessons however good he was}
+    % Ondrej removed the width=\textwidth from the options below to keep font
+	% unenlarged:
+  \adjustbox{center,margin=.5em}{\it His old daughter loved tennis, and took daily lessons however good he was}
 

### 2016-03-18 paper/main.tex
edits

@@
 
-In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME annotation, with the translation shown in English for ease of comprehension.
-When evaluating ``to NYC'' the annotator looks at the translation and sees the word ``stateside''.
-This word captures the whole phrase and so we mark the structural node with an atomic label.
-Here we choose ``Orange'' but the annotator could equally have decided that this translation is wrong and marked it as ``Red''.
-The ability to mark structural nodes with atomic labels allows the annotator to account for translations which only correspond at the phrase level.
-Another feature highlighted in this example is that the root node is ``Adequate''.
-This might seem confusing, as the relationship between the scenes has changed in the translation.
-However, an important feature of HUME is that penalises errors where they occur.
-In HUME, the error is caused by a wrong translation of the linker ``After'' which translates as ``by'' and this is marked as ``Red''.
-The root node still has two scene linked by one linker and there relationship between child nodes is maintained and therefore it is marked ``Adequate''.
+In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME
+annotation, with the translation shown in English for ease of comprehension.
+When evaluating ``to NYC'' the annotator looks at the translation and sees the
+word ``stateside''. This word captures the whole phrase and so we mark the
+structural node with an atomic label. Here we choose ``Orange'' but the
+annotator could equally have decided that this translation is wrong and marked
+it as ``Red''. The ability to mark structural nodes with atomic labels allows
+the annotator to account for translations which only correspond at the phrase
+level. Another feature highlighted in this example is that the root node is
+``Adequate''. This might seem confusing, as the relationship between the scenes
+has changed in the translation. However, an important feature of HUME is that
+penalises errors where they occur. In HUME, the error is caused by a wrong
+translation of the linker ``After'' which translates as ``by'' and this is
+marked as ``Red''.
+The root node still has two scenes linked by one linker and there relationship between child nodes is maintained and therefore it is marked ``Adequate''.
 In HMEANT it is not clear whether errors should be propegated up the SRL hierarchy.

### 2016-03-18 paper/main.tex
edits

--- 2016-03-18 paper/main.tex
@@
 
-In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME annotation, with the translation shown in English for ease of comprehension.
-When evaluating ``to NYC'' the annotator looks at the translation and sees the word ``stateside''.
-This word captures the whole phrase and so we mark the structural node with an atomic label.
-Here we choose ``Orange'' but the annotator could equally have decided that this translation is wrong and marked it as ``Red''.
-The ability to mark structural nodes with atomic labels allows the annotator to account for translations which only correspond at the phrase level.
-Another feature highlighted in this example is that the root node is ``Adequate''.
-This might seem confusing, as the relationship between the scenes has changed in the translation.
-However, an important feature of HUME is that penalises errors where they occur.
-In HUME, the error is caused by a wrong translation of the linker ``After'' which translates as ``by'' and this is marked as ``Red''.
-The root node still has two scene linked by one linker and there relationship between child nodes is maintained and therefore it is marked ``Adequate''.
-In HMEANT it is not clear whether errors should be propegated up the SRL hierarchy.
+In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME
+annotation, with the translation shown in English for ease of comprehension.
+When evaluating ``to NYC'' the annotator looks at the translation and sees the
+word ``stateside''. This word captures the whole phrase and so we mark the
+structural node with an atomic label. Here we choose ``Orange'' but the
+annotator could equally have decided that this translation is wrong and marked
+it as ``Red''. The ability to mark structural nodes with atomic labels allows
+the annotator to account for translations which only correspond at the phrase
+level. Another feature highlighted in this example is that the root node is
+``Adequate''. This might seem confusing, as the relationship between the scenes
+has changed in the translation. However, an important feature of HUME is that
+penalises errors where they occur. In HUME, the error is caused by a wrong
+translation of the linker ``After'' which translates as ``by'' and this is
+marked as ``Red''. The root node still has two scenes linked by one linker and
+the relationship between child nodes is maintained and therefore it is marked
+``Adequate''.
+% Ondrej removes the following, he thinks that errors are not supposed to be
+% propagated in HMEANT either:
+%In HMEANT, it is not clear whether errors should be propegated up the SRL hierarchy.
 

### 2016-03-18 paper/main.tex
Commented out older graphics and heatmaps and removed comments to see how much space we now have

--- 2016-03-18 paper/main.tex
@@
 \newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
-\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
-\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
+%\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
+\newcommand{\oa}[1]{}
+%\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
+\newcommand{\bh}[1]{}
 
@@
 \def\inparcite#1{\newcite{#1}}
-
-\def\equo#1{``{#1}''}  % curly quotes
 
@@
 
-  Recent interest in semantics-aware MT systems underscores the importance of
-  semantic evaluation for Machine Translation (MT) systems. 
-  A semantic measure, sensitive to possible semantic discrepancies
-  between the translation output and the source, is likely support the construction
-  and tuning of semantics-based MT, much in the same way that string-based measures,
-  such as BLEU, enabled progress in shallower MT.
-  We present a novel human semantic evaluation measure for MT, Human
-  UCCA-based MT Evaluation (HUME), building on the UCCA semantic representation scheme.
-  Making progress over previous work, the proposed measure takes into account
-  a wider range of semantic phenomena and does not rely on semantic annotation
-  of the MT output itself, whose interpretation is often unclear.
-  We experiment with four language pairs, demonstrating HUME's broad applicability,
-  as well as its reliability, reflected in its inter-annotator agreement rates.
+%  Recent interest in semantics-aware MT systems underscores the importance of
+%  semantic evaluation for Machine Translation (MT) systems. 
+%  A semantic measure, sensitive to possible semantic discrepancies
+%  between the translation output and the source, is likely to support the construction
+%  and tuning of semantics-based MT, much in the same way that string-based measures,
+%  such as BLEU, enabled progress in shallower MT.
+%  We present a novel human semantic evaluation measure for MT, Human
+%  UCCA-based MT Evaluation (HUME), building on the UCCA semantic representation scheme.
+%  Making progress over previous work, the proposed measure takes into account
+%  a wider range of semantic phenomena and does not rely on semantic annotation
+%  of the MT output itself, whose interpretation is often unclear.
+%  We experiment with four language pairs, demonstrating HUME's broad applicability,
+%  as well as its reliability, reflected in its inter-annotator agreement rates.
+
+Human evaluation of machine translation is a cognitively difficult task and when annotators rate or rank complete sentences we see low levels of agreement especially
+for longer sentences.  
+In order to extract more reliable annotations, previous research broke down evaluation into smaller components by asking humans to annotate semantic roles on the source and translation, and then to align them.
+We propose a new evaluation method which is based on semantic annotation called Universal Conceptual Cognitive Annotation(UCCA).
+UCCA is applicable across many languages, it provides a complete structure over sentences, and it is grounded in the words in the sentence.
+We apply UCCA only over the source or reference sentence, and not over the possibly garbled translation, 
+and this makes annotation more reliable and also efficient as we can reuse labor intensive semantic annotations.
+We show that HUME can be reliably applied to four different target languages and that sentence level inter annotor agreement does not decrease over longer sentences.
 
@@
 
-\begin{figure}
-
-\begin{tikzpicture}[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
-    inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
-  \node (ROOT) [fill=black, circle] {}
-  child {node {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
-  child {node (H1)  [fill=black,circle] {}
-    child {node {graduation} edge from parent node[right] {$P$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    }
-    child {node (H2)  [fill=black,circle] {}
-      child {node  {} edge from parent[draw=none]}
-      child {node  {} edge from parent[draw=none]}
-      child {node  (Tom) {Tom} edge from parent node[right] {$A$}}
-      child {node  {moved} edge from parent node[right] {$P$}}
-      child {node  [fill=black,circle] {}
-        child {node  {to} edge from parent node[right] {$R$}}
-        child {node  {NYC} edge from parent node[right] {$C$}}
-        edge from parent node[right] {$A$}}
-      %child {node {to Paris} edge from parent node[right] {$A$}}
-      edge from parent node[left] {$H$}
-    };
-    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
-  \end{tikzpicture}
-\caption{\label{fig:ucca_example}
-  Sample UCCA annotation of the sentence ``After graduation, Tom moved
-  to NYC''. Leaves correspond to words, whereas nodes correspond to units.
-  Edge labels mark UCCA categories, which are described in the table.}
-%which are largely ignored in this work
-\end{figure}
+%\begin{figure}
+%
+%\begin{tikzpicture}[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
+%    inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
+%  \node (ROOT) [fill=black, circle] {}
+%  child {node {After} edge from parent node[left]  {$L$\hspace{1.5cm}}}
+%  child {node (H1)  [fill=black,circle] {}
+%    child {node {graduation} edge from parent node[right] {$P$}}
+%      %child {node {to Paris} edge from parent node[right] {$A$}}
+%      edge from parent node[left] {$H$}
+%    }
+%    child {node (H2)  [fill=black,circle] {}
+%      child {node  {} edge from parent[draw=none]}
+%      child {node  {} edge from parent[draw=none]}
+%      child {node  (Tom) {Tom} edge from parent node[right] {$A$}}
+%      child {node  {moved} edge from parent node[right] {$P$}}
+%      child {node  [fill=black,circle] {}
+%        child {node  {to} edge from parent node[right] {$R$}}
+%        child {node  {NYC} edge from parent node[right] {$C$}}
+%        edge from parent node[right] {$A$}}
+%      %child {node {to Paris} edge from parent node[right] {$A$}}
+%      edge from parent node[left] {$H$}
+%    };
+%    \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+%  \end{tikzpicture}
+%\caption{\label{fig:ucca_example}
+%  Sample UCCA annotation of the sentence ``After graduation, Tom moved
+%  to NYC''. Leaves correspond to words, whereas nodes correspond to units.
+%  Edge labels mark UCCA categories, which are described in the table.}
+%%which are largely ignored in this work
+%\end{figure}
 
@@
 
-\begin{figure*}
-  \begin{adjustbox}{trim=2.3cm 0 0 0}
-   \begin{tikzpicture}
-     [ level distance=12mm,
-	   level 1/.style={sibling distance=7.5em,level distance=7mm},
-       level 2/.style={sibling distance=5em,level distance=12mm},
-       level 3/.style={sibling distance=5em},
-       level 4/.style={sibling distance=3em}]
-     %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
-     %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
-   \node (ROOT) [label={B}] [fill=blue, circle] {}
-   child {node (H3) [label={A}] [fill=yellow,circle] {}
-     child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
-     child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
-       child {node [label={}] {became} edge from parent node[left]  {}}
-       child {node [label={}] {a} edge from parent node[left]  {}}
-       child {node [label={}] {passion} edge from parent node[left]  {}}
-     }
-     child {node (H2) [label={A}] [fill=yellow,circle] {}
-       child {node {} edge from parent[draw=none] {}}
-       child {node {} edge from parent[draw=none] {}}
-       child {node {} edge from parent[draw=none] {}}
-       child {node {} edge from parent[draw=none] {}}
-       child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
-       child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
-       child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
-       child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
-     }
-   }
-   child {node {} edge from parent[draw=none] {}}
-   child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
-     child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
-     child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
-     child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
-     child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
-   }
-   child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
-   child {node (H5) [label={A}] [fill=yellow,circle] {}  
-     child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
-     child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
-   };
-   \draw[dashed] (H5) -> (she) ;
-
-   \com{
-   child {node (H1) [label={\#8}] [fill=black,circle] {}
-     child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
-       %child {node {to Paris} edge from parent node[right] {$A$}}
-       edge from parent node[left] {$H$}
-     }
-     child {node (H2) [label={\#9}] [fill=black,circle] {}
-       child {node  {} edge from parent[draw=none]}
-       child {node  {} edge from parent[draw=none]}
-       child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
-       child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
-       child {node [label={\#10}] [fill=black,circle] {}
-         child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
-         child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
-         edge from parent node[right] {$A$}}
-       %child {node {to Paris} edge from parent node[right] {$A$}}
-       edge from parent node[left] {$H$}
-     };
-     \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
-     }
-   \end{tikzpicture}
-  \\
-  \end{adjustbox}
-    % Ondrej removed the width=\textwidth from the options below to keep font
-	% unenlarged:
-  \adjustbox{center,margin=.5em}{\it His old daughter loved tennis, and took daily lessons however good he was}
-
-  \caption{\label{fig:hume_tree}
-    Sample UCCA annotation of a source sentence, translation (below in italics), 
-    and translation judgements over the source units, marked with A,B,R,O,G labels. Yellow/blue marks
-    nodes annotated with A/B, respectively.
-    In our experiments, the translation is in German, Romaninan, Polish or Czech.
-    ``became a passion'' is marked as atomic, as it is a multi-word expression.
-    The dashed line indicates that ``she'' is also a child of the unit ``greatly improved''.
-    UCCA categories, and alignment between source and translation are omitted for brevity.}
-
-\end{figure*}
+%\begin{figure*}
+%  \begin{adjustbox}{trim=2.3cm 0 0 0}
+%   \begin{tikzpicture}
+%     [level 1/.style={sibling distance=7.5em},
+%       level 2/.style={sibling distance=5em},
+%       level 3/.style={sibling distance=5em},
+%       level 4/.style={sibling distance=3em}]
+%     %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
+%     %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
+%   \node (ROOT) [label={B}] [fill=blue, circle] {}
+%   child {node (H3) [label={A}] [fill=yellow,circle] {}
+%     child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
+%     child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
+%       child {node [label={}] {became} edge from parent node[left]  {}}
+%       child {node [label={}] {a} edge from parent node[left]  {}}
+%       child {node [label={}] {passion} edge from parent node[left]  {}}
+%     }
+%     child {node (H2) [label={A}] [fill=yellow,circle] {}
+%       child {node {} edge from parent[draw=none] {}}
+%       child {node {} edge from parent[draw=none] {}}
+%       child {node {} edge from parent[draw=none] {}}
+%       child {node {} edge from parent[draw=none] {}}
+%       child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
+%       child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
+%       child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
+%       child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
+%     }
+%   }
+%   child {node {} edge from parent[draw=none] {}}
+%   child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
+%     child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
+%     child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
+%     child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
+%     child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
+%   }
+%   child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
+%   child {node (H5) [label={A}] [fill=yellow,circle] {}  
+%     child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
+%     child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
+%   };
+%   \draw[dashed] (H5) -> (she) ;
+% 
+%   \com{
+%   child {node (H1) [label={\#8}] [fill=black,circle] {}
+%     child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
+%       %child {node {to Paris} edge from parent node[right] {$A$}}
+%       edge from parent node[left] {$H$}
+%     }
+%     child {node (H2) [label={\#9}] [fill=black,circle] {}
+%       child {node  {} edge from parent[draw=none]}
+%       child {node  {} edge from parent[draw=none]}
+%       child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
+%       child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
+%       child {node [label={\#10}] [fill=black,circle] {}
+%         child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
+%         child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
+%         edge from parent node[right] {$A$}}
+%       %child {node {to Paris} edge from parent node[right] {$A$}}
+%       edge from parent node[left] {$H$}
+%     };
+%     \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
+%     }
+%   \end{tikzpicture}
+%  \\
+%  \end{adjustbox}
+%%  \adjustbox{width=\textwidth,center,margin=.5em}{\it His old daughter loved tennis, and took daily lessons however good he was}
+%
+%  \caption{\label{fig:hume_tree}
+%    Sample UCCA annotation of a source sentence, translation (below in italics), 
+%    and translation judgements over the source units, marked with A,B,R,O,G labels. Yellow/blue marks
+%    nodes annotated with A/B, respectively.
+%    In our experiments, the translation is in German, Romaninan, Polish or Czech.
+%    ``became a passion'' is marked as atomic, as it is a multi-word expression.
+%    The dashed line indicates that ``she'' is also a child of the unit ``greatly improved''.
+%    UCCA categories, and alignment between source and translation are omitted for brevity.}
+%  
+%\end{figure*}
 
@@
 \end{inparadesc}
-
-It is also relatively common in translation that verbal constructions get
-nominalized or vice versa. An example from our dataset:
-  ``Two {\bf review} authors assessed the studies independently\dots{}'' --
-  ``Zwei Autoren bewertete die Studien unabh\"angig zu {\bf
-  \"uberpr\"ufen}\dots{}'' (lit. ``Two authors assessed the studies
-  independently to be reviewed\dots").\XXX{This is Omri's example, with
-  Ondrej's gloss for German. Ondrej is
-  not totally sure that the meaning is well preserved.}
-
-Here is \XXX{another} example with a missing linker: the source \equo{However,
-this review\dots} was translated as \equo{Diese
-Überprüfung\dots} (lit. \equo{This review}). The omission is not by HMEANT at
-all.
 
@@
 
-\paragraph{Annotation of Predicates}
-
-HUME allows to annotate multi-word and even
-non-contiguous elements as a single unit. HMEANT, on the other hand, always
-required exactly one word to be annotated as the predicate (``Action''), which
-is often linguistically inadequate.
-
-Another problem with HMEANT identified by \perscite{bojar:wu:ssst:2012} was the
-inability to indicate that the predicate is somewhat mistranslated (e.g.
-accidentally negated or in a slightly wrong tense). HMEANT
-annotation allows either to mark the source and reference frames as matching
-(which implies that the corresponding predicates are correct translations of each
-other), or as non-matching (which leads to a zero score for the whole frame) but
-nothing in between. In HUME, the verb is still present as a sub-unit of the
-frame and has the traffic-lights annotation.
-
-\XXX{Omri: I would consider adding further examples from the Google Doc. For instnace
-``your risk of falling'' -- ``Ihr Risiko''.
-Ondrej: I don't see any HMEANT problem here, the whole NP 'your risk of falling'
-is aligned with 'Ihr Risiko' and HMEANT allows to indicate that the match was
-partial, to that's probably what everyone would choose. I've put the other
-example higher, but please have a look, I'm not sure if we should keep it there
-at all.}
+\paragraph{Technical Differences}
+
++ allows for multi-word and even non-contiguous elements
+
++ allows to indicate the quality of the predicates ("Actions")
+
+\oa{I would consider adding further examples from the Google Doc. For instnace ``your risk of falling'' -- ``Ihr Risiko'' or
+  ``Two {\bf review} authors assessed the studies independently and extracted outcome data'' -- ``Zwei Autoren bewertete die Studien unabh\"angig zu {\bf \"uberpr\"ufen}''}
 
@@
 
-\begin{figure*}[ht!]
-\begin{tabular}{cc}
-
-\subfloat[English-Czech]{
-  \iaafig{cs}
-}
-&
-\subfloat[English-German]{
-  \iaafig{de}
-
-}
-\\
-
-\subfloat[English-Polish]{
-  \iaafig{pl}
-
-}
-&
-\subfloat[English-Romanian]{
-  \iaafig{ro}
-
-}
-\end{tabular}
-\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
-\label{fig:heatmap}
-\end{figure*}
+%\begin{figure*}[ht!]
+%\begin{tabular}{cc}
+%
+%
+%\subfloat[English-Czech]{
+%  \iaafig{cs}
+%}
+%&
+%\subfloat[English-German]{
+%  \iaafig{de}
+%
+%}
+%\\
+%
+%\subfloat[English-Polish]{
+%  \iaafig{pl}
+%  
+%}
+%&
+%\subfloat[English-Romanian]{
+%  \iaafig{ro}
+%
+%}
+%\end{tabular}
+%\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
+%\label{fig:heatmap}
+%\end{figure*}
 

### 2016-03-18 scripts/iaa.py
scaling of text

@@ -60,7 +60,11 @@ def print_overall_iaa(by_lang, lang, args):
     if args.create_heatmaps and group == ("A", "B","R", "O", "G"):
       import matplotlib.pyplot as plt
       import seaborn
+      seaborn.set(font_scale=2.2)
+      plt.rcParams.update({'figure.autolayout': True})
       heatmap = seaborn.heatmap(df)
+      heatmap.set_xlabel("Annotator 1")
+      heatmap.set_ylabel("Annotator 2")
       plt.savefig("iaa_heatmap_{}.png".format(lang))
       plt.clf()
 

### 2016-03-18 paper/main.tex

@@
 %   \begin{tikzpicture}
-%     [level 1/.style={sibling distance=7.5em},
-%       level 2/.style={sibling distance=5em},
+%     [ level distance=12mm,
+%	   level 1/.style={sibling distance=7.5em,level distance=7mm},
+%       level 2/.style={sibling distance=5em,level distance=12mm},
 %       level 3/.style={sibling distance=5em},
@@
 %  \end{adjustbox}
-%%  \adjustbox{width=\textwidth,center,margin=.5em}{\it His old daughter loved tennis, and took daily lessons however good he was}
+%    % Ondrej removed the width=\textwidth from the options below to keep font
+%	% unenlarged:
+%  \adjustbox{center,margin=.5em}{\it His old daughter loved tennis, and took daily lessons however good he was}
 %
@@
 \end{inparadesc}
+
+It is also relatively common in translation that verbal constructions get
+nominalized or vice versa. An example from our dataset:
+  ``Two {\bf review} authors assessed the studies independently\dots{}'' --
+  ``Zwei Autoren bewertete die Studien unabh\"angig zu {\bf
+  \"uberpr\"ufen}\dots{}'' (lit. ``Two authors assessed the studies
+  independently to be reviewed\dots").\XXX{This is Omri's example, with
+  Ondrej's gloss for German. Ondrej is
+  not totally sure that the meaning is well preserved.}
+
+Here is \XXX{another} example with a missing linker: the source \equo{However,
+this review\dots} was translated as \equo{Diese
+Überprüfung\dots} (lit. \equo{This review}). The omission is not by HMEANT at
+all.
 
@@
 
-\paragraph{Technical Differences}
-
-+ allows for multi-word and even non-contiguous elements
-
-+ allows to indicate the quality of the predicates ("Actions")
-
-\oa{I would consider adding further examples from the Google Doc. For instnace ``your risk of falling'' -- ``Ihr Risiko'' or
-  ``Two {\bf review} authors assessed the studies independently and extracted outcome data'' -- ``Zwei Autoren bewertete die Studien unabh\"angig zu {\bf \"uberpr\"ufen}''}
+\paragraph{Annotation of Predicates}
+
+HUME allows to annotate multi-word and even
+non-contiguous elements as a single unit. HMEANT, on the other hand, always
+required exactly one word to be annotated as the predicate (``Action''), which
+is often linguistically inadequate.
+
+Another problem with HMEANT identified by \perscite{bojar:wu:ssst:2012} was the
+inability to indicate that the predicate is somewhat mistranslated (e.g.
+accidentally negated or in a slightly wrong tense). HMEANT
+annotation allows either to mark the source and reference frames as matching
+(which implies that the corresponding predicates are correct translations of each
+other), or as non-matching (which leads to a zero score for the whole frame) but
+nothing in between. In HUME, the verb is still present as a sub-unit of the
+frame and has the traffic-lights annotation.
+
+\XXX{Omri: I would consider adding further examples from the Google Doc. For instnace
+``your risk of falling'' -- ``Ihr Risiko''.
+Ondrej: I don't see any HMEANT problem here, the whole NP 'your risk of falling'
+is aligned with 'Ihr Risiko' and HMEANT allows to indicate that the match was
+partial, to that's probably what everyone would choose. I've put the other
+example higher, but please have a look, I'm not sure if we should keep it there
+at all.}
 

### 2016-03-18 paper/main.tex
edits

--- 2016-03-18 paper/main.tex
@@
 this review\dots} was translated as \equo{Diese
-Überprüfung\dots} (lit. \equo{This review}). The omission is not by HMEANT at
+\"Uberpr\"ufung\dots} (lit. \equo{This review}). The omission is not by HMEANT at
 all.
@@
 annotations also at the alignment stage
-to avoid unduly penalization due to such a spurious diverge.
+to avoid unduly penalization due to such a spurious divergence.
 The same issue was observed by HMEANT authors themselves

### 2016-03-18 scripts/iaa_binned.py
reinstate heatmap, figure scale

@@ -71,12 +71,12 @@ def do_bins(agree, bins, field_name, csv_file, graph_file, name):
         offset += width
         for rect, count in zip(rects[-1], counts[group]):
           ax.text(rect.get_x() + rect.get_width()/2, rect.get_height(), str(count), 
-              ha='center', va='bottom')
+              ha='center', va='bottom', size='small')
       ax.set_yticks(np.arange(0,1.1,0.1))
       ax.set_xticks(index + 0.5)
       ax.set_xlim(0,len(bins))
       ax.set_xticklabels(["{}-{}".format(n,m) for n,m in bins])
-      lgd = ax.legend([r[0] for r in rects],[g for g,_ in groups], loc='upper left')#, bbox_to_anchor=(1.0,1.1))
+      lgd = ax.legend([r[0] for r in rects],[g for g,_ in groups], loc='upper left', borderpad=0.2)#, bbox_to_anchor=(1.0,1.1))
       #ax.set_title("Kappa by {}: {}".format(name, lang_name))
       ax.set_ylabel("Kappa")
       ax.set_xlabel(name)
@@ -110,7 +110,8 @@ def main():
 
 
   # font size
-  plt.rcParams.update({'font.size': 18})
+  plt.rcParams.update({'font.size': 22})
+  plt.rcParams.update({'figure.autolayout': True})
   
   sentences = pandas.read_csv(args.sentence_file)
   nodes = pandas.read_csv(args.node_file, converters={'node_id': str, 'parent' : str})

### 2016-03-18 paper/main.tex
edits
reinstate heatmap, figure scale

@@
 \newcommand{\tabref}[1]{Table~\ref{#1}}
-\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
+%\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
+\newcommand{\XXX}[1]{}
 %\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
@@
 
-Human evaluation of machine translation is a cognitively difficult task and when annotators rate or rank complete sentences we see low levels of agreement especially
+Human evaluation of machine translation is a cognitively difficult task and when
+annotators rate or rank complete sentences, we see low levels of agreement especially
 for longer sentences.  
-In order to extract more reliable annotations, previous research broke down evaluation into smaller components by asking humans to annotate semantic roles on the source and translation, and then to align them.
-We propose a new evaluation method which is based on semantic annotation called Universal Conceptual Cognitive Annotation(UCCA).
+%In order to extract more reliable annotations, previous research broke down evaluation into smaller components by asking humans to annotate semantic roles on the source and translation, and then to align them.
+In order to extract more reliable annotations, previous research broke down
+evaluation into smaller components in various ways. 
+We propose a new evaluation method
+HUME
+which is based on semantic annotation called Universal Conceptual Cognitive Annotation (UCCA).
 UCCA is applicable across many languages, it provides a complete structure over sentences, and it is grounded in the words in the sentence.
-We apply UCCA only over the source or reference sentence, and not over the possibly garbled translation, 
-and this makes annotation more reliable and also efficient as we can reuse labor intensive semantic annotations.
-We show that HUME can be reliably applied to four different target languages and that sentence level inter annotor agreement does not decrease over longer sentences.
+We apply UCCA only over the source or reference sentence, and not over the
+possibly garbled translation.
+%, 
+%and this makes annotation more reliable and also efficient as we can reuse labor
+%intensive semantic annotations.
+We show that HUME can be reliably applied to
+four different target languages and that sentence level inter-annotator agreement
+does not decrease over longer sentences.
 
@@
 
-\paragraph{Annotation of Predicates}
+\paragraph{Annotation of Predicates Themselves}
 
@@
 
-\def\iaafig #1{\includegraphics[width=7cm]{iaa_length_#1.png}}
-
-\begin{figure*}[ht!]
+\def\iaafig #1{\includegraphics[width=3.8cm]{iaa_length_#1.png}}
+
+\begin{figure}[ht!]
+\renewcommand{\tabcolsep}{0pt}
 \begin{tabular}{cc}
@@
 \label{fig:iaalength}
-\end{figure*}
-
-\def\iaafig #1{\includegraphics[width=7cm]{iaa_heatmap_#1.png}}
-
-%\begin{figure*}[ht!]
-%\begin{tabular}{cc}
-%
-%
-%\subfloat[English-Czech]{
-%  \iaafig{cs}
-%}
-%&
-%\subfloat[English-German]{
-%  \iaafig{de}
-%
-%}
-%\\
-%
-%\subfloat[English-Polish]{
-%  \iaafig{pl}
-%  
-%}
-%&
-%\subfloat[English-Romanian]{
-%  \iaafig{ro}
-%
-%}
-%\end{tabular}
-%\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
-%\label{fig:heatmap}
-%\end{figure*}
+\end{figure}
 
@@
 one annotator marked correct, and the other (a Cochrane employee)  as incorrect.
+
+\def\iaafig #1{\includegraphics[width=3.8cm]{iaa_heatmap_#1.png}}
+
+\begin{figure}[ht!]
+\renewcommand{\tabcolsep}{0pt}
+\begin{tabular}{cc}
+
+\subfloat[English-Czech]{
+  \iaafig{cs}
+}
+&
+\subfloat[English-German]{
+  \iaafig{de}
+
+}
+\\
+
+\subfloat[English-Polish]{
+  \iaafig{pl}
+
+}
+&
+\subfloat[English-Romanian]{
+  \iaafig{ro}
+
+}
+\end{tabular}
+\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
+\label{fig:heatmap}
+\end{figure}
 

### 2016-03-18 paper/main.tex
edits
removed larger image

--- 2016-03-18 paper/main.tex
@@
 
-\begin{figure*}
-  \begin{adjustbox}{trim=2.3cm 0 0 0}
-   \begin{tikzpicture}
-     [ level distance=12mm,
-	   level 1/.style={sibling distance=7.5em,level distance=7mm},
-       level 2/.style={sibling distance=5em,level distance=12mm},
-       level 3/.style={sibling distance=5em},
-       level 4/.style={sibling distance=3em}]
-     %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
-     %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
-   \node (ROOT) [label={B}] [fill=blue, circle] {}
-   child {node (H3) [label={A}] [fill=yellow,circle] {}
-     child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
-     child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
-       child {node [label={}] {became} edge from parent node[left]  {}}
-       child {node [label={}] {a} edge from parent node[left]  {}}
-       child {node [label={}] {passion} edge from parent node[left]  {}}
-     }
-     child {node (H2) [label={A}] [fill=yellow,circle] {}
-       child {node {} edge from parent[draw=none] {}}
-       child {node {} edge from parent[draw=none] {}}
-       child {node {} edge from parent[draw=none] {}}
-       child {node {} edge from parent[draw=none] {}}
-       child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
-       child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
-       child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
-       child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
-     }
-   }
-   child {node {} edge from parent[draw=none] {}}
-   child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
-     child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
-     child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
-     child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
-     child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
-   }
-   child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
-   child {node (H5) [label={A}] [fill=yellow,circle] {}  
-     child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
-     child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
-   };
-   \draw[dashed] (H5) -> (she) ;
-
-   \com{
-   child {node (H1) [label={\#8}] [fill=black,circle] {}
-     child {node [label={\#2}] {graduation} edge from parent node[right] {$P$}}
-       %child {node {to Paris} edge from parent node[right] {$A$}}
-       edge from parent node[left] {$H$}
-     }
-     child {node (H2) [label={\#9}] [fill=black,circle] {}
-       child {node  {} edge from parent[draw=none]}
-       child {node  {} edge from parent[draw=none]}
-       child {node [label={\#3}] (Tom) {Tom} edge from parent node[right] {$A$}}
-       child {node [label={\#4}] {moved} edge from parent node[right] {$P$}}
-       child {node [label={\#10}] [fill=black,circle] {}
-         child {node [label={\#5}] {to} edge from parent node[right] {$R$}}
-         child {node [label={\#6}] {NYC} edge from parent node[right] {$C$}}
-         edge from parent node[right] {$A$}}
-       %child {node {to Paris} edge from parent node[right] {$A$}}
-       edge from parent node[left] {$H$}
-     };
-     \draw (H1) -> (Tom) node[midway, right] {$A$ \hspace{0.2cm}};
-     }
-   \end{tikzpicture}
-  \\
-  \end{adjustbox}
-    % Ondrej removed the width=\textwidth from the options below to keep font
-	% unenlarged:
-  \adjustbox{center,margin=.5em}{\it His old daughter loved tennis, and took daily lessons however good he was}
-
-  \caption{\label{fig:hume_tree}
-    Sample UCCA annotation of a source sentence, translation (below in italics), 
-    and translation judgements over the source units, marked with A,B,R,O,G labels. Yellow/blue marks
-    nodes annotated with A/B, respectively.
-    In our experiments, the translation is in German, Romaninan, Polish or Czech.
-    ``became a passion'' is marked as atomic, as it is a multi-word expression.
-    The dashed line indicates that ``she'' is also a child of the unit ``greatly improved''.
-    UCCA categories, and alignment between source and translation are omitted for brevity.}
-
-\end{figure*}
-
 \paragraph{Translation Evaluation.}
@@
 this review\dots} was translated as \equo{Diese
-\"Uberpr\"ufung\dots} (lit. \equo{This review}). The omission is not by HMEANT at
+Überprüfung\dots} (lit. \equo{This review}). The omission is not by HMEANT at
 all.
@@
 annotations also at the alignment stage
-to avoid unduly penalization due to such a spurious divergence.
+to avoid unduly penalization due to such a spurious diverge.
 The same issue was observed by HMEANT authors themselves

### 2016-03-18 paper/main.tex
edits

--- 2016-03-18 paper/main.tex
@@
 \newcommand{\tabref}[1]{Table~\ref{#1}}
-%\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
-\newcommand{\XXX}[1]{}
+\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
 %\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
@@
 to assign a single number that quantifies the quality of the entire sentence. This
-is indeed reflected in the diminishing inter-annotator agreement rates of human ranking measures
-as a function of the sentence length \cite{Bojar:2011}.
+is indeed reflected in the diminishing inter-annotator agreement (IAA) rates of human ranking measures
+%as a function of
+with
+the sentence length \cite{Bojar:2011}.
 Second, a sentence-level quality score does not indicate what parts of the sentence
@@
 
-These problems were partially addressed by measures that decompose over sub-parts of the evaluated
-translation sentence (henceforth, {\it translation}),
-most commonly over its words or n-grams,
-quantifying the overlap of its sub-parts relative a reference translation.
-More recent work on MT evaluation proposed measures
-that decompose over semantically defined units,
+These problems are partially addressed by measures that decompose over sub-parts of the evaluated
+translation sentence (henceforth, {\it translation}). For automatic measures,
+these are commonly words or n-grams, for manual measures some structural
+information is taken into account \parcite{machacek:bojar:segranks:2015}.
+% the following is specific to automatic measures:
+%quantifying the overlap of its sub-parts relative a reference translation.
+A promising line of research decomposes
+%More recent work on MT evaluation proposed measures
+%that decompose
+over semantically defined units,
 quantifying the similarity of the output and the reference in terms of
@@
 HMEANT \cite{lo2011structured}.
+\XXX{Do we want to cut the following shorter?}
 These measures can localize semantic errors in the translation, but
@@
 
-\def\iaafig #1{\includegraphics[width=3.8cm]{iaa_length_#1.png}}
-
-\begin{figure}[ht!]
-\renewcommand{\tabcolsep}{0pt}
+\def\iaafig #1{\includegraphics[width=7cm]{iaa_length_#1.png}}
+
+\begin{figure*}[ht!]
 \begin{tabular}{cc}
@@
 \label{fig:iaalength}
-\end{figure}
+\end{figure*}
+
+\def\iaafig #1{\includegraphics[width=7cm]{iaa_heatmap_#1.png}}
+
+%\begin{figure*}[ht!]
+%\begin{tabular}{cc}
+%
+%
+%\subfloat[English-Czech]{
+%  \iaafig{cs}
+%}
+%&
+%\subfloat[English-German]{
+%  \iaafig{de}
+%
+%}
+%\\
+%
+%\subfloat[English-Polish]{
+%  \iaafig{pl}
+%  
+%}
+%&
+%\subfloat[English-Romanian]{
+%  \iaafig{ro}
+%
+%}
+%\end{tabular}
+%\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
+%\label{fig:heatmap}
+%\end{figure*}
 
@@
 one annotator marked correct, and the other (a Cochrane employee)  as incorrect.
-
-\def\iaafig #1{\includegraphics[width=3.8cm]{iaa_heatmap_#1.png}}
-
-\begin{figure}[ht!]
-\renewcommand{\tabcolsep}{0pt}
-\begin{tabular}{cc}
-
-\subfloat[English-Czech]{
-  \iaafig{cs}
-}
-&
-\subfloat[English-German]{
-  \iaafig{de}
-
-}
-\\
-
-\subfloat[English-Polish]{
-  \iaafig{pl}
-
-}
-&
-\subfloat[English-Romanian]{
-  \iaafig{ro}
-
-}
-\end{tabular}
-\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
-\label{fig:heatmap}
-\end{figure}
 

### 2016-03-18 paper/main.tex
edits

@@
 \newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
-%\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
-\newcommand{\oa}[1]{}
-%\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
-\newcommand{\bh}[1]{}
+%\newcommand{\XXX}[1]{}
+\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
+%\newcommand{\oa}[1]{}
+\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
+%\newcommand{\bh}[1]{}
 
@@
 that the evaluator be proficient in both the source and target languages.
-An elaborate comparision of HUME and HMEANT is given in \secref{sec:hmeant_comp}.
-
-We conduct experiments with four language pairs: English-German, English-Polish,
-English-Romanian and English-Czech, and report efficiency (time of annotation) and
-inter-annotator agreement scores (\secref{sec:experiments}). Efficiency results that HUME
-can be annotated in 2--4 minutes, once the UCCA annotation of the source is given,
-which is only about five times slower than MEANT, and hence still practical to
-employ on a moderate scale. HUME's agreement results are on par with those
-of standard sentence-level ranking measures, but unlike them,
-HUME's scores do not decrease with longer sentences.\bh{We may
-  have to focus more on qualitative aspects, if we just try to sell
-  the measure on IAA, then the figures are not stunning. OA: is that OK?}
+%An elaborate comparison of HUME and HMEANT is given in \secref{sec:hmeant_comp}.
+
+After a brief review (\secref{sec:background}), we describe HUME in detail
+(\secref{sec:hume}) and highlight its benefits over HMEANT
+(\secref{sec:hmeant_comp}).
+Our experiments with four language pairs: English-German, English-Polish,
+English-Romanian and English-Czech (\secref{sec:experiments}) document the
+efficiency (time of annotation) and inter-annotator agreement scores for HUME.
+
+% We conduct experiments with four language pairs: English-German, English-Polish,
+% English-Romanian and English-Czech, and report efficiency (time of annotation) and
+% inter-annotator agreement scores (\secref{sec:experiments}).
+% 
+% \XXX{The following should go rather to the conclusion}
+% Efficiency results that HUME
+% can be annotated in 2--4 minutes, once the UCCA annotation of the source is given,
+% which is only about five times slower than MEANT, and hence still practical to
+% employ on a moderate scale. HUME's agreement results are on par with those
+% of standard sentence-level ranking measures, but unlike them,
+% HUME's scores do not decrease with longer sentences.\bh{We may
+%   have to focus more on qualitative aspects, if we just try to sell
+%   the measure on IAA, then the figures are not stunning. OA: is that OK?}
 

### 2016-03-18 paper/supp.tex

% File ucca_parsing.tex
%

\documentclass[11pt]{article}
\usepackage{colacl-onecolumn}
\usepackage{times}
\usepackage{url}
\usepackage{amsmath}
\usepackage{breqn}
\usepackage{latexsym}
\usepackage{pgfplotstable}
\usepackage{algorithm2e}
\usepackage{hhline}
\usepackage{multirow}
\usepackage[font=small]{caption}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{color}
\usepackage{lipsum,adjustbox}
\usepackage{tikz}
\usepackage{tikz-dependency}
\usepackage{rotating}
\usepackage{float}
\usetikzlibrary{shapes,fit,calc,er,positioning,intersections,decorations.shapes,mindmap,trees}
\tikzset{decorate sep/.style 2 args={decorate,decoration={shape backgrounds,shape=circle,
      shape size=#1,shape sep=#2}}}

\flushbottom \onecolumn

\newcommand{\com}[1]{}
\newcommand{\secref}[1]{Section~\ref{#1}}
\newcommand{\figref}[1]{Figure~\ref{#1}}
\newcommand{\tabref}[1]{Table~\ref{#1}}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}
\SetKwRepeat{Do}{do}{while}
\renewcommand\AlCapFnt{\normalfont\small}

\makeatletter
\renewcommand{\paragraph}{
  \@startsection{paragraph}{4}
  {\z@}{.5ex \@plus .5ex \@minus .2ex}{-1em}
  {\normalfont\normalsize\bfseries}
}
\makeatother

\newcommand\BibTeX{B\textsc{ib}\TeX}

\title{HUME: Human UCCA-Based Evaluation of Machine Translation \\
  Supplementary Matrerial}

\begin{document}
\maketitle

\section{Another Example of HUME Annotation}

Sample UCCA annotation of a source sentence, translation (below in italics), 
and translation judgements over the source units, marked with A,B,R,O,G labels. Yellow/blue marks nodes annotated with A/B, respectively.
In the experiments reported in the paper,
the translation is in German, Romaninan, Polish or Czech.
``became a passion'' is marked as atomic, as it is a multi-word expression.
The dashed line indicates that ``she'' is also a child of the unit ``greatly improved''.
UCCA categories, and alignment between source and translation are omitted for brevity.

\begin{figure*}[h]
  \begin{adjustbox}{trim=2.3cm 0 0 0}
   \begin{tikzpicture}
     [ level distance=12mm,
	   level 1/.style={sibling distance=7.5em,level distance=7mm},
       level 2/.style={sibling distance=5em,level distance=12mm},
       level 3/.style={sibling distance=5em},
       level 4/.style={sibling distance=3em}]
     %[semithick, x=2.4cm,y=0.8cm, >=latex, baseline=7ex,
     %  inner sep=.3ex, outer sep=.7ex, minimum size=1.2ex,->]
   \node (ROOT) [label={B}] [fill=blue, circle] {}
   child {node (H3) [label={A}] [fill=yellow,circle] {}
     child {node [label={[red]:R}] {Golf} edge from parent node[left] {}}
     child {node (H1) [label={[orange]30:O}] [fill=orange,circle] {}
       child {node [label={}] {became} edge from parent node[left]  {}}
       child {node [label={}] {a} edge from parent node[left]  {}}
       child {node [label={}] {passion} edge from parent node[left]  {}}
     }
     child {node (H2) [label={A}] [fill=yellow,circle] {}
       child {node {} edge from parent[draw=none] {}}
       child {node {} edge from parent[draw=none] {}}
       child {node {} edge from parent[draw=none] {}}
       child {node {} edge from parent[draw=none] {}}
       child {node [label={[red]:R}] {for} edge from parent node[left]  {}}
       child {node [label={[green]:G}] {his} edge from parent node[left]  {}}
       child {node [label={[orange]:O}] {oldest} edge from parent node[left]  {}}
       child {node [label={[green]:G}] {daughter} edge from parent node[left]  {}}
     }
   }
   child {node {} edge from parent[draw=none] {}}
   child {node (H4) [label={[]30:A}] [fill=yellow,circle] {}
     child {node (she) [label={[green]:G}] {she} edge from parent node[left]  {}}
     child {node [label={[green]:G}] {attended} edge from parent node[left]  {}}
     child {node [label={[green]:G}] {daily} edge from parent node[left]  {}}
     child {node [label={[green]:G}] {lessons} edge from parent node[left]  {}}
   }
   child {node [label={[red]:R}] {and} edge from parent node[left]  {}}
   child {node (H5) [label={A}] [fill=yellow,circle] {}  
     child {node [label={[red]:R}] {greatly} edge from parent node[left] {}}
     child {node [label={[orange]:O}] {improved} edge from parent node[left] {}}
   };
   \draw[dashed] (H5) -> (she) ;
   \end{tikzpicture}
   \\
  \end{adjustbox}
    % Ondrej removed the width=\textwidth from the options below to keep font
	% unenlarged:
  \adjustbox{center,margin=.5em}{\it His old daughter loved tennis, and took daily lessons however good he was}

  \label{fig:hume_tree}

\end{figure*}

\end{document}

### 2016-03-18 paper/main.tex
edits

@@
 by \newcite{graham2015improving}.
-However, while providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
-and is infeasible where a wide variety of systems or parameter settings are to be evaluated.
+However, while providing the gold standard for MT evaluation, human evaluation
+is not a scalable solution.
+%and is infeasible where a wide variety of systems or parameter settings are to be evaluated.
 % has motivated the construction of cheaper methods that correlate with human
@@
 Scalability is addressed by employing automatic and semi-automatic approximations of human
-judgements. Most commonly, such scores decompose over the sub-parts of
+judgements. Commonly, such scores decompose over the sub-parts of
 the translation, and quantify how many of these sub-parts appear in a manually created reference translation.

### 2016-03-18 paper/main.tex
abstract

--- 2016-03-18 paper/main.tex
@@
 evaluation into smaller components in various ways. 
-We propose a new evaluation method
-HUME
-which is based on semantic annotation called Universal Conceptual Cognitive Annotation (UCCA).
-UCCA is applicable across many languages, it provides a complete structure over sentences, and it is grounded in the words in the sentence.
+We propose a new evaluation method, HUME, 
+which is based on the Universal Conceptual Cognitive Annotation (UCCA)
+semantic annotation.
+UCCA is applicable across many languages, it has better coverage of important
+semantic phenomena than previously used semantic annotation,
+semantic role labeling, and it is grounded in the words in the sentence. 
 We apply UCCA only over the source or reference sentence, and not over the
@@
 We show that HUME can be reliably applied to
-four different target languages and that sentence level inter-annotator agreement
+four different target languages and that sentence-level inter-annotator agreement
 does not decrease over longer sentences.

### 2016-03-18 paper/main.tex
edits

--- 2016-03-18 paper/main.tex
@@
 The most basic notion is the Scene, which describes a movement, an
-action or a state which is persistent in time. Each Scene contains one main relation, as well
-as one or more participants. For example, the sentence ``After graduation, Tom moved to NYC''
+action or a state which is persistent in time. Each Scene contains one main
+relation and zero or more participants. For example, the sentence ``After graduation, Tom moved to NYC''
 contains two Scenes, whose main relations are ``graduation'' and ``moved''.
 The participant ``Tom'' is a part of both Scenes, while ``NYC'' only of the
-latter (\figref{fig:ucca_example}). Further categories account for
+latter (\figref{fig:ucca_example_v2}). Further categories account for
 relations between Scenes and the internal structures of complex participants and relations.

### 2016-03-18 paper/main.tex
New abstract trying to get best of both previous abstracts

--- 2016-03-18 paper/main.tex
@@
 \newcommand{\tabref}[1]{Table~\ref{#1}}
-\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
-%\newcommand{\XXX}[1]{}
-\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
-%\newcommand{\oa}[1]{}
-\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
-%\newcommand{\bh}[1]{}
+%\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
+\newcommand{\XXX}[1]{}
+%\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
+\newcommand{\oa}[1]{}
+%\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
+\newcommand{\bh}[1]{}
 
@@
 
-Human evaluation of machine translation is a cognitively difficult task and when
-annotators rate or rank complete sentences, we see low levels of agreement especially
-for longer sentences.  
-%In order to extract more reliable annotations, previous research broke down evaluation into smaller components by asking humans to annotate semantic roles on the source and translation, and then to align them.
-In order to extract more reliable annotations, previous research broke down
-evaluation into smaller components in various ways. 
-We propose a new evaluation method, HUME, 
-which is based on the Universal Conceptual Cognitive Annotation (UCCA)
-semantic annotation.
-UCCA is applicable across many languages, it has better coverage of important
-semantic phenomena than previously used semantic annotation,
-semantic role labeling, and it is grounded in the words in the sentence. 
-We apply UCCA only over the source or reference sentence, and not over the
-possibly garbled translation.
+Human evaluation of machine translation normally uses sentence-level measures such as relative ranking
+or adequacy/fluency scales. However these provide no insight into the MT errors,
+ and do not scale well with sentence 
+length. We argue for a semantics-based evaluation, which aims to capture how much of the meaning
+is retained in the MT output, and can support the construction and tuning of semantics-based MT. 
+We present a novel human semantic evaluation measure for MT, Human
+UCCA-based MT Evaluation (HUME), building on the Universal Conceptual Cognitive Annotation (UCCA) semantic representation scheme.
+Making progress over previous work, the proposed measure takes into account
+a wider range of semantic phenomena and does not rely on semantic annotation
+of the MT output itself, whose interpretation is often unclear.
+We experiment with four language pairs, demonstrating HUME's broad applicability,
+as well as its reliability, reflected in its inter-annotator agreement rates.
+
+%Human evaluation of machine translation is a cognitively difficult task and when
+%annotators rate or rank complete sentences, we see low levels of agreement especially
+%for longer sentences.  
+%In order to extract more reliable annotations, previous research broke down evaluation into smaller 
+%components by asking humans to annotate semantic roles on the source and translation, and then to align them. 
+%In order to extract more reliable annotations, previous research broke down
+%evaluation into smaller components in various ways. 
+%We propose a new evaluation method
+%HUME
+%which is based on semantic annotation called Universal Conceptual Cognitive Annotation (UCCA). UCCA
+%is applicable across many languages, it provides a complete structure over sentences, and it is 
+%grounded in the words in the sentence. 
+%We apply UCCA only over the source or reference sentence, and not over the
+%possibly garbled translation.
 %, 
@@
 %intensive semantic annotations.
-We show that HUME can be reliably applied to
-four different target languages and that sentence-level inter-annotator agreement
-does not decrease over longer sentences.
+%We show that HUME can be reliably applied to
+%four different target languages and that sentence level inter-annotator agreement
+%does not decrease over longer sentences.
 
@@
 to assign a single number that quantifies the quality of the entire sentence. This
-is indeed reflected in the diminishing inter-annotator agreement (IAA) rates of human ranking measures
-%as a function of
-with
-the sentence length \cite{Bojar:2011}.
+is indeed reflected in the diminishing inter-annotator agreement rates of human ranking measures
+as a function of the sentence length \cite{Bojar:2011}.
 Second, a sentence-level quality score does not indicate what parts of the sentence
@@
 
-These problems are partially addressed by measures that decompose over sub-parts of the evaluated
-translation sentence (henceforth, {\it translation}). For automatic measures,
-these are commonly words or n-grams, for manual measures some structural
-information is taken into account \parcite{machacek:bojar:segranks:2015}.
-% the following is specific to automatic measures:
-%quantifying the overlap of its sub-parts relative a reference translation.
-A promising line of research decomposes
-%More recent work on MT evaluation proposed measures
-%that decompose
-over semantically defined units,
+These problems were partially addressed by measures that decompose over sub-parts of the evaluated
+translation sentence (henceforth, {\it translation}),
+most commonly over its words or n-grams,
+quantifying the overlap of its sub-parts relative a reference translation.
+More recent work on MT evaluation proposed measures
+that decompose over semantically defined units,
 quantifying the similarity of the output and the reference in terms of
@@
 HMEANT \cite{lo2011structured}.
-\XXX{Do we want to cut the following shorter?}
 These measures can localize semantic errors in the translation, but
@@
 that the evaluator be proficient in both the source and target languages.
-%An elaborate comparison of HUME and HMEANT is given in \secref{sec:hmeant_comp}.
-
-After a brief review (\secref{sec:background}), we describe HUME in detail
-(\secref{sec:hume}) and highlight its benefits over HMEANT
-(\secref{sec:hmeant_comp}).
-Our experiments with four language pairs: English-German, English-Polish,
-English-Romanian and English-Czech (\secref{sec:experiments}) document the
-efficiency (time of annotation) and inter-annotator agreement scores for HUME.
-
-% We conduct experiments with four language pairs: English-German, English-Polish,
-% English-Romanian and English-Czech, and report efficiency (time of annotation) and
-% inter-annotator agreement scores (\secref{sec:experiments}).
-% 
-% \XXX{The following should go rather to the conclusion}
-% Efficiency results that HUME
-% can be annotated in 2--4 minutes, once the UCCA annotation of the source is given,
-% which is only about five times slower than MEANT, and hence still practical to
-% employ on a moderate scale. HUME's agreement results are on par with those
-% of standard sentence-level ranking measures, but unlike them,
-% HUME's scores do not decrease with longer sentences.\bh{We may
-%   have to focus more on qualitative aspects, if we just try to sell
-%   the measure on IAA, then the figures are not stunning. OA: is that OK?}
+An elaborate comparision of HUME and HMEANT is given in \secref{sec:hmeant_comp}.
+
+We conduct experiments with four language pairs: English-German, English-Polish,
+English-Romanian and English-Czech, and report efficiency (time of annotation) and
+inter-annotator agreement scores (\secref{sec:experiments}). Efficiency results that HUME
+can be annotated in 2--4 minutes, once the UCCA annotation of the source is given,
+which is only about five times slower than MEANT, and hence still practical to
+employ on a moderate scale. HUME's agreement results are on par with those
+of standard sentence-level ranking measures, but unlike them,
+HUME's scores do not decrease with longer sentences.\bh{We may
+  have to focus more on qualitative aspects, if we just try to sell
+  the measure on IAA, then the figures are not stunning. OA: is that OK?}
 
@@
 by \newcite{graham2015improving}.
-However, while providing the gold standard for MT evaluation, human evaluation
-is not a scalable solution.
-%and is infeasible where a wide variety of systems or parameter settings are to be evaluated.
+However, while providing the gold standard for MT evaluation, human evaluation is not a scalable solution,
+and is infeasible where a wide variety of systems or parameter settings are to be evaluated.
 % has motivated the construction of cheaper methods that correlate with human
@@
 Scalability is addressed by employing automatic and semi-automatic approximations of human
-judgements. Commonly, such scores decompose over the sub-parts of
+judgements. Most commonly, such scores decompose over the sub-parts of
 the translation, and quantify how many of these sub-parts appear in a manually created reference translation.

### 2016-03-18 paper/main.tex

--- 2016-03-18 paper/main.tex
@@
 
-Human evaluation of machine translation is a cognitively difficult task and when
-annotators rate or rank complete sentences, we see low levels of agreement especially
-for longer sentences.  
-%In order to extract more reliable annotations, previous research broke down evaluation into smaller components by asking humans to annotate semantic roles on the source and translation, and then to align them.
-In order to extract more reliable annotations, previous research broke down
-evaluation into smaller components in various ways. 
-We propose a new evaluation method
-HUME
-which is based on semantic annotation called Universal Conceptual Cognitive Annotation (UCCA).
-UCCA is applicable across many languages, it provides a complete structure over sentences, and it is grounded in the words in the sentence.
-We apply UCCA only over the source or reference sentence, and not over the
-possibly garbled translation.
+Human evaluation of machine translation normally uses sentence-level measures such as relative ranking
+or adequacy/fluency scales. However these provide no insight into the MT errors,
+ and do not scale well with sentence 
+length. We argue for a semantics-based evaluation, which aims to capture how much of the meaning
+is retained in the MT output, and can support the construction and tuning of semantics-based MT. 
+We present a novel human semantic evaluation measure for MT, Human
+UCCA-based MT Evaluation (HUME), building on the Universal Conceptual Cognitive Annotation (UCCA) semantic representation scheme.
+Making progress over previous work, the proposed measure takes into account
+a wider range of semantic phenomena and does not rely on semantic annotation
+of the MT output itself, whose interpretation is often unclear.
+We experiment with four language pairs, demonstrating HUME's broad applicability,
+as well as its reliability, reflected in its inter-annotator agreement rates.
+
+%Human evaluation of machine translation is a cognitively difficult task and when
+%annotators rate or rank complete sentences, we see low levels of agreement especially
+%for longer sentences.  
+%In order to extract more reliable annotations, previous research broke down evaluation into smaller 
+%components by asking humans to annotate semantic roles on the source and translation, and then to align them. 
+%In order to extract more reliable annotations, previous research broke down
+%evaluation into smaller components in various ways. 
+%We propose a new evaluation method
+%HUME
+%which is based on semantic annotation called Universal Conceptual Cognitive Annotation (UCCA). UCCA
+%is applicable across many languages, it provides a complete structure over sentences, and it is 
+%grounded in the words in the sentence. 
+%We apply UCCA only over the source or reference sentence, and not over the
+%possibly garbled translation.
 %, 
@@
 %intensive semantic annotations.
-We show that HUME can be reliably applied to
-four different target languages and that sentence level inter-annotator agreement
-does not decrease over longer sentences.
+%We show that HUME can be reliably applied to
+%four different target languages and that sentence level inter-annotator agreement
+%does not decrease over longer sentences.
 

### 2016-03-18 paper/main.tex
shortened section 1

@@
 
-Human judgement is the central, and perhaps most fundamental, criterion for
+Human judgement is the most important criterion for
 estimating the quality of an MT system.
-Nevertheless, common measures for MT evaluation, such as adequacy and fluency judgements
-or the relative ranking of several possible translations, are problematic in two ways.
+Nevertheless, common measures for human MT evaluation, such as adequacy and fluency judgements
+or the relative ranking of possible translations, are problematic in two ways.
 First, as the quality of translation is determined by multiple factors, it is difficult
@@
 Second, a sentence-level quality score does not indicate what parts of the sentence
-are erroneously translated, and hence cannot inform developers in overcoming these errors.
-
-These problems are partially addressed by measures that decompose over sub-parts of the evaluated
-translation sentence (henceforth, {\it translation}). For automatic measures,
+are badly translated, and so cannot inform developers in repairing these errors.
+
+These problems are partially addressed by measures that decompose over parts of the evaluated
+translation. For automatic measures,
 these are commonly words or n-grams, for manual measures some structural
@@
 \XXX{Do we want to cut the following shorter?}
-These measures can localize semantic errors in the translation, but
-their focus on verb argument structures misses many pervasive semantic phenomena
+However, 
+their focus on verb argument structures misses many semantic phenomena
 (such as inter-clause linkage and nominal argument structures) and thus they only provide
-a limited perspective on the semantic similarity between the output and reference.
-Moreover, assigning semantic structure to MT output, especially when low in quality,
+a limited perspective on the semantic similarity between the translation and reference.
+Also, assigning semantic structure to an MT output, especially a poor one,
 is difficult and arguably ill-defined.
@@
 of many fundamental semantic phenomena, such as verbal, nominal and adjectival
-argument structures and the inter-relations between them.
+argument structures and their inter-relations.
 
@@
 semantic units in the source sentence.
-Importantly, HUME only requires semantically annotating the source sentence,
+HUME only requires semantically annotating the source sentence,
 thus avoiding the annotation of machine-generated text 
@@
 By basing the measure on a direct comparison
-of the source and translation we avoid the difficulties engendered by using
-reference translations, albeit at the cost of imposing the additional requirement
-that the evaluator be proficient in both the source and target languages.
+of the source and translation we avoid the difficulties created by using
+reference translations, albeit at the cost of requiring the
+evaluator to be proficient in both the source and target languages.
 %An elaborate comparison of HUME and HMEANT is given in \secref{sec:hmeant_comp}.
@@
 (\secref{sec:hmeant_comp}).
-Our experiments with four language pairs: English-German, English-Polish,
-English-Romanian and English-Czech (\secref{sec:experiments}) document the
+Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments}) document the
 efficiency (time of annotation) and inter-annotator agreement scores for HUME.
@@
 
-\section{Conclusion}
-
 %\section*{Acknowledgments}

### 2016-03-18 paper/main.tex
edits

@@
 to compute the HUME measure. 
-We mark the source sentence as $s$ and the translation sentence as $t$. 
+We denote the source sentence as $s$ and the translation sentence as $t$. 
 The procedure involves two manual steps: (1) UCCA-annotating $s$, 
@@
 the evaluation could be based on the UCCA structure for the \emph{reference}
-translation. This would allow monolingual annotators to take part in the
-evaluation. On the downside, the translation would have to be aligned to the
-reference and more importantly, we would risk inaccurate judgements due to
-the naturally occurring differences between the source and its reference
-translations.}.
+translation.
+More details in the discussion in \secref{src-vs-ref}.
+% This would allow monolingual annotators to take part in the
+% evaluation. On the downside, the translation would have to be aligned to the
+% reference and more importantly, we would risk inaccurate judgements due to
+% the naturally occurring differences between the source and its reference
+% translations.
+}.
 
@@
 (this holds especially for the verb and its modifiers,
-less so for components of noun phrases). Since errors in morphology
-should be expressed on the atomic units, we would expect
-a higher proportion of ``Adequate'' structural units for Czech and Polish
-compared to German and Romanian\oa{but doesn't German also have a case system? Ondrej:
-  could you pinpoint what the difference exactly is?}.
-
-An example HUME annotation is given in \figref{fig:hume_tree},
-which presents an annotation of an UCCA tree, and a corresponding translation.
-We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
-in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
-individually.
+less so for components of noun phrases). Errors in morphology
+should be expressed on the atomic units, so HUME can behave slightly differently
+on configurational vs. non-configurational languages.
+
+% An example HUME annotation is given in \figref{fig:hume_tree},
+% which presents an annotation of an UCCA tree, and a corresponding translation.
+% We see that ``became a passion'' is deemed atomic, as it is translated as a single unit
+% in $t$, and thus its sub-units, ``became'', ``a'' and ``passion'' are not evaluated
+% individually.
 
@@
 \paragraph{Monolingual vs. Bilingual Evaluation}
+\label{src-vs-ref}
 HUME diverges from HMEANT (and also from the shallow automatic measures

### 2016-03-18 paper/main.tex
section 4

@@
 generally does not convey the meaning of these clauses. They appear in 21.7\% of the sentences,
-according to a conservative estimates that only considers non-auxilliary instances of ``be''.
+according to conservative estimates that only considers non-auxilliary instances of ``be''.
 \item[Nominative argument structures,] which are completely ignored under HMEANT are in fact highly
@@
 clauses (mainly discourse markers and coordinating conjunctions) appear in 56\% of the
-sentences\footnote{Argument structures and linkers are explicitly marked in UCCA. We identified
-  non-auxilliary instances of ``be'' and nouns using the NLTK standard tagger.}.
+sentences\footnote{Argument structures and linkers are explicitly marked in UCCA.
+We identified non-auxilliary instances of ``be'' and nouns using the NLTK standard tagger.
+Nominative argument structures are here defined as Scenes whose main relation is headed by a noun.}.
 \end{inparadesc}
@@
   \"uberpr\"ufen}\dots{}'' (lit. ``Two authors assessed the studies
-  independently to be reviewed\dots").\XXX{This is Omri's example, with
+  independently to review\dots").\XXX{This is Omri's example, with
   Ondrej's gloss for German. Ondrej is
@@
 
-Here is \XXX{another} example with a missing linker: the source \equo{However,
-this review\dots} was translated as \equo{Diese
-\"Uberpr\"ufung\dots} (lit. \equo{This review}). The omission is not by HMEANT at
-all.
+Another example where a linker is omitted in translation:
+the source \equo{However, this review\dots}, which was translated as \equo{Diese
+\"Uberpr\"ufung\dots} (lit. \equo{This review}).
+This omission will not be taken into account by HMEANT at all.
 
@@
 The coverage of
-these and other phenomena by HUME will allow to determine empirically
+these and other phenomena by HUME will allow to empirically determine 
 (rather than stipulate), which semantic structures are important
@@
 The translation is correct, except that the main verb ``sein'' (``be'') is omitted.
-While this may be interpreted as a minor error, HMEANT will assign it a very low score,
-as it failed to translate the main verb.
+While this may be interpreted as a minor error, HMEANT will assign the
+sentence a very low score, as it failed to translate the main verb.
 
@@
 \parcite{bojar:wu:ssst:2012,chuchunkov-tarelkin-galinskaya:2014:SSST-8} have
-suggested that sometimes, the structures of the reference and candidate
-translations diverge without any good reason, for instance due to a different
-interpretation of a PP-attachment, the verb can have one additional modifier in
+suggested that sometimes, the structures of the reference and of an accurate translation
+still diverge, for instance due to a different
+interpretation of a PP-attachment, or the verb having an additional modifier in
+%suggested that sometimes, the structures of the reference and candidate
+%translations diverge without any good reason, for instance due to a different
+%interpretation of a PP-attachment, the verb can have one additional modifier in
 one of the structures. It would be desirable to allow updates to the structural
-annotations also at the alignment stage
-to avoid unduly penalization due to such a spurious divergence.
+annotations at the alignment stage,
+to avoid unduly penalization due to such spurious divergences.
+%annotations also at the alignment stage
+%to avoid unduly penalization due to such a spurious divergence.
 The same issue was observed by HMEANT authors themselves
@@
 
-HUME uses only automatic (word-level) alignment, which only serves as a cue for directing
-the attention of the annotators.
-The user is expected to mentally correct the alignment as needed.
+%HUME uses only automatic (word-level) alignment, which only serves as a cue for directing
+%the attention of the annotators.
+%The user is expected to mentally correct the alignment as needed.
+HUME uses automatic (word-level) alignment, which only
+serves as a cue for directing the attention of the annotators.
+The user is expected to mentally correct the
+alignment as needed, thus circumventing this difficulty.
 

### 2016-03-18 paper/main.tex
section 2

--- 2016-03-18 paper/main.tex
@@
 Human evaluation is generally done by ranking the outputs of multiple systems
-e.g., in the WMT shared tasks \cite{bojar2015findings}, or by assigning
-individual adequacy/fluency scores to each translation, a procedure recently improved
+e.g., in the WMT tasks \cite{bojar2015findings}, or by giving
+adequacy/fluency scores to each translation, a procedure recently improved
 by \newcite{graham2015improving}.
@@
 the translation, and quantify how many of these sub-parts appear in a manually created reference translation.
-This decomposition further allows system developers
+This decomposition allows system developers
 to localize the errors, or at least unconfirmed parts within the translation.
@@
 While these measures stimulated much progress in MT research by allowing
-the evaluation of massive-scale experiments,
-their decomposition over words and n-grams is better suited for localizing errors
-in the output of shallow MT methods, and is less instructive in identifying the semantic
-nature of the errors.
+the evaluation of massive-scale experiments, the focus on words and n-grams does
+not provide a good estimate of semantic correctness, and may favour shallow string-based
+MT models.
 
@@
 of their structure. \newcite{liu2005syntactic} took a syntactic approach, 
-quantifying the similarity between the syntactic structures of the reference and translation.
+using dependency grammar, and
 \newcite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
-\newcite{gimenez2007linguistic} proposed to aggregate multiple types of information,
-capturing the overlap between the output and reference in terms of their
+\newcite{gimenez2007linguistic} proposed to combine multiple types of information,
+capturing the overlap between the translation and reference in terms of their
 semantic (predicate-argument structures), lexical and morphosyntactic features.
@@
 
+{\small
 \begin{tabular}{|l|l|l|l|}
@@
 \end{tabular}
+}
     \end{center}
 \caption{\label{fig:ucca_example_v2}
-  Sample UCCA annotation where leaves correspond to words and nodes correspond to units.
-  Edge labels mark UCCA categories, which are described in the table.}
+  Sample UCCA annotation where leaves correspond to words and nodes to units.
+  Edge labels mark UCCA categories, listed  in the table.}
 %which are largely ignored in this work
@@
 cross-linguistically applicable, lightweight
-scheme for semantic annotation. Formally, a UCCA structure is a directed acyclic graph (DAG),
+scheme for semantic annotation. Formally, an UCCA structure is a directed acyclic graph (DAG),
 whose leaves correspond to the words of the text.
@@
 The most basic notion is the Scene, which describes a movement, an
-action or a state which is persistent in time. Each Scene contains one main
+action or a state which persists in time. Each Scene contains one main
 relation and zero or more participants. For example, the sentence ``After graduation, Tom moved to NYC''
@@
 by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
-\cite{Dixon:10a,Dixon:10b,Dixon:12}, one of the most widely used frameworks for linguistic description.
+\cite{Dixon:10a,Dixon:10b,Dixon:12}.
 The cross-linguistic applicability of UCCA has been tested in annotations of
@@
 in translation: UCCA annotations of translated text usually contain the same set of relationships
-\cite{sulem2015conceptual}. This is an indication that the scheme reflects
+\cite{sulem2015conceptual}. This indicates that the scheme reflects
 a common layer of representation that in a correct translation
@@
 argument structure may be translated or paraphrased to a nominal or adjectival
-argument structure, and are thus less suitable as a basis for a semantic measure.
+argument structure, and is thus less suitable as a basis for a semantic measure.
 E.g., ``after he graduated, Tom moved to NYC'' contains two verbs,
@@
 thus does not provide clear decomposition of the sentence into sub-units.
-Also, AMR is more fine-grained than UCCA and consequently more difficult to annotate.
+Also, AMR is more fine-grained than UCCA and consequently harder to annotate.
 Other approaches pursued a grounded representation of semantics,
@@
 these approaches require linguistic expertise for their annotation, due to the intricate and
-fine-grained distinctions they provide, and are thus in our view less suitable for basing
-an interpretable MT evaluation on.
+fine-grained distinctions they provide, and so we argue they are less suitable as a basis
+for an interpretable MT evaluation.
 
@@
 
-This section summarises the manual annotation procedure performed in order 
+This section summarises the manual annotation procedure used
 to compute the HUME measure. 
-We mark the source sentence as $s$ and the translation sentence as $t$. 
+We mark the source sentence as $s$ and the translation as $t$. 
 The procedure involves two manual steps: (1) UCCA-annotating $s$, 

### 2016-03-18 paper/main.tex

--- 2016-03-18 paper/main.tex
@@
 
-This section summarises the manual annotation procedure performed in order 
+This section summarises the manual annotation procedure used
 to compute the HUME measure. 
-We denote the source sentence as $s$ and the translation sentence as $t$. 
+We denote the source sentence as $s$ and the translation as $t$. 
 The procedure involves two manual steps: (1) UCCA-annotating $s$, 

### 2016-03-18 paper/main.tex
Cutting down and editing 3.2 and 3.3

--- 2016-03-18 paper/main.tex
@@
 We proceed to detailing how judgements on the semantic units
-of the source are aggregated into a composite score.
-While future extensive empirical evaluation with HUME and its application to MT
-system tuning may suggest more complex definition,
-as a starting point we take a parsimonious approach,
+of the source are aggregated into a composite score. We start by taking a parsiminous approach
+%While future extensive empirical evaluation with HUME and its application to MT
+%system tuning may suggest more complex definition,
+%as a starting point we take a parsimonious approach,
 and compute an accuracy score.
@@
 Let $\Partial(s,t)$ be the number of atomic units in $s$ marked as Orange relative
-to $t$. Let $\Units(s)$ be the total number of units marked with any of the five labels.
+to $t$. Let $\setsize{\Units(s)}$ be the total number of units marked with any of the five labels.
 Then HUME's composite score is:
@@
 
-We turn to discuss the interface for collecting human judgements on the semantic units
-of the source, relating to the translation. UCCA annotations of the source are done offline,
+We now describe the HUME annotation interface, shown in Figure~\figref{fig:interface}. 
+%We turn to discuss the interface for collecting human judgements on the semantic units
+%of the source, relating to the translation. 
+UCCA annotations of the source are done offline,
 using the UCCA web application.
-
-\figref{fig:interface} presents the user interface of HUME. 
+%\figref{fig:interface} presents the user interface of HUME. 
 The translation appears at the top, and
 the source sentence directly below it. Underneath the source sentence appears
-its list of semantic units. For each semantic unit, including the one for the
-entire source sentence, there is a textbox, displaying the yield of that unit.
-Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
+its list of semantic units. 
+%For each semantic unit, including the one for the
+%entire source sentence, there is a textbox, displaying the yield of that unit.
+%Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
 The user is asked to select a label for each unit,
-by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the left of the unit's box.
+by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the right of the unit's box.
+Note that the UCCA labels are not shown.
 
@@
 
-For example, consider the sentence $s=$``We're driving back on Friday'' and the German
-translation t=``Am Freitag fahren wir zur\"uck'' (lit. ``on Friday driving we back'').
-``driving back'' will be aligned to ``fahren ... zur\"uck'',
-and ``wir'' will be marked as intervening word. Another example
-appears in \figref{fig:interface}, where ``ongoing pregnancy'' is translated to
-``Schwangerschaft ... laufenden'' (lit. ``pregnancy ... ongoing'').
+For example, 
+%consider the sentence $s=$``We're driving back on Friday'' and the German
+%translation t=``Am Freitag fahren wir zur\"uck'' (lit. ``on Friday driving we back'').
+%``driving back'' will be aligned to ``fahren ... zur\"uck'',
+%and ``wir'' will be marked as intervening word. Another example
+in \figref{fig:interface}, ``ongoing pregnancy'' is translated to
+``Schwangerschaft ... laufenden'' (lit. ``pregnancy ... ongoing''). The interleaving 
+German words shown in red should not occur there, and the annotator should take this
+into account and mark this structural node as ``B'', bad.
 
@@
     \caption{The HUME annotation tool, based on the UCCA tool. The faded orange box at the top
-      presents the translation, the large box below it presents the source. The tree of semantic units
-      for the source is given underneath, where the aligned text appears in italics, and intervening words
-      appear in red with a smaller font (e.g., in the first box).}
+      contains the translation, and directly below it is the source sentence. The tree of semantic units
+      is shown below the source. Alignments between the source and translation appear in italics and unaligned intervening words appear in red.}
     \label{fig:interface}

### 2016-03-18 paper/main.tex
edits
shorten section 3

--- 2016-03-18 paper/main.tex
@@
 where units are defined according to the UCCA annotation.
-We note that UCCA annotation is performed once for every source sentence,
+Note that UCCA annotation is performed once for every source sentence,
 irrespective of the number of its translations we wish
@@
 is the set of yields of nodes in $V$. 
-We note that the translation is not UCCA-annotated.
+The translation is not UCCA-annotated.
 
@@
 \end{figure}
-
-In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME
-annotation, with the translation shown in English for ease of comprehension.
-When evaluating ``to NYC'' the annotator looks at the translation and sees the
-word ``stateside''. This word captures the whole phrase and so we mark the
-structural node with an atomic label. Here we choose ``Orange'' but the
-annotator could equally have decided that this translation is wrong and marked
-it as ``Red''. The ability to mark structural nodes with atomic labels allows
-the annotator to account for translations which only correspond at the phrase
-level. Another feature highlighted in this example is that the root node is
-``Adequate''. This might seem confusing, as the relationship between the scenes
-has changed in the translation. However, an important feature of HUME is that
-penalises errors where they occur. In HUME, the error is caused by a wrong
-translation of the linker ``After'' which translates as ``by'' and this is
-marked as ``Red''. The root node still has two scenes linked by one linker and
-the relationship between child nodes is maintained and therefore it is marked
-``Adequate''.
 % Ondrej removes the following, he thinks that errors are not supposed to be
@@
 \paragraph{Translation Evaluation.}
-The basic strategy for computing HUME is to traverse the semantic units
-of the source sentence, which correspond to the different arguments and relations expressed
+The basic strategy for annotating HUME is to traverse the semantic units
+of the source sentence, which correspond to the arguments and relations expressed
 in the text, and to mark the extent to which they have been correctly translated.
@@
 
-Human annotation of the semantic units requires first deciding whether
+Annotation of the semantic units requires first deciding whether
 a unit is {\it structural}, i.e., has meaning-bearing sub-units,
@@
 (e.g., idioms like ``takes two to tango'').
-When a multi-word unit is labeled as atomic, all its sub-units' annotations are discarded
+When a multi-word unit is labeled as atomic, its sub-units' annotations are discarded
 from the evaluation.
@@
 the relation between the sub-units is preserved in $t$.
-For instance, a Scene unit represents the relation between the
-Scene elements, such as the predicate (main relation), participants and adverbials.
+\XXX{(BH: don't follow this) For instance, a Scene unit represents the relation between the
+Scene elements, such as the predicate (main relation), participants and adverbials.}
 Structural units are labeled as ``Adequate'' or ``Bad'', meaning
@@
 \item The components of the translation are ordered differently from the source,
-  such that the relation between them changes. For instance, where
+  such that the relation between them changes. For instance, if
   the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
@@
 
-We note that HUME labels reflect adequacy, rather than fluency judgements.
+HUME labels reflect adequacy, rather than fluency judgements.
 Specifically, annotators are instructed to
@@
 In non-configurational languages, the relationship between the
-governing and the dependent units is often expressed using morphological
+governing and dependent units is often expressed using morphological
 properties of the dependents instead of their ordering
@@
 less so for components of noun phrases). Errors in morphology
-should be expressed on the atomic units, so HUME can behave slightly differently
+should be expressed on the atomic units, so HUME can behave differently
 on configurational vs. non-configurational languages.
@@
 %as their relation in $t$ is reversed.
+
+In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME
+annotation, with the translation shown in English for ease of comprehension.
+When evaluating ``to NYC'' the annotator looks at the translation and sees the
+word ``stateside''. This word captures the whole phrase and so we mark the
+structural node with an atomic label. Here we choose ``Orange'' since it approximately captures the meaning
+in this context. The ability to mark structural nodes with atomic labels allows
+the annotator to account for translations which only correspond at the phrase
+level. Another feature highlighted in this example is that the root node is
+``Adequate''. This is because the root node still has two scenes linked by one linker and
+so the relationship between child nodes is maintained. There is an error caused by a wrong
+translation of the linker ``After'' which translates as ``by'' and this is
+marked as ``Red'', but this error is marked at the point at which it occurs.
 

### 2016-03-18 paper/main.tex

@@
 }
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Experiments}\label{sec:experiments}
+
+In order to validate the HUME metric, we ran an annotation experiment with one source language (English),
+and four target languages (Czech, German, Polish and Romanian), using text from the public health domain.
+The domain is particularly suitable for experimenting on MT evaluation, as semantically
+accurate translation is paramount here, underscoring the importance of accurate
+semantic evaluation measures.
+
+%particularly interested in this domain since our ultimate goal is to provide high-quality translations of health
+%information for speakers of languages other than English, and in this domain accurate transfer of meaning is
+%paramount.
+
+The experiment seeks to address two questions:
+
+\begin{compactitem}
+\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences?
+\item Can HUME capture translation issues that other annotation schemes (such as
+adequacy judgements and HMEANT)
+cannot?
+%\item What can HUME annotation tell us about the strengths and weaknesses of MT?
+%\bh{Maybe we should remove the third one, we
+%probably don't address it in this paper}
+\end{compactitem}
+
+\subsection{UCCA Annotation}
+Source sentences are all in English, and the UCCA annotation was performed by four
+computational linguists and one linguist.
+
+\subsection{Data Sets and Translation Systems}
+
+For each of the four language pairs under consideration  we built phrase-based MT systems
+using Moses \cite{Koehn:2007}. We used large training sets extracted from OPUS \cite{tiedemann:2009}, and
+the data sets released for the WMT14 medical translation task \cite{bojar-EtAl:2014:W14-33}, to provide between
+45 and 85 million sentences of training data, depending on language pair\oa{I think we should probably
+  say how many in each.}.
+
+These translation systems were used to translate texts derived from both NHS 24\footnote{\url{http://www.nhs24.com/}} and Cochrane\footnote{\url{http://www.cochrane.org/}} into the four languages.
+The first organisation is a public body providing healthcare and health-service related information in Scotland, whereas the second is an international NGO which
+provides independent systematic reviews on health-related research. From NHS 24, we selected the segments from their
+``Health A-Z'' which is part of the NHS Inform website, and the Cochrane texts come from their plain language summaries
+and abstracts.\bh{How much more do we need? Perhaps in the intro we could justify our focus on public health text
+(accuracy!). The test corpora should have a public URL by the time we submit, although of course the source texts were 
+already public anyway, and Lexi did some selection of examples for use in this expt. -- do we have to explain this?}
+
+\subsection{Annotation Statistics}
+\label{sec:annot_stats}
+
+For the annotation of the MT output, we recruited two annotators for each of German, Romanian
+and Polish and one main annotator for Czech. In order to allow us to calculate IAA also for
+the Czech annotations we recruited several further annotators to work on a small number of 
+sentences each. We treat these further annotators as one annotator, giving us two annotators
+for each language pair.
+The annotators were all native  speakers of the respective target languages and fluent in English.
+
+In \tabref{tab:annot},
+we show the total number of sentences and nodes annotated by each annotator. Note that not all nodes in all sentences were annotated. Sometimes this was because the
+annotator did not finish a node, or accidentally missed a node, but in many cases it was also because it was an implicit 
+UCCA node, which was not shown to the annotators since there is no corresponding source segment.
+\begin{table}
+\begin{center}
+{\small
+\begin{tabular}{ll|cccc}
+& & cs & de & pl & ro \\
+\hline
+Sentences &  Annot. 1 & 324   & 339  & 351  & 230  \\
+ & Annot. 2 & 205 & 104  & 340  & 337 \\
+\hline
+Nodes & Annot. 1 & 8794  & 9253 & 9557  & 6152 \\
+ &Annot. 2 & 5553 & 2906  & 9303  & 9228  \\
+\end{tabular}
+}
+\caption{Total numbers of sentences and nodes annotated on the MT output.}
+\label{tab:annot}
+\end{center}
+\end{table}
+
+The annotators worked remotely, at their own pace, so we did not collect precise timing
+information. Nevertheless we are able to estimate the annotation time using the timestamps
+provided by the annotation tool, which are recorded whenever an annotated sentence is
+submitted. Annotators are not able to re-open a sentence once submitted.
+
+To estimate the annotation time, we calculate the time difference between successive 
+sentences, and discard outlying times since we assume there was a break taken. From inspection
+of histograms of annotation times, we set the upper threshold at 500 seconds. We
+show median annotation times in Table~\ref{tab:annot_times}, indicating that the annotation
+of a sentence takes around 2--4 minutes, with some variation between annotators.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{l|cccc}
+& cs & de & pl & ro \\
+\hline
+Annot. 1 & 255 & 140  & 138 & 96 \\
+Annot. 2 & -- & 162 & 229 & 207 \\
+\end{tabular}
+\caption{Median annotation times per sentence, in seconds.
+We do not give times for the second Czech 
+``annotator'', since this was a collection of annotators, working in parallel.}
+\label{tab:annot_times}
+\end{center}
+\end{table}
+
+\emph{Describe the UCCA annotation}
+
+\subsection{Inter-Annotator Agreement}
+\label{sec:iaa}
+
+\oa{what is the purpose of the BLEU binning? is it because low quality and high
+quality translations often diverge in their agreement scores?}
+
+In order to assess the consistency of the annotation, we measure the Inter-Annotator
+Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
+For German, Polish and Romanian we used two annotators, with a significant overlap
+between the sentences they annotated. For Czech we had one main annotator, who
+annotated the entire batch, as well multiple secondary annotators, where
+each annotated a sub-part of the batch. \oa{i editted it. is what i wrote correct?}
+
+In order to compute Kappa, we consider the annotation task as a classification task on 
+nodes, and compute agreement on nodes which have two annotations from different annotators.
+We report the overall Kappa, as well as separate Kappas on atomic
+nodes (annotated as Red, Orange or Green) and the structural nodes (annotated
+as Adequate or Bad). In general there should be little confusion between the two
+types of nodes, and in future work the separation between the nodes should be enforced
+by the tool\oa{but there may be some disagreement in the margins on what nodes are atomic and
+  what are not}. The Kappas are shown in \tabref{tab:iaa}.
+
+\begin{table}[!ht]
+\begin{center}
+\begin{tabular}{l|cccc}
+ & cs & de & pl & ro \\
+\hline
+Sentences & 181 & 102 & 334 & 217 \\
+\hline
+All nodes & 4686   & 2793   & 8384   & 5604  \\
+Kappa & 0.64   & 0.61   & 0.58   & 0.69  \\
+\hline
+Atomic nodes & 2982 & 1724 & 5386 & 3570 \\
+Kappa & 0.54 & 0.29 & 0.54 & 0.50 \\
+\hline
+Structural nodes & 1602 & 1040 & 2655 & 1989 \\
+Kappa & 0.31 & 0.44 & 0.33 & 0.58 \\
+\end{tabular}
+\caption{Inter-annotator agreement for the multiply-annotated portions of the data, as
+measured by Cohen's Kappa. }
+\label{tab:iaa}
+\end{center}
+\end{table}
+
+To assess whether, as we hoped, HUME would make it easier to reliably annotate long sentences,
+we binned the sentences according to length and measured Kappa on each of the bins. Each bin covers
+a length range of 10, and we show the variance of kappa with length in \figref{fig:iaalength}.
+
+\def\iaafig #1{\includegraphics[width=3.8cm]{iaa_length_#1.png}}
+
+\begin{figure}[ht!]
+\renewcommand{\tabcolsep}{0pt}
+\begin{tabular}{cc}
+
+\subfloat[English-Czech]{
+  \iaafig{cs}
+}
+&
+\subfloat[English-German]{
+  \iaafig{de}
+
+}
+\\
+
+\subfloat[English-Polish]{
+  \iaafig{pl}
+
+}
+&
+\subfloat[English-Romanian]{
+  \iaafig{ro}
+
+}
+\end{tabular}
+\caption{Inter-annotator Agreement (Kappa) versus sentence length, separately for
+structural and atomic nodes. The numbers are the node counts in each bin. }
+\label{fig:iaalength}
+\end{figure}
+
+We see from \figref{fig:iaalength} that there is no reduction of IAA with sentence
+length, except perhaps in German. Also, from \tabref{tab:iaa} the overall IAA
+is quite similar (0.6-0.7) for all languages, showing good agreement.
+However, there are differences observed when we break down by node type.
+In particular, we see a contrast  between
+Czech and Polish, where the IAA is higher for atomic than for structural nodes, and German and Romanian,
+where the reverse is true. We also observe low IAA (around 0.3) in the cases of
+German atomic nodes, and Polish and Czech structural nodes.
+
+Looking more closely at the areas of disagreement, we see that for the Polish structural nodes, the 
+proportion of As was quite different between the two annotators (53\% vs. 71\%), whereas for other
+languages the annotators agree in the proportions. We believe that this was because one of the Polish
+annotators did not fully understand the guidelines for structural nodes, and percolated
+errors up the tree, creating more Bs. For the German lexical nodes, and Czech structural nodes, where
+Kappa is also around 0.3, the proportion of such nodes being marked as ``correct'' is relatively 
+high, meaning that the class distribution is more skewed, so the expected agreement used in the
+Kappa calculation is high, lowering Kappa. Essentially, the annotation problem is harder when the classes
+are more skewed. Finally we note some evidence of domain-specific disagreements, for instance the German
+MT system normally translated ``review'' (as in ``systematic review'' -- a frequent term in the 
+Cochrane texts) as ``\"uberpr\"ufung'', which 
+one annotator marked correct, and the other (a Cochrane employee)  as incorrect.
+
+\def\iaafig #1{\includegraphics[width=3.8cm]{iaa_heatmap_#1.png}}
+
+\begin{figure}[ht!]
+\renewcommand{\tabcolsep}{0pt}
+\begin{tabular}{cc}
+
+\subfloat[English-Czech]{
+  \iaafig{cs}
+}
+&
+\subfloat[English-German]{
+  \iaafig{de}
+
+}
+\\
+
+\subfloat[English-Polish]{
+  \iaafig{pl}
+
+}
+&
+\subfloat[English-Romanian]{
+  \iaafig{ro}
+
+}
+\end{tabular}
+\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
+\label{fig:heatmap}
+\end{figure}
+
+%\subsection{Discussion of Disagreements}\label{sec:disagreements}
+
+%\subsection{Comparison with Crowdsourced Adequacy Judgements}\label{sec:adequacy}
 
@@
 %The more frames can people annotate in the hypothesis, the better the hypothesis overall.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experiments}\label{sec:experiments}
-
-In order to validate the HUME metric, we ran an annotation experiment with one source language (English),
-and four target languages (Czech, German, Polish and Romanian), using text from the public health domain.
-The domain is particularly suitable for experimenting on MT evaluation, as semantically
-accurate translation is paramount here, underscoring the importance of accurate
-semantic evaluation measures.
-
-%particularly interested in this domain since our ultimate goal is to provide high-quality translations of health
-%information for speakers of languages other than English, and in this domain accurate transfer of meaning is
-%paramount.
-
-The experiment seeks to address two questions:
-
-\begin{compactitem}
-\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences?
-\item Can HUME capture translation issues that other annotation schemes (such as
-adequacy judgements and HMEANT)
-cannot?
-%\item What can HUME annotation tell us about the strengths and weaknesses of MT?
-%\bh{Maybe we should remove the third one, we
-%probably don't address it in this paper}
-\end{compactitem}
-
-\subsection{UCCA Annotation}
-Source sentences are all in English, and the UCCA annotation was performed by four
-computational linguists and one linguist.
-
-\subsection{Data Sets and Translation Systems}
-
-For each of the four language pairs under consideration  we built phrase-based MT systems
-using Moses \cite{Koehn:2007}. We used large training sets extracted from OPUS \cite{tiedemann:2009}, and
-the data sets released for the WMT14 medical translation task \cite{bojar-EtAl:2014:W14-33}, to provide between
-45 and 85 million sentences of training data, depending on language pair\oa{I think we should probably
-  say how many in each.}.
-
-These translation systems were used to translate texts derived from both NHS 24\footnote{\url{http://www.nhs24.com/}} and Cochrane\footnote{\url{http://www.cochrane.org/}} into the four languages.
-The first organisation is a public body providing healthcare and health-service related information in Scotland, whereas the second is an international NGO which
-provides independent systematic reviews on health-related research. From NHS 24, we selected the segments from their
-``Health A-Z'' which is part of the NHS Inform website, and the Cochrane texts come from their plain language summaries
-and abstracts.\bh{How much more do we need? Perhaps in the intro we could justify our focus on public health text
-(accuracy!). The test corpora should have a public URL by the time we submit, although of course the source texts were 
-already public anyway, and Lexi did some selection of examples for use in this expt. -- do we have to explain this?}
-
-\subsection{Annotation Statistics}
-\label{sec:annot_stats}
-
-For the annotation of the MT output, we recruited two annotators for each of German, Romanian
-and Polish and one main annotator for Czech. In order to allow us to calculate IAA also for
-the Czech annotations we recruited several further annotators to work on a small number of 
-sentences each. We treat these further annotators as one annotator, giving us two annotators
-for each language pair.
-The annotators were all native  speakers of the respective target languages and fluent in English.
-
-In \tabref{tab:annot},
-we show the total number of sentences and nodes annotated by each annotator. Note that not all nodes in all sentences were annotated. Sometimes this was because the
-annotator did not finish a node, or accidentally missed a node, but in many cases it was also because it was an implicit 
-UCCA node, which was not shown to the annotators since there is no corresponding source segment.
-\begin{table}
-\begin{center}
-{\small
-\begin{tabular}{ll|cccc}
-& & cs & de & pl & ro \\
-\hline
-Sentences &  Annot. 1 & 324   & 339  & 351  & 230  \\
- & Annot. 2 & 205 & 104  & 340  & 337 \\
-\hline
-Nodes & Annot. 1 & 8794  & 9253 & 9557  & 6152 \\
- &Annot. 2 & 5553 & 2906  & 9303  & 9228  \\
-\end{tabular}
-}
-\caption{Total numbers of sentences and nodes annotated on the MT output.}
-\label{tab:annot}
-\end{center}
-\end{table}
-
-The annotators worked remotely, at their own pace, so we did not collect precise timing
-information. Nevertheless we are able to estimate the annotation time using the timestamps
-provided by the annotation tool, which are recorded whenever an annotated sentence is
-submitted. Annotators are not able to re-open a sentence once submitted.
-
-To estimate the annotation time, we calculate the time difference between successive 
-sentences, and discard outlying times since we assume there was a break taken. From inspection
-of histograms of annotation times, we set the upper threshold at 500 seconds. We
-show median annotation times in Table~\ref{tab:annot_times}, indicating that the annotation
-of a sentence takes around 2--4 minutes, with some variation between annotators.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{l|cccc}
-& cs & de & pl & ro \\
-\hline
-Annot. 1 & 255 & 140  & 138 & 96 \\
-Annot. 2 & -- & 162 & 229 & 207 \\
-\end{tabular}
-\caption{Median annotation times per sentence, in seconds.
-We do not give times for the second Czech 
-``annotator'', since this was a collection of annotators, working in parallel.}
-\label{tab:annot_times}
-\end{center}
-\end{table}
-
-\emph{Describe the UCCA annotation}
-
-\subsection{Inter-Annotator Agreement}
-\label{sec:iaa}
-
-\oa{what is the purpose of the BLEU binning? is it because low quality and high
-quality translations often diverge in their agreement scores?}
-
-In order to assess the consistency of the annotation, we measure the Inter-Annotator
-Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
-For German, Polish and Romanian we used two annotators, with a significant overlap
-between the sentences they annotated. For Czech we had one main annotator, who
-annotated the entire batch, as well multiple secondary annotators, where
-each annotated a sub-part of the batch. \oa{i editted it. is what i wrote correct?}
-
-In order to compute Kappa, we consider the annotation task as a classification task on 
-nodes, and compute agreement on nodes which have two annotations from different annotators.
-We report the overall Kappa, as well as separate Kappas on atomic
-nodes (annotated as Red, Orange or Green) and the structural nodes (annotated
-as Adequate or Bad). In general there should be little confusion between the two
-types of nodes, and in future work the separation between the nodes should be enforced
-by the tool\oa{but there may be some disagreement in the margins on what nodes are atomic and
-  what are not}. The Kappas are shown in \tabref{tab:iaa}.
-
-\begin{table}[!ht]
-\begin{center}
-\begin{tabular}{l|cccc}
- & cs & de & pl & ro \\
-\hline
-Sentences & 181 & 102 & 334 & 217 \\
-\hline
-All nodes & 4686   & 2793   & 8384   & 5604  \\
-Kappa & 0.64   & 0.61   & 0.58   & 0.69  \\
-\hline
-Atomic nodes & 2982 & 1724 & 5386 & 3570 \\
-Kappa & 0.54 & 0.29 & 0.54 & 0.50 \\
-\hline
-Structural nodes & 1602 & 1040 & 2655 & 1989 \\
-Kappa & 0.31 & 0.44 & 0.33 & 0.58 \\
-\end{tabular}
-\caption{Inter-annotator agreement for the multiply-annotated portions of the data, as
-measured by Cohen's Kappa. }
-\label{tab:iaa}
-\end{center}
-\end{table}
-
-To assess whether, as we hoped, HUME would make it easier to reliably annotate long sentences,
-we binned the sentences according to length and measured Kappa on each of the bins. Each bin covers
-a length range of 10, and we show the variance of kappa with length in \figref{fig:iaalength}.
-
-\def\iaafig #1{\includegraphics[width=3.8cm]{iaa_length_#1.png}}
-
-\begin{figure}[ht!]
-\renewcommand{\tabcolsep}{0pt}
-\begin{tabular}{cc}
-
-\subfloat[English-Czech]{
-  \iaafig{cs}
-}
-&
-\subfloat[English-German]{
-  \iaafig{de}
-
-}
-\\
-
-\subfloat[English-Polish]{
-  \iaafig{pl}
-
-}
-&
-\subfloat[English-Romanian]{
-  \iaafig{ro}
-
-}
-\end{tabular}
-\caption{Inter-annotator Agreement (Kappa) versus sentence length, separately for
-structural and atomic nodes. The numbers are the node counts in each bin. }
-\label{fig:iaalength}
-\end{figure}
-
-We see from \figref{fig:iaalength} that there is no reduction of IAA with sentence
-length, except perhaps in German. Also, from \tabref{tab:iaa} the overall IAA
-is quite similar (0.6-0.7) for all languages, showing good agreement.
-However, there are differences observed when we break down by node type.
-In particular, we see a contrast  between
-Czech and Polish, where the IAA is higher for atomic than for structural nodes, and German and Romanian,
-where the reverse is true. We also observe low IAA (around 0.3) in the cases of
-German atomic nodes, and Polish and Czech structural nodes.
-
-Looking more closely at the areas of disagreement, we see that for the Polish structural nodes, the 
-proportion of As was quite different between the two annotators (53\% vs. 71\%), whereas for other
-languages the annotators agree in the proportions. We believe that this was because one of the Polish
-annotators did not fully understand the guidelines for structural nodes, and percolated
-errors up the tree, creating more Bs. For the German lexical nodes, and Czech structural nodes, where
-Kappa is also around 0.3, the proportion of such nodes being marked as ``correct'' is relatively 
-high, meaning that the class distribution is more skewed, so the expected agreement used in the
-Kappa calculation is high, lowering Kappa. Essentially, the annotation problem is harder when the classes
-are more skewed. Finally we note some evidence of domain-specific disagreements, for instance the German
-MT system normally translated ``review'' (as in ``systematic review'' -- a frequent term in the 
-Cochrane texts) as ``\"uberpr\"ufung'', which 
-one annotator marked correct, and the other (a Cochrane employee)  as incorrect.
-
-\def\iaafig #1{\includegraphics[width=3.8cm]{iaa_heatmap_#1.png}}
-
-\begin{figure}[ht!]
-\renewcommand{\tabcolsep}{0pt}
-\begin{tabular}{cc}
-
-\subfloat[English-Czech]{
-  \iaafig{cs}
-}
-&
-\subfloat[English-German]{
-  \iaafig{de}
-
-}
-\\
-
-\subfloat[English-Polish]{
-  \iaafig{pl}
-
-}
-&
-\subfloat[English-Romanian]{
-  \iaafig{ro}
-
-}
-\end{tabular}
-\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
-\label{fig:heatmap}
-\end{figure}
-
-%\subsection{Discussion of Disagreements}\label{sec:disagreements}
-
-%\subsection{Comparison with Crowdsourced Adequacy Judgements}\label{sec:adequacy}
 

### 2016-03-18 paper/main.tex
more shortening

@@
 correspond to individual words, but they may also correspond to unanalyzable
-multi-word expressions, i.e., expressions that do not have meaningful sub-units
-(e.g., idioms like ``takes two to tango'').
-When a multi-word unit is labeled as atomic, its sub-units' annotations are discarded
-from the evaluation.
+multi-word expressions.
+When a multi-word unit is labeled as atomic, its sub-units' annotations are ignored
+in the evaluation.
 

### 2016-03-18 paper/main.tex
edits

--- 2016-03-18 paper/main.tex
@@
 \newcommand{\tabref}[1]{Table~\ref{#1}}
-%\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
-\newcommand{\XXX}[1]{}
+\newcommand{\XXX}[1]{{\color{red}XXX #1}} % a general todo or comment
+%\newcommand{\XXX}[1]{}
 %\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
@@
 translation. For automatic measures,
-these are commonly words or n-grams, for manual measures some structural
+these are often words or n-grams, for manual measures some structural
 information is taken into account \parcite{machacek:bojar:segranks:2015}.
@@
 their focus on verb argument structures misses many semantic phenomena
-(such as inter-clause linkage and nominal argument structures) and thus they only provide
+(such as inter-clause linkage and nominal argument structures), providing
+only
 a limited perspective on the semantic similarity between the translation and reference.
@@
 After a brief review (\secref{sec:background}), we describe HUME in detail
-(\secref{sec:hume}) and highlight its benefits over HMEANT
-(\secref{sec:hmeant_comp}).
+(\secref{sec:hume}).
 Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments}) document the
 efficiency (time of annotation) and inter-annotator agreement scores for HUME.
+Finally, the differences from HMEANT are discussed 
+(\secref{sec:hmeant_comp}).
 
@@
     \begin{center}
-    \includegraphics[width=0.45\textwidth]{ucca-tree-v2.png}
+    \includegraphics[width=0.35\textwidth]{ucca-tree-v2.png}
 
@@
     \begin{center}
-    \includegraphics[width=0.45\textwidth]{ucca-tree-mteval-v2.png}
+    \includegraphics[width=0.35\textwidth]{ucca-tree-mteval-v2.png}
     \end{center}

### 2016-03-18 paper/main.tex
edits

@@
 %\newcommand{\XXX}[1]{}
-%\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
-\newcommand{\oa}[1]{}
-%\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
-\newcommand{\bh}[1]{}
+\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
+%\newcommand{\oa}[1]{}
+\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
+%\newcommand{\bh}[1]{}
 
@@
 Annotation of the semantic units requires first deciding whether
-a unit is {\it structural}, i.e., has meaning-bearing sub-units,
+a unit is {\it structural}, i.e., has meaning-bearing sub-units also in the
+target language,
 or {\it atomic}. In most cases, atomic units
@@
 Structural units have sub-units (children in the UCCA graph), which may themselves
-either be atomic or structural. The labels assigned to such nodes reflect the extent to which
-the relation between the sub-units is preserved in $t$.
-\XXX{(BH: don't follow this) For instance, a Scene unit represents the relation between the
-Scene elements, such as the predicate (main relation), participants and adverbials.}
+either be atomic or structural.
+% Ondrej thinks the following can be simply removed:
+%The labels assigned to such nodes reflect the extent to which
+%the relation between the sub-units is preserved in $t$.
+%\XXX{(BH: don't follow this) For instance, a Scene unit represents the relation between the
+%Scene elements, such as the predicate (main relation), participants and adverbials.}
 Structural units are labeled as ``Adequate'' or ``Bad'', meaning
@@
 There are many ways that the relationship between the children might go
-wrong in translation
+wrong in translation:
 

### 2016-03-18 paper/main.tex
edits

@@
 
-We now describe the HUME annotation interface, shown in Figure~\figref{fig:interface}. 
+We now describe the HUME annotation interface, shown in \figref{fig:interface}. 
 %We turn to discuss the interface for collecting human judgements on the semantic units
@@
 the alignment if it is misleading, and instead 
-make a judgement according to the full translation sentence.
+make a judgement according to the full translation.
 

### 2016-03-18 paper/main.tex
edits

@@
 in \figref{fig:interface}, ``ongoing pregnancy'' is translated to
-``Schwangerschaft ... laufenden'' (lit. ``pregnancy ... ongoing''). The interleaving 
-German words shown in red should not occur there, and the annotator should take this
-into account and mark this structural node as ``B'', bad.
+``Schwangerschaft ... laufenden'' (lit. ``pregnancy ... ongoing''). This alone
+seems acceptable but the interleaving words in red notify the annotator to check
+the whole translation. The meaning of the expression is not preserved in the
+translation
+and the annotator should mark this structural node as ``B'', bad. 
+%The interleaving 
+%German words shown in red should not occur there, and the annotator should take this
+%into account and mark this structural node as ``B'', bad.
 

### 2016-03-18 paper/main.tex
edits

@@
 The experiment seeks to address two questions:
-
-\begin{compactitem}
+\begin{inparaenum}
 \item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences?
@@
 %probably don't address it in this paper}
-\end{compactitem}
-
-\subsection{UCCA Annotation}
-Source sentences are all in English, and the UCCA annotation was performed by four
-computational linguists and one linguist.
+\end{inparaenum}
 
@@
 the data sets released for the WMT14 medical translation task \cite{bojar-EtAl:2014:W14-33}, to provide between
-45 and 85 million sentences of training data, depending on language pair\oa{I think we should probably
-  say how many in each.}.
+45 and 85 million sentences of training data, depending on language pair.
+%\oa{I think we should probably
+%  say how many in each.}.
+% ... if someone asks us to do so...
 
@@
 ``Health A-Z'' which is part of the NHS Inform website, and the Cochrane texts come from their plain language summaries
-and abstracts.\bh{How much more do we need? Perhaps in the intro we could justify our focus on public health text
-(accuracy!). The test corpora should have a public URL by the time we submit, although of course the source texts were 
-already public anyway, and Lexi did some selection of examples for use in this expt. -- do we have to explain this?}
+and abstracts.
+
+%\bh{How much more do we need? Perhaps in the intro we could justify our focus on public health text
+%(accuracy!). The test corpora should have a public URL by the time we submit, although of course the source texts were 
+%already public anyway, and Lexi did some selection of examples for use in this expt. -- do we have to explain this?}
+% ... todo for the camera ready, if we get through
 
@@
 \label{sec:annot_stats}
+
+Source sentences are all in English, and the UCCA annotation was performed by four
+computational linguists and one linguist.
 

### 2016-03-18 paper/main.tex
edits

@@
 
-Source sentences are all in English, and the UCCA annotation was performed by four
+Source sentences are all in English, and their UCCA annotation was performed by four
 computational linguists and one linguist.
@@
 For the annotation of the MT output, we recruited two annotators for each of German, Romanian
-and Polish and one main annotator for Czech. In order to allow us to calculate IAA also for
-the Czech annotations we recruited several further annotators to work on a small number of 
+and Polish and one main annotator for Czech.
+%In order to allow us to calculate
+For IAA also for
+the Czech, we recruited several further annotators to work on a small number of 
 sentences each. We treat these further annotators as one annotator, giving us two annotators
@@
 
-In \tabref{tab:annot},
-we show the total number of sentences and nodes annotated by each annotator. Note that not all nodes in all sentences were annotated. Sometimes this was because the
-annotator did not finish a node, or accidentally missed a node, but in many cases it was also because it was an implicit 
+\tabref{tab:annot}
+shows the total number of sentences and nodes annotated by each annotator. Note
+that not all nodes in all sentences were annotated. Sometimes,
+%this was because
+the annotator did not finish a sentence or accidentally missed a node, but in many cases it was also because it was an implicit 
 UCCA node, which was not shown to the annotators since there is no corresponding source segment.
@@
 
-\oa{what is the purpose of the BLEU binning? is it because low quality and high
-quality translations often diverge in their agreement scores?}
+%\oa{what is the purpose of the BLEU binning? is it because low quality and high
+%quality translations often diverge in their agreement scores?}
 
@@
   Ondrej's gloss for German. Ondrej is
-  not totally sure that the meaning is well preserved.}
+  not totally sure that the meaning is well preserved. So Ondrej would actually
+  *remove* this example, because it would illustrate HMEANT's problems only if
+  the meaning was correct.}
 
@@
 frame and has the traffic-lights annotation.
-
-\XXX{Omri: I would consider adding further examples from the Google Doc. For instnace
-``your risk of falling'' -- ``Ihr Risiko''.
-Ondrej: I don't see any HMEANT problem here, the whole NP 'your risk of falling'
-is aligned with 'Ihr Risiko' and HMEANT allows to indicate that the match was
-partial, to that's probably what everyone would choose. I've put the other
-example higher, but please have a look, I'm not sure if we should keep it there
-at all.}
 

### 2016-03-18 paper/main.tex
edits

@@
 
+\begin{figure*}[t]
+    \begin{center}
+    \includegraphics[width=1\textwidth]{hume_interface2.jpg}
+    \caption{The HUME annotation tool, based on the UCCA tool. The faded orange box at the top
+      contains the translation, and directly below it is the source sentence. The tree of semantic units
+      is shown below the source. Alignments between the source and translation appear in italics and unaligned intervening words appear in red.}
+    \label{fig:interface}
+    \end{center}
+\end{figure*}
+
 In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME
@@
 
-We now describe the HUME annotation interface, shown in \figref{fig:interface}. 
-%We turn to discuss the interface for collecting human judgements on the semantic units
-%of the source, relating to the translation. 
-UCCA annotations of the source are done offline,
-using the UCCA web application.
-%\figref{fig:interface} presents the user interface of HUME. 
-The translation appears at the top, and
-the source sentence directly below it. Underneath the source sentence appears
-its list of semantic units. 
-%For each semantic unit, including the one for the
-%entire source sentence, there is a textbox, displaying the yield of that unit.
-%Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
-The user is asked to select a label for each unit,
+\figref{fig:interface} shows the HUME annotation interface.
+% We now describe the HUME annotation interface, shown in \figref{fig:interface}. 
+% %We turn to discuss the interface for collecting human judgements on the semantic units
+% %of the source, relating to the translation. 
+% UCCA annotations of the source are done offline,
+% using the UCCA web application.
+% %\figref{fig:interface} presents the user interface of HUME. 
+% The translation appears at the top, and
+% the source sentence directly below it. Underneath the source sentence appears
+% its list of semantic units. 
+% %For each semantic unit, including the one for the
+% %entire source sentence, there is a textbox, displaying the yield of that unit.
+% %Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
+The user is asked to select a label for each source semantic unit,
 by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the right of the unit's box.
-Note that the UCCA labels are not shown.
+%Note that the UCCA labels are not shown.
 
@@
 %only partially capture the semantics of the source.
-
-\begin{figure*}[t]
-    \begin{center}
-    \includegraphics[width=1\textwidth]{hume_interface2.jpg}
-    \caption{The HUME annotation tool, based on the UCCA tool. The faded orange box at the top
-      contains the translation, and directly below it is the source sentence. The tree of semantic units
-      is shown below the source. Alignments between the source and translation appear in italics and unaligned intervening words appear in red.}
-    \label{fig:interface}
-    \end{center}
-\end{figure*}
 
@@
 
-\emph{Describe the UCCA annotation}
+%\XXX{Describe the UCCA annotation}
+% ... not enough room or time
 
@@
 For German, Polish and Romanian we used two annotators, with a significant overlap
-between the sentences they annotated. For Czech we had one main annotator, who
-annotated the entire batch, as well multiple secondary annotators, where
-each annotated a sub-part of the batch. \oa{i editted it. is what i wrote correct?}
+between the sentences they annotated. For Czech we had the entire batch
+annotated by one main annotator and multiple secondary annotators, 
+each annotating a sub-part of the batch.
 
@@
 nodes (annotated as Red, Orange or Green) and the structural nodes (annotated
-as Adequate or Bad). In general there should be little confusion between the two
-types of nodes, and in future work the separation between the nodes should be enforced
-by the tool\oa{but there may be some disagreement in the margins on what nodes are atomic and
-  what are not}. The Kappas are shown in \tabref{tab:iaa}.
-
-\begin{table}[!ht]
+as Adequate or Bad).
+As expected and confirmed by confusion matrices in \figref{fig:heatmap}, there
+is generally little confusion between the two types of nodes.
+
+%In general there should be little confusion between the two
+%types of nodes, and in future work the separation between the nodes should be enforced
+%by the tool.
+%\oa{but there may be some disagreement in the margins on what nodes are atomic and
+%  what are not}
+%The Kappas are shown in \tabref{tab:iaa}.
+
+\begin{table}[t]
 \begin{center}
@@
 
+\def\iaafig #1{\includegraphics[width=3.8cm]{iaa_heatmap_#1.png}}
+
+\begin{figure}[t]
+\renewcommand{\tabcolsep}{0pt}
+\begin{tabular}{cc}
+
+\subfloat[English-Czech]{
+  \iaafig{cs}
+}
+&
+\subfloat[English-German]{
+  \iaafig{de}
+
+}
+\\
+
+\subfloat[English-Polish]{
+  \iaafig{pl}
+
+}
+&
+\subfloat[English-Romanian]{
+  \iaafig{ro}
+
+}
+\end{tabular}
+\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
+\label{fig:heatmap}
+\end{figure}
+
 To assess whether, as we hoped, HUME would make it easier to reliably annotate long sentences,
@@
 
-\begin{figure}[ht!]
+\begin{figure}[t]
 \renewcommand{\tabcolsep}{0pt}

### 2016-03-18 paper/main.tex
edits

@@
 HMEANT \cite{lo2011structured}.
-\XXX{Do we want to cut the following shorter?}
-However, 
-their focus on verb argument structures misses many semantic phenomena
-(such as inter-clause linkage and nominal argument structures), providing
-only
-a limited perspective on the semantic similarity between the translation and reference.
-Also, assigning semantic structure to an MT output, especially a poor one,
-is difficult and arguably ill-defined.
+% SAVING SPACE:
+% \XXX{Do we want to cut the following shorter?}
+% However, 
+% their focus on verb argument structures misses many semantic phenomena
+% (such as inter-clause linkage and nominal argument structures), providing
+% only
+% a limited perspective on the semantic similarity between the translation and reference.
+% Also, assigning semantic structure to an MT output, especially a poor one,
+% is difficult and arguably ill-defined.
 
@@
 \item One of the sub-units is missing from the translation.
-  For instance, if ``dog bites man'' is translated into ``dog bites''. \bh{Note that in this case we also 
-  get an error in the atomic unit, assuming that ``man'' is a semantic unit in the source, but for the other
-  two cases there is no atomic unit error.}
+  For instance, if ``dog bites man'' is translated into ``dog bites''.
+% XXX let's leave the reviewers to notice this:
+%  \bh{Note that in this case we also 
+%  get an error in the atomic unit, assuming that ``man'' is a semantic unit in the source, but for the other
+%  two cases there is no atomic unit error.}
 \end{compactitem}
@@
 are presented in a smaller red font.
-Intervening words are likely change the meaning of the translation
+Intervening words are likely to change the meaning of the translation
 of $u$, and thus should be paid attention to when considering whether the translation
@@
 
-In order to compute Kappa, we consider the annotation task as a classification task on 
-nodes, and compute agreement on nodes which have two annotations from different annotators.
+\tabref{tab:iaa} reports the number of nodes which have two annotations from
+different annotators and the corresponding Kappas.
+%
+%In order to compute Kappa, we consider the annotation task as a classification task on 
+%nodes, and compute agreement on nodes which have two annotations from different annotators.
 We report the overall Kappa, as well as separate Kappas on atomic
-nodes (annotated as Red, Orange or Green) and the structural nodes (annotated
+nodes (annotated as Red, Orange or Green) and structural nodes (annotated
 as Adequate or Bad).

### 2016-03-18 paper/main.tex
edits
HMEANT comparison section reduced and edited

--- 2016-03-18 paper/main.tex
@@
 
-First of all, the annotation process is different. We semantically annotate the
-source, then automatically project it to the translation, and ask bilingual
-annotators to assess whether this projection shows a good or bad translation. In
-HMEANT, they semantically annotate both the reference and the translation, then
-align these manually, and assess to what extent they match.
-
 \paragraph{Verbal Structures Only?}
@@
 
-We conducted an analysis of the English UCCA Wikipedia
+This focus on verbal predicates has a number of unfortunate consequences. We show an example below:
+
+\begin{center}
+\begin{tabular}{lp{5.4cm}}
+Source & \small a coronary angioplasty may not be technically possible \\
+Transl.& \small eine koronare Angioplastie kann nicht technisch m{\"o}glich \\
+Gloss & \small a coronary angioplasty can not technically possible \\
+\end{tabular}
+\end{center}
+
+The German machine translation 
+is largely correct, except that the main verb ``sein'' (``be'') is omitted.
+While this may be interpreted as a minor error, HMEANT will assign the
+sentence a very low score, as it failed to translate the main verb.
+
+There are however a number of other serious consequences on focussing on verbal
+predicates. We conducted an analysis of the English UCCA Wikipedia
 corpus, comprising
 of 5324 sentences,
-in order to assess their frequency: \begin{inparadesc}
+in order to assess the impact of the following problems: 
+\\
+\begin{inparadesc}
 \item[Copula clauses] are treated in HMEANT simply as instances of the main verb ``be'', which
@@
 according to conservative estimates that only considers non-auxilliary instances of ``be''.
+\\
 \item[Nominative argument structures,] which are completely ignored under HMEANT are in fact highly
 pervasive, appearing in 48.7\% of the sentences.
+\\
 \item[Linkers] expressing an inter-relation between
@@
 
-It is also relatively common in translation that verbal constructions get
-nominalized or vice versa. An example from our dataset:
-  ``Two {\bf review} authors assessed the studies independently\dots{}'' --
-  ``Zwei Autoren bewertete die Studien unabh\"angig zu {\bf
-  \"uberpr\"ufen}\dots{}'' (lit. ``Two authors assessed the studies
-  independently to review\dots").\XXX{This is Omri's example, with
-  Ondrej's gloss for German. Ondrej is
-  not totally sure that the meaning is well preserved. So Ondrej would actually
-  *remove* this example, because it would illustrate HMEANT's problems only if
-  the meaning was correct.}
-
-Another example where a linker is omitted in translation:
-the source \equo{However, this review\dots}, which was translated as \equo{Diese
-\"Uberpr\"ufung\dots} (lit. \equo{This review}).
-This omission will not be taken into account by HMEANT at all.
+% No space
+%It is also relatively common in translation that verbal constructions get
+%nominalized or vice versa. An example from our dataset:
+%  ``Two {\bf review} authors assessed the studies independently\dots{}'' --
+%  ``Zwei Autoren bewertete die Studien unabh\"angig zu {\bf
+%  \"uberpr\"ufen}\dots{}'' (lit. ``Two authors assessed the studies
+%  independently to review\dots").\XXX{This is Omri's example, with
+%  Ondrej's gloss for German. Ondrej is
+%  not totally sure that the meaning is well preserved. So Ondrej would actually
+%  *remove* this example, because it would illustrate HMEANT's problems only if
+%  the meaning was correct.}
+
+%Another example where a linker is omitted in translation:
+%the source \equo{However, this review\dots}, which was translated as \equo{Diese
+%\"Uberpr\"ufung\dots} (lit. \equo{This review}).
+%This omission will not be taken into account by HMEANT at all.
 
@@
 taken alone, captures the crux of the sentence semantics meanings.
-The coverage of
-these and other phenomena by HUME will allow to empirically determine 
-(rather than stipulate), which semantic structures are important
-to preserve for MT evaluation, and which are less important.
+% I am not sure that this is true
+%The coverage of
+%these and other phenomena by HUME will allow to empirically determine 
+%(rather than stipulate), which semantic structures are important
+%to preserve for MT evaluation, and which are less important.
 
@@
 two independently constructed structural annotations (for both the reference and
-the candidate translation).
-We see this difference as crucial where semantic analysis is concerned, as it
-is often impossible to assign a semantic structure to a low quality MT output sentence.
-
-A design principle of HMEANT states that
-if the translation cannot be annotated with verbal argument structure
-(for whatever reason, including incomprehensibility), it cannot convey the meaning,
-while HUME does not share this theoretical commitment.
-For instance, consider the sentence ``a coronary angioplasty may not be technically possible'',
-and the translation ``eine koronare Angioplastie kann nicht technisch m{\"o}glich''
-(lit. ``a coronary angioplasty can not technically possible''),
-produced by an MT system in our experiments.
-The translation is correct, except that the main verb ``sein'' (``be'') is omitted.
-While this may be interpreted as a minor error, HMEANT will assign the
-sentence a very low score, as it failed to translate the main verb.
+the candidate translation)
+%First of all, the annotation process is different. We semantically annotate the
+%source, then automatically project it to the translation, and ask bilingual
+%annotators to assess whether this projection shows a good or bad translation. In
+%HMEANT, they semantically annotate both the reference and the translation, then
+%align these manually, and assess to what extent they match.
+%We see this difference as crucial where semantic analysis is concerned, as it
+and it is often impossible to assign a semantic structure to a low quality MT output sentence.
 
@@
 always literal enough and the MT output semantics can be composed differently
-than the semantics of the reference. While HUME avoids this problem,
-imposing a structure only on the source does prevent any fine-grained semantic
-analysis of the translation.
-
-Perhaps more importantly, references are often not created by translating the source
-sentences. In some cases, the manual translation was actually carried out in the
-other direction, from ``reference'' to ``source''
-and in others the ``source'' and ``reference'' are independent translations
-from a third language. ``References'' can thus easily contain information
-that does not appear in the ``source''.
-
-\paragraph{Annotation of Predicates Themselves}
-
-HUME allows to annotate multi-word and even
-non-contiguous elements as a single unit. HMEANT, on the other hand, always
-required exactly one word to be annotated as the predicate (``Action''), which
-is often linguistically inadequate.
-
-Another problem with HMEANT identified by \perscite{bojar:wu:ssst:2012} was the
-inability to indicate that the predicate is somewhat mistranslated (e.g.
-accidentally negated or in a slightly wrong tense). HMEANT
-annotation allows either to mark the source and reference frames as matching
-(which implies that the corresponding predicates are correct translations of each
-other), or as non-matching (which leads to a zero score for the whole frame) but
-nothing in between. In HUME, the verb is still present as a sub-unit of the
-frame and has the traffic-lights annotation.
+than the semantics of the reference.
+
+If only monolingual annotators are available, the HUME evaluation can be performed
+with a reference sentence instead of with the source. 
+% While HUME avoids this problem,
+%imposing a structure only on the source does prevent any fine-grained semantic
+%analysis of the translation.
+
+%Perhaps more importantly, references are often not created by translating the source
+%sentences. In some cases, the manual translation was actually carried out in the
+%other direction, from ``reference'' to ``source''
+%and in others the ``source'' and ``reference'' are independent translations
+%from a third language. ``References'' can thus easily contain information
+%that does not appear in the ``source''.
+
+\paragraph{Localisation of the Error}
+
+In HMEANT there are no guidelines as to when to propagate and error.
+In HUME errors are localised as far as possible to where they occur, by the separation of atomic and structural nodes.
+This allows us to calculate more acurately the total score of the sentence.
+
+%I see this a falling under the verbal structures point
+%\paragraph{Annotation of Predicates Themselves}
+
+%HUME allows to annotate multi-word and even
+%non-contiguous elements as a single unit. HMEANT, on the other hand, always
+%required exactly one word to be annotated as the predicate, which
+%is often linguistically inadequate.
+
+%Another problem with HMEANT identified by \perscite{bojar:wu:ssst:2012} was the
+%inability to indicate that the predicate is somewhat mistranslated (e.g.
+%accidentally negated or in a slightly wrong tense). HMEANT
+%annotation allows either to mark the source and reference frames as matching
+%(which implies that the corresponding predicates are correct translations of each
+%other), or as non-matching (which leads to a zero score for the whole frame) but
+%nothing in between. In HUME, the verb is still present as a sub-unit of the
+%frame and has the traffic-lights annotation.
 

### 2016-03-18 paper/main.tex

@@
 
-\begin{table}
+\begin{table}[t]
 \begin{center}
@@
 
-%\oa{what is the purpose of the BLEU binning? is it because low quality and high
-%quality translations often diverge in their agreement scores?}
-
-In order to assess the consistency of the annotation, we measure the Inter-Annotator
-Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
-For German, Polish and Romanian we used two annotators, with a significant overlap
-between the sentences they annotated. For Czech we had one main annotator, who
-annotated the entire batch, as well multiple secondary annotators, where
-each annotated a sub-part of the batch. \oa{i editted it. is what i wrote correct?}
-
-In order to compute Kappa, we consider the annotation task as a classification task on 
-nodes, and compute agreement on nodes which have two annotations from different annotators.
-We report the overall Kappa, as well as separate Kappas on atomic
-nodes (annotated as Red, Orange or Green) and the structural nodes (annotated
-as Adequate or Bad). In general there should be little confusion between the two
-types of nodes, and in future work the separation between the nodes should be enforced
-by the tool\oa{but there may be some disagreement in the margins on what nodes are atomic and
-  what are not}. The Kappas are shown in \tabref{tab:iaa}.
-
-\begin{table}[!ht]
+\begin{table}[t]
 \begin{center}
@@
 
+%\oa{what is the purpose of the BLEU binning? is it because low quality and high
+%quality translations often diverge in their agreement scores?}
+
+In order to assess the consistency of the annotation, we measure the Inter-Annotator
+Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
+For German, Polish and Romanian we used two annotators, with a significant overlap
+between the sentences they annotated. For Czech we had the entire batch
+annotated by one main annotator and multiple secondary annotators, 
+each annotating a sub-part of the batch.
+
+\tabref{tab:iaa} reports the number of nodes which have two annotations from
+different annotators and the corresponding Kappas.
+%
+%In order to compute Kappa, we consider the annotation task as a classification task on 
+%nodes, and compute agreement on nodes which have two annotations from different annotators.
+We report the overall Kappa, as well as separate Kappas on atomic
+nodes (annotated as Red, Orange or Green) and structural nodes (annotated
+as Adequate or Bad).
+As expected and confirmed by confusion matrices in \figref{fig:heatmap}, there
+is generally little confusion between the two types of nodes.
+
+%In general there should be little confusion between the two
+%types of nodes, and in future work the separation between the nodes should be enforced
+%by the tool.
+%\oa{but there may be some disagreement in the margins on what nodes are atomic and
+%  what are not}
+%The Kappas are shown in \tabref{tab:iaa}.
+
+\def\iaafig #1{\includegraphics[width=3.8cm]{iaa_heatmap_#1.png}}
+
+\begin{figure}[t]
+\renewcommand{\tabcolsep}{0pt}
+\begin{tabular}{cc}
+
+\subfloat[English-Czech]{
+  \iaafig{cs}
+}
+&
+\subfloat[English-German]{
+  \iaafig{de}
+
+}
+\\
+
+\subfloat[English-Polish]{
+  \iaafig{pl}
+
+}
+&
+\subfloat[English-Romanian]{
+  \iaafig{ro}
+
+}
+\end{tabular}
+\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
+\label{fig:heatmap}
+\end{figure}
+
 To assess whether, as we hoped, HUME would make it easier to reliably annotate long sentences,

### 2016-03-18 paper/main.tex
edits to section 5

--- 2016-03-18 paper/main.tex
@@
 \end{tabular}
-}
 \caption{Total numbers of sentences and nodes annotated on the MT output.}
 \label{tab:annot}
+}
 \end{center}
@@
 
-\begin{table}
+\begin{table}[t]
 \begin{center}
+{\small
 \begin{tabular}{l|cccc}
@@
 \label{tab:annot_times}
+}
 \end{center}
@@
 
-%\oa{what is the purpose of the BLEU binning? is it because low quality and high
-%quality translations often diverge in their agreement scores?}
-
-In order to assess the consistency of the annotation, we measure the Inter-Annotator
-Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
-For German, Polish and Romanian we used two annotators, with a significant overlap
-between the sentences they annotated. For Czech we had the entire batch
-annotated by one main annotator and multiple secondary annotators, 
-each annotating a sub-part of the batch.
-
-\tabref{tab:iaa} reports the number of nodes which have two annotations from
-different annotators and the corresponding Kappas.
-%
-%In order to compute Kappa, we consider the annotation task as a classification task on 
-%nodes, and compute agreement on nodes which have two annotations from different annotators.
-We report the overall Kappa, as well as separate Kappas on atomic
-nodes (annotated as Red, Orange or Green) and structural nodes (annotated
-as Adequate or Bad).
-As expected and confirmed by confusion matrices in \figref{fig:heatmap}, there
-is generally little confusion between the two types of nodes.
-
-%In general there should be little confusion between the two
-%types of nodes, and in future work the separation between the nodes should be enforced
-%by the tool.
-%\oa{but there may be some disagreement in the margins on what nodes are atomic and
-%  what are not}
-%The Kappas are shown in \tabref{tab:iaa}.
-
 \begin{table}[t]
 \begin{center}
+{\small
 \begin{tabular}{l|cccc}
@@
 \label{tab:iaa}
+}
 \end{center}
 \end{table}
+
+%\oa{what is the purpose of the BLEU binning? is it because low quality and high
+%quality translations often diverge in their agreement scores?}
+
+In order to assess the consistency of the annotation, we measure the Inter-Annotator
+Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
+For German, Polish and Romanian we used two annotators, with a significant overlap
+between the sentences they annotated. For Czech we had the entire batch
+annotated by one main annotator and multiple secondary annotators, 
+each annotating a sub-part of the batch.
+
+\tabref{tab:iaa} reports the number of nodes which have two annotations from
+different annotators and the corresponding Kappas.
+%
+%In order to compute Kappa, we consider the annotation task as a classification task on 
+%nodes, and compute agreement on nodes which have two annotations from different annotators.
+We report the overall Kappa, as well as separate Kappas on atomic
+nodes (annotated as Red, Orange or Green) and structural nodes (annotated
+as Adequate or Bad).
+As expected and confirmed by confusion matrices in \figref{fig:heatmap}, there
+is generally little confusion between the two types of nodes.
+
+%In general there should be little confusion between the two
+%types of nodes, and in future work the separation between the nodes should be enforced
+%by the tool.
+%\oa{but there may be some disagreement in the margins on what nodes are atomic and
+%  what are not}
+%The Kappas are shown in \tabref{tab:iaa}.
 
@@
 
-HMEANT shares much of HUME's motivation in defining a
-measure that decomposes over semantic units, rather than strings. On the other
-hand, there are important differences in design decisions, which we summarize in this
+Like HMEANT, HUME defines  a
+measure that decomposes over semantic units, rather than strings. 
+However, there are important differences in design decisions, which we summarize in this
 section.
 
-First of all, the annotation process is different. We semantically annotate the
+First, the annotation process is different. In HUME, we semantically annotate the
 source, then automatically project it to the translation, and ask bilingual
-annotators to assess whether this projection shows a good or bad translation. In
-HMEANT, they semantically annotate both the reference and the translation, then
+annotators to assess the quality of the projected translation. In
+HMEANT, they semantically annotate both the reference and the translation,
 align these manually, and assess to what extent they match.
@@
 Czech and by \perscite{chuchunkov-tarelkin-galinskaya:2014:SSST-8} for Russian)
-but the problems are apparent in English as well.
+but the problems are also seen in English.
 
@@
 We are not aware of any empirical evaluation that demonstrates that verb argument structure,
-taken alone, captures the crux of the sentence semantics meanings.
+taken alone, captures the crux of the sentence semantics.
 The coverage of
-these and other phenomena by HUME will allow to empirically determine 
-(rather than stipulate), which semantic structures are important
-to preserve for MT evaluation, and which are less important.
-
-\paragraph{Cost of Structural Annotation}
-
-HMEANT structural annotation (semantic role labelling, SRL), similar to e.g.
-PropBank
-\parcite{palmer2010semantic}, is designed to be very simple and easy to explain
-to lay annotators.
-HUME relies on UCCA, which is arguably more abstract, harder to
-explain, and thus more expensive to obtain.
-
-On the other hand, HUME 
-needs the semantic annotation
-done once for a given source sentence, whereas for the HMEANT set-up, every new
-MT output has to be structurally annotated.\oa{we can give some empirical data here.
-  In the WMT 13 experiment, Lexi and Barry report a little over 3 mins of annotation
-  per sentence + 3 translations. HUME annotation is about 2-4 minutes per sentence.
-  UCCA annotation by trained annotators is about 700 tokens/hour (we didn't
-  check it for individual sentences), but that's done only once so it should be amortized.}
+these and other phenomena by HUME  make it possible to 
+empirically determine 
+(rather than stipulate), the relative importance of 
+semantic structures for MT evaluation.
+
+%
+%\paragraph{Cost of Structural Annotation}
+%
+%HMEANT structural annotation (semantic role labelling, SRL), similar to e.g.
+%PropBank
+%\parcite{palmer2010semantic}, is designed to be very simple and easy to explain
+%to lay annotators.
+%HUME relies on UCCA, which is arguably more abstract, harder to
+%explain, and thus more expensive to obtain.
+%
+%On the other hand, HUME 
+%needs the semantic annotation
+%done once for a given source sentence, whereas for the HMEANT set-up, every new
+%MT output has to be structurally annotated.\oa{we can give some empirical data here.
+%  In the WMT 13 experiment, Lexi and Barry report a little over 3 mins of annotation
+%  per sentence + 3 translations. HUME annotation is about 2-4 minutes per sentence.
+%  UCCA annotation by trained annotators is about 700 tokens/hour (we didn't
+%  check it for individual sentences), but that's done only once so it should be amortized.}
 
@@
 
-In HMEANT, the alignment between the (reference and candidate) structures is a key
-part of the manual annotation. If the alignment cannot be created, the candidate
-translation is heavily penalized in score.
+In HMEANT, the alignment between the (reference and translation) structures is a key
+part of the manual annotation. If the alignment cannot be created, the 
+translation is heavily penalized.
 Previous experiments with HMEANT
@@
 %to avoid unduly penalization due to such a spurious divergence.
-The same issue was observed by HMEANT authors themselves
+The same issue was seen by HMEANT authors themselves
 \parcite{lo:wu:reliability:2014}:
@@
 inability to indicate that the predicate is somewhat mistranslated (e.g.
-accidentally negated or in a slightly wrong tense). HMEANT
-annotation allows either to mark the source and reference frames as matching
-(which implies that the corresponding predicates are correct translations of each
-other), or as non-matching (which leads to a zero score for the whole frame) but
-nothing in between. In HUME, the verb is still present as a sub-unit of the
+accidentally negated or in a slightly wrong tense). HMEANT annotation only allows
+for a match or non-match between reference and translation predicates, with non-match
+giving zero score for the whole frame.
+In HUME, the verb is still present as a sub-unit of the
 frame and has the traffic-lights annotation.

### 2016-03-18 paper/main.tex
updated style, natbib citations

--- 2016-03-18 paper/main.tex
@@
 \usepackage{acl2016}
+\usepackage[round]{natbib}
 \usepackage{amsmath}
@@
 
-\def\perscite#1{\newcite{#1}}
-\def\parcite#1{\cite{#1}}
-\def\inparcite#1{\newcite{#1}}
+\def\perscite#1{\citet{#1}}
+\def\parcite#1{\citep{#1}}
+\def\inparcite#1{\citealp{#1}}
 
@@
 with
-the sentence length \cite{Bojar:2011}.
+the sentence length \parcite{Bojar:2011}.
 Second, a sentence-level quality score does not indicate what parts of the sentence
@@
 their verb argument structure; the most notable of these measures is
-HMEANT \cite{lo2011structured}.
+HMEANT \parcite{lo2011structured}.
 % SAVING SPACE:
@@
 Semantic units are defined according to the 
-UCCA scheme \cite{abend2013universal}, an appealing candidate for semantic analysis,
+UCCA scheme \parcite{abend2013universal}, an appealing candidate for semantic analysis,
 due to its cross-linguistic applicability, support for rapid annotation, and coverage
@@
 %this type is the MEANT measure, and its human variant HMEANT
-%\cite{lo2010evaluating,lo2011structured,Lo2011meant,lo2013meant}, which quantifies the
+%\parcite{lo2010evaluating,lo2011structured,Lo2011meant,lo2013meant}, which quantifies the
 %similarity between the output translation and a reference translation in terms of the similarity
@@
 %out of the the huge number of possible translations a sentence may have
-%\cite{dreyer2012hyter}.
+%\parcite{dreyer2012hyter}.
 
@@
 %evaluation that quantifies the adequacy of the translation in preserving the meaning 
-%of the source sentence, and its fluency as a sentence in the target language \cite{lopez2008statistical}.
+%of the source sentence, and its fluency as a sentence in the target language
+%\parcite{lopez2008statistical}.
 Human evaluation is generally done by ranking the outputs of multiple systems
-e.g., in the WMT tasks \cite{bojar2015findings}, or by giving
+e.g., in the WMT tasks \parcite{bojar2015findings}, or by giving
 adequacy/fluency scores to each translation, a procedure recently improved
-by \newcite{graham2015improving}.
+by \perscite{graham2015improving}.
 However, while providing the gold standard for MT evaluation, human evaluation
@@
 The most commonly used measures decompose over n-grams or individual words, e.g., 
-BLEU \cite{Papineni:2002}, NIST \cite{Doddington:2002} and METEOR  \cite{Banerjee:2005}.
+BLEU \parcite{Papineni:2002}, NIST \parcite{Doddington:2002} and METEOR
+\parcite{Banerjee:2005}.
 Another common approach is to determine the similarity between the reference and translation
-by applying string edit distance \cite{snover2006study}.
+by applying string edit distance \parcite{snover2006study}.
 While these measures stimulated much progress in MT research by allowing
@@
 the similarity of the reference and translation in terms
-of their structure. \newcite{liu2005syntactic} took a syntactic approach, 
+of their structure. \perscite{liu2005syntactic} took a syntactic approach, 
 using dependency grammar, and
-\newcite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
-\newcite{gimenez2007linguistic} proposed to combine multiple types of information,
+\perscite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
+\perscite{gimenez2007linguistic} proposed to combine multiple types of information,
 capturing the overlap between the translation and reference in terms of their
@@
 that can be reliably annotated by non-experts, after as little as two hours
-of training \cite{marinotti2014}.
+of training \parcite{marinotti2014}.
 Second, UCCA is a cross-linguistically applicable scheme, seeking to
@@
 by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
-\cite{Dixon:10a,Dixon:10b,Dixon:12}.
+\parcite{Dixon:10a,Dixon:10b,Dixon:12}.
 The cross-linguistic applicability of UCCA has been tested in annotations of
@@
 in translation: UCCA annotations of translated text usually contain the same set of relationships
-\cite{sulem2015conceptual}. This indicates that the scheme reflects
+\parcite{sulem2015conceptual}. This indicates that the scheme reflects
 a common layer of representation that in a correct translation
@@
 
-The Abstract Meaning Representation (AMR) scheme \cite{banarescu2013abstract}
+The Abstract Meaning Representation (AMR) scheme \parcite{banarescu2013abstract}
 shares much of UCCA's motivation for defining a more complete semantic annotation.
@@
 
-%\newcite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
+%\perscite{Basile:12} attempted to make some aspects of the semantic annotation more accessible,
 %but to our knowledge no results of this brand have been reported.
@@
 For each of the four language pairs under consideration  we built phrase-based MT systems
-using Moses \cite{Koehn:2007}. We used large training sets extracted from OPUS \cite{tiedemann:2009}, and
-the data sets released for the WMT14 medical translation task \cite{bojar-EtAl:2014:W14-33}, to provide between
+using Moses \parcite{Koehn:2007}. We used large training sets extracted from
+OPUS \parcite{tiedemann:2009}, and
+the data sets released for the WMT14 medical translation task \parcite{bojar-EtAl:2014:W14-33}, to provide between
 45 and 85 million sentences of training data, depending on language pair.
@@
 done once for a given source sentence, whereas for the HMEANT set-up, every new
-MT output has to be structurally annotated.\oa{we can give some empirical data here.
-  In the WMT 13 experiment, Lexi and Barry report a little over 3 mins of annotation
-  per sentence + 3 translations. HUME annotation is about 2-4 minutes per sentence.
-  UCCA annotation by trained annotators is about 700 tokens/hour (we didn't
-  check it for individual sentences), but that's done only once so it should be amortized.}
+MT output has to be structurally annotated.
+
+%\oa{we can give some empirical data here.
+%  In the WMT 13 experiment, Lexi and Barry report a little over 3 mins of annotation
+%  per sentence + 3 translations. HUME annotation is about 2-4 minutes per sentence.
+%  UCCA annotation by trained annotators is about 700 tokens/hour (we didn't
+%  check it for individual sentences), but that's done only once so it should be amortized.}
 
@@
 %(AMR; \inparcite{banarescu2013amr}). 
-\cite{banarescu2013amr}. 
+\parcite{banarescu2013amr}. 
 AMR is being actively developed with 
@@
 %(SRL; \inparcite{palmer2010semantic}).
-\cite{palmer2010semantic}.
-SRL has been used in a human translation metric called HMEANT~\cite{lo2011structured}. 
+\parcite{palmer2010semantic}.
+SRL has been used in a human translation metric called HMEANT~\parcite{lo2011structured}. 
 HMEANT uses semantic role labels
@@
  When looking at an automatically parsed  English-Chinese  parallel corpora,
- it was shown that 8.7\% of the arguments do not preserve their semantic roles~\cite{fung2006automatic}. 
+ it was shown that 8.7\% of the arguments do not preserve their semantic
+ roles~\parcite{fung2006automatic}. 
 \end{itemize}
@@
 \bibliography{main}
-\bibliographystyle{acl2016}
+\bibliographystyle{acl_natbib}
 

### 2016-03-18 paper/main.tex

--- 2016-03-18 paper/main.tex
@@
 \begin{center}
+{\small
 \begin{tabular}{l|cccc}
@@
 \label{tab:iaa}
+}
 \end{center}
@@
 
-HMEANT shares much of HUME's motivation in defining a
-measure that decomposes over semantic units, rather than strings. On the other
-hand, there are important differences in design decisions, which we summarize in this
+Like HMEANT, HUME defines  a
+measure that decomposes over semantic units, rather than strings. 
+However, there are important differences in design decisions, which we summarize in this
 section.
@@
 
-\paragraph{Cost of Structural Annotation}
-
-HMEANT structural annotation (semantic role labelling, SRL), similar to e.g.
-PropBank
-\parcite{palmer2010semantic}, is designed to be very simple and easy to explain
-to lay annotators.
-HUME relies on UCCA, which is arguably more abstract, harder to
-explain, and thus more expensive to obtain.
-
-On the other hand, HUME 
-needs the semantic annotation
-done once for a given source sentence, whereas for the HMEANT set-up, every new
-MT output has to be structurally annotated.\oa{we can give some empirical data here.
-  In the WMT 13 experiment, Lexi and Barry report a little over 3 mins of annotation
-  per sentence + 3 translations. HUME annotation is about 2-4 minutes per sentence.
-  UCCA annotation by trained annotators is about 700 tokens/hour (we didn't
-  check it for individual sentences), but that's done only once so it should be amortized.}
+%
+%\paragraph{Cost of Structural Annotation}
+%
+%HMEANT structural annotation (semantic role labelling, SRL), similar to e.g.
+%PropBank
+%\parcite{palmer2010semantic}, is designed to be very simple and easy to explain
+%to lay annotators.
+%HUME relies on UCCA, which is arguably more abstract, harder to
+%explain, and thus more expensive to obtain.
+%
+%On the other hand, HUME 
+%needs the semantic annotation
+%done once for a given source sentence, whereas for the HMEANT set-up, every new
+%MT output has to be structurally annotated.\oa{we can give some empirical data here.
+%  In the WMT 13 experiment, Lexi and Barry report a little over 3 mins of annotation
+%  per sentence + 3 translations. HUME annotation is about 2-4 minutes per sentence.
+%  UCCA annotation by trained annotators is about 700 tokens/hour (we didn't
+%  check it for individual sentences), but that's done only once so it should be amortized.}
 

### 2016-03-18 paper/main.tex

--- 2016-03-18 paper/main.tex
@@
 
-\paragraph{Cost of Structural Annotation}
-
-HMEANT structural annotation (semantic role labelling, SRL), similar to e.g.
-PropBank
-\parcite{palmer2010semantic}, is designed to be very simple and easy to explain
-to lay annotators.
-HUME relies on UCCA, which is arguably more abstract, harder to
-explain, and thus more expensive to obtain.
-
-On the other hand, HUME 
-needs the semantic annotation
-done once for a given source sentence, whereas for the HMEANT set-up, every new
-MT output has to be structurally annotated.
+%
+%\paragraph{Cost of Structural Annotation}
+%
+%HMEANT structural annotation (semantic role labelling, SRL), similar to e.g.
+%PropBank
+%\parcite{palmer2010semantic}, is designed to be very simple and easy to explain
+%to lay annotators.
+%HUME relies on UCCA, which is arguably more abstract, harder to
+%explain, and thus more expensive to obtain.
+%
+%On the other hand, HUME 
+%needs the semantic annotation
+%done once for a given source sentence, whereas for the HMEANT set-up, every new
+%MT output has to be structurally annotated.\oa{we can give some empirical data here.
 

### 2016-03-18 paper/main.tex
squashing space

@@
 
-These translation systems were used to translate texts derived from both NHS 24\footnote{\url{http://www.nhs24.com/}} and Cochrane\footnote{\url{http://www.cochrane.org/}} into the four languages.
-The first organisation is a public body providing healthcare and health-service related information in Scotland, whereas the second is an international NGO which
+These translation systems were used to translate texts derived from both NHS
+24\footnote{\url{http://www.nhs24.com/}} and
+Cochrane\footnote{\url{http://www.cochrane.org/}} into the four languages.
+NHS~24 is a public body providing healthcare and health-service
+related information in Scotland, Cochrane is an international NGO
+which 
 provides independent systematic reviews on health-related research. From NHS 24, we selected the segments from their
@@
 For IAA also for
-the Czech, we recruited several further annotators to work on a small number of 
+Czech, several further annotators worked on a small number of 
 sentences each. We treat these further annotators as one annotator, giving us two annotators
@@
 \end{tabular}
-\caption{Inter-annotator Agreement (Kappa) versus sentence length, separately for
+\caption{Kappa versus sentence length, separately for
 structural and atomic nodes. The numbers are the node counts in each bin. }
@@
 
-This focus on verbal predicates has a number of unfortunate consequences. We show an example below:
+%This focus on verbal predicates has a number of unfortunate consequences. We show an example below:
+Consider the example:
+
+% \noindent
+% \begin{inparadesc}
+% \item[Source ] a coronary angioplasty may not be technically possible \\
+% \item[Transl.] eine koronare Angioplastie kann nicht technisch m{\"o}glich \\
+% \item[Gloss ]  a coronary angioplasty can not technically possible \\
+% \end{inparadesc}
 
@@
 predicates. We conducted an analysis of the English UCCA Wikipedia
-corpus, comprising
-of 5324 sentences,
+corpus (5324 sentences)
 in order to assess the impact of the following problems: 
@@
 \\
-\item[Nominative argument structures,] which are completely ignored under HMEANT are in fact highly
+\item[Nominal argument structures,] which are completely ignored under HMEANT, are in fact highly
 pervasive, appearing in 48.7\% of the sentences.
@@
 the 
-inter-annotator agreement on SRL dropped from 90\% to 61\% 
+IAA on SRL dropped from 90\% to 61\% 
 when the two aligned structures were from two different annotators.

### 2016-03-18 paper/main.tex
edits

@@
 
-In HMEANT there are no guidelines as to when to propagate and error.
-In HUME errors are localised as far as possible to where they occur, by the separation of atomic and structural nodes.
-This allows us to calculate more acurately the total score of the sentence.
+In HMEANT, there are no guidelines as to when to propagate and error. In HUME,
+errors are localised as much as possible to where they occur, by the separation of atomic and structural nodes. This allows us to calculate more acurately the total score of the sentence.
 

### 2016-03-18 paper/main.tex
weak attempt at a conclusion

--- 2016-03-18 paper/main.tex
@@
 provides independent systematic reviews on health-related research. From NHS 24, we selected the segments from their
-``Health A-Z'' which is part of the NHS Inform website, and the Cochrane texts come from their plain language summaries
+``Health A-Z'' in the NHS Inform website, and the Cochrane texts come from their plain language summaries
 and abstracts.
@@
 \\
-\item[Nominative argument structures,] which are completely ignored under HMEANT are in fact highly
+\item[Nominative argument structures,] which are ignored by  HMEANT are in fact highly
 pervasive, appearing in 48.7\% of the sentences.
@@
 
-The paper concludes by outlining ongoing work to construct semi- and
-fully-automatic variants of HUME.
+We have introduced HUME, a human semantics based MT evaluation measure which addresses
+a wide range of semantic phenomena. We have shown that it can be reliably annotated in 
+several languages, and that annotation quality is robust to sentence length.
+We believe that HUME and a future automated version of HUME will be able to guide the
+development of a more semantically aware approach to MT.
 

### 2016-03-18 paper/main.tex

--- 2016-03-18 paper/main.tex
@@
 \\
-\item[Nominal argument structures,] which are completely ignored under HMEANT, are in fact highly
+\item[Nominal argument structures,] which are ignored by  HMEANT are in fact highly
 pervasive, appearing in 48.7\% of the sentences.

### 2016-03-18 paper/main.tex
spellcheck

@@
 governing and dependent units is often expressed using morphological
-properties of the dependents instead of their ordering
+properties of the dependants instead of their ordering
 (this holds especially for the verb and its modifiers,
@@
 We proceed to detailing how judgements on the semantic units
-of the source are aggregated into a composite score. We start by taking a parsiminous approach
+of the source are aggregated into a composite score. We start by taking a parsimonious approach
 %While future extensive empirical evaluation with HUME and its application to MT
@@
 $t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
-Where $a(u)$ is discontiguous in the text, words between the left
+Where $a(u)$ is discontinuous in the text, words between the left
 and right boundaries of $a(u)$ which are not contained in it (intervening words)
@@
 
-There are however a number of other serious consequences on focussing on verbal
+There are however a number of other serious consequences on focusing on verbal
 predicates. We conducted an analysis of the English UCCA Wikipedia
@@
 generally does not convey the meaning of these clauses. They appear in 21.7\% of the sentences,
-according to conservative estimates that only considers non-auxilliary instances of ``be''.
+according to conservative estimates that only considers non-auxiliary instances of ``be''.
 \\
@@
 sentences\footnote{Argument structures and linkers are explicitly marked in UCCA.
-We identified non-auxilliary instances of ``be'' and nouns using the NLTK standard tagger.
+We identified non-auxiliary instances of ``be'' and nouns using the NLTK standard tagger.
 Nominative argument structures are here defined as Scenes whose main relation is headed by a noun.}.
@@
 In HMEANT, there are no guidelines as to when to propagate and error. In HUME,
-errors are localised as much as possible to where they occur, by the separation of atomic and structural nodes. This allows us to calculate more acurately the total score of the sentence.
+errors are localised as much as possible to where they occur, by the separation of atomic and structural nodes. This allows us to calculate more accurately the total score of the sentence.
 

### 2016-03-18 paper/main.tex
minor corrections

@@
 The domain is particularly suitable for experimenting on MT evaluation, as semantically
-accurate translation is paramount here, underscoring the importance of accurate
+accurate translation is paramount here, underscoring the importance of reliable
 semantic evaluation measures.
@@
 For each of the four language pairs under consideration  we built phrase-based MT systems
-using Moses \parcite{Koehn:2007}. We used large training sets extracted from
+using Moses \parcite{Koehn:2007}.  These were trained on large parallel data sets extracted from
 OPUS \parcite{tiedemann:2009}, and
-the data sets released for the WMT14 medical translation task \parcite{bojar-EtAl:2014:W14-33}, to provide between
+the data sets released for the WMT14 medical translation task \parcite{bojar-EtAl:2014:W14-33}, 
+giving between 
 45 and 85 million sentences of training data, depending on language pair.
@@
 
-Source sentences are all in English, and their UCCA annotation was performed by four
+The source sentences are all in English, and their UCCA annotation was performed by four
 computational linguists and one linguist.

### 2016-03-18 paper/main.tex
minor changes, mostly to figure1

@@
   Sample UCCA annotation where leaves correspond to words and nodes to units.
-  Edge labels mark UCCA categories, listed  in the table.}
+  The dashed edge indicates that ``Tom'' is also a participant in the ``moved to NYC''
+  Scene. Edge labels mark UCCA categories, listed  in the table.}
 %which are largely ignored in this work
@@
 When evaluating ``to NYC'' the annotator looks at the translation and sees the
-word ``stateside''. This word captures the whole phrase and so we mark the
-structural node with an atomic label. Here we choose ``Orange'' since it approximately captures the meaning
-in this context. The ability to mark structural nodes with atomic labels allows
+word ``stateside''. This word captures the whole phrase and so we mark this
+non-leaf node with an atomic label. Here we choose ``Orange'' since
+it approximately captures the meaning
+in this context. The ability to mark non-leaves with atomic labels allows
 the annotator to account for translations which only correspond at the phrase
 level. Another feature highlighted in this example is that the root node is
-``Adequate''. This is because the root node still has two scenes linked by one linker and
+``Adequate''. This is because the root node still has two Scenes linked by one linker and
 so the relationship between child nodes is maintained. There is an error caused by a wrong
 translation of the linker ``After'' which translates as ``by'' and this is
-marked as ``Red'', but this error is marked at the point at which it occurs.
+marked as ``Red'', but this error is marked at the level at which it occurs.
 

### 2016-03-18 paper/main.tex
minor

@@
 clauses (mainly discourse markers and coordinating conjunctions) appear in 56\% of the
-sentences\footnote{Argument structures and linkers are explicitly marked in UCCA.
+sentences, but ignored by HMEANT.
+An example where a linker is omitted in translation in our experiments is
+with the source \equo{However, this review\dots}, which was translated to
+the German \equo{Diese \"Uberpr\"ufung\dots} (lit. \equo{This review}).
+This omission will not be taken into account by HMEANT at all\footnote{Argument structures and linkers are explicitly marked in UCCA.
 We identified non-auxiliary instances of ``be'' and nouns using the NLTK standard tagger.
@@
 %  the meaning was correct.}
-
-%Another example where a linker is omitted in translation:
-%the source \equo{However, this review\dots}, which was translated as \equo{Diese
-%\"Uberpr\"ufung\dots} (lit. \equo{This review}).
-%This omission will not be taken into account by HMEANT at all.
 
@@
 
-In HMEANT, there are no guidelines as to when to propagate and error. In HUME,
-errors are localised as much as possible to where they occur, by the separation of atomic and structural nodes. This allows us to calculate more accurately the total score of the sentence.
+In HMEANT, there are no guidelines as to when to propagate an error. In HUME,
+errors are localised as much as possible to where they occur,
+by the separation of atomic and structural nodes.
+This allows us to calculate more accurately the total score of the sentence.
 

### 2016-03-18 paper/main.tex
Edits to Intro

--- 2016-03-18 paper/main.tex
@@
 
-We propose the Human UCCA-based MT Evaluation (HUME) metric,
-a human evaluation measure that decomposes over the semantic units of the sentence.
-Semantic units are defined according to the 
-UCCA scheme \parcite{abend2013universal}, an appealing candidate for semantic analysis,
+We propose the HUME metric,
+a human evaluation measure that decomposes over the UCCA semantic units of the sentence.
+UCCA\parcite{abend2013universal} is an appealing candidate for semantic analysis,
 due to its cross-linguistic applicability, support for rapid annotation, and coverage
@@
 HUME operates by aggregating human assessments of the translation quality of individual
-semantic units in the source sentence.
-HUME only requires semantically annotating the source sentence,
-thus avoiding the annotation of machine-generated text 
-and allowing re-use of the source semantic annotation for measuring the quality
+semantic units in the source sentence. We 
+%HUME only requires semantically annotating the source sentence,
+thus avoiding the semantic annotation of machine-generated text. 
+This allows the re-use of the source semantic annotation for measuring the quality
 of different translations of the same source sentence.
@@
 efficiency (time of annotation) and inter-annotator agreement scores for HUME.
-Finally, the differences from HMEANT are discussed 
+Finally, the differences with HMEANT are discussed 
 (\secref{sec:hmeant_comp}).

### 2016-03-18 paper/main.tex
aaa

@@
 \usepackage{lipsum,adjustbox}
+
+\makeatletter
+\renewcommand{\paragraph}{
+  \@startsection{paragraph}{4}
+  {\z@}{.3ex \@plus .3ex \@minus .2ex}{-1em}
+  {\normalfont\normalsize\bfseries}
+}
+\makeatother
+
 \listfiles
@@
 by clicking the ``A'', ``B'', Green, Orange, or Red buttons to the right of the unit's box.
+Units with multiple parents (as with ``Tom'' in \figref{fig:hume_tree_v2}) are displayed
+twice, once under each of their parents, but are only annotatable in one of
+their instances, to avoid double counting.
 %Note that the UCCA labels are not shown.
@@
 shows the total number of sentences and nodes annotated by each annotator. Note
-that not all nodes in all sentences were annotated. Sometimes,
+that not all nodes in all sentences were annotated, often
+due to
 %this was because
-the annotator did not finish a sentence or accidentally missed a node, but in many cases it was also because it was an implicit 
-UCCA node, which was not shown to the annotators since there is no corresponding source segment.
+the annotator not finishing a sentence or accidentally missing a node.
+%, but in many cases it was also because it was an implicit 
+%UCCA node, which was not shown to the annotators since there is no corresponding source segment.
 \begin{table}

### 2016-03-18 paper/main.tex
Edits to Section 2

--- 2016-03-18 paper/main.tex
@@
 This decomposition allows system developers
-to localize the errors, or at least unconfirmed parts within the translation.
+to localize the errors.
+%, or at least unconfirmed parts within the translation.
 The most commonly used measures decompose over n-grams or individual words, e.g., 
@@
 the reference and translation in terms of the overlap in
-their verbal argument structures structures and associated semantic roles.
+their verbal argument structures and associated semantic roles.
 We discuss the differences between HMEANT and HUME in \secref{sec:hmeant_comp}.

### 2016-03-18 paper/main.tex
Final edits, this version is hopefully the final submitted version

@@
 L & Linker &  A & Participant \\
-H & Scene & R & Relator \\
+H & Scene & R & Relater \\
 P & Process & C & Centre \\
@@
 %\end{figure*}
-
-\begin{figure}
-    \begin{center}
-    \includegraphics[width=0.35\textwidth]{ucca-tree-mteval-v2.png}
-    \end{center}
-  \caption{\label{fig:hume_tree_v2}
-     HUME annotation of an UCCA tree with a word aligned example translation shown below. 
-Atomic units are labelled using traffic lights and structural nodes are marked A or B.}
-\end{figure}
-% Ondrej removes the following, he thinks that errors are not supposed to be
-% propagated in HMEANT either:
-%In HMEANT, it is not clear whether errors should be propegated up the SRL hierarchy.
 
@@
 
+\begin{figure}
+    \begin{center}
+    \includegraphics[width=0.35\textwidth]{ucca-tree-mteval-v2.png}
+    \end{center}
+  \caption{\label{fig:hume_tree_v2}
+     HUME annotation of an UCCA tree with a word aligned example translation shown below. 
+Atomic units are labelled using traffic lights and structural nodes are marked A or B.}
+\end{figure}
+% Ondrej removes the following, he thinks that errors are not supposed to be
+% propagated in HMEANT either:
+%In HMEANT, it is not clear whether errors should be propegated up the SRL hierarchy.
+
+In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME
+annotation, with the translation shown in English for ease of comprehension.
+When evaluating ``to NYC'' the annotator looks at the translation and sees the
+word ``stateside''. This word captures the whole phrase and so we mark this
+non-leaf node with an atomic label. Here we choose ``Orange'' since
+it approximately captures the meaning
+in this context. The ability to mark non-leaves with atomic labels allows
+the annotator to account for translations which only correspond at the phrase
+level. Another feature highlighted in this example is that by separating structural
+and atomic nodes, we are able to define where an error occurs, and localise
+the error to its point of origin. The linker ``After'' is translated incorrectly as ``by''
+which changes the meaning of the entire sentence. This error is captured at
+the atomic level, and it is labelled Red. The sentence still contains two scenes and
+a linker and therefore we mark the root node as structurally correct, Adequate.
+
 \begin{figure*}[t]
@@
 \end{figure*}
-
-In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME
-annotation, with the translation shown in English for ease of comprehension.
-When evaluating ``to NYC'' the annotator looks at the translation and sees the
-word ``stateside''. This word captures the whole phrase and so we mark this
-non-leaf node with an atomic label. Here we choose ``Orange'' since
-it approximately captures the meaning
-in this context. The ability to mark non-leaves with atomic labels allows
-the annotator to account for translations which only correspond at the phrase
-level. Another feature highlighted in this example is that the root node is
-``Adequate''. This is because the root node still has two Scenes linked by one linker and
-so the relationship between child nodes is maintained. There is an error caused by a wrong
-translation of the linker ``After'' which translates as ``by'' and this is
-marked as ``Red'', but this error is marked at the level at which it occurs.
 
@@
 Aligned translations are also shown in black alongside the source node.
-If the node is aligned to a set of discontiguous words in the translation, then the unaligned words that appear in between the aligned words are shown in red.
+If the node is aligned to a set of dis-contiguous words in the translation, then the unaligned words that appear in between the aligned words are shown in red.
 Even if these words are not directly aligned to the source node, they
@@
 Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
-For German, Polish and Romanian we used two annotators, with a significant overlap
-between the sentences they annotated. For Czech we had the entire batch
-annotated by one main annotator and multiple secondary annotators, 
-each annotating a sub-part of the batch.
+%For German, Polish and Romanian we used two annotators, with a significant overlap
+%between the sentences they annotated. For Czech we had the entire batch
+%annotated by one main annotator and multiple secondary annotators, 
+%each annotating a sub-part of the batch.
 
@@
 generally does not convey the meaning of these clauses. They appear in 21.7\% of the sentences,
-according to conservative estimates that only considers non-auxiliary instances of ``be''.
+according to conservative estimates that only consider non-auxiliary instances of ``be''.
 \\
@@
 clauses (mainly discourse markers and coordinating conjunctions) appear in 56\% of the
-sentences, but ignored by HMEANT.
+sentences, but are ignored by HMEANT.
 An example where a linker is omitted in translation in our experiments is
@@
 
-In HMEANT, there are no guidelines as to when to propagate an error. In HUME,
-errors are localised as much as possible to where they occur,
-by the separation of atomic and structural nodes.
-This allows us to calculate more accurately the total score of the sentence.
+In HMEANT, an error in a child node often results in the parent node
+being penalised as well. This makes it harder to quantify the true scale of  
+the original error, as its effect gets propagated up the tree. 
+In HUME, errors are localised as much as possible to where they occur,
+by the separation of atomic and structural nodes (See Fig \ref{fig:hume_tree_v2} 
+for an example).
+This allows us to calculate the total score of the sentence more accurately.
 
@@
 We have introduced HUME, a human semantics based MT evaluation measure which addresses
-a wide range of semantic phenomena. We have shown that it can be reliably annotated in 
+a wide range of semantic phenomena. We have shown that it can be efficiently and 
+reliably annotated in 
 several languages, and that annotation quality is robust to sentence length.
-We believe that HUME and a future automated version of HUME will be able to guide the
+We believe that HUME, and a future automated version of HUME, will be able to guide the
 development of a more semantically aware approach to MT.

### 2016-03-18 paper/main.tex
Typo

@@
 %HUME only requires semantically annotating the source sentence,
-thus avoiding the semantic annotation of machine-generated text. 
+are thus avoiding the semantic annotation of machine-generated text. 
 This allows the re-use of the source semantic annotation for measuring the quality
 of different translations of the same source sentence.
-By basing the measure on a direct comparison
-of the source and translation we avoid the difficulties created by using
+%By basing the measure on a direct comparison
+%of the source and translation we avoid the difficulties created by using
+We also avoid creating and relying on possibly suboptimal
 reference translations, albeit at the cost of requiring the

### 2016-05-29 scripts/compute_correlation_sent_len_diff_scores.py
ddd

@@ -1,4 +1,5 @@
-import sys, pdb
+import sys, pdb, math
+import numpy as np
 import scipy.stats
 
 
@@ -16,8 +17,11 @@ if __name__ == "__main__":
     
     ucca_ids = [int(x.strip()) for x in open(sys.argv[1]).readlines()]
     hume = [float(x.strip()) for x in open(sys.argv[2]).readlines()]
-    source_sent_lens = [len(x.strip().split()) for x in open(sys.argv[3]).readlines()]
-    trans_sent_lens = [len(x.strip().split()) for x in open(sys.argv[4]).readlines()]
+    source_sents = [x.strip() for x in open(sys.argv[3]).readlines()]
+    trans_sents = [x.strip() for x in open(sys.argv[4]).readlines()]
+    source_sent_lens = [len(x.split()) for x in source_sents]
+    trans_sent_lens = [len(x.split()) for x in trans_sents]
+    
     da_scores = []
     for ind,l in enumerate(open(sys.argv[5])):
         if ind == 0:
@@ -30,7 +34,7 @@ if __name__ == "__main__":
     # bin the sents by sent length
     bins_source_len = [[] for x in threshs] + [[]]
     for sid, score in da_scores:
-        bins_source_len[find_bin(threshs,source_sent_lens[sid])].append((sid,score,hume[sid]))
+        bins_source_len[find_bin(threshs,source_sent_lens[sid])].append((sid,score,hume[sid],ucca_ids[sid])) # (sid, DA score, hume score, ucca ID)
 
     pearson_by_bins_source = []
     for bin in bins_source_len:
@@ -38,11 +42,11 @@ if __name__ == "__main__":
             pearson_by_bins_source.append((None,0))
         else:
             pearson_by_bins_source.append((\
-                scipy.stats.pearsonr([x[1] for x in bin], [x[2] for x in bin])[0],len(bin)))
+                scipy.stats.spearmanr([x[1] for x in bin], [x[2] for x in bin])[0],len(bin)))
 
     bins_trans_len = [[] for x in threshs] + [[]]
     for sid, score in da_scores:
-        bins_trans_len[find_bin(threshs,trans_sent_lens[sid])].append((sid,score,hume[sid]))
+        bins_trans_len[find_bin(threshs,trans_sent_lens[sid])].append((sid,score,hume[sid],ucca_ids[sid]))
 
     pearson_by_bins_trans = []
     for bin in bins_trans_len:
[…]

### 2016-05-29 scripts/compute_len_correlations.sh

@@ -1,18 +1,18 @@
 
-python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-5.en-ro.csv 20
+python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-5.en-ro.csv 15:20:25
 
-python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-10.en-ro.csv 20
+python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-10.en-ro.csv 15:20:25
 
-python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-15.en-ro.csv 20
-
-
+python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-15.en-ro.csv 15:20:25
 
 
[…]

### 2016-05-30 scripts/calculate_correspondance_hume_DA.py
Added DA variance

@@ -55,12 +55,18 @@ def main():
 
       print ("Correference HUME DA for " + str(judgements) + " judge, en-" + code + ", size " + str(len(df['DA'].tolist())) + ": " 
                    + str(np.corrcoef(df['UCCA'].tolist(), df['DA'].tolist())[1,0]))
+
+      A = np.vstack([  df['UCCA'].tolist()  , np.ones(len(df['UCCA'].tolist()))]).T
+      m,c = np.linalg.lstsq(A, df['DA'].tolist())[0]
+      print (str(m) + " " + str(c))
+
       
       #print(df.head())
       plt.plot(df['UCCA'], df['DA'], '.')
       plt.xlabel('HUME scores')
       plt.ylabel('Averaged Direct Assesment scores')
-      plt.title('Human judgements for English-' + lang + ' system output')
+      plt.title('Human judgements for English-' + lang)
+      plt.plot(df['UCCA'], m*df['UCCA'] + c, 'r', label='Fitted line')
 
       
       fname=HUME_dir + '/humevsDA.' + str(judgements) + 'en-' + code + '.pdf'

### 2016-05-30 scripts/calculate_correspondance_hume_DA.py
Correlating with standardised scores

@@ -20,6 +20,8 @@ LANGCODES = ("Romanian","ro"),("German", "de"),
 HUME_dir = "data/mturkDA"
 DA_dir = "/Users/alexandrabirch/work/2016-YvetteGraham/segment-mteval/proc-hits/analysis"
 fileVersion = "5"
+TYPE = "raw"
+TYPE = "stnd"
 
 
 def main():
@@ -28,7 +30,7 @@ def main():
     for judgements in 5, 10, 15:
 
       UCCAfileName = HUME_dir + "/himl2015.en-" + code + ".uccascores" 
-      DAfileName = HUME_dir + "/ad-raw-seg-scores-" + str(judgements) + ".en-" + code + ".csv" 
+      DAfileName = HUME_dir + "/ad-" + TYPE + "-seg-scores-" + str(judgements) + ".en-" + code + ".csv" 
 
       UCCAresults = pandas.read_csv(UCCAfileName, header=None)
       DAresults = pandas.read_csv(DAfileName, sep='\s+', converters={'SID': int, 'SCR' : float})

### 2016-05-31 scripts/compute_correlation_sent_len_diff_scores.py
d

@@ -1,7 +1,7 @@
 import sys, pdb, math
 import numpy as np
 import scipy.stats
-
+import scipy.spatial.distance as distance
 
 def find_bin(threshs,val):
     """
@@ -112,12 +112,13 @@ if __name__ == "__main__":
                     print('\t'.join([str(x) for x in t]))
     """
     
-    # All sentences
+    # All sentences, ordinal outliers
     ordinal_thresh = 0.5
     all_sents = []
     for bin in bins_source_len:
         all_sents.extend(bin)
-    #print('===== All Bins ======== '+str(len(all_sents)))
+    
+    #print('===== All Bins Ordinal Outliers ======== '+str(len(all_sents)))
     ords1 = np.array([x[1] for x in all_sents]).argsort().argsort()
     ords2 = np.array([x[2] for x in all_sents]).argsort().argsort()
     print('\t'.join(['ordinal by DA', 'ordinal by HUME', 'source sentnece', 'translation','SID','DA score','HUME score','UCCA ID']))
@@ -127,3 +128,19 @@ if __name__ == "__main__":
             print('\t'.join([str(z) for z in t]))
 
 
+    # All sentences, 3-NN outliers
+
+    # compute the 3 NNs for each point (self excluded), and their average distance
+    # find the points with the highest / lowest distance
+    print('\nk-NN analysis\n')
+    k = 3
+    X = np.transpose(np.array([[x[1] for x in all_sents],[x[2] for x in all_sents]]))
+    D = distance.squareform(distance.pdist(X, 'euclidean'))
+    proximity = np.sum(np.partition(D,k+1)[:,:(k+1)],axis=1)
+    proximity_ords = proximity.argsort().argsort()
+    num_outliers = 5
+    for outlier_ind in range(num_outliers):
+        list_index = list(proximity_ords).index(outlier_ind)
+        x = all_sents[list_index]
+        t = (source_sents[x[0]],trans_sents[x[0]],x[0],x[1],x[2],x[3])
+        print('\t'.join([str(z) for z in t]))

### 2016-05-31 scripts/compute_len_correlations.sh

@@ -1,18 +1,18 @@
 
-python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-5.en-ro.csv 15:20:25
+#python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-5.en-ro.csv 15:25
 
 python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-10.en-ro.csv 15:20:25
 
-python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-15.en-ro.csv 15:20:25
+#python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-15.en-ro.csv 15:25
 
 
 
 
-python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-de.uccaids data/mturkDA/himl2015.en-de.uccascores data/mturkDA/himl2015.en-de.en data/mturkDA/himl2015.en-de.trans.de data/mturkDA/ad-stnd-seg-scores-5.en-de.csv  15:20:25
[…]

### 2016-05-31 scripts/compute_correlation_sent_len_diff_scores.py
ddd

@@ -11,9 +11,12 @@ def find_bin(threshs,val):
     
 
 if __name__ == "__main__":
-    if len(sys.argv) != 7:
-        print("compute_correlation_sent_len_diff_scores.py <uccaids> <hume scores> <source sents> <trans sents> <DA scores> <length thresholds (:-delimited)>")
+    if len(sys.argv) != 9:
+        print("compute_correlation_sent_len_diff_scores.py <uccaids> <hume scores> <source sents> <trans sents> <DA scores> <length thresholds (:-delimited)> <ordinal threshold> <num outliers>")
         sys.exit(-1)
+
+    ordinal_thresh = float(sys.argv[7])
+    num_outliers = int(sys.argv[8])
     
     ucca_ids = [int(x.strip()) for x in open(sys.argv[1]).readlines()]
     hume = [float(x.strip()) for x in open(sys.argv[2]).readlines()]
@@ -27,7 +30,8 @@ if __name__ == "__main__":
         if ind == 0:
             continue
         fields = l.strip().split()
-        da_scores.append((int(fields[0]),float(fields[2])))
+        da_scores.append((int(fields[0]),float(fields[2]))) # (sid,da score)
+        #da_scores.append((int(fields[8]),float(fields[9])))
 
     threshs = [int(x) for x in sys.argv[6].split(':')]
 
@@ -42,8 +46,8 @@ if __name__ == "__main__":
             pearson_by_bins_source.append((None,0))
         else:
             pearson_by_bins_source.append((\
-                scipy.stats.spearmanr([x[1] for x in bin], [x[2] for x in bin])[0],len(bin)))
-
+                scipy.stats.pearsonr([x[1] for x in bin], [x[2] for x in bin])[0],len(bin)))
+            
     bins_trans_len = [[] for x in threshs] + [[]]
     for sid, score in da_scores:
         bins_trans_len[find_bin(threshs,trans_sent_lens[sid])].append((sid,score,hume[sid],ucca_ids[sid]))
@@ -54,16 +58,11 @@ if __name__ == "__main__":
             pearson_by_bins_trans.append((None,0))
         else:
             pearson_by_bins_trans.append((\
-                scipy.stats.spearmanr([x[1] for x in bin], [x[2] for x in bin])[0],len(bin)))
+                scipy.stats.pearsonr([x[1] for x in bin], [x[2] for x in bin])[0],len(bin)))
     
-    print(pearson_by_bins_source)
[…]

### 2016-05-31 scripts/compute_len_correlations.sh

@@ -1,25 +1,0 @@
-
-#python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-5.en-ro.csv 15:25
-
-python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-10.en-ro.csv 15:20:25
-
-#python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-15.en-ro.csv 15:25
-
-
-
-
-#python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-de.uccaids data/mturkDA/himl2015.en-de.uccascores data/mturkDA/himl2015.en-de.en data/mturkDA/himl2015.en-de.trans.de data/mturkDA/ad-stnd-seg-scores-5.en-de.csv  15:25
-
-python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-de.uccaids data/mturkDA/himl2015.en-de.uccascores data/mturkDA/himl2015.en-de.en data/mturkDA/himl2015.en-de.trans.de data/mturkDA/ad-stnd-seg-scores-10.en-de.csv 15:20:25
-
-#python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-de.uccaids data/mturkDA/himl2015.en-de.uccascores data/mturkDA/himl2015.en-de.en data/mturkDA/himl2015.en-de.trans.de data/mturkDA/ad-stnd-seg-scores-15.en-de.csv 15:25
-
-
-
-
[…]

### 2016-05-31 paper/main.tex
emnlp format
Added the correlation graphs

@@
-\documentclass[11pt]{article}
-\usepackage{acl2016}
-\usepackage[round]{natbib}
+\documentclass[11pt,letterpaper]{article}
+%\usepackage[round]{natbib}
+\usepackage{emnlp2016}
 \usepackage{amsmath}
@@
 
+\def\emnlppaperid{***}
+
 \makeatletter
@@
 
-\def\perscite#1{\citet{#1}}
-\def\parcite#1{\citep{#1}}
-\def\inparcite#1{\citealp{#1}}
+\def\perscite#1{\newcite{#1}}
+\def\parcite#1{\cite{#1}}
+\def\inparcite#1{\cite{#1}}
 
@@
 We experiment with four language pairs, demonstrating HUME's broad applicability,
-as well as its reliability, reflected in its inter-annotator agreement rates.
+as well as its reliability, reflected in its inter-annotator agreement rates and 
+good correlation with human adequacy scores.
 
@@
 efficiency (time of annotation) and inter-annotator agreement scores for HUME. 
+We also report on the correlation scores of HUME with direct assesment of human 
+adequacy ratings. 
 Finally, the differences with HMEANT are discussed 
@@
 
+\subsection{Comparison with Crowdsourced Adequacy Judgements}\label{sec:adequacy}
+
+\def\iaafig #1{\includegraphics[width=0.4\textwidth]{humevsDA_10en-#1.pdf}}
+
+\begin{figure}[t]
+\renewcommand{\tabcolsep}{0pt}
+\begin{tabular}{c}
+\subfloat[English-German]{
+  \iaafig{de}
+}
+\\
+\subfloat[English-Romanian]{
+  \iaafig{ro}
+}
+\end{tabular}
+\caption{Plots of HUME scores vs Direct Assesment scores which have been standardised for each crowdsourcing annotator. The regression is calculated using linear least squares.}
+\label{fig:dacorrelation}
+\end{figure}
+Recent research~\cite{TODO} has proposed a new method for collecting 
+accuracy ratings. The idea is that collecting a large number of adequacy judgements along a 
+scale of 0 to 100 allows us to apply statistical measures which make them reliable. 
+It has been shown that mean  judgements correlate very strongly with each other. 
+We followed~\cite{TODO} and collected up to 15 adequacy judgements on most of the
+HUME annotated translations for German and Romainian. To ensure high quality judgments
+we restricted the IP of the annotators to those of the relevant country. We found that there were too
+few Czech and Polish annotators, and removing the IP restriction led to very poor quality results. 
+We can see that HUME and Direct Assesment scores do correlate. 
+We also removed unreliable crowdsourcing annotators
+according to the method discussed in ~\cite{TODO}. The Pearsons correlation for en-de is 0.58 and for en-ro is 0.70 which shows that they two measures do correlate. This confirms that 
+the HUME metric is consistent with previously published methods in human evaluation, even though these
+two approaches are very different in their conception.
+We argue that although annotating HUME requires more expertise and effort, 
+the results are more valuable because a HUME evaluation can tell you what parts
+of the source sentence have been retained.
+Direct assesment scores require a crowd, which is sometimes not available or reliable, and it returns a single number which gives little insight into the quality of
+the translated sentence.
+
 %\subsection{Discussion of Disagreements}\label{sec:disagreements}
-
-%\subsection{Comparison with Crowdsourced Adequacy Judgements}\label{sec:adequacy}
 
@@
 \bibliography{main}
-\bibliographystyle{acl_natbib}
+\bibliographystyle{emnlp2016}
 

### 2016-06-01 scripts/compute_correlation_sent_len_diff_scores.py
sss

@@ -11,8 +11,8 @@ def find_bin(threshs,val):
     
 
 if __name__ == "__main__":
-    if len(sys.argv) != 9:
-        print("compute_correlation_sent_len_diff_scores.py <uccaids> <hume scores> <source sents> <trans sents> <DA scores> <length thresholds (:-delimited)> <ordinal threshold> <num outliers>")
+    if len(sys.argv) not in [9,10]:
+        print("compute_correlation_sent_len_diff_scores.py <uccaids> <hume scores> <source sents> <trans sents> <DA scores> <length thresholds (:-delimited)> <ordinal threshold> <num outliers> <overlap sents>")
         sys.exit(-1)
 
     ordinal_thresh = float(sys.argv[7])
@@ -24,6 +24,10 @@ if __name__ == "__main__":
     trans_sents = [x.strip() for x in open(sys.argv[4]).readlines()]
     source_sent_lens = [len(x.split()) for x in source_sents]
     trans_sent_lens = [len(x.split()) for x in trans_sents]
+    if len(sys.argv) > 9:
+        overlap_sentids = [int(x.strip()) for x in open(sys.argv[9]).readlines()]
+    else:
+        overlap_sentids = None
     
     da_scores = []
     for ind,l in enumerate(open(sys.argv[5])):
@@ -31,14 +35,15 @@ if __name__ == "__main__":
             continue
         fields = l.strip().split()
         da_scores.append((int(fields[0]),float(fields[2]))) # (sid,da score)
-        #da_scores.append((int(fields[8]),float(fields[9])))
+
 
     threshs = [int(x) for x in sys.argv[6].split(':')]
 
     # bin the sents by sent length
     bins_source_len = [[] for x in threshs] + [[]]
     for sid, score in da_scores:
-        bins_source_len[find_bin(threshs,source_sent_lens[sid])].append((sid,score,hume[sid],ucca_ids[sid])) # (sid, DA score, hume score, ucca ID)
+        if overlap_sentids is None or ucca_ids[sid] in overlap_sentids :
+            bins_source_len[find_bin(threshs,source_sent_lens[sid])].append((sid,score,hume[sid],ucca_ids[sid])) # (sid, DA score, hume score, ucca ID)
 
     pearson_by_bins_source = []
     for bin in bins_source_len:
@@ -50,7 +55,8 @@ if __name__ == "__main__":
             
     bins_trans_len = [[] for x in threshs] + [[]]
[…]

### 2016-06-01 scripts/outlier_detection.sh
sssd

@@ -1,0 +1,14 @@
+
+echo 'Romanian'
+echo ''
+
+python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-ro.uccaids data/mturkDA/himl2015.en-ro.uccascores data/mturkDA/himl2015.en-ro.en data/mturkDA/himl2015.en-ro.trans.ro data/mturkDA/ad-stnd-seg-scores-10.en-ro.csv 15:20:25 0.34 10 ro_overlap_sents
+
+echo ''
+echo 'German'
+echo ''
+
+python compute_correlation_sent_len_diff_scores.py data/mturkDA/himl2015.en-de.uccaids data/mturkDA/himl2015.en-de.uccascores data/mturkDA/himl2015.en-de.en data/mturkDA/himl2015.en-de.trans.de data/mturkDA/ad-stnd-seg-scores-10.en-de.csv 15:20:25 0.34 10 de_overlap_sents
+
+
+

### 2016-06-02 scripts/calculate_correspondance_hume_DA.py
Added correlations for different subconditions

@@ -11,8 +11,13 @@ import logging
 import sys
 import pandas 
 import matplotlib.pyplot as plt
+from matplotlib import rcParams
+rcParams.update({'figure.autolayout': True})
 
 import numpy as np
+
+import extract_sentences_for_WMT_metrics_task
+
 
 LANGCODES = ("Romanian","ro"), ("Polish", "pl"), ("German", "de"), ("Czech", "cs")
 LANGCODES = ("Romanian","ro"),("German", "de"),
@@ -27,17 +32,22 @@ TYPE = "stnd"
 def main():
 
   for lang, code in LANGCODES:
-    for judgements in 5, 10, 15:
+    #for judgements in 5, 10, 15:
+    corrs = []
+    multicorrs = []
+    for judgements in [10]:
 
       UCCAfileName = HUME_dir + "/himl2015.en-" + code + ".uccascores" 
+      MultUCCAfileName = HUME_dir + "/himl2015.en-" + code + ".multuccascores" 
       DAfileName = HUME_dir + "/ad-" + TYPE + "-seg-scores-" + str(judgements) + ".en-" + code + ".csv" 
 
-      UCCAresults = pandas.read_csv(UCCAfileName, header=None)
+      UCCAresults = pandas.read_csv(UCCAfileName, index_col=False )
+      MultUCCAresults = pandas.read_csv(MultUCCAfileName, index_col=False, skip_blank_lines=False)
       DAresults = pandas.read_csv(DAfileName, sep='\s+', converters={'SID': int, 'SCR' : float})
       
       DAlist = []
       sizeDA = len(DAresults['SCR'])
-      sizeUCCA = len(UCCAresults[0])
+      sizeUCCA = len(UCCAresults["all"])
       for i in range (0, sizeUCCA):
         x = DAresults.loc[DAresults['SID'] == i,'SCR']
         if len(x) > 0:
@@ -45,37 +55,67 @@ def main():
         else:
           DAlist.append("")
       
+      count = -1
+      for score_type in extract_sentences_for_WMT_metrics_task.SCORE_TYPES:
+        count += 1
 
-      df = pandas.DataFrame({ 'UCCA' : UCCAresults[0],
-                          'DA' : DAlist }) 
+        df = pandas.DataFrame({ 'UCCA' : UCCAresults[score_type],
+                            'DA' : DAlist }) 
+        #df.corr(method='pearson', min_periods=1)
+        # df.plot(x='UCCA', y='DA')
[…]

### 2016-06-02 scripts/calculate_correspondance_hume_DA.py
sss

@@ -45,23 +45,22 @@ def main():
         else:
           DAlist.append("")
       
-
       df = pandas.DataFrame({ 'UCCA' : UCCAresults[0],
                           'DA' : DAlist }) 
-
+      
       
       df.corr(method='pearson', min_periods=1)
-
+      
      # df.plot(x='UCCA', y='DA')
       df = df[df['DA'] != ""]
-
+      
       print ("Correference HUME DA for " + str(judgements) + " judge, en-" + code + ", size " + str(len(df['DA'].tolist())) + ": " 
                    + str(np.corrcoef(df['UCCA'].tolist(), df['DA'].tolist())[1,0]))
 
       A = np.vstack([  df['UCCA'].tolist()  , np.ones(len(df['UCCA'].tolist()))]).T
       m,c = np.linalg.lstsq(A, df['DA'].tolist())[0]
       print (str(m) + " " + str(c))
-
+      
       
       #print(df.head())
       plt.plot(df['UCCA'], df['DA'], '.')

### 2016-06-02 paper/main.tex

@@
 %quantifying the overlap of its sub-parts relative a reference translation.
-A promising line of research decomposes
+A promising line of research decomposes metrics
 %More recent work on MT evaluation proposed measures
@@
 a human evaluation measure that decomposes over the UCCA semantic units of the sentence.
-UCCA\parcite{abend2013universal} is an appealing candidate for semantic analysis,
+UCCA \parcite{abend2013universal} is an appealing candidate for semantic analysis,
 due to its cross-linguistic applicability, support for rapid annotation, and coverage
@@
 We followed~\cite{TODO} and collected up to 15 adequacy judgements on most of the
-HUME annotated translations for German and Romainian. To ensure high quality judgments
+HUME annotated translations for German and Romanian. To ensure high quality judgments
 we restricted the IP of the annotators to those of the relevant country. We found that there were too
@@
 
-If only monolingual annotators are available, the HUME evaluation can be performed
+If only monolingual annotators are available, the HUME evaluation could be performed
 with a reference sentence instead of with the source. 
+\bh{But this would be a different metric}
 % While HUME avoids this problem,

### 2016-06-02 paper/main.tex
editting paper

@@
 \usepackage{lipsum,adjustbox}
+\usepackage[font={small}]{caption}
 
@@
 \newcommand{\com}[1]{}
-\newcommand{\secref}[1]{Section~\ref{#1}}
+\newcommand{\secref}[1]{\S\ref{#1}}
 \newcommand{\figref}[1]{Figure~\ref{#1}}
@@
 %\newcommand{\XXX}[1]{}
-\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
-%\newcommand{\oa}[1]{}
-\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
-%\newcommand{\bh}[1]{}
+%\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
+\newcommand{\oa}[1]{}
+%\newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
+\newcommand{\bh}[1]{}
 
@@
 \def\HUME{\func{HUME}}
-\def\Correct{\func{Correct}}
-\def\Partial{\func{Partial}}
+\def\Adequate{\func{Adequate}}
+\def\Green{\func{Green}}
+\def\Orange{\func{Orange}}
 \def\Units{\func{Units}}
@@
 or adequacy/fluency scales. However these provide no insight into the MT errors,
- and do not scale well with sentence 
-length. We argue for a semantics-based evaluation, which aims to capture how much of the meaning
-is retained in the MT output, and can support the construction and tuning of semantics-based MT. 
+and do not scale well with sentence length.
+We argue for a semantics-based evaluation, which captures what meaning components
+are retained in the MT output, providing a more fine-grained analysis of
+translation quality, and enables the construction and tuning of semantics-based MT. 
 We present a novel human semantic evaluation measure for MT, Human
-UCCA-based MT Evaluation (HUME), building on the Universal Conceptual Cognitive Annotation (UCCA) semantic representation scheme.
+UCCA-based MT Evaluation (HUME), building on the Universal Conceptual Cognitive Annotation (UCCA)
+semantic representation scheme.
 Making progress over previous work, the proposed measure takes into account
@@
 
-Human judgement is the most important criterion for
-estimating the quality of an MT system.
+Human judgement is the cornerstone for estimating the quality of an MT system.
 Nevertheless, common measures for human MT evaluation, such as adequacy and fluency judgements
 or the relative ranking of possible translations, are problematic in two ways.
-First, as the quality of translation is determined by multiple factors, it is difficult
-to assign a single number that quantifies the quality of the entire sentence. This
+First, as the quality of translation is multi-faceted, it is difficult
+to quantify the quality of the entire sentence in a single number. This
 is indeed reflected in the diminishing inter-annotator agreement (IAA) rates of human ranking measures
-%as a function of
-with
-the sentence length \parcite{Bojar:2011}.
+with the sentence length \cite{Bojar:2011}.
 Second, a sentence-level quality score does not indicate what parts of the sentence
@@
 quantifying the similarity of the output and the reference in terms of
-their verb argument structure; the most notable of these measures is
-HMEANT \parcite{lo2011structured}.
+their verb argument structure; the most notable of these measures is HMEANT
+\parcite{lo2011structured}.
+
 % SAVING SPACE:
@@
 %HUME only requires semantically annotating the source sentence,
-are thus avoiding the semantic annotation of machine-generated text. 
-This allows the re-use of the source semantic annotation for measuring the quality
-of different translations of the same source sentence.
+are thus avoiding the semantic annotation of machine-generated text,
+which is often garbled or semantically unclear.
+This also allows the re-use of the source semantic annotation for
+measuring the quality of different translations of the same source sentence,
+and 
 %By basing the measure on a direct comparison
 %of the source and translation we avoid the difficulties created by using
-We also avoid creating and relying on possibly suboptimal
-reference translations, albeit at the cost of requiring the
+avoids relying on possibly suboptimal reference translations,
+albeit at the cost of requiring the
 evaluator to be proficient in both the source and target languages.
@@
 (\secref{sec:hume}).
-Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments}) document the
-efficiency (time of annotation) and inter-annotator agreement scores for HUME. 
-We also report on the correlation scores of HUME with direct assesment of human 
-adequacy ratings. 
-Finally, the differences with HMEANT are discussed 
-(\secref{sec:hmeant_comp}).
+Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments})
+document HUME's inter-annotator agreement and efficiency (time of annotation). 
+We further empirically compare HUME with direct assesment of human adequacy ratings,
+and conclude by discussing the differences with HMEANT (\secref{sec:hmeant_comp}).
 
@@
 
-\paragraph{Machine Translation Evaluation.}
+\paragraph{MT Evaluation.}
 %MT evaluation has been traditionally pursued through human-based
@@
 latter (\figref{fig:ucca_example_v2}). Further categories account for
-relations between Scenes and the internal structures of complex participants and relations.
+inter-scene relations and the sub-structures of participants and relations.
 
 % UCCA Motivation
-UCCA is a natural candidate for defining a semantic
-MT evaluation measure for a number of reasons. 
-First, it defines an intuitive set of categories
-that can be reliably annotated by non-experts, after as little as two hours
+The use of UCCA for semantic MT evaluation measure is motivated by several reasons. 
+First, UCCA's set of categories can
+be reliably annotated by non-experts after as little as two hours
 of training \parcite{marinotti2014}.
-Second, UCCA is a cross-linguistically applicable scheme, seeking to
-represent what is shared between languages
-by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
-\parcite{Dixon:10a,Dixon:10b,Dixon:12}.
-The cross-linguistic applicability of UCCA has been tested in annotations of
+Second, UCCA is cross-linguistically applicable, seeking to
+represent what is shared between languages by building on
+linguistic typological theory \parcite{Dixon:10a,Dixon:10b,Dixon:12}.
+Its cross-linguistic applicability has so far been tested in annotations of
 English, French, German and Czech.
-Third, the scheme has been shown to be semantically stable
-in translation: UCCA annotations of translated text usually contain the same set of relationships
-\parcite{sulem2015conceptual}. This indicates that the scheme reflects
+Third, the scheme has been shown to be stable across translations:
+UCCA annotations of translated text usually contain the same set of relations
+\parcite{sulem2015conceptual}, indicating that the scheme reflects
 a common layer of representation that in a correct translation
-is mostly shared between the translation and the source.
-
-We note that verbal argument structures are less stable in this sense, as a verbal
-argument structure may be translated or paraphrased to a nominal or adjectival
-argument structure, and is thus less suitable as a basis for a semantic measure.
-E.g., ``after he graduated, Tom moved to NYC'' contains two verbs,
-but is similar in meaning to the one-verbed
-``after graduation, Tom moved to NYC''.
+is mostly shared between the translation and the source\oa{verb argument structure note?}.
+
+%Verbal argument structures are less stable in this sense, as they
+%may be translated or paraphrased to a nominal or adjectival
+%argument structure, and is thus less suitable as a basis for a semantic measure.
+%E.g., ``after he graduated, Tom moved to NYC'' contains two verbs,
+%but is similar in meaning to the one-verbed
+%``after graduation, Tom moved to NYC''.
 
@@
 
-The Abstract Meaning Representation (AMR) scheme \parcite{banarescu2013abstract}
-shares much of UCCA's motivation for defining a more complete semantic annotation.
+The Abstract Meaning Representation (AMR) \parcite{banarescu2013abstract} project
+shares UCCA's motivation for defining a more complete semantic annotation.
 However, using AMR is not optimal for defining a decomposition of a sentence into semantic
 units as it does not ground its semantic symbols in the text,
-thus does not provide clear decomposition of the sentence into sub-units.
+and thus does not provide clear decomposition of the sentence into sub-units.
 Also, AMR is more fine-grained than UCCA and consequently harder to annotate.
-Other approaches pursued a grounded representation of semantics,
-and have generally represented semantic structures as bi-lexical dependencies.
-Examples include the tectogrammatical layer \parcite{sgallhp:1986} used in the Prague dependency
-treebanks, e.g. the Czech-English one \parcite{hajic2012announcing}, as well
-as dependencies derived from Minimal Recursion Semantics representations \parcite{oepen2006discriminant}.
-Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
-these approaches require linguistic expertise for their annotation, due to the intricate and
-fine-grained distinctions they provide, and so we argue they are less suitable as a basis
-for an interpretable MT evaluation.
+Other approaches represent semantic structures as bi-lexical dependencies
+\parcite{sgallhp:1986,hajic2012announcing,oepen2006discriminant},
+which are indeed grounded in the text, but are less suitable for MT evaluation
+as they require linguistic expertise for their annotation.
+
+%, and so we argue they are less suitable 
+%for an interpretable MT evaluation.
+
+%Examples include the tectogrammatical layer  used in the Prague dependency
+%treebanks, e.g. the Czech-English one \parcite{}, as well
+%as dependencies derived from Minimal Recursion Semantics representations \parcite{}.
+%Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
 
@@
 where units are defined according to the UCCA annotation.
-Note that UCCA annotation is performed once for every source sentence,
+UCCA annotation is performed once for every source sentence,
 irrespective of the number of its translations we wish
@@
   here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}.
-Let $L$ be the set of possible UCCA labels.
-Given a sentence $s = (w_1, \ldots,w_n)$, its UCCA annotation is a DAG $G=(V,E,\ell)$,
-where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
-For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
-The semantic units for a sentence $s$ according to $G$
-is the set of yields of nodes in $V$. 
-The translation is not UCCA-annotated.
+A UCCA annotation for a sentence $s$ is a labeled DAG $G$, whose leaves
+are the words of $s$.
+For every node in $G$, we define its {\it yield} to be its leaf descendants.
+The semantic units for $s$ according to $G$ are the yields of nodes in $G$.
 
@@
 \paragraph{Translation Evaluation.}
-The basic strategy for annotating HUME is to traverse the semantic units
+HUME annotation is done by traversing the semantic units
 of the source sentence, which correspond to the arguments and relations expressed
-in the text, and to mark the extent to which they have been correctly translated.
+in the text, and marking the extent to which they have been correctly translated.
 HUME aggregates the judgements of the users into a composite score, 
@@
 and ``Red'' (incorrect). 
-A Green label means that the meaning of the word or phrase has been largely preserved.
-An Orange label means that the essential meaning of the unit has been preserved,
+Green means that the meaning of the word or phrase has been largely preserved.
+Orange means that the essential meaning of the unit has been preserved,
 but some part of the translation is wrong.
-This could often be due to the translated word having the wrong tense,
-or the wrong morphology, in a way that impacts little on the understandability of the sentence.
-A Red label means that the essential meaning of the unit has not been captured.
+This is often be due to the translated word having the the wrong inflection,
+in a way that impacts little on the understandability of the sentence.
+Red means that the essential meaning of the unit has not been captured.
 
@@
 Structural units are labeled as ``Adequate'' or ``Bad'', meaning
-that the relation between the sub-units went wrong.
-We use a three label system with atomic units, as
-opposed to a two label one in structural units,
-as atomic units often have slight errors, in which cases we do want to provide partial credit.
-
-There are many ways that the relationship between the children might go
-wrong in translation:
-
-\begin{compactitem}
-\item The components of the translation are ordered differently from the source,
+that the relation between the sub-units went wrong\footnote{
+  Three labels are used with atomic units, as opposed to two labels with structural units,
+as atomic units are more susceptible to slight errors.}.
+There are three major cases where the relationship between the children might go
+wrong in translation.
+First, the components of the translation may be ordered differently from the source,
 such that the relation between them changes. For instance, if
@@
 the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
-\item A superfluous word or phrase has been inserted into
+Second, A superfluous word or phrase may be inserted into
 the translation. For instance, if ``dog bites man'' is translated into ``dog viciously bites man''.
-\item One of the sub-units is missing from the translation.
+Third, one of the sub-units may be missing from the translation.
 For instance, if ``dog bites man'' is translated into ``dog bites''.
-% XXX let's leave the reviewers to notice this:
-%  \bh{Note that in this case we also 
-%  get an error in the atomic unit, assuming that ``man'' is a semantic unit in the source, but for the other
-%  two cases there is no atomic unit error.}
-\end{compactitem}
 
@@
 
+\com{
 In non-configurational languages, the relationship between the
@@
 on configurational vs. non-configurational languages.
+}
 
@@
      HUME annotation of an UCCA tree with a word aligned example translation shown below. 
-Atomic units are labelled using traffic lights and structural nodes are marked A or B.}
+Atomic units are labelled using traffic lights and structural units are marked A or B.}
 \end{figure}
@@
 
-In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME
-annotation, with the translation shown in English for ease of comprehension.
+Figure~\ref{fig:hume_tree_v2} presents an example of a HUME
+annotation, where the translation is in English for ease of comprehension.
 When evaluating ``to NYC'' the annotator looks at the translation and sees the
 word ``stateside''. This word captures the whole phrase and so we mark this
-non-leaf node with an atomic label. Here we choose ``Orange'' since
-it approximately captures the meaning
-in this context. The ability to mark non-leaves with atomic labels allows
+non-leaf node with an atomic label. Here we choose Orange since
+it approximately captures the meaning in this context.
+The ability to mark non-leaves with atomic labels allows
 the annotator to account for translations which only correspond at the phrase
 level. Another feature highlighted in this example is that by separating structural
-and atomic nodes, we are able to define where an error occurs, and localise
+and atomic units, we are able to define where an error occurs, and localise
 the error to its point of origin. The linker ``After'' is translated incorrectly as ``by''
@@
     \includegraphics[width=1\textwidth]{hume_interface2.jpg}
-    \caption{The HUME annotation tool, based on the UCCA tool. The faded orange box at the top
-      contains the translation, and directly below it is the source sentence. The tree of semantic units
-      is shown below the source. Alignments between the source and translation appear in italics and unaligned intervening words appear in red.}
+    \caption{The HUME annotation tool. The top orange box
+      contains the translation. The source sentence is directly below it, followed by the tree of the source
+      semantic units. Alignments between the source and translation are in italics and
+      unaligned intervening words are in red (see text).}
     \label{fig:interface}
@@
 We proceed to detailing how judgements on the semantic units
-of the source are aggregated into a composite score. We start by taking a parsimonious approach
+of the source are aggregated into a composite score. 
 %While future extensive empirical evaluation with HUME and its application to MT
@@
 %as a starting point we take a parsimonious approach,
-and compute an accuracy score.
-
-Let $\Correct(s,t)$ be the number of atomic units in $s$ marked as Green relative to $t$ plus
-the number of structural units in $s$ marked as Adequate relative to $t$.
-Let $\Partial(s,t)$ be the number of atomic units in $s$ marked as Orange relative
-to $t$. Let $\setsize{\Units(s)}$ be the total number of units marked with any of the five labels.
+We start by taking a very simple approach and compute an accuracy score.
+Let $\Green(s,t)$, $\Adequate(s,t)$ and $\Orange(s,t)$ be the number of Green, Adequate and Orange
+units respectively. Let $\Units(s)$ be the number of units marked with any of the labels.
 Then HUME's composite score is:
@@
 \vspace{-.5cm}
+{\scriptsize
 \begin{equation}
-  \small
-  \HUME(s,t) = \frac{\Correct(s,t) + 0.5\cdot \Partial(s,t)}{\setsize{\Units(s)}}
+  \HUME(s,t) = \frac{\Green(s,t) + \Adequate(s,t) + 0.5\cdot \Orange(s,t)}{\Units(s)}
   \nonumber
 \end{equation}
+}
+\vspace{-.5cm}
 
@@
 make a judgement according to the full translation.
-
-Concretely, let $s = (w_1,\ldots,w_n)$ be a source sentence, $t = (w'_1,\ldots,w'_n)$ a translation,
-and $A \subset 2^{s \times t}$ a many-to-many word alignment between them.
+Concretely, let $s$ be a source sentence, $t$ a translation,
+and $A \subset 2^{s \times t}$ a many-to-many word alignment.
 If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
-$t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
-Where $a(u)$ is discontinuous in the text, words between the left
-and right boundaries of $a(u)$ which are not contained in it (intervening words)
+$t$ to be $\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t$.
+
+Where the aligned text is discontinuous in $t$, words between the left
+and right boundaries which are not contained in it (intervening words)
 are presented in a smaller red font. 
@@
 seems acceptable but the interleaving words in red notify the annotator to check
-the whole translation. The meaning of the expression is not preserved in the
-translation
-and the annotator should mark this structural node as ``B'', bad. 
+the whole translation, in which the meaning of the expression is not preserved.
+The annotator should thus mark this structural node as Bad.
+
 %The interleaving 
@@
 and four target languages (Czech, German, Polish and Romanian), using text from the public health domain.
-The domain is particularly suitable for experimenting on MT evaluation, as semantically
-accurate translation is paramount here, underscoring the importance of reliable
-semantic evaluation measures.
+Semantically accurate translation is paramount in this domain, which makes
+it particularly suitable for experimenting on MT evaluation. HUME is both evalauted in terms of its
+consistency (inter-annotator agreement), efficiency (time of annotation) and validity (through a comparison
+with crowd-sourced adequacy judgements).
 
@@
 
-The experiment seeks to address two questions:
-\begin{inparaenum}
-\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences?
-\item Can HUME capture translation issues that other annotation schemes (such as
-adequacy judgements and HMEANT)
-cannot?
+%The experiment seeks to address two questions:
+%\begin{inparaenum}
+%\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences?
+%\item Can HUME capture translation issues that other annotation schemes (such as adequacy judgements
+%  and HMEANT) cannot?
+%\end{inparaenum}
+
 %\item What can HUME annotation tell us about the strengths and weaknesses of MT?
@@
 %probably don't address it in this paper}
-\end{inparaenum}
-
-\subsection{Data Sets and Translation Systems}
+
+\subsection{Datasets and Translation Systems}
 
@@
 computational linguists and one linguist.
-
 For the annotation of the MT output, we recruited two annotators for each of German, Romanian
-and Polish and one main annotator for Czech.
-%In order to allow us to calculate
-For IAA also for
-Czech, several further annotators worked on a small number of 
-sentences each. We treat these further annotators as one annotator, giving us two annotators
+and Polish and one main annotator for Czech. For Czech IAA,
+several further annotators worked on a small number of 
+sentences each. We treat these further annotators as one annotator, resulting in two annotators
 for each language pair.
@@
 \tabref{tab:annot}
-shows the total number of sentences and nodes annotated by each annotator. Note
-that not all nodes in all sentences were annotated, often
+shows the total number of sentences and units annotated by each annotator. Note
+that not all units in all sentences were annotated, often
 due to
@@
 \hline
-Nodes & Annot. 1 & 8794  & 9253 & 9557  & 6152 \\
+Units & Annot. 1 & 8794  & 9253 & 9557  & 6152 \\
  &Annot. 2 & 5553 & 2906  & 9303  & 9228  \\
 \end{tabular}
-\caption{Total numbers of sentences and nodes annotated on the MT output.}
+\caption{Number of HUME-annotated sentences and units.}
 \label{tab:annot}
@@
 
-The annotators worked remotely, at their own pace, so we did not collect precise timing
-information. Nevertheless we are able to estimate the annotation time using the timestamps
+\paragraph{Efficiency.}
+We estimate the annotation time using the timestamps
 provided by the annotation tool, which are recorded whenever an annotated sentence is
 submitted. Annotators are not able to re-open a sentence once submitted. 
-
-To estimate the annotation time, we calculate the time difference between successive 
-sentences, and discard outlying times since we assume there was a break taken. From inspection
-of histograms of annotation times, we set the upper threshold at 500 seconds. We
-show median annotation times in Table~\ref{tab:annot_times}, indicating that the annotation
+To estimate the annotation time, we compute the time difference between successive 
+sentences, and discard outlying times since we assume annotation was not continuous.
+From inspection of histograms of annotation times, we set the upper threshold at 500 seconds.
+Median annotation times are presented in Table~\ref{tab:annot_times},
+indicating that the annotation
 of a sentence takes around 2--4 minutes, with some variation between annotators.
+
+%The annotators worked remotely at their own pace, so no precise timing information is available.
+%Nevertheless
 
@@
 \caption{Median annotation times per sentence, in seconds.
-We do not give times for the second Czech 
-``annotator'', since this was a collection of annotators, working in parallel.}
+  In the case of the second Czech 
+  ``annotator'', no timing information is available, as
+  this was a collection of annotators, working in parallel.}
 \label{tab:annot_times}
@@
 
-\subsection{Inter-Annotator Agreement}
+\paragraph{Inter-Annotator Agreement.}
 \label{sec:iaa}
@@
 \hline
-All nodes & 4686   & 2793   & 8384   & 5604  \\
+All units & 4686   & 2793   & 8384   & 5604  \\
 Kappa & 0.64   & 0.61   & 0.58   & 0.69  \\
 \hline
-Atomic nodes & 2982 & 1724 & 5386 & 3570 \\
+Atomic units & 2982 & 1724 & 5386 & 3570 \\
 Kappa & 0.54 & 0.29 & 0.54 & 0.50 \\
 \hline
-Structural nodes & 1602 & 1040 & 2655 & 1989 \\
+Structural units & 1602 & 1040 & 2655 & 1989 \\
 Kappa & 0.31 & 0.44 & 0.33 & 0.58 \\
@@
 %each annotating a sub-part of the batch.
-
-\tabref{tab:iaa} reports the number of nodes which have two annotations from
+\tabref{tab:iaa} reports the number of units which have two annotations from
 different annotators and the corresponding Kappas.
@@
 We report the overall Kappa, as well as separate Kappas on atomic
-nodes (annotated as Red, Orange or Green) and structural nodes (annotated
+units (annotated as Red, Orange or Green) and structural units (annotated
 as Adequate or Bad).
 As expected and confirmed by confusion matrices in \figref{fig:heatmap}, there
-is generally little confusion between the two types of nodes.
+is generally little confusion between the two types of units.
 
@@
 
-To assess whether, as we hoped, HUME would make it easier to reliably annotate long sentences,
-we binned the sentences according to length and measured Kappa on each of the bins. Each bin covers
-a length range of 10, and we show the variance of kappa with length in \figref{fig:iaalength}.
+%presents the variance of kappa with length in each bin covers
+%a length range of 10. and we show .
 
@@
 \caption{Kappa versus sentence length, separately for
-structural and atomic nodes. The numbers are the node counts in each bin. }
+structural and atomic units. The numbers are the node counts in each bin. }
 \label{fig:iaalength}
@@
 
-We see from \figref{fig:iaalength} that there is no reduction of IAA with sentence
+To assess whether, HUME supports the reliable annotation of long sentences,
+we binned the sentences according to length and measured Kappa on each of the bins.
+The Kappas for each bin (bins are 10 words wide) is presented in 
+\figref{fig:iaalength}.
+Results show that there is no reduction of IAA with sentence
 length, except perhaps in German. Also, from \tabref{tab:iaa} the overall IAA
@@
 In particular, we see a contrast  between
-Czech and Polish, where the IAA is higher for atomic than for structural nodes, and German and Romanian,
+Czech and Polish, where the IAA is higher for atomic than for structural units, and German and Romanian,
 where the reverse is true. We also observe low IAA (around 0.3) in the cases of
-German atomic nodes, and Polish and Czech structural nodes.
-
-Looking more closely at the areas of disagreement, we see that for the Polish structural nodes, the 
+German atomic units, and Polish and Czech structural units.
+
+Looking more closely at the areas of disagreement, we see that for the Polish structural units, the 
 proportion of As was quite different between the two annotators (53\% vs. 71\%), whereas for other
 languages the annotators agree in the proportions. We believe that this was because one of the Polish
-annotators did not fully understand the guidelines for structural nodes, and percolated
-errors up the tree, creating more Bs. For the German lexical nodes, and Czech structural nodes, where
-Kappa is also around 0.3, the proportion of such nodes being marked as ``correct'' is relatively 
+annotators did not fully understand the guidelines for structural units, and percolated
+errors up the tree, creating more Bs. For the German lexical units, and Czech structural units, where
+Kappa is also around 0.3, the proportion of such units being marked as ``correct'' is relatively 
 high, meaning that the class distribution is more skewed, so the expected agreement used in the
@@
 In HUME, errors are localised as much as possible to where they occur,
-by the separation of atomic and structural nodes (See Fig \ref{fig:hume_tree_v2} 
+by the separation of atomic and structural units (See Fig \ref{fig:hume_tree_v2} 
 for an example).

### 2016-06-02 paper/main.tex
flagging iaa

@@
 these are often words or n-grams, for manual measures some structural
-information is taken into account \parcite{machacek:bojar:segranks:2015}.
+information is taken into account \parcite{machacek:bojar:segranks:2015}, or the
+annotators are explicitly asked to mark errors, which however suffers from even
+lower agreement than ranking \parcite{lommel:etal:mqm-iaa:2014}.
 % the following is specific to automatic measures:

### 2016-06-02 paper/main.tex
spelling of assesment

@@
 document HUME's inter-annotator agreement and efficiency (time of annotation). 
-We further empirically compare HUME with direct assesment of human adequacy ratings,
+We further empirically compare HUME with direct assessment of human adequacy ratings,
 and conclude by discussing the differences with HMEANT (\secref{sec:hmeant_comp}).
@@
 \end{tabular}
-\caption{Plots of HUME scores vs Direct Assesment scores which have been standardised for each crowdsourcing annotator. The regression is calculated using linear least squares.}
+\caption{Plots of HUME scores vs Direct Assessment scores which have been standardised for each crowdsourcing annotator. The regression is calculated using linear least squares.}
 \label{fig:dacorrelation}
@@
 few Czech and Polish annotators, and removing the IP restriction led to very poor quality results. 
-We can see that HUME and Direct Assesment scores do correlate. 
+We can see that HUME and Direct Assessment scores do correlate. 
 We also removed unreliable crowdsourcing annotators
@@
 of the source sentence have been retained.
-Direct assesment scores require a crowd, which is sometimes not available or reliable, and it returns a single number which gives little insight into the quality of
+Direct assessment scores require a crowd, which is sometimes not available or reliable, and it returns a single number which gives little insight into the quality of
 the translated sentence.

### 2016-06-02 paper/main.tex
is_scene

--- 2016-06-02 paper/main.tex
@@
 
-\subsection{Annotation Statistics}
+\subsection{HUME Annotation Statistics}
 \label{sec:annot_stats}
@@
 \end{tabular}
-\caption{Inter-annotator agreement for the multiply-annotated portions of the data, as
+\caption{Inter-annotator agreement for the multiply-annotated portions of the data,
 measured by Cohen's Kappa. }
@@
 length, except perhaps in German. Also, from \tabref{tab:iaa} the overall IAA
-is quite similar (0.6-0.7) for all languages, showing good agreement.
+is quite similar for all languages, showing good agreement (0.6--0.7).
 However, there are differences observed when we break down by node type.
@@
 We also removed unreliable crowdsourcing annotators
-according to the method discussed in ~\cite{TODO}. The Pearsons correlation for en-de is 0.58 and for en-ro is 0.70 which shows that they two measures do correlate. This confirms that 
+according to the method discussed in ~\cite{TODO}. The Pearsons correlation
+for en-de is 0.58 and for en-ro is 0.70 which shows that they two measures do correlate. This confirms that 
 the HUME metric is consistent with previously published methods in human evaluation, even though these
@@
 argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
-relations and copula clauses. Some of the uncovered cases were already observed
-in previous applications of HMEANT to languages other than English (see \perscite{birch-EtAl:2013:WMT} for German, \perscite{bojar:wu:ssst:2012} for
-Czech and by \perscite{chuchunkov-tarelkin-galinskaya:2014:SSST-8} for Russian)
-but the problems are also seen in English.
-
-%This focus on verbal predicates has a number of unfortunate consequences. We show an example below:
-Consider the example:
-
-% \noindent
-% \begin{inparadesc}
-% \item[Source ] a coronary angioplasty may not be technically possible \\
-% \item[Transl.] eine koronare Angioplastie kann nicht technisch m{\"o}glich \\
-% \item[Gloss ]  a coronary angioplasty can not technically possible \\
-% \end{inparadesc}
+relations and copula clauses.
+We are not aware of any empirical evaluation that demonstrates that verb argument structure,
+taken alone, captures the crux of the sentence semantics meanings.\oa{do we mention here that
+  Scenes in their own right correlate less with DA? maybe we should distinguish verbal and non-verbal
+  scenes?} Consider the example:
 
@@
 
-The German machine translation 
+%``a coronary angioplasty may not be technically possible'', which was translated by our en-de system
+%to ``eine koronare Angioplastie kann nicht technisch m{\"o}glich'' (gloss: 
+%a coronary angioplasty can not technically possible).
+
+The German translation 
 is largely correct, except that the main verb ``sein'' (``be'') is omitted.
@@
 
+%This focus on verbal predicates has a number of unfortunate consequences. We show an example below:
+
+% \noindent
+% \begin{inparadesc}
+% \item[Source ] a coronary angioplasty may not be technically possible \\
+% \item[Transl.] eine koronare Angioplastie kann nicht technisch m{\"o}glich \\
+% \item[Gloss ]  a coronary angioplasty can not technically possible \\
+% \end{inparadesc}
+
+\com{
+}
+
 There are however a number of other serious consequences on focusing on verbal
 predicates. We conducted an analysis of the English UCCA Wikipedia
-corpus (5324 sentences)
-in order to assess the impact of the following problems: 
-\\
-\begin{inparadesc}
-\item[Copula clauses] are treated in HMEANT simply as instances of the main verb ``be'', which
+corpus (5324 sentences) in order to assess the pervasiveness of three phenomena that are not
+well supported by HMEANT.
+First, copula clauses are treated in HMEANT simply as instances of the main verb ``be'', which
 generally does not convey the meaning of these clauses. They appear in 21.7\% of the sentences,
 according to conservative estimates that only consider non-auxiliary instances of ``be''.
-\\
-\item[Nominal argument structures,] which are ignored by  HMEANT are in fact highly
+Second, nominal argument structures, which are ignored by  HMEANT are in fact highly
 pervasive, appearing in 48.7\% of the sentences.
-\\
-\item[Linkers] expressing an inter-relation between
+Third, linkers, that express an inter-relation between
 clauses (mainly discourse markers and coordinating conjunctions) appear in 56\% of the
-sentences, but are ignored by HMEANT.
+sentences, but are again ignored by HMEANT.
 An example where a linker is omitted in translation in our experiments is
-with the source \equo{However, this review\dots}, which was translated to
+\equo{However, this review\dots}, which was translated to
 the German \equo{Diese \"Uberpr\"ufung\dots} (lit. \equo{This review}).
 This omission will not be taken into account by HMEANT at all\footnote{Argument structures and linkers are explicitly marked in UCCA.
-We identified non-auxiliary instances of ``be'' and nouns using the NLTK standard tagger.
-Nominative argument structures are here defined as Scenes whose main relation is headed by a noun.}.
-\end{inparadesc}
+Non-auxiliary instances of ``be'' and nouns are identified using the NLTK standard tagger.
+Nominative argument structures are defined as Scenes whose main relation is headed by a noun.}.
 
@@
 
-We are not aware of any empirical evaluation that demonstrates that verb argument structure,
-taken alone, captures the crux of the sentence semantics meanings.
+We note that some of these issues were already observed
+in previous applications of HMEANT to languages other than
+English. See \perscite{birch-EtAl:2013:WMT} for German, \perscite{bojar:wu:ssst:2012} for
+Czech and \perscite{chuchunkov-tarelkin-galinskaya:2014:SSST-8} for Russian.
+
 % I am not sure that this is true
@@
 
-\paragraph{One vs. Two Structures}
-
-HUME uses just one structural annotation (for the source) while HMEANT relies on
-two independently constructed structural annotations (for both the reference and
-the candidate translation)
+\paragraph{One Structure or Two.}
+HUME only annotates the source, while HMEANT relies on
+two independently constructed structural annotations, one for the reference and one
+for the translation.
 %First of all, the annotation process is different. We semantically annotate the
@@
 %We see this difference as crucial where semantic analysis is concerned, as it
-and it is often impossible to assign a semantic structure to a low quality MT output sentence.
-
-However, HMEANT's independent annotation of the source and translation makes a stricter check
-of whether the translation can be understood
-than we do. HUME may be artificially boosting the perceived understandability of
-the sentence by allowing access to the source.
-
-\paragraph{Importance of Alignment}
-
-In HMEANT, the alignment between the (reference and translation) structures is a key
+
+Not annotating the translation is appealing as it is often impossible to assign a
+semantic structure to a low quality translation.
+On the other hand, HUME's annotation may be artificially boosting the perceived understandability
+of the sentence by allowing access to the source.
+
+%%%% UNTIL HERE %%%%
+\paragraph{Alignment.}
+In HMEANT, the alignment between the reference and translation structures is a key
 part of the manual annotation. If the alignment cannot be created, the 

### 2016-06-02 scripts/get_csv.py

@@ -65,7 +65,6 @@ def main():
     mt_colkeys = ["mteval_{}".format(key) for key in mt_keys]
     with open(args.sentenceFile, "w") as csvfh,\
         open(args.nodeFile, "w") as ncsvfh:
-      print("sent_id,annot_id,ucca_annot_id,lang,timestamp,source,target,reference,align,bleu,ucca_node_count,", file=csvfh, end="")
       print("ucca_H,", file=csvfh, end="")
       print(",".join(mt_colkeys), file=csvfh,end="")
       print(file=csvfh)
@@ -98,6 +97,8 @@ def main():
         for node in sent.ucca_tree.nodes:
           node = sent.ucca_tree.nodes[node]
           if node.tag != "FN": continue
+          is_scene = 'T' if node.is_scene() else 'F'
+          
           mt_annot = sent.mt_annotation.get(node.ID, "M")
           mt_eval_counts[mt_annot] += 1
           ucca_counts[node.ftag] += 1
@@ -118,13 +119,13 @@ def main():
                 return get_src_tgt(node.children[0])
             source,target,pos = [" ".join(l) for l in list(zip(*[get_src_tgt(node) for node in node.children]))]
           ncsv_writer.writerow((node.ID, str(sent.sent_id), sent.annot_id, sent.lang, mt_annot, \
-                str(child_count), children, parent, tag, pos, source, target ))
+                str(child_count), children, parent, tag, pos, source, target, is_scene ))
         fields.append(sum(mt_eval_counts.values()))
         fields.append(ucca_counts["H"])
         fields += [mt_eval_counts[key] for key in mt_keys]
-
+        
         csv_writer.writerow(fields)
- 
+
 if __name__ == "__main__":
   main()
 

### 2016-06-02 paper/main.tex
sss

@@
 \usepackage{pgfplotstable}
+\usepackage{titlesec}
 %\usepackage{hyperref}
@@
 \def\emnlppaperid{***}
+\setlength{\belowcaptionskip}{-10pt}
 
 \makeatletter
+
+\titlespacing*{\section}
+{0pt}{1ex plus 1ex minus .2ex}{1ex plus .2ex}
+\titlespacing*{\subsection}
+{0pt}{1ex plus 1ex minus .2ex}{1ex plus .2ex}
+
 \renewcommand{\paragraph}{
   \@startsection{paragraph}{4}
-  {\z@}{.3ex \@plus .3ex \@minus .2ex}{-1em}
+  {\z@}{.1ex \@plus .1ex \@minus .1ex}{-1.5em}
   {\normalfont\normalsize\bfseries}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\vspace{-.1em}
 \section{The HUME Measure}\label{sec:hume}
+\vspace{-.2em}
 
@@
     \begin{center}
-    \includegraphics[width=1\textwidth]{hume_interface2.jpg}
+    \includegraphics[width=.8\textwidth]{hume_interface2.jpg}
     \caption{The HUME annotation tool. The top orange box
@@
 using Moses \parcite{Koehn:2007}.  These were trained on large parallel data sets extracted from
-OPUS \parcite{tiedemann:2009}, and
-the data sets released for the WMT14 medical translation task \parcite{bojar-EtAl:2014:W14-33}, 
-giving between 
-45 and 85 million sentences of training data, depending on language pair.
+OPUS \parcite{tiedemann:2009}, and the data sets released for the WMT14
+medical translation task \parcite{bojar-EtAl:2014:W14-33}, 
+giving between 45 and 85 million sentences of training data, depending on language pair.
 %\oa{I think we should probably
@@
 % ... if someone asks us to do so...
-
 These translation systems were used to translate texts derived from both NHS
@@
 related information in Scotland, Cochrane is an international NGO
-which 
-provides independent systematic reviews on health-related research. From NHS 24, we selected the segments from their
-``Health A-Z'' in the NHS Inform website, and the Cochrane texts come from their plain language summaries
+which provides independent systematic reviews on health-related research.
+NHS~24 texts come from the ``Health A-Z'' section in the NHS Inform
+website, and Cochrane texts come from their plain language summaries
 and abstracts.
@@
 The annotators were all native speakers of the respective target languages and fluent in English.
+\oa{maybe we should put in what training they've received}
 
 \tabref{tab:annot}
-shows the total number of sentences and units annotated by each annotator. Note
-that not all units in all sentences were annotated, often
-due to
-%this was because
-the annotator not finishing a sentence or accidentally missing a node.
+shows the total number of sentences and units annotated by each annotator.
+Not all units in all sentences were annotated, often due to
+the annotator %not finishing a sentence or
+accidentally missing a node.
 %, but in many cases it was also because it was an implicit 
@@
 \hline
-Sentences &  Annot. 1 & 324   & 339  & 351  & 230  \\
+\#Sentences &  Annot. 1 & 324   & 339  & 351  & 230  \\
  & Annot. 2 & 205 & 104  & 340  & 337 \\
 \hline
-Units & Annot. 1 & 8794  & 9253 & 9557  & 6152 \\
+\#Units & Annot. 1 & 8794  & 9253 & 9557  & 6152 \\
  &Annot. 2 & 5553 & 2906  & 9303  & 9228  \\
 \end{tabular}
-\caption{Number of HUME-annotated sentences and units.}
+\caption{HUME-annotated \#sentences and \#units.}
 \label{tab:annot}
@@
 \end{tabular}
-\caption{Inter-annotator agreement for the multiply-annotated portions of the data,
+\caption{IAA for the multiply-annotated units,
 measured by Cohen's Kappa. }
@@
 In order to assess the consistency of the annotation, we measure the Inter-Annotator
-Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
+Agreement (IAA) using Cohen's Kappa on the multiply-annotated units.
 %For German, Polish and Romanian we used two annotators, with a significant overlap
@@
 
-%%%% UNTIL HERE %%%%
 \paragraph{Alignment.}
@@
 %interpretation of a PP-attachment, the verb can have one additional modifier in
-one of the structures. It would be desirable to allow updates to the structural
-annotations at the alignment stage,
+one of the structures. It would be desirable to allow modifications to
+the SRL annotations at the alignment stage,
 to avoid unduly penalization due to such spurious divergences.
@@
 %to avoid unduly penalization due to such a spurious divergence.
-The same issue was seen by HMEANT authors themselves
-\parcite{lo:wu:reliability:2014}:
-the 
-IAA on SRL dropped from 90\% to 61\% 
+The same issue is noted in \parcite{lo:wu:reliability:2014}:
+the IAA on SRL dropped from 90\% to 61\% 
 when the two aligned structures were from two different annotators.
-
-%HUME uses only automatic (word-level) alignment, which only serves as a cue for directing
-%the attention of the annotators.
-%The user is expected to mentally correct the alignment as needed.
 HUME uses automatic (word-level) alignment, which only
@@
 
-\paragraph{Monolingual vs. Bilingual Evaluation}
+%HUME uses only automatic (word-level) alignment, which only serves as a cue for directing
+%the attention of the annotators.
+%The user is expected to mentally correct the alignment as needed.
+
+\paragraph{Monolingual vs. Bilingual Evaluation.}
 \label{src-vs-ref}
-HUME diverges from HMEANT (and also from the shallow automatic measures
-like BLEU or METEOR) in not requiring a reference.
+HUME diverges from both HMEANT and from shallower automatic measures
+like BLEU or METEOR, in not requiring a reference.
 Instead, it compares the source directly with the output translation.
-
 This requires us to employ bilingual annotators, but has the benefit of avoiding
-the reference translation. It is known that the reference translations are not
-always literal enough and the MT output semantics can be composed differently
-than the semantics of the reference.
-
-If only monolingual annotators are available, the HUME evaluation could be performed
-with a reference sentence instead of with the source. 
-\bh{But this would be a different metric}
+using a reference. Reference translations may often be phrased in many structurally different
+ways, which may again lead unduly penalization of the translation in cases where
+it is a paraphrase of the reference.
+
+%If only monolingual annotators are available, the HUME evaluation could be performed
+%with a reference sentence instead of with the source. 
+%\bh{But this would be a different metric}
+
 % While HUME avoids this problem,
@@
 
-\paragraph{Localisation of the Error}
-
+\com{
+\paragraph{Error Localisation.}
 In HMEANT, an error in a child node often results in the parent node
@@
 In HUME, errors are localised as much as possible to where they occur,
-by the separation of atomic and structural units (See Fig \ref{fig:hume_tree_v2} 
-for an example).
-This allows us to calculate the total score of the sentence more accurately.
+by the separation of atomic and structural units,
+which supports a more accurate aggregation of the translation errors
+to a composite score.
+}
+\oa{Omitted point here}
 
@@
 
-We have introduced HUME, a human semantics based MT evaluation measure which addresses
-a wide range of semantic phenomena. We have shown that it can be efficiently and 
-reliably annotated in 
-several languages, and that annotation quality is robust to sentence length.
-We believe that HUME, and a future automated version of HUME, will be able to guide the
-development of a more semantically aware approach to MT.
+We have introduced HUME, a human semantic MT evaluation measure which addresses
+a wide range of semantic phenomena. We have shown that it can be reliably and 
+efficiently annotated in multiple languages,
+and that annotation quality is robust to sentence length.
+We believe that
+HUME, and a future automated version of HUME, allows for a more fine-grained analysis of translation quality,
+and will be to guide the development of a more semantically aware approach to MT.
 

### 2016-06-02 scripts/get_csv.py

@@ -69,7 +69,7 @@ def main():
       print(",".join(mt_colkeys), file=csvfh,end="")
       print(file=csvfh)
 
-      print("node_id,sent_id,annot_id,lang,mt_label,child_count,children,parent,ucca_label,pos,source,target", file=ncsvfh)
+      print("node_id,sent_id,annot_id,lang,mt_label,child_count,children,parent,ucca_label,pos,source,target,is_scene", file=ncsvfh)
 
       csv_writer = csv.writer(csvfh,  lineterminator=os.linesep)
       ncsv_writer = csv.writer(ncsvfh,  lineterminator=os.linesep)

### 2016-06-02 paper/main.tex
abbreviated

@@
 information is taken into account \parcite{machacek:bojar:segranks:2015}, or the
-annotators are explicitly asked to mark errors, which however suffers from even
+annotators are explicitly asked to mark errors, which however suffer from even
 lower agreement than ranking \parcite{lommel:etal:mqm-iaa:2014}.
@@
 A promising line of research decomposes metrics
-%More recent work on MT evaluation proposed measures
-%that decompose
 over semantically defined units,
@@
 \parcite{lo2011structured}.
+%More recent work on MT evaluation proposed measures
+%that decompose
 
@@
 % UCCA Motivation
-The use of UCCA for semantic MT evaluation measure is motivated by several reasons. 
+The use of UCCA for semantic MT evaluation measure is motivated by two main reasons. 
 First, UCCA's set of categories can
@@
 English, French, German and Czech.
-Third, the scheme has been shown to be stable across translations:
-UCCA annotations of translated text usually contain the same set of relations
-\parcite{sulem2015conceptual}, indicating that the scheme reflects
-a common layer of representation that in a correct translation
-is mostly shared between the translation and the source\oa{verb argument structure note?}.
+%Third, the scheme has been shown to be stable across translations:
+%UCCA annotations of translated text usually contain the same set of relations
+%\parcite{sulem2015conceptual}, indicating that the scheme reflects
+%a common layer of representation that in a correct translation
+%is mostly shared between the translation and the source.
 
@@
 
-Like HMEANT, HUME defines  a
-measure that decomposes over semantic units, rather than strings. 
-However, there are important differences in design decisions, which we summarize in this
-section.
+We turn to discussing the main differences between HUME and HMEANT,
+a semantic measure that measures the translation quality according to its similarity
+with a reference, measures in terms of the overlap between their SRL annotations.
+%Like HMEANT, HUME defines  a
+%measure that decomposes over semantic units, rather than strings. 
+%However, there are important differences in design decisions, which we summarize in this
+%section.
 
@@
 We are not aware of any empirical evaluation that demonstrates that verb argument structure,
-taken alone, captures the crux of the sentence semantics meanings.\oa{do we mention here that
-  Scenes in their own right correlate less with DA? maybe we should distinguish verbal and non-verbal
-  scenes?} Consider the example:
+taken alone, captures the crux of the sentence semantics meanings.
+Moreover, relying only on verbal argument structures is less stable across paraphrases
+and translations, as a verbal argument structure
+may be translated to a nominal or adjectival argument structure
+(e.g., ``after graduation'' may be translated into ``after he graduated'').
+This may lead to unjustified penalties low HMEANT score, as the verb
+in one structure has nothing to align to in the other.
+On the other hand, HUME has been shown to be reasonably stable in an English-French
+corpus study \cite{sulem2015conceptual}.
+
+Consider the following example:
 
@@
 Non-auxiliary instances of ``be'' and nouns are identified using the NLTK standard tagger.
-Nominative argument structures are defined as Scenes whose main relation is headed by a noun.}.
+Nominal argument structures are here Scenes whose main relation is headed by a noun.}.
 
@@
 %We see this difference as crucial where semantic analysis is concerned, as it
-
 Not annotating the translation is appealing as it is often impossible to assign a

### 2016-06-02 paper/main.tex
minor

@@
 Let $\Green(s,t)$, $\Adequate(s,t)$ and $\Orange(s,t)$ be the number of Green, Adequate and Orange
-units respectively. Let $\Units(s)$ be the number of units marked with any of the labels.
+units, respectively. Let $\Units(s)$ be the number of units marked with any of the labels.
 Then HUME's composite score is:
@@
 \com{
-In Figure~\ref{mttool} you we show the MT evaluation tool.
+In Figure~\ref{mttool}, we show the MT evaluation tool.
 The sentence at the top shows the complete MT system output.

### 2016-06-02 paper/main.tex
minor
Latest graphic

--- 2016-06-02 paper/main.tex
@@
 \usepackage{pgfplotstable}
-\usepackage{titlesec}
 %\usepackage{hyperref}
@@
 \usepackage{lipsum,adjustbox}
-\usepackage[font={small}]{caption}
 
 \def\emnlppaperid{***}
-\setlength{\belowcaptionskip}{-10pt}
 
 \makeatletter
-
-\titlespacing*{\section}
-{0pt}{1ex plus 1ex minus .2ex}{1ex plus .2ex}
-\titlespacing*{\subsection}
-{0pt}{1ex plus 1ex minus .2ex}{1ex plus .2ex}
-
 \renewcommand{\paragraph}{
   \@startsection{paragraph}{4}
-  {\z@}{.1ex \@plus .1ex \@minus .1ex}{-1.5em}
+  {\z@}{.3ex \@plus .3ex \@minus .2ex}{-1em}
   {\normalfont\normalsize\bfseries}
@@
 \newcommand{\com}[1]{}
-\newcommand{\secref}[1]{\S\ref{#1}}
+\newcommand{\secref}[1]{Section~\ref{#1}}
 \newcommand{\figref}[1]{Figure~\ref{#1}}
@@
 \def\HUME{\func{HUME}}
-\def\Adequate{\func{Adequate}}
-\def\Green{\func{Green}}
-\def\Orange{\func{Orange}}
+\def\Correct{\func{Correct}}
+\def\Partial{\func{Partial}}
 \def\Units{\func{Units}}
@@
 or adequacy/fluency scales. However these provide no insight into the MT errors,
-and do not scale well with sentence length.
-We argue for a semantics-based evaluation, which captures what meaning components
-are retained in the MT output, providing a more fine-grained analysis of
-translation quality, and enables the construction and tuning of semantics-based MT. 
+ and do not scale well with sentence 
+length. We argue for a semantics-based evaluation, which aims to capture how much of the meaning
+is retained in the MT output, and can support the construction and tuning of semantics-based MT. 
 We present a novel human semantic evaluation measure for MT, Human
-UCCA-based MT Evaluation (HUME), building on the Universal Conceptual Cognitive Annotation (UCCA)
-semantic representation scheme.
+UCCA-based MT Evaluation (HUME), building on the Universal Conceptual Cognitive Annotation (UCCA) semantic representation scheme.
 Making progress over previous work, the proposed measure takes into account
@@
 
-Human judgement is the cornerstone for estimating the quality of an MT system.
+Human judgement is the most important criterion for
+estimating the quality of an MT system.
 Nevertheless, common measures for human MT evaluation, such as adequacy and fluency judgements
 or the relative ranking of possible translations, are problematic in two ways.
-First, as the quality of translation is multi-faceted, it is difficult
-to quantify the quality of the entire sentence in a single number. This
+First, as the quality of translation is determined by multiple factors, it is difficult
+to assign a single number that quantifies the quality of the entire sentence. This
 is indeed reflected in the diminishing inter-annotator agreement (IAA) rates of human ranking measures
-with the sentence length \cite{Bojar:2011}.
+%as a function of
+with
+the sentence length \parcite{Bojar:2011}.
 Second, a sentence-level quality score does not indicate what parts of the sentence
@@
 these are often words or n-grams, for manual measures some structural
-information is taken into account \parcite{machacek:bojar:segranks:2015}, or the
-annotators are explicitly asked to mark errors, which however suffers from even
-lower agreement than ranking \parcite{lommel:etal:mqm-iaa:2014}.
+information is taken into account \parcite{machacek:bojar:segranks:2015}.
 % the following is specific to automatic measures:
@@
 quantifying the similarity of the output and the reference in terms of
-their verb argument structure; the most notable of these measures is HMEANT
-\parcite{lo2011structured}.
-
+their verb argument structure; the most notable of these measures is
+HMEANT \parcite{lo2011structured}.
 % SAVING SPACE:
@@
 %HUME only requires semantically annotating the source sentence,
-are thus avoiding the semantic annotation of machine-generated text,
-which is often garbled or semantically unclear.
-This also allows the re-use of the source semantic annotation for
-measuring the quality of different translations of the same source sentence,
-and 
+are thus avoiding the semantic annotation of machine-generated text. 
+This allows the re-use of the source semantic annotation for measuring the quality
+of different translations of the same source sentence.
 %By basing the measure on a direct comparison
 %of the source and translation we avoid the difficulties created by using
-avoids relying on possibly suboptimal reference translations,
-albeit at the cost of requiring the
+We also avoid creating and relying on possibly suboptimal
+reference translations, albeit at the cost of requiring the
 evaluator to be proficient in both the source and target languages.
@@
 (\secref{sec:hume}).
-Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments})
-document HUME's inter-annotator agreement and efficiency (time of annotation). 
-We further empirically compare HUME with direct assessment of human adequacy ratings,
-and conclude by discussing the differences with HMEANT (\secref{sec:hmeant_comp}).
+Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments}) document the
+efficiency (time of annotation) and inter-annotator agreement scores for HUME. 
+We also report on the correlation scores of HUME with direct assesment of human 
+adequacy ratings. 
+Finally, the differences with HMEANT are discussed 
+(\secref{sec:hmeant_comp}).
 
@@
 
-\paragraph{MT Evaluation.}
+\paragraph{Machine Translation Evaluation.}
 %MT evaluation has been traditionally pursued through human-based
@@
 latter (\figref{fig:ucca_example_v2}). Further categories account for
-inter-scene relations and the sub-structures of participants and relations.
+relations between Scenes and the internal structures of complex participants and relations.
 
 % UCCA Motivation
-The use of UCCA for semantic MT evaluation measure is motivated by several reasons. 
-First, UCCA's set of categories can
-be reliably annotated by non-experts after as little as two hours
+UCCA is a natural candidate for defining a semantic
+MT evaluation measure for a number of reasons. 
+First, it defines an intuitive set of categories
+that can be reliably annotated by non-experts, after as little as two hours
 of training \parcite{marinotti2014}.
-Second, UCCA is cross-linguistically applicable, seeking to
-represent what is shared between languages by building on
-linguistic typological theory \parcite{Dixon:10a,Dixon:10b,Dixon:12}.
-Its cross-linguistic applicability has so far been tested in annotations of
+Second, UCCA is a cross-linguistically applicable scheme, seeking to
+represent what is shared between languages
+by building on linguistic typological theory, and specifically on ``Basic Linguistic Theory''
+\parcite{Dixon:10a,Dixon:10b,Dixon:12}.
+The cross-linguistic applicability of UCCA has been tested in annotations of
 English, French, German and Czech.
-Third, the scheme has been shown to be stable across translations:
-UCCA annotations of translated text usually contain the same set of relations
-\parcite{sulem2015conceptual}, indicating that the scheme reflects
+Third, the scheme has been shown to be semantically stable
+in translation: UCCA annotations of translated text usually contain the same set of relationships
+\parcite{sulem2015conceptual}. This indicates that the scheme reflects
 a common layer of representation that in a correct translation
-is mostly shared between the translation and the source\oa{verb argument structure note?}.
-
-%Verbal argument structures are less stable in this sense, as they
-%may be translated or paraphrased to a nominal or adjectival
-%argument structure, and is thus less suitable as a basis for a semantic measure.
-%E.g., ``after he graduated, Tom moved to NYC'' contains two verbs,
-%but is similar in meaning to the one-verbed
-%``after graduation, Tom moved to NYC''.
+is mostly shared between the translation and the source.
+
+We note that verbal argument structures are less stable in this sense, as a verbal
+argument structure may be translated or paraphrased to a nominal or adjectival
+argument structure, and is thus less suitable as a basis for a semantic measure.
+E.g., ``after he graduated, Tom moved to NYC'' contains two verbs,
+but is similar in meaning to the one-verbed
+``after graduation, Tom moved to NYC''.
 
@@
 
-The Abstract Meaning Representation (AMR) \parcite{banarescu2013abstract} project
-shares UCCA's motivation for defining a more complete semantic annotation.
+The Abstract Meaning Representation (AMR) scheme \parcite{banarescu2013abstract}
+shares much of UCCA's motivation for defining a more complete semantic annotation.
 However, using AMR is not optimal for defining a decomposition of a sentence into semantic
 units as it does not ground its semantic symbols in the text,
-and thus does not provide clear decomposition of the sentence into sub-units.
+thus does not provide clear decomposition of the sentence into sub-units.
 Also, AMR is more fine-grained than UCCA and consequently harder to annotate.
-Other approaches represent semantic structures as bi-lexical dependencies
-\parcite{sgallhp:1986,hajic2012announcing,oepen2006discriminant},
-which are indeed grounded in the text, but are less suitable for MT evaluation
-as they require linguistic expertise for their annotation.
-
-%, and so we argue they are less suitable 
-%for an interpretable MT evaluation.
-
-%Examples include the tectogrammatical layer  used in the Prague dependency
-%treebanks, e.g. the Czech-English one \parcite{}, as well
-%as dependencies derived from Minimal Recursion Semantics representations \parcite{}.
-%Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
+Other approaches pursued a grounded representation of semantics,
+and have generally represented semantic structures as bi-lexical dependencies.
+Examples include the tectogrammatical layer \parcite{sgallhp:1986} used in the Prague dependency
+treebanks, e.g. the Czech-English one \parcite{hajic2012announcing}, as well
+as dependencies derived from Minimal Recursion Semantics representations \parcite{oepen2006discriminant}.
+Nevertheless, despite their appeal as deep and intuitive means for semantic representation,
+these approaches require linguistic expertise for their annotation, due to the intricate and
+fine-grained distinctions they provide, and so we argue they are less suitable as a basis
+for an interpretable MT evaluation.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\vspace{-.1em}
 \section{The HUME Measure}\label{sec:hume}
-\vspace{-.2em}
 
@@
 where units are defined according to the UCCA annotation.
-UCCA annotation is performed once for every source sentence,
+Note that UCCA annotation is performed once for every source sentence,
 irrespective of the number of its translations we wish
@@
   here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}.
-A UCCA annotation for a sentence $s$ is a labeled DAG $G$, whose leaves
-are the words of $s$.
-For every node in $G$, we define its {\it yield} to be its leaf descendants.
-The semantic units for $s$ according to $G$ are the yields of nodes in $G$.
+Let $L$ be the set of possible UCCA labels.
+Given a sentence $s = (w_1, \ldots,w_n)$, its UCCA annotation is a DAG $G=(V,E,\ell)$,
+where $\ell: V \rightarrow L$, and where $W = \{w_1,\ldots,w_n\}$ correspond to the leaves of $G$.
+For every node $v \in V$, we define its {\it yield} $yld(v)$ as $v$'s descendants in $W$.
+The semantic units for a sentence $s$ according to $G$
+is the set of yields of nodes in $V$. 
+The translation is not UCCA-annotated.
 
@@
 \paragraph{Translation Evaluation.}
-HUME annotation is done by traversing the semantic units
+The basic strategy for annotating HUME is to traverse the semantic units
 of the source sentence, which correspond to the arguments and relations expressed
-in the text, and marking the extent to which they have been correctly translated.
+in the text, and to mark the extent to which they have been correctly translated.
 HUME aggregates the judgements of the users into a composite score, 
@@
 and ``Red'' (incorrect). 
-Green means that the meaning of the word or phrase has been largely preserved.
-Orange means that the essential meaning of the unit has been preserved,
+A Green label means that the meaning of the word or phrase has been largely preserved.
+An Orange label means that the essential meaning of the unit has been preserved,
 but some part of the translation is wrong.
-This is often be due to the translated word having the the wrong inflection,
-in a way that impacts little on the understandability of the sentence.
-Red means that the essential meaning of the unit has not been captured.
+This could often be due to the translated word having the wrong tense,
+or the wrong morphology, in a way that impacts little on the understandability of the sentence.
+A Red label means that the essential meaning of the unit has not been captured.
 
@@
 Structural units are labeled as ``Adequate'' or ``Bad'', meaning
-that the relation between the sub-units went wrong\footnote{
-  Three labels are used with atomic units, as opposed to two labels with structural units,
-as atomic units are more susceptible to slight errors.}.
-There are three major cases where the relationship between the children might go
-wrong in translation.
-First, the components of the translation may be ordered differently from the source,
+that the relation between the sub-units went wrong.
+We use a three label system with atomic units, as
+opposed to a two label one in structural units,
+as atomic units often have slight errors, in which cases we do want to provide partial credit.
+
+There are many ways that the relationship between the children might go
+wrong in translation:
+
+\begin{compactitem}
+\item The components of the translation are ordered differently from the source,
   such that the relation between them changes. For instance, if
@@
   the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
-Second, A superfluous word or phrase may be inserted into
+\item A superfluous word or phrase has been inserted into
   the translation. For instance, if ``dog bites man'' is translated into ``dog viciously bites man''.
-Third, one of the sub-units may be missing from the translation.
+\item One of the sub-units is missing from the translation.
   For instance, if ``dog bites man'' is translated into ``dog bites''.
+% XXX let's leave the reviewers to notice this:
+%  \bh{Note that in this case we also 
+%  get an error in the atomic unit, assuming that ``man'' is a semantic unit in the source, but for the other
+%  two cases there is no atomic unit error.}
+\end{compactitem}
 
@@
 
-\com{
 In non-configurational languages, the relationship between the
@@
 on configurational vs. non-configurational languages.
-}
 
@@
      HUME annotation of an UCCA tree with a word aligned example translation shown below. 
-Atomic units are labelled using traffic lights and structural units are marked A or B.}
+Atomic units are labelled using traffic lights and structural nodes are marked A or B.}
 \end{figure}
@@
 
-Figure~\ref{fig:hume_tree_v2} presents an example of a HUME
-annotation, where the translation is in English for ease of comprehension.
+In Figure~\ref{fig:hume_tree_v2} we can see an example of a possible HUME
+annotation, with the translation shown in English for ease of comprehension.
 When evaluating ``to NYC'' the annotator looks at the translation and sees the
 word ``stateside''. This word captures the whole phrase and so we mark this
-non-leaf node with an atomic label. Here we choose Orange since
-it approximately captures the meaning in this context.
-The ability to mark non-leaves with atomic labels allows
+non-leaf node with an atomic label. Here we choose ``Orange'' since
+it approximately captures the meaning
+in this context. The ability to mark non-leaves with atomic labels allows
 the annotator to account for translations which only correspond at the phrase
 level. Another feature highlighted in this example is that by separating structural
-and atomic units, we are able to define where an error occurs, and localise
+and atomic nodes, we are able to define where an error occurs, and localise
 the error to its point of origin. The linker ``After'' is translated incorrectly as ``by''
@@
     \begin{center}
-    \includegraphics[width=.8\textwidth]{hume_interface2.jpg}
-    \caption{The HUME annotation tool. The top orange box
-      contains the translation. The source sentence is directly below it, followed by the tree of the source
-      semantic units. Alignments between the source and translation are in italics and
-      unaligned intervening words are in red (see text).}
+    \includegraphics[width=1\textwidth]{hume_interface2.jpg}
+    \caption{The HUME annotation tool, based on the UCCA tool. The faded orange box at the top
+      contains the translation, and directly below it is the source sentence. The tree of semantic units
+      is shown below the source. Alignments between the source and translation appear in italics and unaligned intervening words appear in red.}
     \label{fig:interface}
@@
 We proceed to detailing how judgements on the semantic units
-of the source are aggregated into a composite score. 
+of the source are aggregated into a composite score. We start by taking a parsimonious approach
 %While future extensive empirical evaluation with HUME and its application to MT
@@
 %as a starting point we take a parsimonious approach,
-We start by taking a very simple approach and compute an accuracy score.
-Let $\Green(s,t)$, $\Adequate(s,t)$ and $\Orange(s,t)$ be the number of Green, Adequate and Orange
-units respectively. Let $\Units(s)$ be the number of units marked with any of the labels.
+and compute an accuracy score.
+
+Let $\Correct(s,t)$ be the number of atomic units in $s$ marked as Green relative to $t$ plus
+the number of structural units in $s$ marked as Adequate relative to $t$.
+Let $\Partial(s,t)$ be the number of atomic units in $s$ marked as Orange relative
+to $t$. Let $\setsize{\Units(s)}$ be the total number of units marked with any of the five labels.
 Then HUME's composite score is:
@@
 \vspace{-.5cm}
-{\scriptsize
 \begin{equation}
-  \HUME(s,t) = \frac{\Green(s,t) + \Adequate(s,t) + 0.5\cdot \Orange(s,t)}{\Units(s)}
+  \small
+  \HUME(s,t) = \frac{\Correct(s,t) + 0.5\cdot \Partial(s,t)}{\setsize{\Units(s)}}
   \nonumber
 \end{equation}
-}
-\vspace{-.5cm}
 
@@
 make a judgement according to the full translation.
-Concretely, let $s$ be a source sentence, $t$ a translation,
-and $A \subset 2^{s \times t}$ a many-to-many word alignment.
+
+Concretely, let $s = (w_1,\ldots,w_n)$ be a source sentence, $t = (w'_1,\ldots,w'_n)$ a translation,
+and $A \subset 2^{s \times t}$ a many-to-many word alignment between them.
 If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
-$t$ to be $\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t$.
-
-Where the aligned text is discontinuous in $t$, words between the left
-and right boundaries which are not contained in it (intervening words)
+$t$ to be $a(u)=\bigcup_{(x_s,x_t) \in A \wedge x_s \cap yld(u) \neq \emptyset} x_t \subset t$.
+Where $a(u)$ is discontinuous in the text, words between the left
+and right boundaries of $a(u)$ which are not contained in it (intervening words)
 are presented in a smaller red font.
@@
 seems acceptable but the interleaving words in red notify the annotator to check
-the whole translation, in which the meaning of the expression is not preserved.
-The annotator should thus mark this structural node as Bad.
-
+the whole translation. The meaning of the expression is not preserved in the
+translation
+and the annotator should mark this structural node as ``B'', bad. 
 %The interleaving 
@@
 and four target languages (Czech, German, Polish and Romanian), using text from the public health domain.
-Semantically accurate translation is paramount in this domain, which makes
-it particularly suitable for experimenting on MT evaluation. HUME is both evalauted in terms of its
-consistency (inter-annotator agreement), efficiency (time of annotation) and validity (through a comparison
-with crowd-sourced adequacy judgements).
+The domain is particularly suitable for experimenting on MT evaluation, as semantically
+accurate translation is paramount here, underscoring the importance of reliable
+semantic evaluation measures.
 
@@
 
-%The experiment seeks to address two questions:
-%\begin{inparaenum}
-%\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences?
-%\item Can HUME capture translation issues that other annotation schemes (such as adequacy judgements
-%  and HMEANT) cannot?
-%\end{inparaenum}
-
+The experiment seeks to address two questions:
+\begin{inparaenum}
+\item Can HUME be annotated quickly and reliably, and does the reliability still hold for longer sentences?
+\item Can HUME capture translation issues that other annotation schemes (such as
+adequacy judgements and HMEANT)
+cannot?
 %\item What can HUME annotation tell us about the strengths and weaknesses of MT?
@@
 %probably don't address it in this paper}
-
-\subsection{Datasets and Translation Systems}
+\end{inparaenum}
+
+\subsection{Data Sets and Translation Systems}
 
@@
 using Moses \parcite{Koehn:2007}.  These were trained on large parallel data sets extracted from
-OPUS \parcite{tiedemann:2009}, and the data sets released for the WMT14
-medical translation task \parcite{bojar-EtAl:2014:W14-33}, 
-giving between 45 and 85 million sentences of training data, depending on language pair.
+OPUS \parcite{tiedemann:2009}, and
+the data sets released for the WMT14 medical translation task \parcite{bojar-EtAl:2014:W14-33}, 
+giving between 
+45 and 85 million sentences of training data, depending on language pair.
 %\oa{I think we should probably
@@
 % ... if someone asks us to do so...
+
 These translation systems were used to translate texts derived from both NHS
@@
 related information in Scotland, Cochrane is an international NGO
-which provides independent systematic reviews on health-related research.
-NHS~24 texts come from the ``Health A-Z'' section in the NHS Inform
-website, and Cochrane texts come from their plain language summaries
+which 
+provides independent systematic reviews on health-related research. From NHS 24, we selected the segments from their
+``Health A-Z'' in the NHS Inform website, and the Cochrane texts come from their plain language summaries
 and abstracts.
@@
 
-\subsection{HUME Annotation Statistics}
+\subsection{Annotation Statistics}
 \label{sec:annot_stats}
@@
 computational linguists and one linguist.
+
 For the annotation of the MT output, we recruited two annotators for each of German, Romanian
-and Polish and one main annotator for Czech. For Czech IAA,
-several further annotators worked on a small number of 
-sentences each. We treat these further annotators as one annotator, resulting in two annotators
+and Polish and one main annotator for Czech.
+%In order to allow us to calculate
+For IAA also for
+Czech, several further annotators worked on a small number of 
+sentences each. We treat these further annotators as one annotator, giving us two annotators
 for each language pair.
 The annotators were all native  speakers of the respective target languages and fluent in English.
-\oa{maybe we should put in what training they've received}
 
 \tabref{tab:annot}
-shows the total number of sentences and units annotated by each annotator.
-Not all units in all sentences were annotated, often due to
-the annotator %not finishing a sentence or
-accidentally missing a node.
+shows the total number of sentences and nodes annotated by each annotator. Note
+that not all nodes in all sentences were annotated, often
+due to
+%this was because
+the annotator not finishing a sentence or accidentally missing a node.
 %, but in many cases it was also because it was an implicit 
@@
 \hline
-\#Sentences &  Annot. 1 & 324   & 339  & 351  & 230  \\
+Sentences &  Annot. 1 & 324   & 339  & 351  & 230  \\
  & Annot. 2 & 205 & 104  & 340  & 337 \\
 \hline
-\#Units & Annot. 1 & 8794  & 9253 & 9557  & 6152 \\
+Nodes & Annot. 1 & 8794  & 9253 & 9557  & 6152 \\
  &Annot. 2 & 5553 & 2906  & 9303  & 9228  \\
 \end{tabular}
-\caption{HUME-annotated \#sentences and \#units.}
+\caption{Total numbers of sentences and nodes annotated on the MT output.}
 \label{tab:annot}
@@
 
-\paragraph{Efficiency.}
-We estimate the annotation time using the timestamps
+The annotators worked remotely, at their own pace, so we did not collect precise timing
+information. Nevertheless we are able to estimate the annotation time using the timestamps
 provided by the annotation tool, which are recorded whenever an annotated sentence is
 submitted. Annotators are not able to re-open a sentence once submitted.
-To estimate the annotation time, we compute the time difference between successive 
-sentences, and discard outlying times since we assume annotation was not continuous.
-From inspection of histograms of annotation times, we set the upper threshold at 500 seconds.
-Median annotation times are presented in Table~\ref{tab:annot_times},
-indicating that the annotation
+
+To estimate the annotation time, we calculate the time difference between successive 
+sentences, and discard outlying times since we assume there was a break taken. From inspection
+of histograms of annotation times, we set the upper threshold at 500 seconds. We
+show median annotation times in Table~\ref{tab:annot_times}, indicating that the annotation
 of a sentence takes around 2--4 minutes, with some variation between annotators.
-
-%The annotators worked remotely at their own pace, so no precise timing information is available.
-%Nevertheless
 
@@
 \caption{Median annotation times per sentence, in seconds.
-  In the case of the second Czech 
-  ``annotator'', no timing information is available, as
-  this was a collection of annotators, working in parallel.}
+We do not give times for the second Czech 
+``annotator'', since this was a collection of annotators, working in parallel.}
 \label{tab:annot_times}
@@
 
-\paragraph{Inter-Annotator Agreement.}
+\subsection{Inter-Annotator Agreement}
 \label{sec:iaa}
@@
 \hline
-All units & 4686   & 2793   & 8384   & 5604  \\
+All nodes & 4686   & 2793   & 8384   & 5604  \\
 Kappa & 0.64   & 0.61   & 0.58   & 0.69  \\
 \hline
-Atomic units & 2982 & 1724 & 5386 & 3570 \\
+Atomic nodes & 2982 & 1724 & 5386 & 3570 \\
 Kappa & 0.54 & 0.29 & 0.54 & 0.50 \\
 \hline
-Structural units & 1602 & 1040 & 2655 & 1989 \\
+Structural nodes & 1602 & 1040 & 2655 & 1989 \\
 Kappa & 0.31 & 0.44 & 0.33 & 0.58 \\
 \end{tabular}
-\caption{IAA for the multiply-annotated units,
+\caption{Inter-annotator agreement for the multiply-annotated portions of the data, as
 measured by Cohen's Kappa. }
@@
 In order to assess the consistency of the annotation, we measure the Inter-Annotator
-Agreement (IAA) using Cohen's Kappa on the multiply-annotated units.
+Agreement (IAA) using Cohen's Kappa on the multiply-annotated portions of the data.
 %For German, Polish and Romanian we used two annotators, with a significant overlap
@@
 %each annotating a sub-part of the batch.
-\tabref{tab:iaa} reports the number of units which have two annotations from
+
+\tabref{tab:iaa} reports the number of nodes which have two annotations from
 different annotators and the corresponding Kappas.
@@
 We report the overall Kappa, as well as separate Kappas on atomic
-units (annotated as Red, Orange or Green) and structural units (annotated
+nodes (annotated as Red, Orange or Green) and structural nodes (annotated
 as Adequate or Bad).
 As expected and confirmed by confusion matrices in \figref{fig:heatmap}, there
-is generally little confusion between the two types of units.
+is generally little confusion between the two types of nodes.
 
@@
 
-%presents the variance of kappa with length in each bin covers
-%a length range of 10. and we show .
+To assess whether, as we hoped, HUME would make it easier to reliably annotate long sentences,
+we binned the sentences according to length and measured Kappa on each of the bins. Each bin covers
+a length range of 10, and we show the variance of kappa with length in \figref{fig:iaalength}.
 
@@
 \caption{Kappa versus sentence length, separately for
-structural and atomic units. The numbers are the node counts in each bin. }
+structural and atomic nodes. The numbers are the node counts in each bin. }
 \label{fig:iaalength}
@@
 
-To assess whether, HUME supports the reliable annotation of long sentences,
-we binned the sentences according to length and measured Kappa on each of the bins.
-The Kappas for each bin (bins are 10 words wide) is presented in 
-\figref{fig:iaalength}.
-Results show that there is no reduction of IAA with sentence
+We see from \figref{fig:iaalength} that there is no reduction of IAA with sentence
 length, except perhaps in German. Also, from \tabref{tab:iaa} the overall IAA
-is quite similar for all languages, showing good agreement (0.6--0.7).
+is quite similar (0.6-0.7) for all languages, showing good agreement.
 However, there are differences observed when we break down by node type.
 In particular, we see a contrast  between
-Czech and Polish, where the IAA is higher for atomic than for structural units, and German and Romanian,
+Czech and Polish, where the IAA is higher for atomic than for structural nodes, and German and Romanian,
 where the reverse is true. We also observe low IAA (around 0.3) in the cases of
-German atomic units, and Polish and Czech structural units.
-
-Looking more closely at the areas of disagreement, we see that for the Polish structural units, the 
+German atomic nodes, and Polish and Czech structural nodes.
+
+Looking more closely at the areas of disagreement, we see that for the Polish structural nodes, the 
 proportion of As was quite different between the two annotators (53\% vs. 71\%), whereas for other
 languages the annotators agree in the proportions. We believe that this was because one of the Polish
-annotators did not fully understand the guidelines for structural units, and percolated
-errors up the tree, creating more Bs. For the German lexical units, and Czech structural units, where
-Kappa is also around 0.3, the proportion of such units being marked as ``correct'' is relatively 
+annotators did not fully understand the guidelines for structural nodes, and percolated
+errors up the tree, creating more Bs. For the German lexical nodes, and Czech structural nodes, where
+Kappa is also around 0.3, the proportion of such nodes being marked as ``correct'' is relatively 
 high, meaning that the class distribution is more skewed, so the expected agreement used in the
@@
 
-\def\iaafig #1{\includegraphics[width=0.4\textwidth]{humevsDA_10en-#1.pdf}}
+\def\iaafig #1{\includegraphics[width=0.45\textwidth]{humevsDA_10en-#1.pdf}}
 
@@
 \end{tabular}
-\caption{Plots of HUME scores vs Direct Assessment scores which have been standardised for each crowdsourcing annotator. The regression is calculated using linear least squares.}
+\caption{Plots of HUME scores vs Direct Assesment scores which have been standardised for each crowdsourcing annotator and averaged for 10 annotators.
+HUME scores are averaged where there were two annotations.
+%The regression is calculated using linear least squares.
+}
 \label{fig:dacorrelation}
 \end{figure}
-Recent research~\cite{TODO} has proposed a new method for collecting 
-accuracy ratings. The idea is that collecting a large number of adequacy judgements along a 
-scale of 0 to 100 allows us to apply statistical measures which make them reliable. 
+Recent research~\cite{graham2015accurate,graham2015crowd} has proposed a new approach for collecting 
+accuracy ratings. By crowdsourcing a large number of adequacy judgements for each candidate 
+translation on a fine-grained 
+scale of 0 to 100, we can apply statistical measures which result in reliable aggregate scores. 
 It has been shown that mean  judgements correlate very strongly with each other. 
-We followed~\cite{TODO} and collected up to 15 adequacy judgements on most of the
+We followed~\citep{graham2015accurate} and collected up to 15 adequacy judgements on most of the
 HUME annotated translations for German and Romanian. To ensure high quality judgments
@@
 few Czech and Polish annotators, and removing the IP restriction led to very poor quality results. 
-We can see that HUME and Direct Assessment scores do correlate. 
 We also removed unreliable crowdsourcing annotators
-according to the method discussed in ~\cite{TODO}. The Pearsons correlation
-for en-de is 0.58 and for en-ro is 0.70 which shows that they two measures do correlate. This confirms that 
+according to the method discussed in ~\cite{TODO}.
+
+We can see in Figure~\ref{fig:dacorrelation} that HUME and Direct Assesment scores do correlate. 
+The Pearsons correlation for en-de is 0.58 and for en-ro is 0.70 which confirms that 
 the HUME metric is consistent with previously published methods in human evaluation, even though these
 two approaches are very different in their conception.
-We argue that although annotating HUME requires more expertise and effort, 
-the results are more valuable because a HUME evaluation can tell you what parts
-of the source sentence have been retained.
-Direct assessment scores require a crowd, which is sometimes not available or reliable, and it returns a single number which gives little insight into the quality of
-the translated sentence.
+We argue that although annotating HUME requires more expertise and effort than direct assesment it has
+two advantages. It returns fine-grained information about the translations and it can use one 
+expert annotator. Direct assesment returns only one aggregate score, and requires a large number
+of crowdsourcing workers who may not be available or reliable.
+
+\begin{figure}[t]
+\includegraphics[width=0.45\textwidth]{humevsDAcorrtypes-10en-dero.pdf}
+\caption{Pearsons correlation HUME scores vs DA for different node types.
+Struct is correlation for all structural nodes, P+S is correlation for all process and state nodes, H are parallel scene nodes, A participant, C centre, E elaborators, and L linker nodes.}
+\label{fig:dacorrelationtypes}
+\end{figure}
+
+In Figure~\ref{fig:dacorrelationtypes} we can see an analysis of HUME correlations with DA by HUME node type.
+Generally correlations are good.
+Some types of nodes seem to predict the DA scores better than others.
+The HUME scores on parallel scene nodes (H) and participant nodes (A) seem to correlate more strongly with DA than scores on process and state nodes (P+S) or linker nodes (L).
 
@@
 argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
-relations and copula clauses.
-We are not aware of any empirical evaluation that demonstrates that verb argument structure,
-taken alone, captures the crux of the sentence semantics meanings.\oa{do we mention here that
-  Scenes in their own right correlate less with DA? maybe we should distinguish verbal and non-verbal
-  scenes?} Consider the example:
+relations and copula clauses. Some of the uncovered cases were already observed
+in previous applications of HMEANT to languages other than English (see \perscite{birch-EtAl:2013:WMT} for German, \perscite{bojar:wu:ssst:2012} for
+Czech and by \perscite{chuchunkov-tarelkin-galinskaya:2014:SSST-8} for Russian)
+but the problems are also seen in English.
+
+%This focus on verbal predicates has a number of unfortunate consequences. We show an example below:
+Consider the example:
+
+% \noindent
+% \begin{inparadesc}
+% \item[Source ] a coronary angioplasty may not be technically possible \\
+% \item[Transl.] eine koronare Angioplastie kann nicht technisch m{\"o}glich \\
+% \item[Gloss ]  a coronary angioplasty can not technically possible \\
+% \end{inparadesc}
 
@@
 
-%``a coronary angioplasty may not be technically possible'', which was translated by our en-de system
-%to ``eine koronare Angioplastie kann nicht technisch m{\"o}glich'' (gloss: 
-%a coronary angioplasty can not technically possible).
-
-The German translation 
+The German machine translation 
 is largely correct, except that the main verb ``sein'' (``be'') is omitted.
@@
 
-%This focus on verbal predicates has a number of unfortunate consequences. We show an example below:
-
-% \noindent
-% \begin{inparadesc}
-% \item[Source ] a coronary angioplasty may not be technically possible \\
-% \item[Transl.] eine koronare Angioplastie kann nicht technisch m{\"o}glich \\
-% \item[Gloss ]  a coronary angioplasty can not technically possible \\
-% \end{inparadesc}
-
-\com{
-}
-
 There are however a number of other serious consequences on focusing on verbal
 predicates. We conducted an analysis of the English UCCA Wikipedia
-corpus (5324 sentences) in order to assess the pervasiveness of three phenomena that are not
-well supported by HMEANT.
-First, copula clauses are treated in HMEANT simply as instances of the main verb ``be'', which
+corpus (5324 sentences)
+in order to assess the impact of the following problems: 
+\\
+\begin{inparadesc}
+\item[Copula clauses] are treated in HMEANT simply as instances of the main verb ``be'', which
 generally does not convey the meaning of these clauses. They appear in 21.7\% of the sentences,
 according to conservative estimates that only consider non-auxiliary instances of ``be''.
-Second, nominal argument structures, which are ignored by  HMEANT are in fact highly
+\\
+\item[Nominal argument structures,] which are ignored by  HMEANT are in fact highly
 pervasive, appearing in 48.7\% of the sentences.
-Third, linkers, that express an inter-relation between
+\\
+\item[Linkers] expressing an inter-relation between
 clauses (mainly discourse markers and coordinating conjunctions) appear in 56\% of the
-sentences, but are again ignored by HMEANT.
+sentences, but are ignored by HMEANT.
 An example where a linker is omitted in translation in our experiments is
-\equo{However, this review\dots}, which was translated to
+with the source \equo{However, this review\dots}, which was translated to
 the German \equo{Diese \"Uberpr\"ufung\dots} (lit. \equo{This review}).
 This omission will not be taken into account by HMEANT at all\footnote{Argument structures and linkers are explicitly marked in UCCA.
-Non-auxiliary instances of ``be'' and nouns are identified using the NLTK standard tagger.
-Nominative argument structures are defined as Scenes whose main relation is headed by a noun.}.
+We identified non-auxiliary instances of ``be'' and nouns using the NLTK standard tagger.
+Nominative argument structures are here defined as Scenes whose main relation is headed by a noun.}.
+\end{inparadesc}
 
@@
 
-We note that some of these issues were already observed
-in previous applications of HMEANT to languages other than
-English. See \perscite{birch-EtAl:2013:WMT} for German, \perscite{bojar:wu:ssst:2012} for
-Czech and \perscite{chuchunkov-tarelkin-galinskaya:2014:SSST-8} for Russian.
-
+We are not aware of any empirical evaluation that demonstrates that verb argument structure,
+taken alone, captures the crux of the sentence semantics meanings.
 % I am not sure that this is true
@@
 
-\paragraph{One Structure or Two.}
-HUME only annotates the source, while HMEANT relies on
-two independently constructed structural annotations, one for the reference and one
-for the translation.
+\paragraph{One vs. Two Structures}
+
+HUME uses just one structural annotation (for the source) while HMEANT relies on
+two independently constructed structural annotations (for both the reference and
+the candidate translation)
 %First of all, the annotation process is different. We semantically annotate the
@@
 %We see this difference as crucial where semantic analysis is concerned, as it
-
-Not annotating the translation is appealing as it is often impossible to assign a
-semantic structure to a low quality translation.
-On the other hand, HUME's annotation may be artificially boosting the perceived understandability
-of the sentence by allowing access to the source.
-
-\paragraph{Alignment.}
-In HMEANT, the alignment between the reference and translation structures is a key
+and it is often impossible to assign a semantic structure to a low quality MT output sentence.
+
+However, HMEANT's independent annotation of the source and translation makes a stricter check
+of whether the translation can be understood
+than we do. HUME may be artificially boosting the perceived understandability of
+the sentence by allowing access to the source.
+
+\paragraph{Importance of Alignment}
+
+In HMEANT, the alignment between the (reference and translation) structures is a key
 part of the manual annotation. If the alignment cannot be created, the 
@@
 %interpretation of a PP-attachment, the verb can have one additional modifier in
-one of the structures. It would be desirable to allow modifications to
-the SRL annotations at the alignment stage,
+one of the structures. It would be desirable to allow updates to the structural
+annotations at the alignment stage,
 to avoid unduly penalization due to such spurious divergences.
@@
 %to avoid unduly penalization due to such a spurious divergence.
-The same issue is noted in \parcite{lo:wu:reliability:2014}:
-the IAA on SRL dropped from 90\% to 61\% 
+The same issue was seen by HMEANT authors themselves
+\parcite{lo:wu:reliability:2014}:
+the 
+IAA on SRL dropped from 90\% to 61\% 
 when the two aligned structures were from two different annotators.
+
+%HUME uses only automatic (word-level) alignment, which only serves as a cue for directing
+%the attention of the annotators.
+%The user is expected to mentally correct the alignment as needed.
 HUME uses automatic (word-level) alignment, which only
@@
 
-%HUME uses only automatic (word-level) alignment, which only serves as a cue for directing
-%the attention of the annotators.
-%The user is expected to mentally correct the alignment as needed.
-
-\paragraph{Monolingual vs. Bilingual Evaluation.}
+\paragraph{Monolingual vs. Bilingual Evaluation}
 \label{src-vs-ref}
-HUME diverges from both HMEANT and from shallower automatic measures
-like BLEU or METEOR, in not requiring a reference.
+HUME diverges from HMEANT (and also from the shallow automatic measures
+like BLEU or METEOR) in not requiring a reference.
 Instead, it compares the source directly with the output translation.
+
 This requires us to employ bilingual annotators, but has the benefit of avoiding
-using a reference. Reference translations may often be phrased in many structurally different
-ways, which may again lead unduly penalization of the translation in cases where
-it is a paraphrase of the reference.
-
-%If only monolingual annotators are available, the HUME evaluation could be performed
-%with a reference sentence instead of with the source. 
-%\bh{But this would be a different metric}
-
+the reference translation. It is known that the reference translations are not
+always literal enough and the MT output semantics can be composed differently
+than the semantics of the reference.
+
+If only monolingual annotators are available, the HUME evaluation could be performed
+with a reference sentence instead of with the source. 
+\bh{But this would be a different metric}
 % While HUME avoids this problem,
@@
 
-\com{
-\paragraph{Error Localisation.}
+\paragraph{Localisation of the Error}
+
 In HMEANT, an error in a child node often results in the parent node
@@
 In HUME, errors are localised as much as possible to where they occur,
-by the separation of atomic and structural units,
-which supports a more accurate aggregation of the translation errors
-to a composite score.
-}
-\oa{Omitted point here}
+by the separation of atomic and structural nodes (See Fig \ref{fig:hume_tree_v2} 
+for an example).
+This allows us to calculate the total score of the sentence more accurately.
 
@@
 
-We have introduced HUME, a human semantic MT evaluation measure which addresses
-a wide range of semantic phenomena. We have shown that it can be reliably and 
-efficiently annotated in multiple languages,
-and that annotation quality is robust to sentence length.
-We believe that
-HUME, and a future automated version of HUME, allows for a more fine-grained analysis of translation quality,
-and will be to guide the development of a more semantically aware approach to MT.
+We have introduced HUME, a human semantics based MT evaluation measure which addresses
+a wide range of semantic phenomena. We have shown that it can be efficiently and 
+reliably annotated in 
+several languages, and that annotation quality is robust to sentence length.
+We believe that HUME, and a future automated version of HUME, will be able to guide the
+development of a more semantically aware approach to MT.
 

### 2016-06-02 paper/main.tex

--- 2016-06-02 paper/main.tex
@@
      HUME annotation of an UCCA tree with a word aligned example translation shown below. 
-Atomic units are labelled using traffic lights and structural units are marked A or B.}
+Atomic units are labelled using traffic lights (Red, Orange, Green) and structural units are marked A or B.}
 \end{figure}

### 2016-06-02 paper/main.tex
Final touches to 4.3

@@
 \end{tabular}
-\caption{Plots of HUME scores vs Direct Assesment scores which have been standardised for each crowdsourcing annotator and averaged for 10 annotators.
-HUME scores are averaged where there were two annotations.
+\caption{HUME vs DA scores. DA score have been standardised for each crowdsourcing annotator and averaged across exactly 10 annotators. HUME scores are averaged where there were two annotations. 
 %The regression is calculated using linear least squares.
@@
 Recent research~\cite{graham2015accurate,graham2015crowd} has proposed a new approach for collecting 
-accuracy ratings. By crowdsourcing a large number of adequacy judgements for each candidate 
+accuracy ratings, direct assessment (DA). By crowdsourcing a large number of adequacy judgements for each candidate 
 translation on a fine-grained 
 scale of 0 to 100, we can apply statistical measures which result in reliable aggregate scores. 
-It has been shown that mean  judgements correlate very strongly with each other. 
-We followed~\citep{graham2015accurate} and collected up to 15 adequacy judgements on most of the
+It has been shown that mean DA judgements correlate very strongly with each other. 
+We followed~\parcite{graham2015accurate} and collected up to 15 adequacy judgements on most of the
 HUME annotated translations for German and Romanian. To ensure high quality judgments
@@
 few Czech and Polish annotators, and removing the IP restriction led to very poor quality results.
-We also removed unreliable crowdsourcing annotators
-according to the method discussed in ~\cite{TODO}.
-
-We can see in Figure~\ref{fig:dacorrelation} that HUME and Direct Assesment scores do correlate. 
-The Pearsons correlation for en-de is 0.58 and for en-ro is 0.70 which confirms that 
-the HUME metric is consistent with previously published methods in human evaluation, even though these
-two approaches are very different in their conception.
-We argue that although annotating HUME requires more expertise and effort than direct assesment it has
-two advantages. It returns fine-grained information about the translations and it can use one 
-expert annotator. Direct assesment returns only one aggregate score, and requires a large number
-of crowdsourcing workers who may not be available or reliable.
+
+We can see in Figure~\ref{fig:dacorrelation} that HUME and Direct Assesment scores correlate reasonably well.  
+The Pearsons correlation for en-de is 0.58 and for en-ro is 0.70 overall which confirms that 
+the HUME metric is consistent with an accepted human evaluation method, namely DA, 
+even though the two metrics are very different in their conception. 
+Although DA is a valuable tool for translation research, HUME has two advantages:
+it returns fine-grained semantic information about the quality of translations and it only requires one 
+annotator. Direct assesment returns a single score, and it needs a large 
+crowd which may not be available or reliable.
 
@@
 \caption{Pearsons correlation HUME scores vs DA for different node types.
-Struct is correlation for all structural nodes, P+S is correlation for all process and state nodes, H are parallel scene nodes, A participant, C centre, E elaborators, and L linker nodes.}
+Struct refers to all structural nodes, P+S are all process and state nodes, H parallel scene nodes, A participant, C centre, E elaborators, and L linker nodes.}
 \label{fig:dacorrelationtypes}
@@
 In Figure~\ref{fig:dacorrelationtypes} we can see an analysis of HUME correlations with DA by HUME node type.
-Generally correlations are good.
+We can see that correlation increases when combining different sematic node types and that correlation is highest for node type ``all''.
 Some types of nodes seem to predict the DA scores better than others.
 The HUME scores on parallel scene nodes (H) and participant nodes (A) seem to correlate more strongly with DA than scores on process and state nodes (P+S) or linker nodes (L).
+Center nodes (C) are also more correlated than elaborator nodes (E), which is what we would expect given that center nodes are by definition more important to the semantics of a sentence than elaborators.
 

### 2016-06-02 scripts/calculate_correspondance_hume_DA.py
Improved the corrtypes bar chart

@@ -31,10 +31,12 @@ TYPE = "stnd"
 
 def main():
 
+  corrs = []
+  nums = []
   for lang, code in LANGCODES:
     #for judgements in 5, 10, 15:
-    corrs = []
-    multicorrs = []
+    langcorrs = []
+    langnums = []
     for judgements in [10]:
 
       UCCAfileName = HUME_dir + "/himl2015.en-" + code + ".uccascores" 
@@ -61,11 +63,19 @@ def main():
 
         df = pandas.DataFrame({ 'UCCA' : UCCAresults[score_type],
                             'DA' : DAlist }) 
+        if (score_type == "P"):
+          df2 = pandas.DataFrame({ 'UCCA' : UCCAresults["S"],
+                            'DA' : DAlist }) 
+          df = pandas.concat([df,df2])
+        elif (score_type == "S"):
+          continue
+        
         #df.corr(method='pearson', min_periods=1)
         # df.plot(x='UCCA', y='DA')
         df = df[df['DA'] != ""]
         corr = np.corrcoef(df['UCCA'].tolist(), df['DA'].tolist())[1,0]
-        corrs.append(corr)
+        langcorrs.append(corr)
+        langnums.append(df['DA'].size)
         print ("All: type " + score_type + " NumDA: " + str(judgements) 
                    + ", en-" + code + ", size " + str(len(df['DA'].tolist())) + ": " 
                    + str(corr))
@@ -87,35 +97,49 @@ def main():
           plt.savefig(fname)
           plt.clf()
 
-        df = pandas.DataFrame({ 'UCCA' : MultUCCAresults[score_type],
-                            'DA' : DAlist }) 
-        #df.corr(method='pearson', min_periods=1)
-        # df.plot(x='UCCA', y='DA')
-        df = df[df['DA'] != ""]
-        df = df[df['UCCA'] >= 0]
-        corr = np.corrcoef(df['UCCA'].tolist(), df['DA'].tolist())[1,0]
-        multicorrs.append(corr)
-        print ("Mullt: type " + score_type + " NumDA: " + str(judgements) 
-                   + ", en-" + code + ", size " + str(len(df['DA'].tolist())) + ": " 
-                   + str(corr))
[…]

### 2016-06-02 paper/main.tex
Shortening the text

@@
 Human evaluation of machine translation normally uses sentence-level measures such as relative ranking
-or adequacy/fluency scales. However these provide no insight into the MT errors,
+or adequacy/fluency scales. However these provide no insight into possible errors,
 and do not scale well with sentence length.
@@
 translation quality, and enables the construction and tuning of semantics-based MT. 
-We present a novel human semantic evaluation measure for MT, Human
+We present a novel human semantic evaluation measure, Human
 UCCA-based MT Evaluation (HUME), building on the Universal Conceptual Cognitive Annotation (UCCA)
 semantic representation scheme.
-Making progress over previous work, the proposed measure takes into account
-a wider range of semantic phenomena and does not rely on semantic annotation
-of the MT output itself, whose interpretation is often unclear.
+%Making progress over previous work, the proposed measure takes into account
+HUME covers
+a wider range of semantic phenomena than previous methods and does not rely on semantic annotation
+of the potentially garbled MT output. 
+%itself, whose interpretation is often unclear.
 We experiment with four language pairs, demonstrating HUME's broad applicability,
-as well as its reliability, reflected in its inter-annotator agreement rates and 
-good correlation with human adequacy scores.
+and report good
+%as well as its reliability, reflected in its 
+inter-annotator agreement rates and 
+ correlation with human adequacy scores.
 
@@
 We propose the HUME metric,
-a human evaluation measure that decomposes over the UCCA semantic units of the sentence.
+a human evaluation measure that decomposes over the UCCA semantic units.
+% of the sentence.
 UCCA \parcite{abend2013universal} is an appealing candidate for semantic analysis,
@@
 %of the source and translation we avoid the difficulties created by using
-avoids relying on possibly suboptimal reference translations,
-albeit at the cost of requiring the
-evaluator to be proficient in both the source and target languages.
+avoids relying on possibly suboptimal reference translations.
+%albeit at the cost of requiring the
+%evaluator to be proficient in both the source and target languages.
 %An elaborate comparison of HUME and HMEANT is given in \secref{sec:hmeant_comp}.
 
-After a brief review (\secref{sec:background}), we describe HUME in detail
-(\secref{sec:hume}).
-Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments})
-document HUME's inter-annotator agreement and efficiency (time of annotation). 
-We further empirically compare HUME with direct assessment of human adequacy ratings,
-and conclude by discussing the differences with HMEANT (\secref{sec:hmeant_comp}).
+%Could reinstate
+%After a brief review (\secref{sec:background}), we describe HUME in detail
+%(\secref{sec:hume}).
+%Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments})
+%document HUME's inter-annotator agreement and efficiency (time of annotation). 
+%We further empirically compare HUME with direct assessment of human adequacy ratings,
+%and conclude by discussing the differences with HMEANT (\secref{sec:hmeant_comp}).
 
@@
 adequacy/fluency scores to each translation, a procedure recently improved
-by \perscite{graham2015improving}.
+by \perscite{graham2015accurate}.
 However, while providing the gold standard for MT evaluation, human evaluation
@@
 as atomic units are more susceptible to slight errors.}.
-There are three major cases where the relationship between the children might go
-wrong in translation.
-First, the components of the translation may be ordered differently from the source,
-such that the relation between them changes. For instance, if
-the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
-the semantic unit corresponding to the Scene will be marked as Bad, while
-the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
-Second, A superfluous word or phrase may be inserted into
-the translation. For instance, if ``dog bites man'' is translated into ``dog viciously bites man''.
-Third, one of the sub-units may be missing from the translation.
-For instance, if ``dog bites man'' is translated into ``dog bites''.
+%here are three major cases where the relationship between the children might go
+%wrong in translation and 
+We will use the example ``man bites dog'' to illustrate typical examples of why a structural node
+should be labelled as ``Bad'':
+incorrect ordering (``dog bites man''), 
+deletion (``man bites'') and insertion (``man bites biscuit dog''). 
+%First, the components of the translation may be ordered differently from the source,
+%such that the relation between them changes. For instance, if
+%the sentence ``man bites dog'' is translated into a sentence meaning ``dog bites man'',
+%the semantic unit corresponding to the Scene will be marked as Bad, while
+%the individual atomic units ``dog'', ``bites'' and ``man'' will be marked as Green.
+%Second, A superfluous word or phrase may be inserted into
+%the translation. For instance, if ``dog bites man'' is translated into ``dog viciously bites man''.
+%Third, one of the sub-units may be missing from the translation.
+%For instance, if ``dog bites man'' is translated into ``dog bites''.
 
@@
 \end{figure}
-Recent research~\cite{graham2015accurate,graham2015crowd} has proposed a new approach for collecting 
+Recent research~\cite{graham2015accurate,graham2015crowd,graham2015improving} has proposed a new approach for collecting 
 accuracy ratings, direct assessment (DA). By crowdsourcing a large number of adequacy judgements for each candidate 
@@
 Center nodes (C) are also more correlated than elaborator nodes (E), which is what we would expect given that center nodes are by definition more important to the semantics of a sentence than elaborators.
+We leave it to future work to construct a HUME score which weights the different node types according to their semantic relevance.
 

### 2016-06-03 paper/main.tex
edited

@@
 %\newcommand{\XXX}[1]{}
-\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
-%\newcommand{\oa}[1]{}
+%\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
+\newcommand{\oa}[1]{}
 \newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
@@
 Human evaluation is generally done by ranking the outputs of multiple systems
-e.g., in the WMT tasks \parcite{bojar2015findings}, or by giving
+e.g., in the WMT tasks \parcite{bojar2015findings}, or by assigning
 adequacy/fluency scores to each translation, a procedure recently improved
@@
 Annot. 1 & 255 & 140  & 138 & 96 \\
-Annot. 2 & -- & 162 & 229 & 207 \\
+Annot. 2 & $^*$ & 162 & 229 & 207 \\
 \end{tabular}
 \caption{Median annotation times per sentence, in seconds.
-  In the case of the second Czech 
-  ``annotator'', no timing information is available, as
+  $^*$: no timing information is available, as
   this was a collection of annotators, working in parallel.}
@@
 \end{tabular}
-\caption{Confusion matrices for each of the language pairs, plotted as heatmaps.}
+\caption{Confusion matrices for each language pair.}
 \label{fig:heatmap}
@@
 annotators did not fully understand the guidelines for structural units, and percolated
-errors up the tree, creating more Bs. For the German lexical units, and Czech structural units, where
-Kappa is also around 0.3, the proportion of such units being marked as ``correct'' is relatively 
+errors up the tree, creating more Bs. For German atomic and Czech structural units, where Kappa is also around 0.3, the proportion of such units being marked as ``correct'' is relatively 
 high, meaning that the class distribution is more skewed, so the expected agreement used in the
-Kappa calculation is high, lowering Kappa. Essentially, the annotation problem is harder when the classes
-are more skewed. Finally we note some evidence of domain-specific disagreements, for instance the German
-MT system normally translated ``review'' (as in ``systematic review'' -- a frequent term in the 
+Kappa calculation is high, lowering Kappa.
+%ssentially, the annotation problem is harder when the classes are more skewed.
+Finally we note some evidence of domain-specific disagreements, for instance
+the German MT system normally translated ``review'' (as in ``systematic review'' -- a frequent term in the 
 Cochrane texts) as ``\"uberpr\"ufung'', which 
@@
 
-\subsection{Comparison with Crowdsourced Adequacy Judgements}\label{sec:adequacy}
+\subsection{Comparison with Direct Assessment}\label{sec:adequacy}
 
@@
 \end{figure}
-Recent research~\cite{graham2015accurate,graham2015crowd,graham2015improving} has proposed a new approach for collecting 
-accuracy ratings, direct assessment (DA). By crowdsourcing a large number of adequacy judgements for each candidate 
-translation on a fine-grained 
-scale of 0 to 100, we can apply statistical measures which result in reliable aggregate scores. 
-It has been shown that mean DA judgements correlate very strongly with each other. 
-We followed~\parcite{graham2015accurate} and collected up to 15 adequacy judgements on most of the
-HUME annotated translations for German and Romanian. To ensure high quality judgments
-we restricted the IP of the annotators to those of the relevant country. We found that there were too
-few Czech and Polish annotators, and removing the IP restriction led to very poor quality results.
-
-We can see in Figure~\ref{fig:dacorrelation} that HUME and Direct Assesment scores correlate reasonably well.  
-The Pearsons correlation for en-de is 0.58 and for en-ro is 0.70 overall which confirms that 
-the HUME metric is consistent with an accepted human evaluation method, namely DA, 
-even though the two metrics are very different in their conception. 
-Although DA is a valuable tool for translation research, HUME has two advantages:
-it returns fine-grained semantic information about the quality of translations and it only requires one 
-annotator. Direct assesment returns a single score, and it needs a large 
-crowd which may not be available or reliable.
+
+Recent research~\cite{graham2015accurate,graham2015crowd,graham2015improving} has proposed a new approach for collecting accuracy ratings, direct assessment (DA).
+By crowdsourcing a large number of adequacy judgements for each candidate translation on a fine-grained scale of 0 to 100, we can apply statistical measures which result in reliable aggregate scores, that correlate very strongly with one other.
+
+We followed~\parcite{graham2015accurate} and collected 10 adequacy
+judgements on most of the HUME annotated translations for
+German and Romanian, using the CrowdFlower platform\footnote{\url{https://www.crowdflower.com}}.
+To ensure high quality judgments we restricted the IP of the annotators to those of the relevant country.
+Czech and Polish judgements were
+excluded, as our experiments only yielded very little high quality data.
+
+Figure~\ref{fig:dacorrelation} plots the HUME score for each sentence against its DA score.
+HUME and Direct Assesment scores correlate reasonably well.
+The Pearson correlation for en-ro (en-de) is 0.70 (0.58), and 0.78 (0.74) where only doubly HUME-annotated points are taken.
+This confirms that HUME is consistent with an accepted human evaluation method, despite the differences in
+their conception.
+While DA is a valuable tool for MT research, HUME has two advantages:
+it returns fine-grained semantic information about 
+the quality of translations and it only requires very few annotators.
+Direct assesment returns a single opaque score, and (as also noted by
+Graham et al.), requires a large crowd which may not be available or reliable.
 
@@
 \includegraphics[width=0.45\textwidth]{humevsDAcorrtypes-10en-dero.pdf}
-\caption{Pearsons correlation HUME scores vs DA for different node types.
+\caption{Pearson correlation HUME scores vs DA for different node types.
 Struct refers to all structural nodes, P+S are all process and state nodes, H parallel scene nodes, A participant, C centre, E elaborators, and L linker nodes.}
@@
 
-In Figure~\ref{fig:dacorrelationtypes} we can see an analysis of HUME correlations with DA by HUME node type.
-We can see that correlation increases when combining different sematic node types and that correlation is highest for node type ``all''.
+Figure~\ref{fig:dacorrelationtypes} presents an analysis of HUME correlations with DA by HUME node type.
+We can see that correlation is highest where all sub-sets are taken into consideration (``all'' bar), supporting our claim for the value of taking into account a rich set of semantic phenomena.
 Some types of nodes seem to predict the DA scores better than others.
-The HUME scores on parallel scene nodes (H) and participant nodes (A) seem to correlate more strongly with DA than scores on process and state nodes (P+S) or linker nodes (L).
+The HUME scores on Parallel Scenes (H) and Participants (A) seem to correlate more strongly with DA than scores on Scene Main Relations (P+S) or linker nodes (L).
 Center nodes (C) are also more correlated than elaborator nodes (E), which is what we would expect given that center nodes are by definition more important to the semantics of a sentence than elaborators.
-We leave it to future work to construct a HUME score which weights the different node types according to their semantic relevance.
+We leave it to future work to construct a HUME score which weights the different node types according to their semantic relevance.\oa{I think we should be careful here.
+It may very well be that these stem from the number of nodes taken into account.
+I think we should also say explicitly that .}
+This is also a demonstration of the more refined analysis (allowing to slice the data by semantic phenomena) enabled by taking a semantic approach to MT evaluation.
 
@@
 
+\vspace{-.3em}
 \begin{center}
@@
 \end{center}
+\vspace{-.5em}
 

### 2016-06-03 paper/main.tex
abbreviating

@@
 \includegraphics[width=0.45\textwidth]{humevsDAcorrtypes-10en-dero.pdf}
-\caption{Pearson correlation HUME scores vs DA for different node types.
-Struct refers to all structural nodes, P+S are all process and state nodes, H parallel scene nodes, A participant, C centre, E elaborators, and L linker nodes.}
+\caption{Pearson correlation HUME scores vs DA for different node types: all
+structural nodes (struct), process and state nodes jointly (P+S), parallel
+scenes (H), participants (P), centres (C), elaborators (E) and linkers
+(L).}
+%Struct refers to all structural nodes, P+S are all process and state nodes, H parallel scene nodes, A participant, C centre, E elaborators, and L linker nodes.}
 \label{fig:dacorrelationtypes}

### 2016-06-03 paper/main.tex
bin size
space

--- 2016-06-03 paper/main.tex
@@
 
-\titlespacing*{\section}
-{0pt}{1ex plus 1ex minus .2ex}{1ex plus .2ex}
-\titlespacing*{\subsection}
-{0pt}{1ex plus 1ex minus .2ex}{1ex plus .2ex}
-
-\renewcommand{\paragraph}{
-  \@startsection{paragraph}{4}
-  {\z@}{.1ex \@plus .1ex \@minus .1ex}{-1.5em}
-  {\normalfont\normalsize\bfseries}
-}
+%\titlespacing*{\section}
+%{0pt}{1ex plus 1ex minus 1ex}{1ex plus .2ex}
+%\titlespacing*{\subsection}
+%              {0pt}{1ex plus 1ex minus 1ex}{1ex plus .2ex}
+
+\titlespacing{\section}{2pt}{*0}{*0}
+\titlespacing{\subsection}{2pt}{*0}{*0}
+\titlespacing{\subsubsection}{2pt}{*0}{*0}
+\titlespacing{\paragraph}{1pt}{*0}{3pt}
+
 \makeatother
@@
 
-Human evaluation of machine translation normally uses sentence-level measures such as relative ranking
-or adequacy/fluency scales. However these provide no insight into possible errors,
+Human evaluation of machine translation normally uses sentence-level measures such as relative ranking or adequacy scales. However these provide no insight into possible errors,
 and do not scale well with sentence length.
@@
 We present a novel human semantic evaluation measure, Human
-UCCA-based MT Evaluation (HUME), building on the Universal Conceptual Cognitive Annotation (UCCA)
-semantic representation scheme.
+UCCA-based MT Evaluation (HUME), building on the UCCA semantic representation scheme.
 %Making progress over previous work, the proposed measure takes into account
@@
 %Could reinstate
-%After a brief review (\secref{sec:background}), we describe HUME in detail
-%(\secref{sec:hume}).
-%Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments})
-%document HUME's inter-annotator agreement and efficiency (time of annotation). 
-%We further empirically compare HUME with direct assessment of human adequacy ratings,
-%and conclude by discussing the differences with HMEANT (\secref{sec:hmeant_comp}).
+After a brief review (\secref{sec:background}), we describe HUME in detail
+(\secref{sec:hume}).
+Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments}) document HUME's inter-annotator agreement and efficiency (time of annotation).
+We further empirically compare HUME with direct assessment of human adequacy ratings, and conclude by discussing the differences with HMEANT (\secref{sec:hmeant_comp}).
 
@@
 L & Linker &  A & Participant \\
-H & Scene & R & Relater \\
+H & Parallel Scene & R & Relater \\
 P & Process & C & Centre \\
@@
 \caption{\label{fig:ucca_example_v2}
-  Sample UCCA annotation where leaves correspond to words and nodes to units.
+  Sample UCCA annotation. Leaves correspond to words and nodes to units.
   The dashed edge indicates that ``Tom'' is also a participant in the ``moved to NYC''
-  Scene. Edge labels mark UCCA categories, listed  in the table.}
+  Scene. Edge labels mark UCCA categories.}
 %which are largely ignored in this work
@@
 irrespective of the number of its translations we wish
-to evaluate\footnote{Instead of relying on the source,
-the evaluation could be based on the UCCA structure for the \emph{reference}
-translation.
-More details in the discussion in \secref{src-vs-ref}.
+to evaluate
+%\footnote{Instead of relying on the source,
+%the evaluation could be based on the UCCA structure for the \emph{reference}
+%translation.
+%More details in the discussion in \secref{src-vs-ref}.
 % This would allow monolingual annotators to take part in the
@@
 % translations.
-}.
+%}.
 
@@
 
-Structural units have sub-units (children in the UCCA graph), which may themselves
-either be atomic or structural.
+Structural units have sub-units (children in the UCCA graph),
+which are themselves atomic or structural.
 % Ondrej thinks the following can be simply removed:
@@
 
-We turn to discussing the main differences between HUME and HMEANT,
-a semantic measure that measures the translation quality according to its similarity
-with a reference, measures in terms of the overlap between their SRL annotations.
+We discuss the main differences between HUME and HMEANT,
+a human MT evaluation metric that measures the overlap between the translation a reference
+in terms of their SRL annotations.
+
 %Like HMEANT, HUME defines  a
@@
 
-HMEANT focuses on verbal
-argument structures, ignoring other pervasive phenomena such as nominal predicates, inter-clause
-relations and copula clauses.
-We are not aware of any empirical evaluation that demonstrates that verb argument structure,
-taken alone, captures the crux of the sentence semantics meanings.
-Moreover, relying only on verbal argument structures is less stable across paraphrases
-and translations, as a verbal argument structure
-may be translated to a nominal or adjectival argument structure
-(e.g., ``after graduation'' may be translated into ``after he graduated'').
-This may lead to unjustified penalties low HMEANT score, as the verb
-in one structure has nothing to align to in the other.
-On the other hand, HUME has been shown to be reasonably stable in an English-French
-corpus study \cite{sulem2015conceptual}.
-
-Consider the following example:
+HMEANT focuses on verbal argument structures, ignoring other pervasive phenomena such as non-verbal predicates and inter-clause relations. Consider the following example:
 
@@
 
-There are however a number of other serious consequences on focusing on verbal
-predicates. We conducted an analysis of the English UCCA Wikipedia
-corpus (5324 sentences) in order to assess the pervasiveness of three phenomena that are not
-well supported by HMEANT.
-First, copula clauses are treated in HMEANT simply as instances of the main verb ``be'', which
-generally does not convey the meaning of these clauses. They appear in 21.7\% of the sentences,
-according to conservative estimates that only consider non-auxiliary instances of ``be''.
-Second, nominal argument structures, which are ignored by  HMEANT are in fact highly
+We conducted an analysis of the English UCCA Wikipedia corpus (5324 sentences) in order to assess the pervasiveness of three phenomena that are not well supported by HMEANT.
+First, copula clauses are treated in HMEANT simply as instances of the main verb ``be'', which generally does not convey the meaning of these clauses.
+They appear in 21.7\% of the sentences, according to conservative estimates that only consider non-auxiliary
+instances of ``be''.
+Second, nominal argument structures, ignored by HMEANT, are in fact highly
 pervasive, appearing in 48.7\% of the sentences.
-Third, linkers, that express an inter-relation between
-clauses (mainly discourse markers and coordinating conjunctions) appear in 56\% of the
-sentences, but are again ignored by HMEANT.
-An example where a linker is omitted in translation in our experiments is
-\equo{However, this review\dots}, which was translated to
-the German \equo{Diese \"Uberpr\"ufung\dots} (lit. \equo{This review}).
-This omission will not be taken into account by HMEANT at all\footnote{Argument structures and linkers are explicitly marked in UCCA.
+Third, linkers that express inter-relations between
+clauses (mainly discourse markers and conjunctions) appear in 56\% of the
+sentences, but are again ignored by HMEANT. As noted in our experiments, linkers
+are sometimes omitted in translation, but these omissions are not taken
+into consideration by HMEANT\footnote{Argument structures and linkers are explicitly marked in UCCA.
 Non-auxiliary instances of ``be'' and nouns are identified using the NLTK standard tagger.
 Nominal argument structures are here Scenes whose main relation is headed by a noun.}.
+
+%An example where a linker is omitted in translation in our experiments is
+%\equo{However, this review\dots}, which was translated to
+%the German \equo{Diese \"Uberpr\"ufung\dots} (lit. \equo{This review}).
+%This omission will not be taken into account by HMEANT at all
+
+We are not aware of any empirical argument suggesting that verb argument structures,
+taken alone, capture the crux of the sentence semantics.
+Moreover, relying only on verbal argument structures is less stable across paraphrases
+and translations, as a verbal argument structure
+may be translated to a nominal or adjectival argument structure
+(e.g., ``after graduation'' may be translated into ``after he graduated'').
+This may lead to an unduly low HMEANT score, as the verb
+in one structure has nothing to align to in the other.
+On the other hand, UCCA has been shown to be reasonably stable in an English-French
+corpus study \cite{sulem2015conceptual}.
 
@@
 semantic structure to a low quality translation.
-On the other hand, HUME's annotation may be artificially boosting the perceived understandability
-of the sentence by allowing access to the source.
+On the other hand, HUME may be artificially boosting the perceived understandability
+of the translation by allowing access to the source.
 
@@
 translation is heavily penalized.
-Previous experiments with HMEANT
-\parcite{bojar:wu:ssst:2012,chuchunkov-tarelkin-galinskaya:2014:SSST-8} have
-suggested that sometimes, the structures of the reference and of an accurate translation
-still diverge, for instance due to a different
+\newcite{bojar:wu:ssst:2012} and \newcite{chuchunkov-tarelkin-galinskaya:2014:SSST-8}
+argue that the structures of the reference and of an accurate translation
+may still diverge, for instance due to a different
 interpretation of a PP-attachment, or the verb having an additional modifier in
@@
 the SRL annotations at the alignment stage,
-to avoid unduly penalization due to such spurious divergences.
+to avoid unduly penalizing such spurious divergences.
 %annotations also at the alignment stage
@@
 \label{src-vs-ref}
-HUME diverges from both HMEANT and from shallower automatic measures
-like BLEU or METEOR, in not requiring a reference.
+HUME diverges from HMEANT and from shallower measures
+like BLEU, in not requiring a reference.
 Instead, it compares the source directly with the output translation.
-This requires us to employ bilingual annotators, but has the benefit of avoiding
-using a reference. Reference translations may often be phrased in many structurally different
-ways, which may again lead unduly penalization of the translation in cases where
-it is a paraphrase of the reference.
+This requires the employment of bilingual annotators, but has the benefit of avoiding
+using a reference, which is never uniquely defined, and may thus 
+lead to unjustly low scores where the translation is a paraphrase of the reference.
 

### 2016-06-03 paper/main.tex
abbreviating

--- 2016-06-03 paper/main.tex
@@
 To assess whether, HUME supports the reliable annotation of long sentences,
-we binned the sentences according to length and measured Kappa on each of the bins.
-The Kappas for each bin (bins are 10 words wide) is presented in 
-\figref{fig:iaalength}.
-Results show that there is no reduction of IAA with sentence
+we binned the sentences according to length and measured Kappa on each of the
+bins (\figref{fig:iaalength}).
+%The Kappas for each bin is presented in 
+%\figref{fig:iaalength}.
+%Results show that there is
+We see
+no reduction of IAA with sentence
 length, except perhaps in German. Also, from \tabref{tab:iaa} the overall IAA

### 2016-06-03 paper/main.tex
space

--- 2016-06-03 paper/main.tex
@@
 %\newcommand{\XXX}[1]{}
-%\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
-\newcommand{\oa}[1]{}
+\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
+%\newcommand{\oa}[1]{}
 \newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}

### 2016-06-03 paper/main.tex
rephrased DA, stressing the issue with getting annotations

--- 2016-06-03 paper/main.tex
@@
 Recent research~\cite{graham2015accurate,graham2015crowd,graham2015improving} has proposed a new approach for collecting accuracy ratings, direct assessment (DA).
-By crowdsourcing a large number of adequacy judgements for each candidate translation on a fine-grained scale of 0 to 100, we can apply statistical measures which result in reliable aggregate scores, that correlate very strongly with one other.
-
-We followed~\parcite{graham2015accurate} and collected 10 adequacy
+%By crowdsourcing a large number of adequacy judgements for each candidate translation on a fine-grained scale of 0 to 100, we can apply statistical measures which result in reliable aggregate scores, that correlate very strongly with one other.
+Statistical interpretation of a large number of crowd-sourced adequacy
+judgements for each candidate translation on a fine-grained scale of 0 to 100
+results in reliable aggregate scores, that correlate very strongly with one
+other.
+
+We attempted to follow \perscite{graham2015accurate} but struggled to get enough
+judgements for our target languages. We ended up with 10 adequacy 
 judgements on most of the HUME annotated translations for
-German and Romanian, using the CrowdFlower platform\footnote{\url{https://www.crowdflower.com}}.
+German and Romanian but insufficient data for Czech and Polish. We see this as a
+severe practical limitation of DA.
+%We followed~\parcite{graham2015accurate} and collected 10 adequacy
+%judgements on most of the HUME annotated translations for
+%German and Romanian, using the CrowdFlower platform\footnote{\url{https://www.crowdflower.com}}.
 To ensure high quality judgments we restricted the IP of the annotators to those of the relevant country.
 Czech and Polish judgements were
-excluded, as our experiments only yielded very little high quality data.
+%excluded, as our experiments only yielded very little high quality data.
 

### 2016-06-03 paper/main.tex
minor
moved footnote 7

@@
 
-Figure~\ref{fig:dacorrelation} plots the HUME score for each sentence against its DA score.
-HUME and Direct Assesment scores correlate reasonably well.
-The Pearson correlation for en-ro (en-de) is 0.70 (0.58), and 0.78 (0.74) where only doubly HUME-annotated points are taken.
-This confirms that HUME is consistent with an accepted human evaluation method, despite the differences in
+Figure~\ref{fig:dacorrelation} plots the HUME score for each sentence against
+its DA score. HUME and Direct Assesment scores correlate reasonably well. The
+Pearson correlation for en-ro (en-de) is 0.70 (0.58), and 0.78 (0.74) if only
+doubly HUME-annotated points are considered. This confirms that HUME is consistent with an accepted human evaluation method, despite the differences in
 their conception.
@@
 
-We conducted an analysis of the English UCCA Wikipedia corpus (5324 sentences) in order to assess the pervasiveness of three phenomena that are not well supported by HMEANT.
+We conducted an analysis of the English UCCA Wikipedia corpus (5324 sentences) in order to assess the pervasiveness of three phenomena that are not well supported by HMEANT.\footnote{Argument structures and linkers are explicitly marked in UCCA.
+Non-auxiliary instances of ``be'' and nouns are identified using the NLTK standard tagger.
+Nominal argument structures are here Scenes whose main relation is headed by a noun.}
 First, copula clauses are treated in HMEANT simply as instances of the main verb ``be'', which generally does not convey the meaning of these clauses.
@@
 are sometimes omitted in translation, but these omissions are not taken
-into consideration by HMEANT\footnote{Argument structures and linkers are explicitly marked in UCCA.
-Non-auxiliary instances of ``be'' and nouns are identified using the NLTK standard tagger.
-Nominal argument structures are here Scenes whose main relation is headed by a noun.}.
+into consideration by HMEANT.
 
@@
 }
-\oa{Omitted point here}
+%\oa{Omitted point here}
 

### 2016-06-03 paper/main.tex
put the contents of the omitted point in one extra sentence

@@
 sentence a very low score, as it failed to translate the main verb.
+%
+Conversely, HMEANT does not allow to penalize errors such as tense or negation
+flip in a correctly aligned predicate.
 
@@
 translation is heavily penalized.
-\newcite{bojar:wu:ssst:2012} and \newcite{chuchunkov-tarelkin-galinskaya:2014:SSST-8}
+\perscite{bojar:wu:ssst:2012} and \perscite{chuchunkov-tarelkin-galinskaya:2014:SSST-8}
 argue that the structures of the reference and of an accurate translation
@@
 %to avoid unduly penalization due to such a spurious divergence.
-The same issue is noted in \parcite{lo:wu:reliability:2014}:
+The same issue is noted by \perscite{lo:wu:reliability:2014}:
 the IAA on SRL dropped from 90\% to 61\% 
@@
 }
-%\oa{Omitted point here}
+%\oa{Omitted point here} % OB: Added a sentence covering it: Conversely,
+%HMEANT does not...
 

### 2016-06-03 paper/main.tex
abbreviating

@@
 
-To assess whether, HUME supports the reliable annotation of long sentences,
+To assess HUME reliability for long sentences,
 we binned the sentences according to length and measured Kappa on each of the

### 2016-06-03 paper/main.tex
abbreviating caption of fig5
on 8 pages with omri's comments out

@@
 
-\begin{figure}[t]
-\renewcommand{\tabcolsep}{0pt}
-\begin{tabular}{cc}
-
-\subfloat[English-Czech]{
-  \iaafig{cs}
-}
-&
-\subfloat[English-German]{
-  \iaafig{de}
-
-}
-\\
-
-\subfloat[English-Polish]{
-  \iaafig{pl}
-
-}
-&
-\subfloat[English-Romanian]{
-  \iaafig{ro}
-
-}
-\end{tabular}
-\caption{Kappa versus sentence length, separately for
-structural and atomic units. The numbers are the node counts in each bin. }
-\label{fig:iaalength}
-\end{figure}
-
 To assess HUME reliability for long sentences,
@@
 
+\begin{figure}[t]
+\renewcommand{\tabcolsep}{0pt}
+\begin{tabular}{cc}
+
+\subfloat[English-Czech]{
+  \iaafig{cs}
+}
+&
+\subfloat[English-German]{
+  \iaafig{de}
+
+}
+\\
+
+\subfloat[English-Polish]{
+  \iaafig{pl}
+
+}
+&
+\subfloat[English-Romanian]{
+  \iaafig{ro}
+
+}
+\end{tabular}
+%\caption{Kappa versus sentence length, separately for
+%structural and atomic units. The numbers are the node counts in each bin. }
+\caption{Kappa versus sentence length for
+structural and atomic units. (Node counts in bins on top of each bar.)
+}
+\label{fig:iaalength}
+\end{figure}
+
 Looking more closely at the areas of disagreement, we see that for the Polish structural units, the 
@@
 Center nodes (C) are also more correlated than elaborator nodes (E), which is what we would expect given that center nodes are by definition more important to the semantics of a sentence than elaborators.
-We leave it to future work to construct a HUME score which weights the different node types according to their semantic relevance.\oa{I think we should be careful here.
-It may very well be that these stem from the number of nodes taken into account.
-I think we should also say explicitly that .}
+We leave it to future work to construct a HUME score which weights the different node types according to their semantic relevance.
+\oa{I think we should be careful here. It may very well be that these stem from the number of nodes taken into account. I think we should also say explicitly that .}
 This is also a demonstration of the more refined analysis (allowing to slice the data by semantic phenomena) enabled by taking a semantic approach to MT evaluation.

### 2016-06-03 paper/main.tex
typo

@@
 information is taken into account \parcite{machacek:bojar:segranks:2015}, or the
-annotators are explicitly asked to mark errors, which however suffer from even
+annotators are explicitly asked to mark errors, which however suffers from even
 lower agreement than ranking \parcite{lommel:etal:mqm-iaa:2014}.

### 2016-06-03 scripts/extract_node_nums.py
Adding node counts for correltaion via node type

@@ -1,0 +1,56 @@
+
+#!/usr/bin/env python
+from __future__ import print_function, division
+
+import sys, string
+import pandas
+import numpy
+from  pandas_confusion import ConfusionMatrix
+import extract_sentences_for_WMT_metrics_task
+
+
+SCORE_TYPES = ['all','atomic','struct','P','S','C','H','E','A','L']
+LANGS = ['ro','de','cs','pl']
+LANGS = ['de','ro']
+
+prefix="data/wmtmetrics/himl2015"
+
+DAsents ={}
+DAsents['ro'] = [641, 711, 639, 195, 575, 11, 721, 659, 101, 785, 625, 687, 213, 291, 729, 381, 435, 185, 375, 45, 485, 63, 611, 457, 59, 405, 1, 47, 767, 401, 31, 233, 647, 475, 391, 737, 327, 449, 469, 169, 299, 257, 17, 689, 129, 395, 263, 289, 145, 617, 349, 249, 369, 357, 131, 555, 55, 559, 43, 345, 419, 561, 265, 79, 201, 223, 759, 261, 171, 339, 459, 217, 583, 733, 465, 331, 619, 663, 749, 677, 651, 675, 727, 609, 313, 589, 67, 603, 507, 207, 231, 85, 247, 597, 719, 649, 461, 593, 65, 613, 97, 139, 443, 761, 633, 367, 199, 541, 297, 567, 177, 411, 181, 203, 255, 753, 743, 635, 545, 359, 205, 445, 489, 135, 713, 717, 229, 535, 127, 91, 509, 661, 83, 599, 585, 173, 227, 73, 513, 483, 453, 269, 377, 493, 107, 159, 373, 533, 325, 491, 517, 577, 521, 329, 9, 655, 347, 383, 403, 117, 573, 81, 765, 311, 783, 49, 211, 735, 219, 23, 275, 307, 75, 397, 701, 143, 547, 183, 571, 433, 549, 303, 187, 113, 605, 165, 19, 155, 321, 429, 225, 631, 95, 133, 189, 607, 121, 251, 793, 591, 241, 441, 273, 193, 341, 315, 209, 421, 563, 27, 167, 741, 285, 543, 769, 115, 531, 703, 109, 499, 335, 601, 351, 277, 681, 309, 243, 473, 515, 35, 51, 437, 645, 179, 197, 683, 137, 775, 467, 175, 37, 301, 705, 99, 385, 725, 279, 671, 271, 337, 799, 237, 21, 71, 667, 149]
[…]

### 2016-06-03 paper/main.tex
xxx

@@
 its DA score. HUME and Direct Assesment scores correlate reasonably well. The
-Pearson correlation for en-ro (en-de) is 0.70 (0.58), and 0.78 (0.74) if only
-doubly HUME-annotated points are considered. This confirms that HUME is consistent with an accepted human evaluation method, despite the differences in
-their conception.
-While DA is a valuable tool for MT research, HUME has two advantages:
+Pearson correlation for en-ro (en-de) is 0.70 (0.58), or 0.78 (0.74) if only
+doubly HUME-annotated points are considered.
+This confirms that HUME is consistent with an accepted human evaluation method,
+despite the differences in their conception.
+While DA is a valuable tool, HUME has two advantages:
 it returns fine-grained semantic information about 
@@
 Direct assesment returns a single opaque score, and (as also noted by
-Graham et al.), requires a large crowd which may not be available or reliable.
+Graham et al.), and requires a large crowd which may not be available or reliable.
 
@@
 
-Figure~\ref{fig:dacorrelationtypes} presents an analysis of HUME correlations with DA by HUME node type.
-We can see that correlation is highest where all sub-sets are taken into consideration (``all'' bar), supporting our claim for the value of taking into account a rich set of semantic phenomena.
+Figure~\ref{fig:dacorrelationtypes} presents an analysis of HUME correlations with DA by HUME unit type.
+Each bar represents a correlation between HUME and an aggregate HUME score
+based on a sub-set of the units (\#nodes for the en-de/en-ro setting in brackets):
+all units ('all', 8726/11087), atomic ('atomic', 5417/6888) and structural units ('struct', 3207/3997),
+and units by UCCA categories: Scene main relations (i.e, Process and State units; 'P+S', / ), Participants ('A', / )
+Centres ('C', / ), elaborators ('E', / ) and linkers ('L', / )
+In both languages correlation is highest in the 'all' case, supporting our claim for the value of taking into account a rich set of semantic phenomena.
 Some types of nodes seem to predict the DA scores better than others.

### 2016-06-03 paper/main.tex
minor correction - edit rate
dd

@@
 \includegraphics[width=0.45\textwidth]{humevsDAcorrtypes-10en-dero.pdf}
-\caption{Pearson correlation HUME scores vs DA for different node types: all
-structural nodes (struct), process and state nodes jointly (P+S), parallel
-scenes (H), participants (P), centres (C), elaborators (E) and linkers
-(L).}
+\caption{Pearson correlation of HUME vs. DA scores for different node types for both
+  the en-de and en-ro experiments. See text for the bar legend.
 %Struct refers to all structural nodes, P+S are all process and state nodes, H parallel scene nodes, A participant, C centre, E elaborators, and L linker nodes.}
-\label{fig:dacorrelationtypes}
+\label{fig:dacorrelationtypes}}
 \end{figure}
 
-Figure~\ref{fig:dacorrelationtypes} presents an analysis of HUME correlations with DA by HUME unit type.
+Figure~\ref{fig:dacorrelationtypes} presents an analysis of HUME's correlations with DA by HUME unit type,
+an analysis enabled by HUME's semantic decomposition. 
 Each bar represents a correlation between HUME and an aggregate HUME score
 based on a sub-set of the units (\#nodes for the en-de/en-ro setting in brackets):
-all units ('all', 8726/11087), atomic ('atomic', 5417/6888) and structural units ('struct', 3207/3997),
-and units by UCCA categories: Scene main relations (i.e, Process and State units; 'P+S', / ), Participants ('A', / )
-Centres ('C', / ), elaborators ('E', / ) and linkers ('L', / )
-In both languages correlation is highest in the 'all' case, supporting our claim for the value of taking into account a rich set of semantic phenomena.
-Some types of nodes seem to predict the DA scores better than others.
-The HUME scores on Parallel Scenes (H) and Participants (A) seem to correlate more strongly with DA than scores on Scene Main Relations (P+S) or linker nodes (L).
-Center nodes (C) are also more correlated than elaborator nodes (E), which is what we would expect given that center nodes are by definition more important to the semantics of a sentence than elaborators.
-We leave it to future work to construct a HUME score which weights the different node types according to their semantic relevance.
-\oa{I think we should be careful here. It may very well be that these stem from the number of nodes taken into account. I think we should also say explicitly that .}
-This is also a demonstration of the more refined analysis (allowing to slice the data by semantic phenomena) enabled by taking a semantic approach to MT evaluation.
+all units ('all', 8624/10885), atomic ('atomic', 5417/6888) and structural units ('struct', 3207/3997),
+and units by UCCA categories: Scene main relations (i.e, Process and State units; 'P and S', 954/1178), Parallel Scenes ('H', 656/784),
+Participants ('A', 1348/1746),
+Centres ('C', 1904/2474), elaborators ('E', 1608/2031) and linkers ('L', 261/315).
+In both setting correlation is highest in the 'all' case, supporting our claim for the value of aggregating over a wide range of semantic phenomena.
+Some types of nodes predict the DA scores better than others.
+HUME scores on As correlate more strongly with DA than scores on Scene Main Relations (P+S).
+Center nodes (C) are also more correlated than elaborator nodes (E), which is expected given that Centers are defined to be more semantically dominant.
+Future work will construct an aggregate HUME score which weights the different node types according to their semantic relevance.
 

### 2016-06-03 paper/main.tex
minor corrections

@@
 Another common approach is to determine the similarity between the reference and translation
-by applying string edit distance \parcite{snover2006study}.
+in terms of string edits \parcite{snover2006study}.
 While these measures stimulated much progress in MT research by allowing
@@
 Intervening words are likely to change the meaning of the translation
-of $u$, and thus should be paid attention to when considering whether the translation
+of $u$, and thus should be attended to when considering whether the translation
 is correct or not.
@@
 Semantically accurate translation is paramount in this domain, which makes
-it particularly suitable for experimenting on MT evaluation. HUME is both evalauted in terms of its
+it particularly suitable for semantic  MT evaluation. HUME is both evalauted in terms of its
 consistency (inter-annotator agreement), efficiency (time of annotation) and validity (through a comparison
@@
 We see
-no reduction of IAA with sentence
-length, except perhaps in German. Also, from \tabref{tab:iaa} the overall IAA
+no discernable reduction of IAA with sentence
+length. Also, from \tabref{tab:iaa} the overall IAA
 is quite similar for all languages, showing good agreement (0.6--0.7).
@@
 results in reliable aggregate scores, that correlate very strongly with one
-other.
+another.
 
 We attempted to follow \perscite{graham2015accurate} but struggled to get enough
-judgements for our target languages. We ended up with 10 adequacy 
+crowd-sourced judgements for our target languages. We ended up with 10 adequacy 
 judgements on most of the HUME annotated translations for
@@
 Centres ('C', 1904/2474), elaborators ('E', 1608/2031) and linkers ('L', 261/315).
-In both setting correlation is highest in the 'all' case, supporting our claim for the value of aggregating over a wide range of semantic phenomena.
+In both settings correlation is highest in the 'all' case, supporting our claim for the value of aggregating over a wide range of semantic phenomena.
 Some types of nodes predict the DA scores better than others.
@@
 %
-Conversely, HMEANT does not allow to penalize errors such as tense or negation
+Conversely, HMEANT does not penalize errors such as tense or negation
 flip in a correctly aligned predicate.

### 2016-06-03 paper/main.tex
shorten a bit

@@
 To assess HUME reliability for long sentences,
-we binned the sentences according to length and measured Kappa on each of the
-bins (\figref{fig:iaalength}).
+we binned the sentences according to length and measured Kappa on each bin
+(\figref{fig:iaalength}).
 %The Kappas for each bin is presented in 
@@
 We see
-no discernable reduction of IAA with sentence
+no discernible reduction of IAA with sentence
 length. Also, from \tabref{tab:iaa} the overall IAA
-is quite similar for all languages, showing good agreement (0.6--0.7).
+is similar for all languages, showing good agreement (0.6--0.7).
 However, there are differences observed when we break down by node type.
-In particular, we see a contrast  between
+Specifically, we see a contrast  between
 Czech and Polish, where the IAA is higher for atomic than for structural units, and German and Romanian,
@@
 HUME scores on As correlate more strongly with DA than scores on Scene Main Relations (P+S).
-Center nodes (C) are also more correlated than elaborator nodes (E), which is expected given that Centers are defined to be more semantically dominant.
+Center nodes (C) are also correlate more than than elaborators (E), which is expected given that Centers are defined to be more semantically dominant.
 Future work will construct an aggregate HUME score which weights the different node types according to their semantic relevance.

### 2016-06-03 paper/main.tex
moved text to caption, fits 8 pages

--- 2016-06-03 paper/main.tex
@@
 \includegraphics[width=0.45\textwidth]{humevsDAcorrtypes-10en-dero.pdf}
-\caption{Pearson correlation of HUME vs. DA scores for different node types for both
-  the en-de and en-ro experiments. See text for the bar legend.
-%Struct refers to all structural nodes, P+S are all process and state nodes, H parallel scene nodes, A participant, C centre, E elaborators, and L linker nodes.}
-\label{fig:dacorrelationtypes}}
-\end{figure}
-
-Figure~\ref{fig:dacorrelationtypes} presents an analysis of HUME's correlations with DA by HUME unit type,
-an analysis enabled by HUME's semantic decomposition. 
-Each bar represents a correlation between HUME and an aggregate HUME score
+\caption{Pearson correlation of HUME vs. DA scores for en-ro and en-de.
+%different node types for both
+%  the en-de and en-ro experiments.
+Each bar represents a correlation between DA and an aggregate HUME score
 based on a sub-set of the units (\#nodes for the en-de/en-ro setting in brackets):
@@
 Centres ('C', 1904/2474), elaborators ('E', 1608/2031) and linkers ('L', 261/315).
-In both settings correlation is highest in the 'all' case, supporting our claim for the value of aggregating over a wide range of semantic phenomena.
+%  See text for the bar legend.
+%Struct refers to all structural nodes, P+S are all process and state nodes, H parallel scene nodes, A participant, C centre, E elaborators, and L linker nodes.}
+\label{fig:dacorrelationtypes}}
+\end{figure}
+
+Figure~\ref{fig:dacorrelationtypes} presents an analysis of HUME's correlations with DA by HUME unit type,
+an analysis enabled by HUME's semantic decomposition. 
+For both target languages, correlation is highest in the 'all' case, supporting our claim for the value of aggregating over a wide range of semantic phenomena.
 Some types of nodes predict the DA scores better than others.

### 2016-06-03 paper/main.tex
typos corrected

@@
 Concretely, let $s$ be a source sentence, $t$ a translation,
-and $A \subset 2^{s \times t}$ a many-to-many word alignment.
+and $A \subset 2^s \times 2^t$ a many-to-many word alignment.
 If $u$ is a semantic unit in $s$, whose yield is $yld(u)$, we define the aligned text in
@@
 Semantically accurate translation is paramount in this domain, which makes
-it particularly suitable for semantic  MT evaluation. HUME is both evalauted in terms of its
+it particularly suitable for semantic  MT evaluation. HUME is both evaluated in terms of its
 consistency (inter-annotator agreement), efficiency (time of annotation) and validity (through a comparison
@@
 Figure~\ref{fig:dacorrelation} plots the HUME score for each sentence against
-its DA score. HUME and Direct Assesment scores correlate reasonably well. The
+its DA score. HUME and Direct Assessment scores correlate reasonably well. The
 Pearson correlation for en-ro (en-de) is 0.70 (0.58), or 0.78 (0.74) if only
@@
 the quality of translations and it only requires very few annotators.
-Direct assesment returns a single opaque score, and (as also noted by
-Graham et al.), and requires a large crowd which may not be available or reliable.
+Direct assessment returns a single opaque score, and (as also noted by
+Graham et al.) requires a large crowd which may not be available or reliable.
 
@@
 and that annotation quality is robust to sentence length.
+Comparison to direct assessments further support HUME's validity.
 We believe that

### 2016-06-30 paper/main.tex
typos
Added authors and changed styles for upload to Arxiv

@@
 %\usepackage[round]{natbib}
-\usepackage{emnlp2016}
+\usepackage{acl2016}
 \usepackage{amsmath}
@@
 \usepackage[font={small}]{caption}
+\aclfinalcopy
 
@@
 
+\newcommand*{\affaddr}[1]{#1} 
+\newcommand*{\affmark}[1][*]{\textsuperscript{#1}}
+\newcommand*{\email}[1]{\texttt{#1}}
+
 \title{HUME: Human UCCA-Based Evaluation of Machine Translation}
 
-\author{Author 1\\
-	    XYZ Company\\
-	    111 Anywhere Street\\
-	    Mytown, NY 10000, USA\\
-	    {\tt author1@xyz.org}
-	  \And
-	Author 2\\
-  	ABC University\\
-  	900 Main Street\\
-  	Ourcity, PQ, Canada A1A 1T2\\
-  {\tt author2@abc.ca}}
+\author{
+Alexandra Birch\textsuperscript{1}, Barry Haddow\textsuperscript{1}, Ond{\v{r}}ej Bojar\textsuperscript{2}, Omri Abend\textsuperscript{3}\\
+\textsuperscript{1}School of Informatics, University of Edinburgh\\
+\textsuperscript{2}Institute of Formal and Applied Linguistics, Charles University in Prague\\
+\textsuperscript{3}School of Computer Science and Engineering, Hebrew University of Jerusalem\\
+\texttt{a.birch@ed.ac.uk, bhaddow@inf.ed.ac.uk}\\
+\texttt{bojar@ufal.mff.cuni.cz, oabend@cs.huji.ac.il}\\
+}
 
@@
 
-Human evaluation of machine translation normally uses sentence-level measures such as relative ranking or adequacy scales. However these provide no insight into possible errors,
+Human evaluation of machine translation normally uses sentence-level measures such as relative ranking or adequacy scales. However, these provide no insight into possible errors,
 and do not scale well with sentence length.

### 2016-06-30 paper/main.tex
Corrected Charles uni and acknowledgements

@@
 
-\newcommand*{\affaddr}[1]{#1} 
-\newcommand*{\affmark}[1][*]{\textsuperscript{#1}}
-\newcommand*{\email}[1]{\texttt{#1}}
+\newcommand{\GrantNo}{644402}
+\newcommand{\ProjectName}{HimL}
+\newcommand{\ProjectType}{Horizon 2020 Innovation Action}
 
@@
 \textsuperscript{1}School of Informatics, University of Edinburgh\\
-\textsuperscript{2}Institute of Formal and Applied Linguistics, Charles University in Prague\\
+\textsuperscript{2} Charles University in Prague, Faculty of Mathematics and Physics\\
 \textsuperscript{3}School of Computer Science and Engineering, Hebrew University of Jerusalem\\
@@
 
-%\section*{Acknowledgments}
+\section*{Acknowledgments}
+{T}his work was conducted within the scope of the \ProjectType, which has received funding from the European Union’s Horizon 2020 research and innovation programme under grant agreement No \GrantNo (\ProjectName).
 

### 2016-09-11 paper/main.tex
for camera ready

--- 2016-06-03 paper/main.tex
@@
 \usepackage{graphicx}
+\usepackage[T1]{fontenc}
 \usepackage{url}
@@
 \usepackage{titlesec}
-%\usepackage{hyperref}
 \usepackage{color}
@@
 \usepackage[font={small}]{caption}
+%\usepackage{hyperref}
+
+\aclfinalcopy
 
 \def\emnlppaperid{***}
-\setlength{\belowcaptionskip}{-10pt}
+%\setlength{\belowcaptionskip}{-10pt}
 
@@
 
-\titlespacing{\section}{2pt}{*0}{*0}
-\titlespacing{\subsection}{2pt}{*0}{*0}
-\titlespacing{\subsubsection}{2pt}{*0}{*0}
-\titlespacing{\paragraph}{1pt}{*0}{3pt}
+%\titlespacing{\section}{2pt}{*0}{*0}
+%\titlespacing{\subsection}{2pt}{*0}{*0}
+%\titlespacing{\subsubsection}{2pt}{*0}{*0}
+%\titlespacing{\paragraph}{1pt}{*0}{3pt}
 
@@
 %\newcommand{\XXX}[1]{}
-%\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
-\newcommand{\oa}[1]{}
+\newcommand{\oa}[1]{\footnote{\color{red}OA: #1}}
+\newcommand{\oamod}[1]{{\color{red}#1}}
+
+%\newcommand{\oa}[1]{}
 \newcommand{\bh}[1]{\footnote{\color{blue}BH: #1}}
@@
 
+\newcommand{\GrantNo}{644402}
+\newcommand{\ProjectName}{HimL}
+\newcommand{\ProjectType}{Horizon 2020 Innovation Action}
+
 \title{HUME: Human UCCA-Based Evaluation of Machine Translation}
 
-\author{Author 1\\
-	    XYZ Company\\
-	    111 Anywhere Street\\
-	    Mytown, NY 10000, USA\\
-	    {\tt author1@xyz.org}
-	  \And
-	Author 2\\
-  	ABC University\\
-  	900 Main Street\\
-  	Ourcity, PQ, Canada A1A 1T2\\
-  {\tt author2@abc.ca}}
+\author{
+  Alexandra Birch\textsuperscript{1}, Omri Abend\textsuperscript{2}, Ond{\v{r}}ej Bojar\textsuperscript{3},
+  Barry Haddow\textsuperscript{1}\\
+\textsuperscript{1}School of Informatics, University of Edinburgh\\
+\textsuperscript{2}School of Computer Science and Engineering, Hebrew University of Jerusalem\\
+\textsuperscript{3} Charles University in Prague, Faculty of Mathematics and Physics\\
+\texttt{a.birch@ed.ac.uk, oabend@cs.huji.ac.il}\\
+\texttt{bojar@ufal.mff.cuni.cz, bhaddow@inf.ed.ac.uk}\\
+}
 
@@
 
-Human evaluation of machine translation normally uses sentence-level measures such as relative ranking or adequacy scales. However these provide no insight into possible errors,
+Human evaluation of machine translation normally uses sentence-level measures such as relative ranking or adequacy scales. However, these provide no insight into possible errors,
 and do not scale well with sentence length.
 We argue for a semantics-based evaluation, which captures what meaning components
-are retained in the MT output, providing a more fine-grained analysis of
-translation quality, and enables the construction and tuning of semantics-based MT. 
+are retained in the MT output, thus providing a more fine-grained analysis of
+translation quality, and enabling the construction and tuning of semantics-based MT. 
 We present a novel human semantic evaluation measure, Human
@@
 These problems are partially addressed by measures that decompose over parts of the evaluated
-translation. For automatic measures,
-these are often words or n-grams, for manual measures some structural
-information is taken into account \parcite{machacek:bojar:segranks:2015}, or the
-annotators are explicitly asked to mark errors, which however suffers from even
-lower agreement than ranking \parcite{lommel:etal:mqm-iaa:2014}.
+translation, often words or n-grams (see \secref{sec:background} for a brief survey of previous
+work). A promising line of research decomposes metrics
+over semantically defined units, quantifying the similarity of the output and the
+reference in terms of their verb argument structure; the most notable of these measures is HMEANT
+\parcite{lo2011structured}.
+
 % the following is specific to automatic measures:
 %quantifying the overlap of its sub-parts relative a reference translation.
-A promising line of research decomposes metrics
-over semantically defined units,
-quantifying the similarity of the output and the reference in terms of
-their verb argument structure; the most notable of these measures is HMEANT
-\parcite{lo2011structured}.
 %More recent work on MT evaluation proposed measures
@@
 semantic units in the source sentence. We 
-%HUME only requires semantically annotating the source sentence,
 are thus avoiding the semantic annotation of machine-generated text,
@@
 measuring the quality of different translations of the same source sentence,
-and 
+and avoids relying on possibly suboptimal reference translations.
+
 %By basing the measure on a direct comparison
 %of the source and translation we avoid the difficulties created by using
-avoids relying on possibly suboptimal reference translations.
 %albeit at the cost of requiring the
@@
 %are necessarily incomplete, as they compare to a small number of references
-%out of the the huge number of possible translations a sentence may have
+%out of the huge number of possible translations a sentence may have
 %\parcite{dreyer2012hyter}.
@@
 by \perscite{graham2015accurate}.
+Some manual measures ask annotators to explicitly mark errors, which however
+suffers from even lower agreement than ranking \parcite{lommel:etal:mqm-iaa:2014}.
+
 However, while providing the gold standard for MT evaluation, human evaluation
 is not a scalable solution.
-%and is infeasible where a wide variety of systems or parameter settings are to be evaluated.
-% has motivated the construction of cheaper methods that correlate with human
-%judgement.
-
 Scalability is addressed by employing automatic and semi-automatic approximations of human
 judgements. Commonly, such scores decompose over the sub-parts of
-the translation, and quantify how many of these sub-parts appear in a manually created reference translation.
-This decomposition allows system developers
-to localize the errors.
-%, or at least unconfirmed parts within the translation.
+the translation, and quantify how many of these sub-parts appear in a manually
+created reference translation.
+This decomposition allows system developers to localize the errors.
 The most commonly used measures decompose over n-grams or individual words, e.g., 
-BLEU \parcite{Papineni:2002}, NIST \parcite{Doddington:2002} and METEOR
-\parcite{Banerjee:2005}.
+BLEU \parcite{Papineni:2002}, NIST \parcite{Doddington:2002} and METEOR \parcite{Banerjee:2005}.
 Another common approach is to determine the similarity between the reference and translation
@@
 MT models.
+
+%and is infeasible where a wide variety of systems or parameter settings are to be evaluated.
+% has motivated the construction of cheaper methods that correlate with human
+%judgement.
 
@@
 semantic (predicate-argument structures), lexical and morphosyntactic features.
+\perscite{machacek:bojar:segranks:2015} divided the source sentences into shorter segments,
+defined using a phrase structure parse, and applied human ranking to the resulting segments.\oa{is this description OK?}
 
@@
 UCCA (Universal Conceptual Cognitive Annotation, \inparcite{abend2013universal}) is a
-cross-linguistically applicable, lightweight
-scheme for semantic annotation. Formally, an UCCA structure is a directed acyclic graph (DAG),
+cross-linguistically applicable scheme for semantic annotation.
+Formally, an UCCA structure is a directed acyclic graph (DAG),
 whose leaves correspond to the words of the text.
@@
 
-UCCA's current inventory of distinctions focuses on argument structures
-(adjectival, nominal, verbal and others) and relations between them.
-The most basic notion is the Scene, which describes a movement, an
+UCCA's basic inventory of distinctions (its {\it foundational layer})
+focuses on argument structures (adjectival, nominal, verbal and others)
+and relations between them. The most basic notion is the Scene, which describes a movement, an
 action or a state which persists in time. Each Scene contains one main
@@
 The use of UCCA for semantic MT evaluation measure is motivated by two main reasons. 
-First, UCCA's set of categories can
+First, UCCA's foundational layer can
 be reliably annotated by non-experts after as little as two hours
@@
 English, French, German and Czech.
-%Third, the scheme has been shown to be stable across translations:
-%UCCA annotations of translated text usually contain the same set of relations
-%\parcite{sulem2015conceptual}, indicating that the scheme reflects
-%a common layer of representation that in a correct translation
-%is mostly shared between the translation and the source.
+Third, the scheme has been shown to be stable across translations:
+UCCA annotations of translated text usually contain the same set of relations
+\parcite{sulem2015conceptual}, indicating that UCCA reflects a layer of
+representation that in a correct translation is mostly shared between
+the translation and the source.
 
@@
 However, using AMR is not optimal for defining a decomposition of a sentence into semantic
-units as it does not ground its semantic symbols in the text,
+units as it does not anchor its semantic symbols in the text,
 and thus does not provide clear decomposition of the sentence into sub-units.
@@
 \parcite{sgallhp:1986,hajic2012announcing,oepen2006discriminant},
-which are indeed grounded in the text, but are less suitable for MT evaluation
+which are indeed anchored in the text, but are less suitable for MT evaluation
 as they require linguistic expertise for their annotation.
@@
 
-\vspace{-.1em}
 \section{The HUME Measure}\label{sec:hume}
-\vspace{-.2em}
 
@@
 The procedure involves two manual steps: (1) UCCA-annotating $s$, 
-(2) human judgements as to the translation quality of each semantic unit of $s$ relative to $t$,
+(2) HUME-annotation: human judgements as to the translation quality of each semantic
+unit of $s$ relative to $t$,
 where units are defined according to the UCCA annotation.
 UCCA annotation is performed once for every source sentence,
-irrespective of the number of its translations we wish
-to evaluate
-%\footnote{Instead of relying on the source,
-%the evaluation could be based on the UCCA structure for the \emph{reference}
-%translation.
-%More details in the discussion in \secref{src-vs-ref}.
-% This would allow monolingual annotators to take part in the
-% evaluation. On the downside, the translation would have to be aligned to the
-% reference and more importantly, we would risk inaccurate judgements due to
-% the naturally occurring differences between the source and its reference
-% translations.
-%}.
+irrespective of the number of its translations we wish to evaluate,
+and requires proficiency in the source language only.
+HUME annotation requires the employment of bilingual annotators
+\footnote{Where bilingual annotators are not available,
+  the evaluation could be based on the UCCA structure for the
+  \emph{reference} translation. See discussion in \secref{sec:hmeant_comp}.}
 
@@
 a unit is {\it structural}, i.e., has meaning-bearing sub-units also in the
-target language,
-or {\it atomic}. In most cases, atomic units
-correspond to individual words, but they may also correspond to unanalyzable
-multi-word expressions.
+target language, or {\it atomic}.
+In most cases, atomic units
+correspond to individual words, but they may also correspond to
+multi-word expressions that translate as one unit. \oamod{For instance,
+the expression ``took a shower'' is translated into the German ``duschte'',
+while its individual words do not correspond to any sub-part of the German translation,
+motivating the labeling the entire expression as an atomic node.}
 When a multi-word unit is labeled as atomic, its sub-units' annotations are ignored
@@
 
-Atomic units can be labelled as ``Green'' (correct), ``Orange'' (partially correct)
-and ``Red'' (incorrect). 
+Atomic units can be labelled as ``Green'' (G, correct), ``Orange'' (O, partially correct)
+and ``Red'' (R, incorrect). 
 Green means that the meaning of the word or phrase has been largely preserved.
@@
 but some part of the translation is wrong.
-This is often be due to the translated word having the the wrong inflection,
+This is often be due to the translated word having the wrong inflection,
 in a way that impacts little on the understandability of the sentence.
@@
 %Scene elements, such as the predicate (main relation), participants and adverbials.}
-Structural units are labeled as ``Adequate'' or ``Bad'', meaning
+Structural units are labeled as ``Adequate'' (A) or ``Bad'' (B), meaning
 that the relation between the sub-units went wrong\footnote{
@@
      HUME annotation of an UCCA tree with a word aligned example translation shown below. 
-Atomic units are labelled using traffic lights (Red, Orange, Green) and structural units are marked A or B.}
+     Atomic units are labelled using traffic lights (Red, Orange, Green) and structural
+     units are marked A or B.}
 \end{figure}
@@
 
-\vspace{-.5cm}
+\vspace{-.3cm}
+
 {\scriptsize
@@
 }
+
 \vspace{-.5cm}
@@
 
-\figref{fig:interface} shows the HUME annotation interface.
+\figref{fig:interface} shows the HUME annotation interface\footnote{A demo of HUME
+  can be found in \url{www.cs.huji.ac.il/~oabend/hume_demo.html}}.
 % We now describe the HUME annotation interface, shown in \figref{fig:interface}. 
@@
 % %Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
+\oamod{One source sentence and one translation are presented at a time.}
 The user is asked to select a label for each source semantic unit,
@@
 their instances, to avoid double counting.
-%Note that the UCCA labels are not shown.
 
@@
 seems acceptable but the interleaving words in red notify the annotator to check
-the whole translation, in which the meaning of the expression is not preserved.
+the whole translation, in which the meaning of the expression is not
+preserved\footnote{The interleaving words are ``... und beide berichtet berichteten ...'' (lit.
+  ``... and both report reported ...''), which doesn't form any coherent relation with the rest
+  of the sentence.}.
 The annotator should thus mark this structural node as Bad.
@@
 NHS~24 is a public body providing healthcare and health-service
-related information in Scotland, Cochrane is an international NGO
+related information in Scotland; Cochrane is an international NGO
 which provides independent systematic reviews on health-related research.
@@
 based on a sub-set of the units (\#nodes for the en-de/en-ro setting in brackets):
-all units ('all', 8624/10885), atomic ('atomic', 5417/6888) and structural units ('struct', 3207/3997),
-and units by UCCA categories: Scene main relations (i.e, Process and State units; 'P and S', 954/1178), Parallel Scenes ('H', 656/784),
-Participants ('A', 1348/1746),
-Centres ('C', 1904/2474), elaborators ('E', 1608/2031) and linkers ('L', 261/315).
+all units (``all'', 8624/10885), atomic (``atomic'', 5417/6888)
+and structural units (``struct'', 3207/3997),
+and units by UCCA categories: Scene main relations
+(i.e, Process and State units; ``P and S'', 954/1178), Parallel Scenes (``H'', 656/784),
+Participants (``A'', 1348/1746),
+Centres (``C'', 1904/2474), elaborators (``E'', 1608/2031) and linkers (``L'', 261/315).
 %  See text for the bar legend.
@@
 
+\oa{This entire section needs reviewing}
+
 We discuss the main differences between HUME and HMEANT,
 a human MT evaluation metric that measures the overlap between the translation a reference
-in terms of their SRL annotations.
+in terms of their SRL annotations. The discussion is acoompanied by examples from our
+experimental data.
 
@@
 
-\vspace{-.3em}
 \begin{center}
@@
 \end{center}
-\vspace{-.5em}
 
@@
 
-The German translation 
-is largely correct, except that the main verb ``sein'' (``be'') is omitted.
+The German translation is largely correct, except that the main verb ``sein'' (``be'') is omitted.
 While this may be interpreted as a minor error, HMEANT will assign the
 sentence a very low score, as it failed to translate the main verb.
-%
-Conversely, HMEANT does not penalize errors such as tense or negation
-flip in a correctly aligned predicate.
-
-%This focus on verbal predicates has a number of unfortunate consequences. We show an example below:
-
-% \noindent
-% \begin{inparadesc}
-% \item[Source ] a coronary angioplasty may not be technically possible \\
-% \item[Transl.] eine koronare Angioplastie kann nicht technisch m{\"o}glich \\
-% \item[Gloss ]  a coronary angioplasty can not technically possible \\
-% \end{inparadesc}
-
-\com{
-}
+
+Conversely, HMEANT does not penalize errors (e.g., partial semantic compatability,
+or negation flip) in a correctly aligned predicate:
+
+\begin{center}
+\begin{tabular}{lp{5.4cm}}
+Source & \small ... don't skip meals ... \\
+Transl.& \small ... nicht {\"u}berspringen Mahlzeiten ... \\
+Gloss & \small don't skip meals ... \\
+\end{tabular}
+\end{center}
+
+While the translation is mostly understandable, the verb {\"u}berspringen is incorrect here
+(should be ``auslassen''). HUME allows marking the main verb as Orange, or even red, while
+giving partial score for the translation of ``don't'' and of ``meals''. HMEANT may only
+mark the predicate as correct, or mark the entire argument structure as an error.
+
+It is also relatively common that verbal constructions are trasnlated
+as non-verbal ones or vice versa. Consider the following example:
+
+\begin{center}
+\begin{tabular}{lp{5.4cm}}
+Source & \small ... tend to be higher in saturated fats \\
+Transl.& \small in der Regel h{\"o}her in ges{\"a}ttigte Fette \\
+Gloss & \small as a rule higher in saturated fats \\
+\end{tabular}
+\end{center}
+
+The German translation is largely correct, although the English verb ``tend''
+was translated into the German prepositional phrase ``in der Regel'' (``as a rule'').
+HMEANT will consider the translation to be of poor quality as there is no German verb
+to align the English one with.
 
@@
 Non-auxiliary instances of ``be'' and nouns are identified using the NLTK standard tagger.
-Nominal argument structures are here Scenes whose main relation is headed by a noun.}
-First, copula clauses are treated in HMEANT simply as instances of the main verb ``be'', which generally does not convey the meaning of these clauses.
-They appear in 21.7\% of the sentences, according to conservative estimates that only consider non-auxiliary
-instances of ``be''.
-Second, nominal argument structures, ignored by HMEANT, are in fact highly
-pervasive, appearing in 48.7\% of the sentences.
-Third, linkers that express inter-relations between
-clauses (mainly discourse markers and conjunctions) appear in 56\% of the
-sentences, but are again ignored by HMEANT. As noted in our experiments, linkers
-are sometimes omitted in translation, but these omissions are not taken
-into consideration by HMEANT.
-
-%An example where a linker is omitted in translation in our experiments is
-%\equo{However, this review\dots}, which was translated to
-%the German \equo{Diese \"Uberpr\"ufung\dots} (lit. \equo{This review}).
-%This omission will not be taken into account by HMEANT at all
-
-We are not aware of any empirical argument suggesting that verb argument structures,
-taken alone, capture the crux of the sentence semantics.
-Moreover, relying only on verbal argument structures is less stable across paraphrases
-and translations, as a verbal argument structure
-may be translated to a nominal or adjectival argument structure
-(e.g., ``after graduation'' may be translated into ``after he graduated'').
-This may lead to an unduly low HMEANT score, as the verb
-in one structure has nothing to align to in the other.
-On the other hand, UCCA has been shown to be reasonably stable in an English-French
-corpus study \cite{sulem2015conceptual}.
-
-% No space
-%It is also relatively common in translation that verbal constructions get
-%nominalized or vice versa. An example from our dataset:
-%  ``Two {\bf review} authors assessed the studies independently\dots{}'' --
-%  ``Zwei Autoren bewertete die Studien unabh\"angig zu {\bf
-%  \"uberpr\"ufen}\dots{}'' (lit. ``Two authors assessed the studies
-%  independently to review\dots").\XXX{This is Omri's example, with
-%  Ondrej's gloss for German. Ondrej is
-%  not totally sure that the meaning is well preserved. So Ondrej would actually
-%  *remove* this example, because it would illustrate HMEANT's problems only if
-%  the meaning was correct.}
+Nominal argument structures are here Scenes whose main relation is headed by a noun.} First, copula clauses are treated in HMEANT simply as instances of the main verb ``be'', which generally does not convey the meaning of these clauses.
+They appear in 21.7\% of the sentences, according to conservative estimates that only consider non-auxiliary instances of ``be''.
+Second, nominal argument structures, ignored by HMEANT, are in fact highly pervasive, appearing in 48.7\% of the sentences.
+Third, linkers that express inter-relations between clauses (mainly discourse markers and conjunctions) appear in 56\% of the sentences, but are again ignored by HMEANT.
+For instance, linkers are sometimes omitted in translation, but these omissions are not penalized by HMEANT.
+The following is such an example from our experimental dataset:
+
+\begin{center}
+\begin{tabular}{lp{5.4cm}}
+Source & \small However, this review was restricted to ... \\
+Transl.& \small Diese {\"U}berpr{\"u}fung bescr{\"a}nkte sich auf ... \\
+Gloss & \small This review was restricted to ... \\
+\end{tabular}
+\end{center}
+
+%We are not aware of any empirical argument suggesting that verb argument structures,
+%taken alone, capture the crux of the sentence semantics.
+%Moreover, as the above example shows, a verbal argument structure
+%may be translated to a nominal or adjectival argument structure,
+%leading to an unduly low HMEANT score.
+%On the other hand, UCCA has been shown to be reasonably stable in an English-French
+%corpus study \cite{sulem2015conceptual}.
 
@@
 for the translation.
-%First of all, the annotation process is different. We semantically annotate the
-%source, then automatically project it to the translation, and ask bilingual
-%annotators to assess whether this projection shows a good or bad translation. In
-%HMEANT, they semantically annotate both the reference and the translation, then
-%align these manually, and assess to what extent they match.
-%We see this difference as crucial where semantic analysis is concerned, as it
 Not annotating the translation is appealing as it is often impossible to assign a
@@
 interpretation of a PP-attachment, or the verb having an additional modifier in
-%suggested that sometimes, the structures of the reference and candidate
-%translations diverge without any good reason, for instance due to a different
-%interpretation of a PP-attachment, the verb can have one additional modifier in
 one of the structures. It would be desirable to allow modifications to
@@
 to avoid unduly penalizing such spurious divergences.
-%annotations also at the alignment stage
-%to avoid unduly penalization due to such a spurious divergence.
+
 The same issue is noted by \perscite{lo:wu:reliability:2014}:
@@
 lead to unjustly low scores where the translation is a paraphrase of the reference.
-
-%If only monolingual annotators are available, the HUME evaluation could be performed
-%with a reference sentence instead of with the source. 
+If only monolingual annotators are available, the HUME evaluation could be performed
+with a reference sentence instead of with the source. This, however,
+would risk inaccurate judgements due to the naturally occurring differences
+between the source and its reference translations.
+
 %\bh{But this would be a different metric}
@@
 Comparison to direct assessments further support HUME's validity.
-We believe that
-HUME, and a future automated version of HUME, allows for a more fine-grained analysis of translation quality,
-and will be to guide the development of a more semantically aware approach to MT.
+We believe that HUME, and a future automated version of HUME,
+allows for a finer-grained analysis of translation quality,
+and will be useful in informing the development of a more semantically aware
+approach to MT.
 
@@
 translation provided by the system. These word
-alignments are used to project the UCCA annotation from the source sentence to the the translation output, and then bilingual
+alignments are used to project the UCCA annotation from the source sentence to the translation output, and then bilingual
 annotators go through each projected UCCA node, assessing how well it is translated.  
@@
 
-%\section*{Acknowledgments}
+\section*{Acknowledgments}
+{T}his work was conducted within the scope of the \ProjectType, which has received funding from the European Union's Horizon 2020 research and innovation programme under grant agreement No \GrantNo (\ProjectName).
 

### 2016-09-11 paper/main.tex
for camera ready

@@
 Transl.& \small ... nicht {\"u}berspringen Mahlzeiten ... \\
-Gloss & \small don't skip meals ... \\
+Gloss & \small ... don't skip meals ... \\
 \end{tabular}
@@
 Source & \small ... tend to be higher in saturated fats \\
-Transl.& \small in der Regel h{\"o}her in ges{\"a}ttigte Fette \\
-Gloss & \small as a rule higher in saturated fats \\
+Transl.& \small ... in der Regel h{\"o}her in ges{\"a}ttigte Fette \\
+Gloss & \small ... as a rule higher in saturated fats \\
 \end{tabular}

### 2016-09-12 sample_annot.py
printing graphic

@@ -1,6 +1,7 @@
 #!/usr/bin/env python3
 from xml.etree.ElementTree import ElementTree, tostring, fromstring
 from xml.dom import minidom
+import pprint
 
 from ucca.convert import *
 
@@ -15,6 +16,7 @@ def main():
   align = dfh.readline()[:-1]
   annot = dfh.readline()[2:-2]
   
+  pprint.pprint (annot)
   elem = fromstring(annot)
   #xmlstr = minidom.parseString(ET.tostring(elem)).toprettyxml(indent="   ")
   #print(xmlstr)

### 2016-09-12 paper/main.tex

@@
 
-Conversely, HMEANT does not penalize errors (e.g., partial semantic compatability,
-or negation flip) in a correctly aligned predicate:
+Conversely, HMEANT does not penalize errors (e.g., incorrect tense or negation flip) in a correctly aligned predicate:
 
@@
 
-While the translation is mostly understandable, the verb {\"u}berspringen is incorrect here
-(should be ``auslassen''). HUME allows marking the main verb as Orange, or even red, while
-giving partial score for the translation of ``don't'' and of ``meals''. HMEANT may only
-mark the predicate as correct, or mark the entire argument structure as an error.
+While the translation is mostly understandable, the verb {\"u}berspringen is incorrectly inflected. HUME allows marking the main verb as Orange, while
+marking the translation of the argument (``Malhzeiten'' or ``meals'') as correct. HMEANT may only mark the predicate as correct, or mark the entire argument structure as an error.
 

### 2016-09-13 paper/main.tex
My changes for camera ready

@@
 This also allows the re-use of the source semantic annotation for
-measuring the quality of different translations of the same source sentence,
-and avoids relying on possibly suboptimal reference translations.
+measuring the quality of different translations of the same source sentence
+and avoids relying on reference translations, 
+which have been shown to bias annotators~\cite{fomicheva-specia_ACL:2016}.
 
@@
 adequacy/fluency scores to each translation, a procedure recently improved
-by \perscite{graham2015accurate}.
+by \perscite{graham2015accurate} and called direct assessment. We use this method to compare and contrast
+with HUME later in the paper. HTER~\parcite{snover2006study} is another widely used human evaluation metric
+which uses edit distance metrics to compare a translation and its human 
+post-edition. HTER suffers from the problem that small edits in the translation
+could in fact be serious flaws in accuracy, eg. deleting a negation. 
 Some manual measures ask annotators to explicitly mark errors, which however
@@
 \perscite{machacek:bojar:segranks:2015} divided the source sentences into shorter segments,
-defined using a phrase structure parse, and applied human ranking to the resulting segments.\oa{is this description OK?}
+defined using a phrase structure parse, and applied human ranking to the resulting segments.
 
@@
     \begin{center}
-    \includegraphics[width=0.35\textwidth]{ucca-tree-v2.png}
+    \includegraphics[width=0.35\textwidth]{ucca-tree-v3.png}
 
@@
   Sample UCCA annotation. Leaves correspond to words and nodes to units.
-  The dashed edge indicates that ``Tom'' is also a participant in the ``moved to NYC''
+  The dashed edge indicates that ``Tom'' is also a participant in the ``moved to America''
   Scene. Edge labels mark UCCA categories.}
@@
 action or a state which persists in time. Each Scene contains one main
-relation and zero or more participants. For example, the sentence ``After graduation, Tom moved to NYC''
+relation and zero or more participants. For example, the sentence ``After graduation, Tom moved to America''
 contains two Scenes, whose main relations are ``graduation'' and ``moved''.
-The participant ``Tom'' is a part of both Scenes, while ``NYC'' only of the
+The participant ``Tom'' is a part of both Scenes, while ``America'' only of the
 latter (\figref{fig:ucca_example_v2}). Further categories account for
@@
 %argument structure, and is thus less suitable as a basis for a semantic measure.
-%E.g., ``after he graduated, Tom moved to NYC'' contains two verbs,
+%E.g., ``after he graduated, Tom moved to America'' contains two verbs,
 %but is similar in meaning to the one-verbed
-%``after graduation, Tom moved to NYC''.
+%``after graduation, Tom moved to America''.
 
@@
 correspond to individual words, but they may also correspond to
-multi-word expressions that translate as one unit. \oamod{For instance,
+multi-word expressions that translate as one unit. For instance,
 the expression ``took a shower'' is translated into the German ``duschte'',
 while its individual words do not correspond to any sub-part of the German translation,
-motivating the labeling the entire expression as an atomic node.}
+motivating the labeling the entire expression as an atomic node.
 When a multi-word unit is labeled as atomic, its sub-units' annotations are ignored
@@
 %For example, consider the sentence ``After graduation, Tom moved
-%to NYC'' as the source $s$, see \figref{fig:ucca_example}.
+%to America'' as the source $s$, see \figref{fig:ucca_example}.
 %The atomic units are the individual words,
@@
 %Then ``after'', ``graduation'', ``Tom'' and ``to'' are marked Green, ``moved'' is marked
-%Orange (as it corresponds to ``will move'' in $t$), and ``NYC''
+%Orange (as it corresponds to ``will move'' in $t$), and ``America''
 %is marked as Red, as it was mistranslated to ``Paris''. Node \#9 is marked Adequate,
@@
     \begin{center}
-    \includegraphics[width=0.35\textwidth]{ucca-tree-mteval-v2.png}
+    \includegraphics[width=0.35\textwidth]{ucca-tree-mteval-v3.png}
     \end{center}
@@
 annotation, where the translation is in English for ease of comprehension.
-When evaluating ``to NYC'' the annotator looks at the translation and sees the
+When evaluating ``to America'' the annotator looks at the translation and sees the
 word ``stateside''. This word captures the whole phrase and so we mark this
@@
 % %Information on the category of the unit (e.g., whether it is a Scene, participant etc.).
-\oamod{One source sentence and one translation are presented at a time.}
+One source sentence and one translation are presented at a time.
 The user is asked to select a label for each source semantic unit,
@@
 The annotators were all native speakers of the respective target languages and fluent in English.
-\oa{maybe we should put in what training they've received}
+They completed a three hour on-line training session where we described UCCA and the HUME task and then
+we worked through a number of examples.
 
@@
 is generally little confusion between the two types of units.
+This results in the Kappa for all units being considerably higher than the kappa
+over the atomic units or structural units, where there is more internal confusion.
 
@@
 
-\def\iaafig #1{\includegraphics[width=3.8cm]{iaa_heatmap_#1.png}}
+\def\iaafig #1{\includegraphics[width=4cm]{iaa_heatmap_#1.png}}
 
@@
 
+We have just presented a quantitative comparison between DA and HUME. DA is however, a very
+different type of metric to HUME. It returns just one score per sentence and requires annotators to make 
+decisions collating many smaller judgements about individual errors in the sentence. 
+Comparing DA and HUME is important in order to see where they agree and differ in order to
+determine under what circumstances it would be appropriate to apply each method. 
+In cases where we specifically want to know about the quality of semantic
+components in a translation, HUME and HMEANT could be considered.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\oa{This entire section needs reviewing}
-
-We discuss the main differences between HUME and HMEANT,
-a human MT evaluation metric that measures the overlap between the translation a reference
-in terms of their SRL annotations. The discussion is acoompanied by examples from our
+HMEANT is a human MT evaluation metric that measures the overlap between the translation a reference
+in terms of their SRL annotations.
+In this section we present a qualitative comparison between HUME and HMEANT
+and the discussion is accompanied by examples from our
 experimental data.
@@
 
-Conversely, HMEANT does not penalize errors (e.g., incorrect tense or negation flip) in a correctly aligned predicate:
-
-\begin{center}
-\begin{tabular}{lp{5.4cm}}
-Source & \small ... don't skip meals ... \\
-Transl.& \small ... nicht {\"u}berspringen Mahlzeiten ... \\
-Gloss & \small ... don't skip meals ... \\
-\end{tabular}
-\end{center}
-
-While the translation is mostly understandable, the verb {\"u}berspringen is incorrectly inflected. HUME allows marking the main verb as Orange, while
-marking the translation of the argument (``Malhzeiten'' or ``meals'') as correct. HMEANT may only mark the predicate as correct, or mark the entire argument structure as an error.
-
-It is also relatively common that verbal constructions are trasnlated
+%This is not entirely true - they do allow an alignment with partial
+%Conversely, HMEANT does not penalize errors (e.g., incorrect tense or negation flip) in a correctly aligned predicate:
+%
+%\begin{center}
+%\begin{tabular}{lp{5.4cm}}
+%Source & \small ... don't skip meals ... \\
+%Transl.& \small ... nicht {\"u}berspringen Mahlzeiten ... \\
+%Gloss & \small ... don't skip meals ... \\
+%\end{tabular}
+%\end{center}
+%While the translation is mostly understandable, the verb {\"u}berspringen is incorrectly inflected. HUME allows marking the %main verb as Orange, while
+%marking the translation of the argument (``Malhzeiten'' or ``meals'') as correct. HMEANT may only mark the predicate as %correct, or mark the entire argument structure as an error.
+
+It is also relatively common that verbal constructions are translated
 as non-verbal ones or vice versa. Consider the following example:

### 2016-09-14 paper/main.tex
All authors equally

@@
   Alexandra Birch\textsuperscript{1}, Omri Abend\textsuperscript{2}, Ond{\v{r}}ej Bojar\textsuperscript{3},
-  Barry Haddow\textsuperscript{1}\\
+  Barry Haddow\textsuperscript{1*}\\
 \textsuperscript{1}School of Informatics, University of Edinburgh\\
@@
 \section{Introduction}\label{sec:intro}
+
+\let\thefootnote\relax\footnotetext{*All authors contributed equally to this work.}
 

### 2016-09-15 paper/main.tex
hmeant comparison

@@
 
-%This is not entirely true - they do allow an alignment with partial
-%Conversely, HMEANT does not penalize errors (e.g., incorrect tense or negation flip) in a correctly aligned predicate:
-%
-%\begin{center}
-%\begin{tabular}{lp{5.4cm}}
-%Source & \small ... don't skip meals ... \\
-%Transl.& \small ... nicht {\"u}berspringen Mahlzeiten ... \\
-%Gloss & \small ... don't skip meals ... \\
-%\end{tabular}
-%\end{center}
-%While the translation is mostly understandable, the verb {\"u}berspringen is incorrectly inflected. HUME allows marking the %main verb as Orange, while
-%marking the translation of the argument (``Malhzeiten'' or ``meals'') as correct. HMEANT may only mark the predicate as %correct, or mark the entire argument structure as an error.
-
 It is also relatively common that verbal constructions are translated
@@
 
-We conducted an analysis of the English UCCA Wikipedia corpus (5324 sentences) in order to assess the pervasiveness of three phenomena that are not well supported by HMEANT.\footnote{Argument structures and linkers are explicitly marked in UCCA.
-Non-auxiliary instances of ``be'' and nouns are identified using the NLTK standard tagger.
-Nominal argument structures are here Scenes whose main relation is headed by a noun.} First, copula clauses are treated in HMEANT simply as instances of the main verb ``be'', which generally does not convey the meaning of these clauses.
+In other cases, HMEANT's does not to penalize errors. For instance, as HMEANT is defined as an aggregate
+of the scores given to the slot fillers, it cannot assign partial scores to the predicates themselves in
+cases of a divergent tense or a negation flip. Consider the following:
+
+\begin{center}
+\begin{tabular}{lp{5.4cm}}
+Source & \small ... don't skip meals ... \\
+Transl.& \small ... nicht {\"u}berspringen Mahlzeiten ... \\
+Gloss & \small ... don't skip meals ... \\
+\end{tabular}
+\end{center}
+
+The translation is mostly understandable, although the verb ``{\"u}berspringen'' is incorrectly inflected.
+HUME allows marking it as Orange (as it was  indeed annotated in our experiments), while
+HMEANT allows either to mark it as correct, or otherwise mark the entire frame as a ``no match''.
+
+We conducted an analysis of the English UCCA Wikipedia corpus (5324 sentences) in order to assess the
+pervasiveness of three phenomena that are not well supported by HMEANT.\footnote{Argument structures and
+linkers are explicitly marked in UCCA. Non-auxiliary instances of ``be'' and nouns are identified using
+the NLTK standard tagger. Nominal argument structures are here Scenes whose main relation is headed by a noun.}
+First, copula clauses are treated in HMEANT simply as instances of the main verb ``be'', which generally does not convey the meaning of these clauses.
 They appear in 21.7\% of the sentences, according to conservative estimates that only consider non-auxiliary instances of ``be''.

### 2016-09-15 paper/main.tex
misc
sss

@@
 \author{
-  Alexandra Birch\textsuperscript{1}, Omri Abend\textsuperscript{2}, Ond{\v{r}}ej Bojar\textsuperscript{3},
+  Alexandra Birch\textsuperscript{1}\thanks{$^*$ All authors contributed equally to this work.} , Omri Abend\textsuperscript{2*}, Ond{\v{r}}ej Bojar\textsuperscript{3*},
   Barry Haddow\textsuperscript{1*}\\
@@
 
-\let\thefootnote\relax\footnotetext{*All authors contributed equally to this work.}
+%\let\thefootnote\relax\footnotetext{*All authors contributed equally to this work.}
 
@@
 
+\com{
 In other cases, HMEANT's does not to penalize errors. For instance, as HMEANT is defined as an aggregate
@@
 HMEANT allows either to mark it as correct, or otherwise mark the entire frame as a ``no match''.
+}
 

### 2016-09-19 paper/main.tex
repository

@@
 
+All annotation data gathered in this project, together with analysis scripts is available
+in github\footnote{\url{https://github.com/bhaddow/hume-emnlp16}}.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section*{Acknowledgments}
-{T}his work was conducted within the scope of the \ProjectType, which has received funding from the European Union's Horizon 2020 research and innovation programme under grant agreement No \GrantNo (\ProjectName).
+{T}his work was conducted within the scope of the \ProjectType, which has received funding from the European Union's Horizon 2020 research and innovation programme under grant agreement No \GrantNo\ (\ProjectName).
 

### 2016-09-21 paper/main.tex
Added eg HUME vs DA

@@
 
+\begin{figure*}[t]
+\begin{tabular}{c}
+\includegraphics[width=1\textwidth]{tree_653_de1_de2.png}\\
+\includegraphics[width=0.6\textwidth]{DAline-extract.png}
+\end{tabular}
+\caption{An example of a doubly annotated HUME en-de sentence.
+Structural nodes which are `A' are shown in yellow, and `B' are shown in blue.
+A sample of 10 of the raw DA assesments for the German machine translation are shown below.}
+\label{fig:humeeg}
+\end{figure*}
+%Original sentence ID = 653
+%HUME SID=287
+%RAW DA
+%82.0+36.0+57.0+32.0+74.0+57.0+24.0+95.0+30.0+0.0+70.0+38.0+41.0+92.0+45.0+15.0+59.0+82.0+24.0+87.0+24.0+24.0+86.0+87.0+8.0+36.0+43.0+61.0+41.0+14.0+0.0+15.0+50.0+16.0+26.0+23.0+83.0+39.0+45.0+1.0+80.0+31.0+53.0+12.0+64.0+22.0+46.0+39.0
+%STD DA
+%-0.345604136518,-3.33742540772,0.423428930981,-0.548184035933,-0.677176346874,-1.65622648693,-1.55926607798,-1.1272965301,-0.0171653191891,-1.02340816821,-0.872524664699,-0.966670725408,-0.280564543666,-3.14230662916,-0.0830045886964,-1.47697478054,0.470426332641,-1.1678743283,-0.432290259946,-1.92671695959,0.141372817859,-1.36373703007,-0.179200976557,-0.427598369803
+%
+%54.8 = sample of 10 raw scores
+%-0.230 = sample of 10 std scores
+%Generate a random sample each time: here is one random sample from score_segs.R: 
+%57+64+0+39+83+50+45+53+32+46
+%57+64+0+39+83+50+45+53+32+46
+%36+64+80+41+1+83+46+50+39+95=535 -> show this one as is closest to 54.8
+
 %\subsection{Discussion of Disagreements}\label{sec:disagreements}
@@
 We have just presented a quantitative comparison between DA and HUME. DA is however, a very
-different type of metric to HUME. It returns just one score per sentence and requires annotators to make 
-decisions collating many smaller judgements about individual errors in the sentence. 
-Comparing DA and HUME is important in order to see where they agree and differ in order to
+different type of metric to HUME. Looking at Figure~\ref{fig:humeeg} we can see an example of
+a German translation annotated with both HUME and DA. DA standardises and averages scores across annotators and returns one score per sentence. 
+HUME allows us to interogate the quality of individual types of semantic components in a translation, and in
+this sense it is more similar to HMEANT.
+Comparing DA and HUME is important in order to see where they agree and to
 determine under what circumstances it would be appropriate to apply each method.
-In cases where we specifically want to know about the quality of semantic
-components in a translation, HUME and HMEANT could be considered.
 

### 2016-09-22 paper/main.tex
small changes

@@
 The use of UCCA for semantic MT evaluation measure is motivated by two main reasons. 
-First, UCCA's foundational layer can
-be reliably annotated by non-experts after as little as two hours
-of training \parcite{marinotti2014}.
+First, UCCA's foundational layer can be annotated by non-experts after a short training
+\parcite{abend2013universal,marinotti2014}.
 Second, UCCA is cross-linguistically applicable, seeking to
@@
 Semantically accurate translation is paramount in this domain, which makes
-it particularly suitable for semantic  MT evaluation. HUME is both evaluated in terms of its
-consistency (inter-annotator agreement), efficiency (time of annotation) and validity (through a comparison
+it particularly suitable for semantic  MT evaluation. HUME is evaluated in terms of its
+consistency (inter-annotator agreement), efficiency (time of annotation) and validity (by comparing it
 with crowd-sourced adequacy judgements).
@@
 \begin{tabular}{c}
-\includegraphics[width=1\textwidth]{tree_653_de1_de2.png}\\
-\includegraphics[width=0.6\textwidth]{DAline-extract.png}
+\includegraphics[width=1\textwidth]{tree_653_ver2.png}\\
+%\includegraphics[width=0.6\textwidth]{DAline-extract.png}
 \end{tabular}
-\caption{An example of a doubly annotated HUME en-de sentence.
-Structural nodes which are `A' are shown in yellow, and `B' are shown in blue.
-A sample of 10 of the raw DA assesments for the German machine translation are shown below.}
+\caption{An example of a doubly annotated HUME en-de sentence. The color of the nodes correspond to their
+  HUME label: yellow (A), blue (B), and atomic nodes with their traffic. Nodes in which the annotators
+  disagree are split in color. The scale at the bottom right shows a sample of 10 raw DA assessments
+  for this example. }
 \label{fig:humeeg}
@@
 
-We have just presented a quantitative comparison between DA and HUME. DA is however, a very
-different type of metric to HUME. Looking at Figure~\ref{fig:humeeg} we can see an example of
-a German translation annotated with both HUME and DA. DA standardises and averages scores across annotators and returns one score per sentence. 
-HUME allows us to interogate the quality of individual types of semantic components in a translation, and in
-this sense it is more similar to HMEANT.
-Comparing DA and HUME is important in order to see where they agree and to
-determine under what circumstances it would be appropriate to apply each method.
+While a quantitative comparison between DA and HUME, as we have just presented, is telling of their
+differences and commonalities, it is important to note that the two measures are fundamentally different.
+Figure~\ref{fig:humeeg} presents an example of a doubly HUME-annotated English-German example
+accompanied by 10 raw DA scores assigned to that sentence. 
+While DA standardises and averages scores across annotators to denoise the raw data, and obtain
+a single aggregate score, HUME decomposes over a combinatorial structure, thus allowing
+to localize the translation errors.
+%Comparing DA and HUME is important in order to see where they agree and to
+%determine under what circumstances it would be appropriate to apply each method.
 
@@
 in terms of their SRL annotations.
-In this section we present a qualitative comparison between HUME and HMEANT
-and the discussion is accompanied by examples from our
-experimental data.
+In this section we present a qualitative comparison between HUME and HMEANT,
+using examples from our experimental data.
 
@@
 like BLEU, in not requiring a reference.
-Instead, it compares the source directly with the output translation.
+Instead, it directly compares the source and the translation.
 This requires the employment of bilingual annotators, but has the benefit of avoiding
@@
 All annotation data gathered in this project, together with analysis scripts is available
-in github\footnote{\url{https://github.com/bhaddow/hume-emnlp16}}.
+online\footnote{\url{https://github.com/bhaddow/hume-emnlp16}}.
 

### 2016-09-22 paper/main.tex
minor

@@
 Our experiments with four language pairs: English to Czech, German, Polish and Romanian (\secref{sec:experiments}) document HUME's inter-annotator agreement and efficiency (time of annotation).
-We further empirically compare HUME with direct assessment of human adequacy ratings, and conclude by discussing the differences with HMEANT (\secref{sec:hmeant_comp}).
+We further empirically compare HUME with direct assessment of human adequacy ratings (\secref{sec:adequacy}), and conclude by discussing the differences with HMEANT (\secref{sec:hmeant_comp}).
 
@@
 adequacy/fluency scores to each translation, a procedure recently improved
-by \perscite{graham2015accurate} and called direct assessment. We use this method to compare and contrast
+by \perscite{graham2015accurate} under the title Direct Assessment. We use this method to compare and contrast
 with HUME later in the paper. HTER~\parcite{snover2006study} is another widely used human evaluation metric
@@
 post-edition. HTER suffers from the problem that small edits in the translation
-could in fact be serious flaws in accuracy, eg. deleting a negation. 
-Some manual measures ask annotators to explicitly mark errors, which however
-suffers from even lower agreement than ranking \parcite{lommel:etal:mqm-iaa:2014}.
+could in fact be serious flaws in accuracy, e.g., deleting a negation. 
+Some manual measures ask annotators to explicitly mark errors, but this has
+been found to have even lower agreement than ranking \parcite{lommel:etal:mqm-iaa:2014}.
 
@@
 using dependency grammar, and
-\perscite{owczarzak2007evaluating} took a similar approach using lexical-functional grammar structures.
+\perscite{owczarzak2007evaluating} took a similar approach using Lexical Functional Grammar structures.
 \perscite{gimenez2007linguistic} proposed to combine multiple types of information,
@@
 \paragraph{Semantic Representation.}
-UCCA (Universal Conceptual Cognitive Annotation, \inparcite{abend2013universal}) is a
+UCCA (Universal Conceptual Cognitive Annotation) \cite{abend2013universal} is a
 cross-linguistically applicable scheme for semantic annotation.
@@
 whose leaves correspond to the words of the text.
-The graph's nodes, called {\sc units}, are either terminals or several elements jointly
+The graph's nodes, called {\it units}, are either terminals or several elements jointly
 viewed as a single entity according to some semantic or cognitive consideration. Edges bear
@@
 focuses on argument structures (adjectival, nominal, verbal and others)
-and relations between them. The most basic notion is the Scene, which describes a movement, an
+and relations between them. The most basic notion is the {\it Scene}, which describes a movement, an
 action or a state which persists in time. Each Scene contains one main
@@
 % UCCA Motivation
-The use of UCCA for semantic MT evaluation measure is motivated by two main reasons. 
+The use of UCCA for semantic MT evaluation has several motivations. 
 First, UCCA's foundational layer can be annotated by non-experts after a short training
@@
 
-The Abstract Meaning Representation (AMR) \parcite{banarescu2013abstract} project
+The Abstract Meaning Representation (AMR) \parcite{banarescu2013abstract}
 shares UCCA's motivation for defining a more complete semantic annotation.
@@
 units as it does not anchor its semantic symbols in the text,
-and thus does not provide clear decomposition of the sentence into sub-units.
+and thus does not provide clear decomposition of the sentence into sub-spans.
 Also, AMR is more fine-grained than UCCA and consequently harder to annotate.
@@
 and requires proficiency in the source language only.
-HUME annotation requires the employment of bilingual annotators
-\footnote{Where bilingual annotators are not available,
+HUME annotation requires the employment of bilingual
+annotators.\footnote{Where bilingual annotators are not available,
   the evaluation could be based on the UCCA structure for the
@@
 We begin by creating UCCA annotations for the source sentence, following the
-UCCA guidelines\footnote{All UCCA-related resources can be found
-  here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}.
+UCCA guidelines.\footnote{All UCCA-related resources can be found
+  here: \url{http://www.cs.huji.ac.il/~oabend/ucca.html}}
 A UCCA annotation for a sentence $s$ is a labeled DAG $G$, whose leaves
@@
 Annotation of the semantic units requires first deciding whether
-a unit is {\it structural}, i.e., has meaning-bearing sub-units also in the
-target language, or {\it atomic}.
+a unit is {\it structural}, i.e., has meaning-bearing
+sub-units in the target language, or {\it atomic}.
 In most cases, atomic units
@@
   \caption{\label{fig:hume_tree_v2}
-     HUME annotation of an UCCA tree with a word aligned example translation shown below. 
+     HUME annotation of an UCCA tree with a word-aligned example translation shown below. 
      Atomic units are labelled using traffic lights (Red, Orange, Green) and structural
@@
 which changes the meaning of the entire sentence. This error is captured at
-the atomic level, and it is labelled Red. The sentence still contains two scenes and
-a linker and therefore we mark the root node as structurally correct, Adequate.
+the atomic level, and it is labelled Red. The sentence still contains two Scenes and
+a Linker and therefore we mark the root node as structurally correct, Adequate.
 
@@
 This allows the user, especially in long sentences, to focus her attention on the parts
-most likely to be relevant for her judgement. As the alignments are automatically derived,
+that are most likely to be relevant for her judgement. As the alignments are automatically derived,
 and therefore noisy, the annotator is instructed to treat the aligned text is a cue, but to ignore
@@
 medical translation task \parcite{bojar-EtAl:2014:W14-33}, 
-giving between 45 and 85 million sentences of training data, depending on language pair.
+giving between 45 and 85 million sentences of training data, depending on the language pair.
 %\oa{I think we should probably
@@
 For the annotation of the MT output, we recruited two annotators for each of German, Romanian
-and Polish and one main annotator for Czech. For Czech IAA,
+and Polish and one main annotator for Czech. For computing Czech IAA,
 several further annotators worked on a small number of 
@@
 The annotators were all native speakers of the respective target languages and fluent in English.
-They completed a three hour on-line training session where we described UCCA and the HUME task and then
-we worked through a number of examples.
+They completed a three hour on-line training session which included a description of
+UCCA and the HUME task, followed by walking through a few examples.
 
@@
 To estimate the annotation time, we compute the time difference between successive 
-sentences, and discard outlying times since we assume annotation was not continuous.
+sentences, and discard outlying times, assuming annotation was not continuous in these cases.
 From inspection of histograms of annotation times, we set the upper threshold at 500 seconds.
@@
 is generally little confusion between the two types of units.
-This results in the Kappa for all units being considerably higher than the kappa
+This results in the Kappa for all units being considerably higher than the Kappa
 over the atomic units or structural units, where there is more internal confusion.
@@
 
-%presents the variance of kappa with length in each bin covers
+%presents the variance of Kappa with length in each bin covers
 %a length range of 10. and we show .
@@
 no discernible reduction of IAA with sentence
-length. Also, from \tabref{tab:iaa} the overall IAA
-is similar for all languages, showing good agreement (0.6--0.7).
+length. \tabref{tab:iaa} also shows that the overall IAA
+is similar for all languages, presenting good agreement (0.6--0.7).
 However, there are differences observed when we break down by node type.
@@
 the German MT system normally translated ``review'' (as in ``systematic review'' -- a frequent term in the 
-Cochrane texts) as ``\"uberpr\"ufung'', which 
+Cochrane texts) as ``\"Uberpr\"ufung'', which 
 one annotator marked correct, and the other (a Cochrane employee) as incorrect.
 
-\subsection{Comparison with Direct Assessment}\label{sec:adequacy}
+\section{Comparison with Direct Assessment}\label{sec:adequacy}
 
@@
 This confirms that HUME is consistent with an accepted human evaluation method,
-despite the differences in their conception.
+despite their conceptual differences.
 While DA is a valuable tool, HUME has two advantages:
@@
 Center nodes (C) are also more correlated than elaborator nodes (E), which is expected given that Centers are defined to be more semantically dominant.
-Future work will construct an aggregate HUME score which weights the different node types according to their semantic relevance.
+Future work will construct an aggregate HUME score which weights the different node types according to their semantic prominence.
 
@@
 
-While a quantitative comparison between DA and HUME, as we have just presented, is telling of their
-differences and commonalities, it is important to note that the two measures are fundamentally different.
+%While a quantitative comparison between DA and HUME, as we have just presented, is telling of their
+%differences and commonalities, it is important to note that the two measures are fundamentally different.
 Figure~\ref{fig:humeeg} presents an example of a doubly HUME-annotated English-German example
 accompanied by 10 raw DA scores assigned to that sentence.
-While DA standardises and averages scores across annotators to denoise the raw data, and obtain
-a single aggregate score, HUME decomposes over a combinatorial structure, thus allowing
-to localize the translation errors.
+The figure illustrates the conceptual difference between the measures:
+while DA standardises and averages scores across annotators to denoise the
+crowd-sourced raw data, thus obtaining a single aggregate score,
+HUME decomposes over a combinatorial structure, thus allowing to
+localize the translation errors. We turn to comparing HUME to a more
+conceptually-related  measure, namely HMEANT.
+
 %Comparing DA and HUME is important in order to see where they agree and to
@@
 
-HMEANT focuses on verbal argument structures, ignoring other pervasive phenomena such as non-verbal predicates and inter-clause relations. Consider the following example:
+HMEANT focuses on verbal argument structures, ignoring other pervasive phenomena such as non-verbal predicates and inter-clausal relations. Consider the following example:
 
@@
 
-The German translation is largely correct, although the English verb ``tend''
-was translated into the German prepositional phrase ``in der Regel'' (``as a rule'').
+The German translation is largely correct, despite the grammatical divergence,
+namely that the English verb ``tend''
+is translated into the German prepositional phrase ``in der Regel'' (``as a rule'').
 HMEANT will consider the translation to be of poor quality as there is no German verb
-to align the English one with.
+to align with the English one.
 
@@
 linkers are explicitly marked in UCCA. Non-auxiliary instances of ``be'' and nouns are identified using
-the NLTK standard tagger. Nominal argument structures are here Scenes whose main relation is headed by a noun.}
+the NLTK standard tagger. Nominal argument structures are here Scenes whose Main Relation is headed by a noun.}
 First, copula clauses are treated in HMEANT simply as instances of the main verb ``be'', which generally does not convey the meaning of these clauses.
@@
 
-All annotation data gathered in this project, together with analysis scripts is available
+All annotation data gathered in this project, together with analysis scripts, is available
 online\footnote{\url{https://github.com/bhaddow/hume-emnlp16}}.
@@
 a word in the sentence at the bottom. A scene must contain a process (P) or a state (S). It can also contain
-participants (A) and it can be linked to other scenes by a linker (Linker). Participants, processes and states can be
+participants (A) and it can be linked to other scenes by a Linker (Linker). Participants, processes and states can be
 further analysed into elaborators (E), centres (C) and relators (R). These labels are very 
@@
 a minimal set of labels. It provides a complete semantic tree which does not rely on syntactic heads and the semantic structure is grounded directly to the words in the sentence. 
- Even though the set of UCCA labels are fairly restricted,  nevertheless they allow us to determine the most important components of the graph (for example it defines centres and linkers which would be likely to carry more weight than elaborators), and we can use this to better calculate the score.
+ Even though the set of UCCA labels are fairly restricted,  nevertheless they allow us to determine the most important components of the graph (for example it defines centres and Linkers which would be likely to carry more weight than elaborators), and we can use this to better calculate the score.
 We think that UCCA is the

### 2016-09-23 paper/main.tex
minor corrections, vectored diagram

@@
 \usepackage[T1]{fontenc}
+\usepackage{tikz}
 \usepackage{url}
@@
 %\usepackage{hyperref}
+\usetikzlibrary{positioning}
 
@@
 
-Human judgement is the cornerstone for estimating the quality of an MT system.
+%Human judgement is the cornerstone for estimating the quality of an MT system.
+Human judgement should be the ultimate test of the quality of an MT system.
 Nevertheless, common measures for human MT evaluation, such as adequacy and fluency judgements
@@
 We propose the HUME metric,
-a human evaluation measure that decomposes over the UCCA semantic units.
+a human evaluation measure that decomposes over UCCA semantic units.
 % of the sentence.
@@
 adequacy/fluency scores to each translation, a procedure recently improved
-by \perscite{graham2015accurate} under the title Direct Assessment. We use this method to compare and contrast
+by \perscite{graham2015accurate} under the title Direct Assessment. We use this latter method to compare and
+contrast
 with HUME later in the paper. HTER~\parcite{snover2006study} is another widely used human evaluation metric
@@
 \begin{figure*}[t]
-\begin{tabular}{c}
-\includegraphics[width=1\textwidth]{tree_653_ver2.png}\\
+\begin{tikzpicture}
+\node(tree) [] {\includegraphics[width=1\textwidth]{tree_653_de1_de2.pdf}};
+\node [below right = -2cm and  -8cm of tree]{\includegraphics[width=0.4\textwidth]{DAline-extract.png}};
+\end{tikzpicture}
+%\begin{tabular}{c}
+%\includegraphics[width=1\textwidth]{tree_653_ver2.png}\\
 %\includegraphics[width=0.6\textwidth]{DAline-extract.png}
-\end{tabular}
+%\end{tabular}
 \caption{An example of a doubly annotated HUME en-de sentence. The color of the nodes correspond to their
@@
 HUME decomposes over a combinatorial structure, thus allowing to
-localize the translation errors. We turn to comparing HUME to a more
+localize the translation errors. We now turn to comparing HUME to a more
 conceptually-related  measure, namely HMEANT.
@@
 \section*{Acknowledgments}
-{T}his work was conducted within the scope of the \ProjectType, which has received funding from the European Union's Horizon 2020 research and innovation programme under grant agreement No \GrantNo\ (\ProjectName).
+%{T}his work was conducted within the scope of the \ProjectType, which has received funding from the European Union's Horizon 2020 research and innovation programme under grant agreement No \GrantNo\ (\ProjectName).
+
+This project has received funding from the European Union's Horizon 2020 research and innovation
+programme under grant agreement \GrantNo\ (\ProjectName).
 

### 2016-10-07 paper/main.tex
Fixed typo, added final camera ready to git
switch to emnlp style
Removed eg. still a paragraph over page limit

@@
 %\usepackage[round]{natbib}
-\usepackage{acl2016}
+\usepackage{emnlp2016}
 \usepackage{amsmath}
@@
 
-\aclfinalcopy
+\emnlpfinalcopy
 
@@
 
-\begin{figure*}[t]
-\begin{tikzpicture}
-\node(tree) [] {\includegraphics[width=1\textwidth]{tree_653_de1_de2.pdf}};
-\node [below right = -2cm and  -8cm of tree]{\includegraphics[width=0.4\textwidth]{DAline-extract.png}};
-\end{tikzpicture}
+%\begin{figure*}[t]
+%\begin{tikzpicture}
+%\node(tree) [] {\includegraphics[width=1\textwidth]{tree_653_de1_de2.pdf}};
+%\node [below right = -2cm and  -8cm of tree]{\includegraphics[width=0.4\textwidth]{DAline-extract.png}};
+%\end{tikzpicture}
 %\begin{tabular}{c}
@@
 %\end{tabular}
-\caption{An example of a doubly annotated HUME en-de sentence. The color of the nodes correspond to their
-  HUME label: yellow (A), blue (B), and atomic nodes with their traffic. Nodes in which the annotators
-  disagree are split in color. The scale at the bottom right shows a sample of 10 raw DA assessments
-  for this example. }
-\label{fig:humeeg}
-\end{figure*}
+%\caption{An example of a doubly annotated HUME en-de sentence. The color of the nodes correspond to their
+%  HUME label: yellow (A), blue (B), and atomic nodes with their traffic light color. Nodes in which the annotators
+%  disagree are split in color. The scale at the bottom right shows a sample of 10 raw DA assessments
+%  for this example. }
+%\label{fig:humeeg}
+%\end{figure*}
 %Original sentence ID = 653
@@
 %differences and commonalities, it is important to note that the two measures are fundamentally different.
-Figure~\ref{fig:humeeg} presents an example of a doubly HUME-annotated English-German example
-accompanied by 10 raw DA scores assigned to that sentence.
-The figure illustrates the conceptual difference between the measures:
+%Figure~\ref{fig:humeeg} presents an example of a doubly HUME-annotated English-German example
+%accompanied by 10 raw DA scores assigned to that sentence.
+%The figure illustrates the conceptual difference between the measures:
+HUME and DA are conceputally very different metrics:
 while DA standardises and averages scores across annotators to denoise the
